### 2021-04-22 ms.Rmd
Setting up repo

---
title: "Main Manuscript"
author: Whiting et al. 
date: "`r Sys.Date()`"
bibliography: refs.bib
csl: science.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: true
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
  ## and rounded to 2 digits      ## and rounded to 2 digits
  options(digits = 2)
```

```{r getlibrary, echo = FALSE, results = "hide"}
# load libraries
devtools::install_github("karthik/wesanderson")
pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, ggarrange)
```

```{r displayModels, echo = FALSE, results = "hide"}
#Load data and clean a bit
Display_1 <- read.table("./data/Display_1.csv", 
                          header=TRUE, sep=",", dec=".", strip.white=TRUE)

#exclude bottom flank.
Display_1 <- Display_1 %>% 
              filter(!BodyRegion == "botflank") %>% as.data.frame()

# create as factor
    Display_1$BodyRegion <- factor(Display_1$BodyRegion)

```

We conducted experiments manipulating both the social and predatory context chameleons encountered in their native Kenyan range, and the introduced Hawaiian range.

### 2021-04-22 supp.Rmd
Setting up repo

---
title: "Supplemental Materials, Methods and Results"
author: Whiting et al. 
date: "`r Sys.Date()`"
bibliography: refs.bib
csl: science.csl
output: 
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: true
    reference_docx: template.docx
editor_options: 
  chunk_output_type: console
---

```{r display}
#Load data and clean a bit
Display_1 <- read.table("../data/Display_1.csv", 
                          header=TRUE, sep=",", dec=".", strip.white=TRUE)

#exclude bottom flank.
Display_1 <- Display_1 %>% 
              filter(!BodyRegion == "botflank") %>% as.data.frame()

# create as factor
    Display_1$BodyRegion <- factor(Display_1$BodyRegion)

# Fit the model that esatimtes changes across body regions
ConspicDisplayAC <- lmer(dL ~ SVL + Population + BodyRegion + Background 
                         + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
ConspicDisplayAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Display_1)
summary(glht(ConspicDisplayAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
                                                          "Hawaii.tailbase - Kenya.tailbase = 0" ))))
```

### 2021-04-22 R/ChamModels_2.0.R

@@ -1,0 +1,724 @@
+# load libraries
+devtools::install_github("karthik/wesanderson")
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson)
+
+
+### CONSPICUOUSNESS OF DISPLAY MODELS
+
+# Load Dataset - Display_1
+setwd("~/Dropbox/chameleon/chemeleon_signal_release")
+Display_1 <- read.table("./data/Display_1.csv", 
+                          header=TRUE, sep=",", dec=".", strip.white=TRUE)
+names(Display_1) # header rows R is reading
+Display_1 <- Display_1 %>% filter(!BodyRegion == "botflank") %>% as.data.frame() #exclude bottom flank.
+Display_1$BodyRegion <- factor(Display_1$BodyRegion) #updates levels for the variable BodyRegion
+str(Display_1) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
+
+## Display - Achromatic contrast (AC or dL)
+## just this next section has Drew's code, which averages the backround (Leaf_B1, Leaf_B2, Leaf_T1, Leaf_T2, Stem1, Stem2) for each lizard body region measurement.
+## This is the full model; however, we do not need background because the way the data is set up R is taking an average. See next model (background dropped)
+ConspicDisplayAC <- lmer(dL ~ SVL + Population + BodyRegion + Background 
+                         + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
+Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
+ConspicDisplayAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Display_1)
+summary(glht(ConspicDisplayAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
+                                                          "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+
+## dropping background from model (it is still averaged)
+## homogeneity of slopes test for SVl-dL by population. To meet test assumptions slopes for each population must be NS.
+## assumptions are met
+ConspicDisplayAC <- lmer(dL ~ SVL + Population + BodyRegion +  
+                           + Population:BodyRegion + (1|ID), data=Display_1)
+anova(ConspicDisplayAC)
+
[…]

### 2021-04-22 ms.Rmd
Updating script for the MS and supple

@@
 ---
-title: "Main Manuscript"
+title: "Evolution of a conspicuous dynamic visual signal in an introduced chameleon"
 author: Whiting et al. 
@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, ggarrange)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp)
+
+p_value <- function(x){
+  if(x <= 0.0001) {tmp = "< 0.0001"}
+  if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
+  if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
+  if(x >= 0.01) {tmp =x}
+  return(tmp)
+}
+
 ```
@@
 ```{r displayModels, echo = FALSE, results = "hide"}
+#########################################################
+# Male-male Competition Displays
+#########################################################
 #Load data and clean a bit
@@
 
+# Display dl
+    ConspicDisplayAC <- lmer(dL ~ SVL + Population + (1|ID), data=Display_1)
+    summary(ConspicDisplayAC)
+    anovaConspicDisplayAC <- anova(ConspicDisplayAC)
+
+## Display dS
+         ConspicDisplayCC <- lmer(dS ~ SVL + Population + (1|ID), data=Display_1)
+    anovaConspicDisplayCC <- anova(ConspicDisplayCC)
+
+#########################################################
+# Courtship Displays
+#########################################################
+
+Courtship <- read.table("./data/Courtship_1.2.csv", 
+                          header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Courtship dl
+ConspicCourtshipAC <- lmer(dL ~ SVL + Population  + (1|ID), data=Courtship)
+courship_anovaAC <- anova(ConspicCourtshipAC)
+
+# Courtship dS
+  ConspicCourtshipCC <- lmer(dS ~ SVL + Population  + (1|ID), data=Courtship)
+courship_anovaCC <- anova(ConspicCourtshipCC)
 ```
 
-We conducted experiments manipulating both the social and predatory context chameleons encountered in their native Kenyan range, and the introduced Hawaiian range.
+# Abstract
+Rapid evolution frequently follows a significant change to an organism’s environment, such as during colonization of a novel habitat.
+Vertebrates, including Caribbean Anolis lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales.
+Traits that are sexually selected can be similarly affected, although well-documented examples from wild populations are rare.
+The Jackson’s chameleon (*Chamaeleo x.
+jacksonii*) from East Africa was inadvertently introduced to the Hawaiian island of Oahu in 1972.
+In its native range, Jackson’s chameleons are preyed upon by a wide range of predators, including snakes and birds, which are thought to shape the dynamics of color change in chameleons.
+We took advantage of the Hawaiian introduction, an environment with only a few potential predators and therefore, very low predation pressure, to test for character release of conspicuous social color signals.
+To test this hypothesis, we measured color change in response to multiple social and predator stimuli in wild chameleons in both Hawai’i and Kenya.
+We used visual modelling of colour signals to estimate signal conspicuousness to both the chameleon and predator (bird, snake) visual systems.
+The Hawai’i and Kenya backgrounds were a similar color, but the major difference was in brightness: the Hawai’i background was darker.
+As a result, chameleons from both Kenya and Hawai’i were brighter against the Hawai’i background.
+However, Hawai’i males were more conspicuous (courtship, display) against their own background for some body regions (mainly gular), providing evidence for character release.
+Also, Hawai’i males were more conspicuous (courtship, display) to snake and bird predators against their own background.
+We suggest that relaxing natural selection can result in signal release in systems experiencing strong sexual selection.
+
+# Introduction
+
+When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods.
+This view understandably persisted for at least another 120 years following the publication of On the Origin of Species (Darwin, 1859).
+We now have numerous examples of evolutionary change following natural selection in both the laboratory and wild, and not only in invertebrates with short generation times.
+Vertebrates, including Caribbean Anolis lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales (e.g.
+@losos97).
+Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in phenotype over short time intervals.
+
+Conspicuous colour signals evolve because of the fitness advantage conferred to the signaller, and frequently becomes increasingly elaborate through a combination of male contest competition and female preference (e.g.
+runaway selection).
+However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly.
+Evolution can respond to this selection pressure in a variety of ways that include shifting conspicuous signals to areas of the body that might be concealed to predators.
+For example, many conspicuous lizards have colourful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards.
+Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-dense streams.
+These examples are from animals with static colour signals.
+A parallel solution that reduces the risk of exposure to predators is dynamic colour change because signals can be switched on and off relatively rapidly through neural or endocrine control.
+A wide range of invertebrates (e.g.
+cephalopods) and vertebrates (e.g.
+lizards and frogs) have convergently evolved this ability.
+
+Unlike animals with static colour signals, we have a poor understanding of how animals with dynamic colour change respond to predation pressure and whether relaxing predation pressure will result in an increasingly conspicuous signal.
+Tackling this question in the wild is particularly difficult unless we can take advantage of unplanned changes to an organism’s environment such as during colonization of a novel habitat.
+Living organisms are occasionally translocated great distances by human action to environments far removed from their native range.
+These novel environments typically have a completely different suite of organisms, including predators, and offer a unique opportunity to study signal evolution in founder populations.
+Jackson’s chameleons (*Chamaeleo j.
+xanthopholus*) are just such a case.
+In 1972, a shipment of chameleons was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
+The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island.
+On Oahu, the number of potential predators of chameleons is few (see methods), while in Kenya they are preyed upon by a wide range bird, snake and occasionally, mammal, predators.
+This accidental introduction represents a unique opportunity to study the evolution of a dynamic visual signal by comparing native (Kenya) and introduced (Oahu) chameleons.
+
+# Results
+
+We conducted experiments manipulating both the social and predatory context chameleons encountered, in both their native Kenyan range and their introduced Hawaiian range, to test three hypotheses about colour signal evolution.
+First, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal.
+Sexual selection favours more conspicious signals and these should evolve to be more conspicuous without predators.
+In suppor of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (**Male-Male Contests**: Acromatic ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; **Courtship displays**: Acromatic ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was true across all body regions measured (See Supplemental Materials, Table S1).
+
+Specifically, we tested the following predictions: 1.) display colouration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; 2.) display colouration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawai’i than Kenya.
+3.) under character release, males will have greater capacity to change colour from a subordinate to a dominant state in Hawai’i compared to Kenya.
+
+We also tested the hypothesis that sensory drive and local adaptation are responsible for driving display coloration.
+Specifically, we predicted that chameleon display colouration in each population would be more conspicuous within its own signaling environment compared to the other population’s signaling environment.
+For example, Hawaiian chameleons were historically found in Kenya but under sensory drive and local adaptation they are predicted to be more conspicuous in their current signaling environment, Hawai’i.
+Likewise, Kenyan chameleons have no history in Hawai’i and are therefore predicted to be more conspicuous within their current signaling environment, Kenya.

### 2021-04-22 supp.Rmd
Updating script for the MS and supple

@@
 output: 
+  bookdown::word_document2:
+    toc: no
+    toc_depth: 6
+    number_sections: true
+    reference_docx: template.docx
   bookdown::html_document2:
@@
     bibliography: refs.bib
-  bookdown::word_document2:
-    toc: no
-    toc_depth: 6
-    number_sections: true
-    reference_docx: template.docx
 editor_options: 
@@
 ---
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
+## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
+  ## and rounded to 2 digits      ## and rounded to 2 digits
+  options(digits = 2)
+```
 
-```{r display}
+```{r getlibrary, echo = FALSE, results = "hide"}
+# load libraries
+devtools::install_github("karthik/wesanderson")
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, pander, kableExtra, flextable)
+```
+
+```{r display, results = "hide"}
+
+#########################################################
+# Male-male Competition Displays
+#########################################################
 #Load data and clean a bit
-Display_1 <- read.table("../data/Display_1.csv", 
+Display_1 <- read.table("./data/Display_1.csv", 
                           header=TRUE, sep=",", dec=".", strip.white=TRUE)
@@
                          + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
+
+# Fit model for linear hypothesis tests
 Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
 ConspicDisplayAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Display_1)
-summary(glht(ConspicDisplayAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
+
+result1 = glht(ConspicDisplayAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
+                                                            "Hawaii.midflank - Kenya.midflank = 0",     
+                                                            "Hawaii.topflank - Kenya.topflank = 0", 
+                                                             "Hawaii.tailbase - Kenya.tailbase = 0" )))
+
+## Display - Chromatic contrast (CC or dS)
+## just this next section has Drew's code, which averages the backround (Leaf_B1, Leaf_B2, Leaf_T1, Leaf_T2, Stem1, Stem2) for each lizard body region measurement.
+## This is the full model; however, we do not need background because the way the data is set up R is taking an average. See next model (background dropped)
+
+# Fit the model that esatimtes changes across body regions
+  ConspicDisplayCC <- lmer(dS ~ SVL + Population + BodyRegion + Background 
+                         + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
+Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
+
+# Fit model for linear hypothesis tests
+ConspicDisplayCC2 <- lmer(dS ~ PopBRInt + (1|ID), data=Display_1)
+summary(glht(ConspicDisplayCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
+                                                          "Hawaii.midflank - Kenya.midflank = 0", 
+                                                          "Hawaii.topflank - Kenya.topflank = 0", 
                                                           "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+
+#########################################################
+# Courtship Displays
+#########################################################
+
+Courtship <- read.table("./data/Courtship_1.2.csv", 
+                          header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Courtship - Chromatic dl
+ConspicCourtshipAC <- lmer(dL ~ SVL + Population + BodyRegion + Background 
+                           + Population:Background + Population:BodyRegion + (1|ID), data=Courtship)
+
+Courtship$PopBRInt <- interaction(Courtship$Population,Courtship$BodyRegion)
+ConspicCourtshipAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Courtship)
+summary(glht(ConspicCourtshipAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
+                                                            "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+
+ConspicCourtshipAC <- lmer(dL ~ SVL + Population + BodyRegion +  
+                             + Population:BodyRegion + (1|ID), data=Courtship)
+anova(ConspicCourtshipAC)
+ranova(ConspicCourtshipAC)
+
+# Courtship - Chromatic dS
+ConspicCourtshipCC <- lmer(dS ~ SVL + Population + BodyRegion + Background 
+                           + Population:Background + Population:BodyRegion + (1|ID), data=Courtship)
+Courtship$PopBRInt <- interaction(Courtship$Population,Courtship$BodyRegion)
+ConspicCourtshipCC2 <- lmer(dS ~ PopBRInt + (1|ID), data=Courtship)
+summary(glht(ConspicCourtshipCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
+                                                            "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+
 ```
+
+# Materials and Methods
+
+## Study area and collection of chameleons
+
+We collected most chameleons at night by spotlighting, and a small proportion of individuals during the day.
+When chameleons were beyond reach, we used 4-m extensible poles that chameleons could be encouraged to step on to, before lowering them to the ground.
+We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
+Field work was conducted in Hawai’i during January-February 2006 and in Kenya during 5-13 April 2006.
+In Hawai’i, chameleons were collected from a single population in the Manoa Valley, Oahu, accessible along Tantalus Drive at an elevation of ca.
+x-x.
+The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth.
+Most of the plants in this area were nonnative, and chameleons also occurred in hedge-grows and stands of bamboo.
+The area experiences high annual rainfall (mean: xx).
+In Kenya, field work was conducted near the foothills of Mt.
+Kenya, close to the town of Runyenjes.
+The area is partially cleared for agriculture, leaving small stands of forest.
+The vegetation was a mix of native trees and shrubs interspersed with exotic species.
+
+## Behaviour trials
+
+In order to quantify display colouration we staged encounters between conspecifics (male-male and male-female) and model (bird and snake) predators and measured their subsequent display colouration (details below).
+All chameleons were used in both predator and social trials although in some cases we were unable to quantify all social signals for a particular individual.
+We conducted all conspecific-social trials first, in case anti-predator trials unduly influenced their social interactions.
+Behavioural trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
+We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand.
+This structure was about 1.5 m high and each arm was approximately 60 cm.
+The triangular setup allowed chameleons to display at a comfortable distance and also to escape from aggressive interactions, particularly during male-male trials.
+All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
+The signalling environment was therefore the same for all animals.
+
+We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
+This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing colour and behaviour.
+In male-male trials both individuals would adopt a lemon-yellow display coloration (Fig.
+x) and advance towards each other.
+This would be accompanied by head shakes and would lead to horn-locking if males were not obviously asymmetric.
+Once dominance was settled, the subordinate would turn brown and flee.
+In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking.
+The day following social trials we presented each chameleon with a snake and bird model predator (separately).
+Lizards were drawn at random and the order of presentation (snake or bird first) was random.
+Both trials were conducted on the same day but were separated by at least three hours.
+A trial began when a chameleon was removed from its cloth bag and placed on the triangular stand.
+At this point the lizard was in a neutral colour state.
+In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon.
+In the case of the bird trials, the bird was attached to a x cm wooden pole and was ‘flown’ past the chameleon from above and below it.
+Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behaviour by chameleons.
+Once chameleons changed colour, we quantified their spectral reflectance following the same procedure described for social trials.
+We also conducted seven trials with a just a branch with no predator attached, as a control for the model predator.
+The snake model was moulded from a dead boomslang (Dispholidus typus) and painted by a professional artist to resemble a typical green form of this species (previously used in Stuart-Fox…).
+The bird was ….
+
+## Colour measurements, habitat background and irradiance
+
+During all behaviour trials, we measured the colour of each body region (Fig.
+x) once only, because chameleons change colour rapidly and we had to reduce the handling stress to the animal.
+Patches were measured in random sequence to avoid any bias of order.
+When a chameleon had completely changed colour (to a human observer) and was also performing typical behaviour associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 optic spectrophotometer and a PX2 light source.
+We used a custom-made probe-holder which we placed in contact with the animal at a standard angle (45; @RN2) and distance (5 mm), thereby excluding ambient light.
+All measurements were relative to a dark, and a certified 99% white reflectance standard.
+When chameleons started to change colour in response to our handling we stopped measuring colour and allowed them to continue a social trial before continuing to measure spectral reflectance.
+In the case of a predator trial we would once again present them with the predator model until they returned to their anti-predator display.
+This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behaviour following handling.
+
+We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signalling environment.
+This included 1.) the upper surface of leaves; 2.) the underside of leaves; and 3.) branches.
+We also measured the habitat light environment by recording irradiance in the chameleons environment.
+
+## Visual modelling
+
+# Supplemental Results
+
+```{r tableS1}
+summary(result1)
+```
+
+```{FigS1, fig.width = 5, fig.height = 10, fig.caption = "test"}
+
+#Making a point plot with error bars split by body size and population 
+#Compute raw means and standard errors for body part and population for dS and assigned this data into Display_plot_dS_dat
+Display_plot_dS_dat <- Display_1 %>% #data for dS Chromatic constrast
+  group_by(Population, BodyRegion) %>% 
+  summarise(Mean = mean(dS, na.rm = T), 
+            SE = sd(dS)/sqrt(length(unique(ID))),
+            Upper = Mean + 1.96*SE,
+            Lower = Mean - 1.96*SE)
+
+Display_plot_dL_dat <- Display_1 %>% #data for dL Achromatic contrast
+  group_by(Population, BodyRegion) %>% 
+  summarise(Mean = mean(dL, na.rm = T), 
+            SE = sd(dL)/sqrt(length(unique(ID))),
+            Upper = Mean + 1.96*SE,
+            Lower = Mean - 1.96*SE)
+
+#Plotting Chromatic contrast (dS)
+p1 <- ggplot(Display_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
+  geom_point(size = 3) +
+  geom_line() +
+  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab(TeX("Chromatic contrast (JND)")) +
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+#Plotting Achromatic contrast (dL)
+p2 <- ggplot(Display_plot_dL_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
+  geom_point(size = 3) +
+  geom_line() +
+  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab(TeX("Achromatic contrast (JND)")) +
+  theme_bw() +
+  theme(#legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+#Making a point plot with error bars split by body size and population 
+#Compute raw means and standard errors for body part and population for dS and assigned this data into Display_plot_dS_dat
+Courtship_plot_dS_dat <- Courtship %>% #data for dS Chromatic constrast
+  group_by(Population, BodyRegion) %>% 
+  summarise(Mean = mean(dS, na.rm = T), 
+            SE = sd(dS)/sqrt(length(unique(ID))),
+            Upper = Mean + 1.96*SE,
+            Lower = Mean - 1.96*SE)
+
+Courtship_plot_dL_dat <- Courtship %>% #data for dL Achromatic contrast
+  group_by(Population, BodyRegion) %>% 
+  summarise(Mean = mean(dL, na.rm = T), 
+            SE = sd(dL)/sqrt(length(unique(ID))),
+            Upper = Mean + 1.96*SE,
+            Lower = Mean - 1.96*SE)
+
+#Plotting Chromatic contrast (dS)
+p3 <- ggplot(Courtship_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
+  geom_point(size = 3) +
+  geom_line() +
+  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab(TeX("Chromatic contrast (JND)")) +
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+#Plotting Achromatic contrast (dL)
+p4 <- ggplot(Courtship_plot_dL_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
+  geom_point(size = 3) +
+  geom_line() +
+  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab(TeX("Achromatic contrast (JND)")) +
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+p1 + p2 + p3 + p4 + plot_annotation(tag_levels = c('A', 'B', 'C', 'D')) & theme(plot.tag = element_text(face = 'bold', size=20))
+
+```
+
+# References

### 2021-04-22 R/ChamModels_2.0.R

@@ -109,24 +109,24 @@ Display_plot_dS_dat <- Display_1 %>% #data for dS Chromatic constrast
   group_by(Population, BodyRegion) %>% 
   summarise(Mean = mean(dS, na.rm = T), 
             SE = sd(dS)/sqrt(length(unique(ID))),
-            Upper = Mean + SE,
-            Lower = Mean - SE) 
+            Upper = Mean + 1.96*SE,
+            Lower = Mean - 1.96*SE) 
             
 Display_plot_dL_dat <- Display_1 %>% #data for dL Achromatic contrast
   group_by(Population, BodyRegion) %>% 
   summarise(Mean = mean(dL, na.rm = T), 
             SE = sd(dL)/sqrt(length(unique(ID))),
-            Upper = Mean + SE,
-            Lower = Mean - SE) 
+            Upper = Mean + 1.96*SE,
+            Lower = Mean - 1.96*SE) 
 
 #Plotting Chromatic contrast (dS)
-ggplot(Display_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
+p1 <- ggplot(Display_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
   geom_point(size = 3) +
   geom_line() +
   geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
   scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
   xlab("Population") + 
-  ylab("Chromatic contrast") +
+  ylab(TeX("Chromatic contrast ($\\Delta s$)")) +
   theme_bw() +
   theme(#legend.position = "none",
     panel.grid.major = element_blank(), 

### 2021-04-22 ms.Rmd
Adding code now. Figure for social context done

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg)
 
@@
 
+# Center SVL to make intercept interpretable
+Display_1$sc_SVL <- with(Display_1, scale(SVL, scale=FALSE))
+
 #exclude bottom flank.
@@
 # Display dl
-    ConspicDisplayAC <- lmer(dL ~ SVL + Population + (1|ID), data=Display_1)
+    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + (1|ID), data=Display_1)
     summary(ConspicDisplayAC)
@@
 ## Display dS
-         ConspicDisplayCC <- lmer(dS ~ SVL + Population + (1|ID), data=Display_1)
+         ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + (1|ID), data=Display_1)
     anovaConspicDisplayCC <- anova(ConspicDisplayCC)
@@
 
+# Load data
 Courtship <- read.table("./data/Courtship_1.2.csv", 
@@
 
+# Center SVL do intercept is at the mean body size
+Courtship$sc_SVL <- with(Courtship, scale(SVL, scale=FALSE))
+
 # Courtship dl
-ConspicCourtshipAC <- lmer(dL ~ SVL + Population  + (1|ID), data=Courtship)
+ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population  + (1|ID), data=Courtship)
+summary(ConspicCourtshipAC)
 courship_anovaAC <- anova(ConspicCourtshipAC)
@@
 # Courtship dS
-  ConspicCourtshipCC <- lmer(dS ~ SVL + Population  + (1|ID), data=Courtship)
+  ConspicCourtshipCC <- lmer(dS ~ sc_SVL + Population  + (1|ID), data=Courtship)
  courship_anovaCC <- anova(ConspicCourtshipCC)
@@
 
-We conducted experiments manipulating both the social and predatory context chameleons encountered, in both their native Kenyan range and their introduced Hawaiian range, to test three hypotheses about colour signal evolution.
+We manipulated both the social and predatory context chameleons encountered, in both their native Kenyan range and their introduced Hawaiian range, to test three hypotheses about colour signal evolution.
 First, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal.
-Sexual selection favours more conspicious signals and these should evolve to be more conspicuous without predators.
-In suppor of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (**Male-Male Contests**: Acromatic ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; **Courtship displays**: Acromatic ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was true across all body regions measured (See Supplemental Materials, Table S1).
+Sexual selection favours more conspicious signals.
+As such, signals should evolve to be more conspicuous when predators are absent.
+In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (**Male-Male Contests**: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; **Courtship displays**: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, Table S1).
+
+```{r Figure1}
+## Make figures that average across body regions, but display Social Context
+
+courtship_data <- Courtship %>% 
+                  select(ID, Population, Colour, SVL, BodyRegion, Background, dS, dL)
+  display_data <- Display_1 %>%
+                  select(ID, Population, Colour, SVL, BodyRegion, Background, dS, dL)
+combined_data <- rbind(courtship_data, display_data)
+
+# Now calculate the averages for each population anc Colour (i.e., context)
+
+summary_data <- combined_data %>%
+                group_by(Population, Colour) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>%
+                mutate(Colour = if_else(Colour == "Courtship", "Courtship", "Male-male contest"))
+
+males <- readPNG("./photos/male_fight.png", native = TRUE)
+#males <- image_scale(males, "200x200")
+
+females <- readPNG("./photos/courtship.png", native = TRUE)
+#females <- image_scale(females, "200x200")
+
+p1 <- ggplot(summary_data, aes(x = Population, y = Mean_dS, group = Colour, color = Colour)) +
+    ylim(2.5, 6.5) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Chromatic contrast (JND)") +
+  annotate("segment", x = 0.85, xend = 1.85, y = 6.1, yend = 6.1, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 5.7, yend = 5.7, colour = "black") + 
+  annotate("text", x = 0.85+(1.85-0.85)/2,  y = 6.2, label = "NS",  size = 6) + 
+  annotate("text", x = 1.12+(2.12-1.12)/2, y = 5.8, label = "NS", size = 6) + 
+  theme_bw() +
+  labs(colour = "Social Context") +
+  theme(legend.position = c(0.28, 0.08),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+   inset_element(males,
+                left = 0.45,
+                right = 0.90,
+                bottom = 0.58,
+                top = 0.75, ignore_tag = TRUE) + 
+    inset_element(females,
+                left = 0.30,
+                right = 0.75,
+                bottom = 0.15,
+                top = 0.28, ignore_tag = TRUE) 
+      #annotation_raster(males, xmin = 1.2, xmax = 1.9, ymin = 4.8, ymax = 5.3) + 
+      #annotation_raster(females, xmin = 1, xmax = 1.7, ymin = 3.5, ymax = 4)
+
+p2 <- ggplot(summary_data, aes(x = Population, y = Mean_dL, group = Colour, color = Colour)) +
+  ylim(15, 45) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Achromatic contrast (JND)") +
+   annotate("segment", x = 0.90, xend = 1.90, y = 40, yend = 40, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 42, yend = 42, colour = "black") + 
+  annotate("text", x = 0.9+((1.90 - 0.90)/2), y = 41, label = "**",  size = 6) + 
+  annotate("text", x = 1.12+((2.12 - 1.12)/ 2), y = 43, label = "**", size = 6) + 
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+    inset_element(males,
+                left = 0.45,
+                right = 0.90,
+                bottom = 0.60,
+                top = 0.75, ignore_tag = TRUE) + 
+    inset_element(females,
+                left = 0.05,
+                right = 0.50,
+                bottom = 0.25,
+                top = 0.38, ignore_tag = TRUE)
+
+plot_grid(p1, p2, labels = c("A", "B"), label_size = 16,
+  label_fontface = "bold", rel_widths=c(1,1))
+
+```
 

### 2021-04-22 supp.Rmd
Adding code now. Figure for social context done

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, pander, kableExtra, flextable)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, pander, kableExtra, flextable, jpeg)
 ```
@@
 
+# Center SVL to make intercept interpretable
+Display_1$sc_SVL <- with(Display_1, scale(SVL, scale=FALSE))
+
 # Fit the model that esatimtes changes across body regions
-ConspicDisplayAC <- lmer(dL ~ SVL + Population + BodyRegion + Background 
+ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Background 
                          + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
@@
 # Fit the model that esatimtes changes across body regions
-  ConspicDisplayCC <- lmer(dS ~ SVL + Population + BodyRegion + Background 
+  ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Background 
                          + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
@@
 
+# Center SVL to make intercept interpretable
+Courtship$sc_SVL <- with(Courtship, scale(SVL, scale=FALSE))
+
+# Fit model
+ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population + BodyRegion +  
+                             + Population:BodyRegion + (1|ID), data=Courtship)
+anova(ConspicCourtshipAC)
+
+ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + (1|ID), data=Courtship)
+summary(ConspicCourtshipAC)
+anova(ConspicCourtshipAC)
+
 # Courtship - Chromatic dl
-ConspicCourtshipAC <- lmer(dL ~ SVL + Population + BodyRegion + Background 
-                           + Population:Background + Population:BodyRegion + (1|ID), data=Courtship)
 
@@
 ConspicCourtshipAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Courtship)
-summary(glht(ConspicCourtshipAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
+summary(glht(ConspicCourtshipAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
+                                                            "Hawaii.midflank - Kenya.midflank = 0", 
+                                                            "Hawaii.topflank - Kenya.topflank = 0", 
                                                             "Hawaii.tailbase - Kenya.tailbase = 0" ))))
-
-ConspicCourtshipAC <- lmer(dL ~ SVL + Population + BodyRegion +  
-                             + Population:BodyRegion + (1|ID), data=Courtship)
-anova(ConspicCourtshipAC)
-ranova(ConspicCourtshipAC)
 
@@
 ConspicCourtshipCC2 <- lmer(dS ~ PopBRInt + (1|ID), data=Courtship)
-summary(glht(ConspicCourtshipCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
+summary(glht(ConspicCourtshipCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
+                                                            "Hawaii.midflank - Kenya.midflank = 0", 
+                                                            "Hawaii.topflank - Kenya.topflank = 0", 
                                                             "Hawaii.tailbase - Kenya.tailbase = 0" ))))
@@
 
-```{FigS1, fig.width = 5, fig.height = 10, fig.caption = "test"}
+```{r FigS1, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Chormatic (A, C) and Achromatic (B, D) contrast in just noticible differences (JND) for Hawaiian and Kenyan lizards relative to their environments.
+All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
+Mean JNDs are provided along with 95% confidence intervals.
+A & B represent male-male contest displays whereas C & D represent courtship displays"}
 
@@
   group_by(Population, BodyRegion) %>% 
-  summarise(Mean = mean(dS, na.rm = T), 
+  summarise( Mean = mean(dS, na.rm = T), 
                SE = sd(dS)/sqrt(length(unique(ID))),
@@
 p1 <- ggplot(Display_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3) +
-  geom_line() +
-  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
+  geom_point(size = 3, position = position_dodge(width=0.2)) +
+  geom_line(position = position_dodge(width=0.2)) +
+  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
   scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
   xlab("") + 
-  ylab(TeX("Chromatic contrast (JND)")) +
+  ylab("Chromatic contrast (JND)") +
   theme_bw() +
@@
 p2 <- ggplot(Display_plot_dL_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3) +
-  geom_line() +
-  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
+  geom_point(size = 3, position = position_dodge(width=0.2)) +
+  geom_line(position = position_dodge(width=0.2)) +
+  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
   scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
   xlab("") + 
-  ylab(TeX("Achromatic contrast (JND)")) +
+  ylab("Achromatic contrast (JND)") +
   theme_bw() +
+  labs(color = "Body Region") + 
   theme(#legend.position = "none",
@@
 p3 <- ggplot(Courtship_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3) +
-  geom_line() +
-  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
+  geom_point(size = 3, position = position_dodge(width=0.2)) +
+  geom_line(position = position_dodge(width=0.2)) +
+  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
   scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
   xlab("") + 
-  ylab(TeX("Chromatic contrast (JND)")) +
+  ylab("Chromatic contrast (JND)") +
   theme_bw() +
@@
 p4 <- ggplot(Courtship_plot_dL_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3) +
-  geom_line() +
-  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0) + 
+  geom_point(size = 3, position = position_dodge(width=0.2)) +
+  geom_line(position = position_dodge(width=0.2)) +
+  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
   scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
   xlab("") + 
-  ylab(TeX("Achromatic contrast (JND)")) +
+  ylab("Achromatic contrast (JND)") +
   theme_bw() +
@@
 
+```{r FigS2}
+
+files <- list.files("./photos/ESM/")
+img1 <- readJPEG(paste0("./photos/ESM/", files[1]))
+img2 <- readJPEG(paste0("./photos/ESM/", files[2]))
+img3 <- readJPEG(paste0("./photos/ESM/", files[3]))
+img4 <- readJPEG(paste0("./photos/ESM/", files[4]))
+img5 <- readJPEG(paste0("./photos/ESM/", files[5]))
+img6 <- readJPEG(paste0("./photos/ESM/", files[6]))
+
+p1 <- cowplot::ggdraw()+
+        cowplot::draw_image(img1)
+p2 <- cowplot::ggdraw()+
+        cowplot::draw_image(img2)
+p3 <- cowplot::ggdraw()+
+        cowplot::draw_image(img3)
+p4 <- cowplot::ggdraw()+
+        cowplot::draw_image(img4)
+p5 <- cowplot::ggdraw()+
+        cowplot::draw_image(img5)
+p6 <- cowplot::ggdraw()+
+        cowplot::draw_image(img6)
+
+(p1 + p3 + p6) / (p2 + p5 + p4) + plot_annotation(tag_levels = c("A", "B", "C", "D", "E", "F"))
+
+```
 # References

### 2021-04-22 R/ChamModels_2.0.R

@@ -286,15 +286,15 @@ Courtship_plot_dS_dat <- Courtship_1.2 %>% #data for dS Chromatic constrast
   group_by(Population, BodyRegion) %>% 
   summarise(Mean = mean(dS, na.rm = T), 
             SE = sd(dS)/sqrt(length(unique(ID))),
-            Upper = Mean + SE,
-            Lower = Mean - SE) 
+            Upper = Mean + 1.96*SE,
+            Lower = Mean - 1.96*SE) 
 
 Courtship_plot_dL_dat <- Courtship_1.2 %>% #data for dL Achromatic contrast
   group_by(Population, BodyRegion) %>% 
   summarise(Mean = mean(dL, na.rm = T), 
             SE = sd(dL)/sqrt(length(unique(ID))),
-            Upper = Mean + SE,
-            Lower = Mean - SE) 
+            Upper = Mean + 1.96*SE,
+            Lower = Mean - 1.96*SE) 
 
             
 Courtship_plot_dL_dat

### 2021-04-22 ms.Rmd
updating MS

@@
 
-```{r displayModels, echo = FALSE, results = "hide"}
+```{r Models, echo = FALSE, results = "hide"}
 #########################################################
@@
     summary(ConspicDisplayAC)
+    CI_AC_display <- data.frame(confint(ConspicDisplayAC))
     anovaConspicDisplayAC <- anova(ConspicDisplayAC)
@@
 summary(ConspicCourtshipAC)
+CI_AC_court <- data.frame(confint(ConspicCourtshipAC))
 courship_anovaAC <- anova(ConspicCourtshipAC)
@@
 We manipulated both the social and predatory context chameleons encountered, in both their native Kenyan range and their introduced Hawaiian range, to test three hypotheses about colour signal evolution.
-First, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal.
-Sexual selection favours more conspicious signals.
-As such, signals should evolve to be more conspicuous when predators are absent.
-In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (**Male-Male Contests**: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; **Courtship displays**: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, Table S1).
+Using spectrophotometry and visual modelling that relied on our understanding of the chameleon visual system we quntified chromatic and achromatic contrast against the backgrounds in each environment (i.e., leaves and stems in Kenya and Hawaii).
 
-```{r Figure1}
+Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals.
+In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (**Male-Male Contests**: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; **Courtship displays**: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, table S1).
+Overall, Kenyan chameleons were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r fixef(ConspicDisplayAC)[3]`, 95% CI = `r CI_AC_display[5,1]` to  `r CI_AC_display[5,2]`, df = `r summary(ConspicDisplayAC)$coefficients[3, "df"]`, t =  `r summary(ConspicDisplayAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicDisplayAC)$coefficients[3, "Pr(>|t|)"])`, Courtship: $\beta$ = `r fixef(ConspicCourtshipAC)[3]`, 95% CI = `r CI_AC_court[5,1]` to  `r CI_AC_court[5,2]`, df = `r summary(ConspicCourtshipAC)$coefficients[3, "df"]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicCourtshipAC)$coefficients[3, "Pr(>|t|)"])`, see figure 1a & b).
+
+Sexual signals are expected to be detected much more easily by predators (ie., snakes and birds) in Hawaii as a result of reduced predation.
+In support of this prediction, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c).
+
+```{r Figure1, fig.width = 15, fig.height= 9.287443, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves).
+A & B) Contrast of male chameleons during male-male contests and courtship.
+JNDs are calculated based on the chameleon visual system.
+'NS' denotes non-significant contasts, whereas '**' indicates significant differences < 0.001.}
 ## Make figures that average across body regions, but display Social Context

### 2021-04-23 ms.Rmd
MS udpated now to include predators

@@
     summary(ConspicDisplayAC)
+    coefAIC_display = fixef(ConspicDisplayAC)[3]
     CI_AC_display <- data.frame(confint(ConspicDisplayAC))
@@
 summary(ConspicCourtshipAC)
+coefAIC_court = fixef(ConspicCourtshipAC)[3]
 CI_AC_court <- data.frame(confint(ConspicCourtshipAC))
@@
  courship_anovaCC <- anova(ConspicCourtshipCC)
+
+ #########################################################
+# Displays to Predators, birds and snake
+#########################################################
+ Predator <- read.table("./data/Predator_2_3.csv", 
+                       header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Clean up
+ Predator <- Predator %>% filter(!BodyRegion == "botflank" & !BodyRegion =='gularbot') %>% as.data.frame() #exclude bottom flank.
+Predator$BodyRegion <- factor(Predator$BodyRegion) #updates levels for the variable BodyRegion
+
+# Transformation
+Predator <- Predator %>% mutate(log_dS = log(dS),
+                                log_dL = log(dL),
+                                sqrt_dS = sqrt(dS),
+                                sqrt_dL = sqrt(dL))
+
+# Models - Achromatic Contrast (dL)
+
+# First, if predator type varies across populations then we would predcit that whether snake and bird actually are different would depend on the population.
+For example, if, say, in HAwaii and Kenya birds were equally likely as predators then dL would not vary for birds across populations, but would so for snakes.
+This is an interaction.
+
+PredatorAC <- lmer(dL ~ Predator + Population + Predator:Population + (1|ID), data = Predator) 
+summary(PredatorAC)
+anova_predators_inter <- anova(PredatorAC)
+
+# No support for an interaction, so can drop interaction.
+Tells us that predation, regardless of whether we are talking snake or bird is just low in Hawaii, otherwise, we would predict that signals would evolve to match the predatory system
+PredatorAC <- lmer(dL ~ Predator + Population + (1|ID), data = Predator) 
+summary(PredatorAC)
+anova_predators <- anova(PredatorAC)
+
+# Chromatic contrast - dS
+
+PredatorCC <- lmer(dS ~ Predator + Population + Predator:Population + (1|ID), data = Predator) 
+summary(PredatorCC)
+anova_predators_interCC <- anova(PredatorCC)
+
+Predator$pred_pop <- with(Predator, interaction(Predator, Population))
+PredatorCC_pred_pop <- lmer(dS ~ pred_pop + (1|ID), data = Predator) 
+summary(PredatorCC_pred_pop)
+
+hypthesis_test <- glht(PredatorCC_pred_pop, df = 101, # df to get t-distribution, use number of lizards not data points
+             linfct = mcp(pred_pop = c("snake.Hawaii - snake.Kenya = 0",
+                                       "bird.Hawaii  - bird.Kenya = 0",
+                                       "snake.Hawaii - bird.Hawaii = 0",
+                                       "snake.Kenya  - bird.Kenya = 0")))
+summary(hypthesis_test)
+CI_hyp_test <- as.data.frame(confint(hypthesis_test)$confint)
+
 ```
@@
 We manipulated both the social and predatory context chameleons encountered, in both their native Kenyan range and their introduced Hawaiian range, to test three hypotheses about colour signal evolution.
-Using spectrophotometry and visual modelling that relied on our understanding of the chameleon visual system we quntified chromatic and achromatic contrast against the backgrounds in each environment (i.e., leaves and stems in Kenya and Hawaii).
+Using spectrophotometry and visual modelling that relied on our understanding of the known chameleon visual system we quntified chromatic and achromatic contrast against the backgrounds in each environment (i.e., leaves and stems in Kenya and Hawaii).
 
-Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals.
+Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals whereas greater conspicuousness of sexual signals should be selected in social contexts.
 In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (**Male-Male Contests**: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; **Courtship displays**: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, table S1).
-Overall, Kenyan chameleons were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r fixef(ConspicDisplayAC)[3]`, 95% CI = `r CI_AC_display[5,1]` to  `r CI_AC_display[5,2]`, df = `r summary(ConspicDisplayAC)$coefficients[3, "df"]`, t =  `r summary(ConspicDisplayAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicDisplayAC)$coefficients[3, "Pr(>|t|)"])`, Courtship: $\beta$ = `r fixef(ConspicCourtshipAC)[3]`, 95% CI = `r CI_AC_court[5,1]` to  `r CI_AC_court[5,2]`, df = `r summary(ConspicCourtshipAC)$coefficients[3, "df"]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicCourtshipAC)$coefficients[3, "Pr(>|t|)"])`, see figure 1a & b).
+Kenyan chameleons were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r coefAIC_display`, 95% CI = `r CI_AC_display[5,1]` to  `r CI_AC_display[5,2]`, df = `r summary(ConspicDisplayAC)$coefficients[3, "df"]`, t =  `r summary(ConspicDisplayAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicDisplayAC)$coefficients[3, "Pr(>|t|)"])`, Courtship: $\beta$ = `r coefAIC_court`, 95% CI = `r CI_AC_court[5,1]` to  `r CI_AC_court[5,2]`, df = `r summary(ConspicCourtshipAC)$coefficients[3, "df"]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicCourtshipAC)$coefficients[3, "Pr(>|t|)"])`, see figure 1a & b).
 
-Sexual signals are expected to be detected much more easily by predators (ie., snakes and birds) in Hawaii as a result of reduced predation.
-In support of this prediction, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c).
+As a consequence of signal release, sexual signals are expected to be more easily detected by predators (i.e., snakes and birds) in Hawaii.
+Differential predation pressure resulting from snake and bird predation varying would also result in populations varying in their conspicuousness to different predators.
+In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c; Achromatic Contrast: ($dl$); $F_{`r anova_predators$NumDF[2]`,`r anova_predators$DenDF[2]`}$ = `r anova_predators[5][2,]`, p = `r p_value(anova_predators[6][2,])`).
+Differences in chromatic contrast between populations, however,  depended on the predator (significant predator by population interaction: $F_{`r anova_predators_interCC$NumDF[3]`,`r anova_predators_interCC$DenDF[3]`}$ = `r anova_predators_interCC[5][3,]`, p = `r p_value(anova_predators_interCC[6][3,])`).
+Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,1]`, 95% CI = `r CI_hyp_test[1,2]` to `r CI_hyp_test[1,3]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,1]`, 95% CI = `r CI_hyp_test[2,2]` to `r CI_hyp_test[2,3]`, p = < 0.001).
 
-```{r Figure1, fig.width = 15, fig.height= 9.287443, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves).
+```{r Figure1, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves).
 A & B) Contrast of male chameleons during male-male contests and courtship.
 JNDs are calculated based on the chameleon visual system.
-'NS' denotes non-significant contasts, whereas '**' indicates significant differences < 0.001.}
+NS denotes non-significant contasts, whereas astericks indicates significant differences is < 0.001."}
 ## Make figures that average across body regions, but display Social Context
@@
 
-# Now calculate the averages for each population anc Colour (i.e., context)
+# Now calculate the averages for each population and Colour (i.e., context) / predation
+
+summary_data_pred <- Predator %>%
+                group_by(Predator, Population) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL)
 
@@
 
+#Grab photos
   males <- readPNG("./photos/male_fight.png", native = TRUE)
-#males <- image_scale(males, "200x200")
+females <- readPNG("./photos/courtship.png",  native = TRUE)
+  snake <- readPNG("./photos/snake_cham.png", native = TRUE)
+   bird <- readPNG("./photos/bird_cham.png",  native = TRUE)
 
-females <- readPNG("./photos/courtship.png", native = TRUE)
-#females <- image_scale(females, "200x200")
-
+#plots
 p1 <- ggplot(summary_data, aes(x = Population, y = Mean_dS, group = Colour, color = Colour)) +
@@
 
-plot_grid(p1, p2, labels = c("A", "B"), label_size = 16,
+p3 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dL, group = Predator, color = Predator)) +
+  ylim(10, 30) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Achromatic contrast (JND)") +
+   annotate("segment", x = 0.90, xend = 1.90, y = 28, yend = 28, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 26, yend = 26, colour = "black") + 
+  annotate("text", x = 0.9+((1.90 - 0.90)/2), y = 28.5, label = "**",  size = 6) + 
+  annotate("text", x = 1.12+((2.12 - 1.12)/ 2), y = 26.5, label = "**", size = 6) + 
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+    inset_element(snake,
+                left = 0.45,
+                right = 0.75,
+                bottom = 0.60,
+                top = 0.70, ignore_tag = TRUE) + 
+    inset_element(bird,
+                left = 0.05,
+                right = 0.50,
+                bottom = 0.25,
+                top = 0.38, ignore_tag = TRUE)
+
+p4 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dS, group = Predator, color = Predator)) +
+  ylim(1, 9) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Chromatic contrast (JND)") +
+   annotate("segment", x = 0.88, xend = 1.88, y = 8.8, yend = 8.8, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y =2.5, yend = 2.5, colour = "black") + 
+  annotate("text", x = 0.9+((1.90 - 0.90)/2), y = 8.9, label = "**",  size = 6) + 
+  annotate("text", x = 1.12+((2.12 - 1.12)/ 2), y = 2.7, label = "NS", size = 6) + 
+  theme_bw() +
+  theme(legend.position = c(0.10, 0.08),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+    inset_element(bird,
+                left = 0.30,
+                right = 0.60,
+                bottom = 0.45,
+                top = 0.60, ignore_tag = TRUE) + 
+    inset_element(snake,
+                left = 0.02,
+                right = 0.32,
+                bottom = 0.20,
+                top = 0.35, ignore_tag = TRUE)
+
+plot_grid(p1, p2, p4, p3, labels = c("A", "B", "C", "D"), label_size = 16,
   label_fontface = "bold", rel_widths=c(1,1))

### 2021-04-23 supp.Rmd
MS udpated now to include predators

@@
                                                             "Hawaii.topflank - Kenya.topflank = 0", 
+                                                            "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+
+#########################################################
+# Displays to Predators, birds and snake
+#########################################################
+ Predator <- read.table("./data/Predator_2_3.csv", 
+                       header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Clean up
+ Predator <- Predator %>% filter(!BodyRegion == "botflank" & !BodyRegion =='gularbot') %>% as.data.frame() #exclude bottom flank.
+Predator$BodyRegion <- factor(Predator$BodyRegion) #updates levels for the variable BodyRegion
+
+# Transformation, just in case residuals are non-normal they may help.
+Remember, models assume normality of residuals not normality of the data.
+None-normality could be driven by effects you wish to model.
+So, you need to account for these first.
+Predator <- Predator %>% mutate(log_dS = log(dS),
+                                log_dL = log(dL),
+                                sqrt_dS = sqrt(dS),
+                                sqrt_dL = sqrt(dL))
+
+## Predator Model - Achromatic contrast (AC or dL)
+## just this next section has Drew's code, which averages the backround (Leaf_B1, Leaf_B2, Leaf_T1, Leaf_T2, Stem1, Stem2) for each lizard body region measurement.
+## This is the full model; however, we do not need background because the way the data is set up R is taking an average. See next model (background dropped)
+PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion + Background 
+                         + Predator:Background + Predator:BodyRegion + (1|ID), data=Predator)
+anova(PredatorAC)
+
+# Posthoc tests
+Predator$PopBRInt <- interaction(Predator$Population,Predator$BodyRegion)
+PredatorAC2 <- lmer(sqrt_dL ~ PopBRInt + (1|ID), data=Predator)
+summary(glht(PredatorAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
+                                                    "Hawaii.midflank - Kenya.midflank = 0", 
+                                                    "Hawaii.topflank - Kenya.topflank = 0", 
+                                                    "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+## dropping background from model (it is still averaged)
+# anova gives us significance of fixed effects
+# ranova tests random effects (note: report just LRT and df)
+# need the SD of residuals and random effects from summary (just report random effects)
+
+PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion +  
+                           + Population:BodyRegion + (1|ID), data= Predator) # we have gularbot for Kenya but not Hawaii, so exclude it
+anova(PredatorAC)
+ranova(PredatorAC)
+
+#for random effects, use ranova (is variance > 0?) report just LRT and df and need the SD of residuals and random effects from summary (just report random effects)
+
+Predator$PopBRInt <- interaction(Predator$Population,Predator$BodyRegion)
+PredatorAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Predator)
+summary(glht(PredatorAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
+                                                    "Hawaii.midflank - Kenya.midflank = 0", 
+                                                    "Hawaii.topflank - Kenya.topflank = 0", 
+                                                    "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+
+# Chromatic Contast - dS
+
+## Predator Model - Chromatic contrast (CC or dS)
+## just this next section has Drew's code, which averages the backround (Leaf_B1, Leaf_B2, Leaf_T1, Leaf_T2, Stem1, Stem2) for each lizard body region measurement.
+## This is the full model; however, we do not need background because the way the data is set up R is taking an average. See next model (background dropped)
+PredatorCC <- lmer(sqrt_dS ~ Predator + Population + BodyRegion + Background 
+                   + Predator:Background + Predator:BodyRegion + (1|ID), data=Predator)
+Predator$PopBRInt <- interaction(Predator$Population,Predator$BodyRegion)
+PredatorCC2 <- lmer(sqrt_dS ~ PopBRInt + (1|ID), data=Predator)
+summary(glht(PredatorCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
+                                                    "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+
+## dropping background from model (it is still averaged)
+# anova gives us significance of fixed effects
+# ranova tests random effects (note: report just LRT and df)
+# need the SD of residuals and random effects from summary (just report random effects)
+
+PredatorCC <- lmer(sqrt_dS ~ Predator + Population + BodyRegion +  
+                     + Population:BodyRegion + (1|ID), data=Predator[Predator$BodyRegion!='gularbot',]) # we have gularbot for Kenya but not Hawaii, so exclude it
+anova(PredatorCC)
+ranova(PredatorCC)
+
+### This is old code for when in the above model, predator was nonsignificant, so we can drop it from the model (below).
+But after sqrt it is significant, so don't use this code.
+Just keeping it in case.
+PredatorCC <- lmer(sqrt_dS ~ Population + BodyRegion +  
+                     + Population:BodyRegion + (1|ID), data=Predator[Predator$BodyRegion!='gularbot',]) # we have gularbot for Kenya but not Hawaii, so exclude it
+anova(PredatorCC)
+ranova(PredatorCC)
+###
+
+#for random effects, use ranova (is variance > 0?) report just LRT and df and need the SD of residuals and random effects from summary (just report random effects)
+
+Predator$PopBRInt <- interaction(Predator$Population,Predator$BodyRegion)
+PredatorCC2 <- lmer(sqrt_dS ~ PopBRInt + (1|ID), data=Predator[Predator$BodyRegion!='gularbot',])
+summary(glht(PredatorCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
                                                     "Hawaii.tailbase - Kenya.tailbase = 0" ))))

### 2021-04-23 ms.Rmd
Adding some sub-headings to make it easier to follow. I worry a little still about non-independence within individuals, so also added code to do sensitivity analysis which uses Robust Variance Estaimtors and small sample corrections to esatimte SEs correctly

@@
 #########################################################
-# Male-male Competition Displays
+# Social Context: Male-male Competition Displays
 #########################################################
@@
 
-# Display dl
+##########
+## Achromatic Contrast - dL
+##########
     ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + (1|ID), data=Display_1)
@@
 
-## Display dS
+##########
+# Chromatic contrast - dS
+##########
          ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + (1|ID), data=Display_1)
@@
 #########################################################
-# Courtship Displays
+# Social Context: Courtship Displays
 #########################################################
@@
 
-# Courtship dl
+##########
+## Achromatic Contrast - dL
+##########
 ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population  + (1|ID), data=Courtship)
@@
 
-# Courtship dS
+##########
+# Chromatic contrast - dS
+##########
   ConspicCourtshipCC <- lmer(dS ~ sc_SVL + Population  + (1|ID), data=Courtship)
@@
  #########################################################
-# Displays to Predators, birds and snake
+# Predator Context: Birds and snakes
 #########################################################
@@
 
-# Transformation
-Predator <- Predator %>% mutate(log_dS = log(dS),
-                                log_dL = log(dL),
-                                sqrt_dS = sqrt(dS),
-                                sqrt_dL = sqrt(dL))
-
-# Models - Achromatic Contrast (dL)
+##########
+## Achromatic Contrast - dL
+##########
 
@@
 anova_predators <- anova(PredatorAC)
-
+##########
 # Chromatic contrast - dS
-
+##########
 PredatorCC <- lmer(dS ~ Predator + Population + Predator:Population + (1|ID), data = Predator) 
@@
 
+#########################################################
+# Local Adaptation: Testing for local adaptation
+#########################################################
+##################
+# Social Context: Male-male competition
+##################
+DisplayBgrd <- read.table("./data/Display_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+DisplayBgrd <- DisplayBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
+
+##########
+# Achromatic Contrast - dL
+##########
+DispBgrdAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
+
+DispBgrdAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
+summary(DispBgrdAC)
+anova(DispBgrdAC_inter, DispBgrdAC)
+
+# Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+clubSandwich::coef_test(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
+DisplayBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = DisplayBgrd)
+
+# df to get t-distribution, use number of lizards not data points
+hypthesis_test_preppop <- glht(DisplayBgrd_pop_back, df = 69-1, 
+             linfct = mcp(pop_back = c("Hawaii.Kenya  -Hawaii.Hawaii = 0")))
+
+summary(hypthesis_test_preppop)
+p_displ_AC <- as.numeric(summary(hypthesis_test_preppop)$test$pvalues)
+CI_dis_AC <- as.data.frame(confint(hypthesis_test_preppop)$confint)
+##########
+# Chromatic Contrast - dS
+##########
+DispBgrdCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
+
+DispBgrdCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
+summary(DispBgrdCC)
+anova(DispBgrdCC_inter, DispBgrdCC)
+
+# Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+clubSandwich::coef_test(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
+DisplayBgrd_pop_backCC <- lmer(dS ~ pop_back + (1|ID), data = DisplayBgrd)
+
+hypthesis_test_preppop_CC <- glht(DisplayBgrd_pop_backCC, df = 69-1, # df to get t-distribution, use number of lizards not data points
+             linfct = mcp(pop_back = c("Hawaii.Kenya  -Hawaii.Hawaii = 0")))
+
+summary(hypthesis_test_preppop_CC)
+p_displ_CC <- as.numeric(summary(hypthesis_test_preppop_CC)$test$pvalues)
+CI_Disp_CC <- as.data.frame(confint(hypthesis_test_preppop_CC)$confint)
+
+###################
+# Social Context: Courtship
+##################
+CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+CourtshipBgrd <- CourtshipBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
+
+##########
+## Achromatic Contrast - dL
+##########
+
+CourtshipAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
+
+CourtshipAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
+summary(CourtshipAC)
+anova(CourtshipAC_inter, CourtshipAC)
+
+CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
+CourtshipBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = CourtshipBgrd)
+
+# Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+clubSandwich::coef_test(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+hypthesis_test_preppopCourtship <- glht(CourtshipBgrd_pop_back, df = 69-1, # df to get t-distribution, use number of lizards not data points
+             linfct = mcp(pop_back = c("Hawaii.Kenya  -Hawaii.Hawaii = 0")))
+
+summary(hypthesis_test_preppopCourtship)
+ p_courtship_AC <- as.numeric(summary(hypthesis_test_preppopCourtship)$test$pvalues)
+CI_AC_Courtship <- data.frame(confint(hypthesis_test_preppopCourtship)$confint)
+
+##########
+# Chromatic - dS
+##########
+CourtshipCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
+
+      CourtshipCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC)
+        anova(CourtshipCC)
+
+# Test whether interaction is supported
+anova(CourtshipCC_inter, CourtshipCC)
+
+# Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+clubSandwich::coef_test(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+# Posthoc comparions
+  CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
+CourtshipBgrd_pop_backCC <- lmer(dS ~ pop_back + (1|ID), data = CourtshipBgrd)
+
+# Specific planned comparisons
+hypthesis_test_preppopCourtshipCC <- glht(CourtshipBgrd_pop_backCC, df = 69-1, # df to get t-distribution, use number of lizards not data points
+             linfct = mcp(pop_back = c("Hawaii.Kenya - Hawaii.Hawaii = 0")))
+
+summary(hypthesis_test_preppopCourtshipCC)
+p_courtship_CC <- as.numeric(summary(hypthesis_test_preppopCourtshipCC)$test$pvalues)
+CI_CC_Courtship <- data.frame(confint(hypthesis_test_preppopCourtshipCC)$confint)
 ```
+
+Martin J. Whiting $^1,*$, Brenden S. Holland $^2$, J. Scott Keogh$^3$, Daniel W.A.Noble$^3$, Katrina J. Rankin$^4$ and Devi M. Stuart-Fox$^4$
+
+$^1$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
 
@@
 Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals whereas greater conspicuousness of sexual signals should be selected in social contexts.
-In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (**Male-Male Contests**: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; **Courtship displays**: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, table S1).
+In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; Courtship displays: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, table S1).
 Kenyan chameleons were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r coefAIC_display`, 95% CI = `r CI_AC_display[5,1]` to  `r CI_AC_display[5,2]`, df = `r summary(ConspicDisplayAC)$coefficients[3, "df"]`, t =  `r summary(ConspicDisplayAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicDisplayAC)$coefficients[3, "Pr(>|t|)"])`, Courtship: $\beta$ = `r coefAIC_court`, 95% CI = `r CI_AC_court[5,1]` to  `r CI_AC_court[5,2]`, df = `r summary(ConspicCourtshipAC)$coefficients[3, "df"]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicCourtshipAC)$coefficients[3, "Pr(>|t|)"])`, see figure 1a & b).
@@
 Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,1]`, 95% CI = `r CI_hyp_test[1,2]` to `r CI_hyp_test[1,3]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,1]`, 95% CI = `r CI_hyp_test[2,2]` to `r CI_hyp_test[2,3]`, p = < 0.001).
+
+Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment.
+While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chamleeons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_value(p_displ_AC)`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_value(p_displ_CC)`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_value(p_courtship_AC)`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_value(p_courtship_CC)`).
 
@@
 
+```{r Figure2, fig.width = 14, fig.height= 9, fig.cap= ""}
+##########
+# Data
+##########
+courtship_summary <- CourtshipBgrd %>% 
+                    group_by(Population, Background_pop) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>% filter(Population == "Hawaii") %>% as.data.frame() %>% mutate(Social_Context = "Courtship")
+
+display_summary <- DisplayBgrd %>%
+                group_by(Population, Background_pop) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>% filter(Population == "Hawaii") %>% as.data.frame() %>% mutate(Social_Context = "Male-male contest")
+
+summary_data_localAdap <- rbind(courtship_summary, display_summary)
+
+##########
+# Plots
+##########
+#Grab photos
+  males <- readPNG("./photos/male_fight.png", native = TRUE)
+females <- readPNG("./photos/courtship.png",  native = TRUE)
+
+#plots
+p1.2 <- ggplot(summary_data_localAdap, aes(x = Background_pop, y = Mean_dS, group = Social_Context, color = Social_Context)) +
+    ylim(2.5, 6.5) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Background Environment") + 
+  ylab("Chromatic contrast (JND)") +
+  annotate("segment", x = 0.85, xend = 1.85, y = 6.1, yend = 6.1, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 5.7, yend = 5.7, colour = "black") + 
+  annotate("text", x = 0.85+(1.85-0.85)/2,  y = 6.2, label = "NS",  size = 6) + 
+  annotate("text", x = 1.12+(2.12-1.12)/2, y = 5.8, label = "NS", size = 6) + 
+  theme_bw() +
+  labs(colour = "Social Context") +
+  theme(legend.position = c(0.28, 0.08),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+   inset_element(males,
+                left = 0.45,
+                right = 0.90,
+                bottom = 0.58,
+                top = 0.75, ignore_tag = TRUE) + 
+    inset_element(females,
+                left = 0.30,
+                right = 0.75,
+                bottom = 0.15,
+                top = 0.28, ignore_tag = TRUE)
+
+p2.2 <- ggplot(summary_data_localAdap, aes(x = Background_pop, y = Mean_dL, group = Social_Context, color = Social_Context)) +
+  ylim(15, 45) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Background Environment") + 
+  ylab("Achromatic contrast (JND)") +
+   annotate("segment", x = 0.90, xend = 1.90, y = 40, yend = 40, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 42, yend = 42, colour = "black") + 
+  annotate("text", x = 0.9+((1.90 - 0.90)/2), y = 41, label = "**",  size = 6) + 
+  annotate("text", x = 1.12+((2.12 - 1.12)/ 2), y = 43, label = "**", size = 6) + 
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+    inset_element(males,
+                left = 0.45,
+                right = 0.90,
+                bottom = 0.60,
+                top = 0.75, ignore_tag = TRUE) + 
+    inset_element(females,
+                left = 0.05,
+                right = 0.50,
+                bottom = 0.25,
+                top = 0.38, ignore_tag = TRUE)
+
+plot_grid(p1.2, p2.2, labels = c("A", "B"), label_size = 16,
+  label_fontface = "bold", rel_widths=c(1,1))
+
+```
+
 Specifically, we tested the following predictions: 1.) display colouration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; 2.) display colouration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawai’i than Kenya.

### 2021-04-23 R/ChamModels_2.0.R

@@ -719,6 +719,8 @@ DisplayBgrdCC <- lmer(dS ~ Population + BodyRegion + Background + Background2
                       + Population*Background + Population*Background2 + BodyRegion*Background + BodyRegion*Background2 +(1|ID), data=DisplayBgrd)
 summary(DisplayBgrdCC)
 anova(DisplayBgrdCC)
+
+
 #View plots of residuals
 qqnorm(residuals(DisplayBgrdCC))
 hist(residuals(DisplayBgrdCC))

### 2021-04-28 ms.Rmd
Cleaning up script and annotating better. Added in RVE to check non-independence is not an issue

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich)
 
@@
 ##########
+
+    # Fit the model
     ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + (1|ID), data=Display_1)
       summary(ConspicDisplayAC)
-    coefAIC_display = fixef(ConspicDisplayAC)[3]
+
+    # Get the coefficients, and ANOVA statistics for the paper
+          coefAIC_display <- fixef(ConspicDisplayAC)[3]
             CI_AC_display <- data.frame(confint(ConspicDisplayAC))
     anovaConspicDisplayAC <- anova(ConspicDisplayAC)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+    robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, cluster = Display_1_IDs$ID, vcov = "CR2")
 
@@
 ##########
+    #Fit the model
          ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + (1|ID), data=Display_1)
+         summary(ConspicDisplayCC)
+
+    # Get ANOVA statistics for model for the paper
          anovaConspicDisplayCC <- anova(ConspicDisplayCC)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+    robust_ConspicDisplayCC <- coef_test(ConspicDisplayCC, cluster = Display_1_IDs$ID, vcov = "CR2")
 
@@
 ##########
+
+  # Fit the model
     ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population  + (1|ID), data=Courtship)
     summary(ConspicCourtshipAC)
-coefAIC_court = fixef(ConspicCourtshipAC)[3]
+
+  # Get model stats for ppaer
+       coefAIC_court <- fixef(ConspicCourtshipAC)[3]
          CI_AC_court <- data.frame(confint(ConspicCourtshipAC))
     courship_anovaAC <- anova(ConspicCourtshipAC)
+
+ # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+    robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, cluster = Courtship_IDs$ID, vcov = "CR2")
 
@@
 ##########
+
+  # Fit the model
   ConspicCourtshipCC <- lmer(dS ~ sc_SVL + Population  + (1|ID), data=Courtship)
+    summary(ConspicCourtshipCC)
+
+  # Get the ANOVA stats for paper
     courship_anovaCC <- anova(ConspicCourtshipCC)
+
+  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+    robust_ConspicCourtshipCC <- coef_test(ConspicCourtshipCC, cluster = Courtship_IDs$ID, vcov = "CR2")
 
@@
 # First, if predator type varies across populations then we would predcit that whether snake and bird actually are different would depend on the population.
-For example, if, say, in HAwaii and Kenya birds were equally likely as predators then dL would not vary for birds across populations, but would so for snakes.
+For example, if, say, in Hawaii and Kenya birds were equally likely as predators then dL would not vary for birds across populations, but would so for snakes.
 This is an interaction.
 
-PredatorAC <- lmer(dL ~ Predator + Population + Predator:Population + (1|ID), data = Predator) 
-summary(PredatorAC)
-anova_predators_inter <- anova(PredatorAC)
+  # Fit model testing this idea above
+  PredatorAC_inter <- lmer(dL ~ Predator + Population + Predator:Population + (1|ID), data = Predator) 
+  summary(PredatorAC_inter)
 
-# No support for an interaction, so can drop interaction.
-Tells us that predation, regardless of whether we are talking snake or bird is just low in Hawaii, otherwise, we would predict that signals would evolve to match the predatory system
+  # Get ANOVA stats for reporting in the paper of interaction
+  anova_predators_inter <- anova(PredatorAC_inter)
+
+# No support for an interaction.
+Drop interaction and fit the main effects model.
+Tells us that predation, regardless of whether we are talking about snakes or birds is just low in Hawaii, otherwise, we would predict that signals would evolve to match the predatory system
+
+  # Fit main effects model
     PredatorAC <- lmer(dL ~ Predator + Population + (1|ID), data = Predator) 
     summary(PredatorAC)
+
+  # Grab the ANOVA summary stats
     anova_predators <- anova(PredatorAC)
+
+  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Predator_IDs <- Predator %>% filter(!is.na(Predator), !is.na(Population)) # filter out misssing data
+    robust_PredatorAC <- coef_test(PredatorAC, cluster = Predator_IDs$ID, vcov = "CR2")
+
 ##########
@@
 ##########
+
+    # First, if predator type varies across populations then we would predcit that whether snake and bird actually are different would depend on the population.
+For example, if, say, in Hawaii and Kenya birds were equally likely as predators then dL would not vary for birds across populations, but would so for snakes.
+This is an interaction.
+
+    # Fit the interaction model.
     PredatorCC <- lmer(dS ~ Predator + Population + Predator:Population + (1|ID), data = Predator) 
     summary(PredatorCC)
+
+    # Get ANOVA statistics for the paper
     anova_predators_interCC <- anova(PredatorCC)
 
+    # Support for an interaction between predator type and population in the chromatic realm. Planned comparisons. First re-fit model
       Predator$pred_pop <- with(Predator, interaction(Predator, Population))
@@
 
-hypthesis_test <- glht(PredatorCC_pred_pop, df = 101, # df to get t-distribution, use number of lizards not data points
+    # Conduct hypothesis tests of interest
+    hypthesis_test <- glht(PredatorCC_pred_pop, df = 101-1, # df to get t-distribution, use number of lizards not data points
              linfct = mcp(pred_pop = c("snake.Hawaii - snake.Kenya = 0",
@@
                                     "snake.Kenya  - bird.Kenya = 0")))
+    # Get the contrasts, and statistical data for these plannned comparions.
       summary(hypthesis_test)
       CI_hyp_test <- as.data.frame(confint(hypthesis_test)$confint)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Predator_IDs <- Predator %>% filter(!is.na(Predator), !is.na(Population)) # filter out misssing data
+    robust_PredatorCC <- coef_test(PredatorCC, cluster = Predator_IDs$ID, vcov = "CR2")
 
@@
 ##########
+
+  # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit the interaction model
   DispBgrdAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC_inter)
 
+  # Fit the main effect model
   DispBgrdAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
   summary(DispBgrdAC)
+
+  # Doesn't look like an interaction, so test.
   anova(DispBgrdAC_inter, DispBgrdAC)
 
-# Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
-clubSandwich::coef_test(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+  robust_localAdapt <- clubSandwich::coef_test(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
+  # Planned contrast. Fit the model in a slightly different way
   DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
@@
 
+  # Grab contrast stats for paper
   summary(hypthesis_test_preppop)
@@
 ##########
+
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+   # Fit the interaction model
     DispBgrdCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
+    summary(DispBgrdCC_inter)
 
+    # Fit the main effect model
           DispBgrdCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
           summary(DispBgrdCC)
+
+    # Looks like weak evidence for interaction    
         anova(DispBgrdCC_inter, DispBgrdCC)
 
-# Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
-clubSandwich::coef_test(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+      robust_DispBgrdCC <- clubSandwich::coef_test(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
+    # Planned contrasts. Refit model in slight different form
       DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
@@
 
-hypthesis_test_preppop_CC <- glht(DisplayBgrd_pop_backCC, df = 69-1, # df to get t-distribution, use number of lizards not data points
+    # Conduct hypothesis test, df to get t-distribution, use number of lizards not data points
+      hypthesis_test_preppop_CC <- glht(DisplayBgrd_pop_backCC, df = 69-1, 
              linfct = mcp(pop_back = c("Hawaii.Kenya  -Hawaii.Hawaii = 0")))
 
+    # Grab the statistics for the paper
       summary(hypthesis_test_preppop_CC)
@@
 
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit the interaction model
     CourtshipAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
 
+  # Fit main effects model
     CourtshipAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
     summary(CourtshipAC)
+
+  # Interaction needed? No
     anova(CourtshipAC_inter, CourtshipAC)
 
+   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+   robust_CourtshipAC <- clubSandwich::coef_test(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+  # Planned contrasts / hypothesis tests. Need to fit model a little differently
     CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
@@
 
-# Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
-clubSandwich::coef_test(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
-
-hypthesis_test_preppopCourtship <- glht(CourtshipBgrd_pop_back, df = 69-1, # df to get t-distribution, use number of lizards not data points
+  # Hypothesis tests
+      hypthesis_test_preppopCourtship <- glht(CourtshipBgrd_pop_back, df = 69-1, 
+                                        # df to get t-distribution, use number of lizards not data points
              linfct = mcp(pop_back = c("Hawaii.Kenya  -Hawaii.Hawaii = 0")))
@@
       summary(hypthesis_test_preppopCourtship)
+
+  # Grab the stats for paper
       p_courtship_AC <- as.numeric(summary(hypthesis_test_preppopCourtship)$test$pvalues)
@@
 ##########
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit interaction model
     CourtshipCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
 
+  # Fit main effect model
       CourtshipCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
@@
   # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
-clubSandwich::coef_test(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+  robust_CourtshipCC <- clubSandwich::coef_test(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
 
-# Posthoc comparions
+  # Planned comparions
     CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
@@
 
-# Specific planned comparisons
-hypthesis_test_preppopCourtshipCC <- glht(CourtshipBgrd_pop_backCC, df = 69-1, # df to get t-distribution, use number of lizards not data points
+  # Specific planned comparisons, # df to get t-distribution, use number of lizards not data points
+      hypthesis_test_preppopCourtshipCC <- glht(CourtshipBgrd_pop_backCC, df = 69-1, 
              linfct = mcp(pop_back = c("Hawaii.Kenya - Hawaii.Hawaii = 0")))
 
+  # Grab summary stats for paper
     summary(hypthesis_test_preppopCourtshipCC)
@@
 ```
-
-Specifically, we tested the following predictions: 1.) display colouration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; 2.) display colouration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawai’i than Kenya.
-3.) under character release, males will have greater capacity to change colour from a subordinate to a dominant state in Hawai’i compared to Kenya.
-
-We also tested the hypothesis that sensory drive and local adaptation are responsible for driving display coloration.
-Specifically, we predicted that chameleon display colouration in each population would be more conspicuous within its own signaling environment compared to the other population’s signaling environment.
-For example, Hawaiian chameleons were historically found in Kenya but under sensory drive and local adaptation they are predicted to be more conspicuous in their current signaling environment, Hawai’i.
-Likewise, Kenyan chameleons have no history in Hawai’i and are therefore predicted to be more conspicuous within their current signaling environment, Kenya.

### 2021-04-28 ms.Rmd
Exporting figures, but fixing them up manually in adobe illustrator to more easily size photos to scale. Hard to get right when hard coding. Added new code to simply bring in the edited figures to the MS as opposed to rerun the figures.

@@
 NS denotes non-significant contasts, whereas astericks indicates significant differences is < 0.001."}
+rerun = FALSE
+if(rerun){
 ## Make figures that average across body regions, but display Social Context
@@
   labs(colour = "Social Context") +
-  theme(legend.position = c(0.28, 0.08),
+  theme(legend.position = c(0.18, 0.08),
     panel.grid.major = element_blank(), 
@@
 
+#pdf(file = "./output/figures/Figure1.1.pdf", width = 14, height = 14)
 plot_grid(p1, p2, p4, p3, labels = c("A", "B", "C", "D"), label_size = 16,
   label_fontface = "bold", rel_widths=c(1,1))
-
+#dev.off()
+} else {
+  plot(image_read_pdf("./output/figures/Figure1.pdf"))
+}
 ```
 
-```{r Figure2, fig.width = 14, fig.height= 9, fig.cap= ""}
+```{r Figure2, fig.width = 14, fig.height= 9, fig.cap= "Local adaptation of chromatic and acromatic contrast for Hawaiian chameleons against Hawaiian and Kenyan backgrounds."}
+rerun=FALSE
+
+if(rerun){
 ##########
@@
   labs(colour = "Social Context") +
-  theme(legend.position = c(0.28, 0.08),
+  theme(legend.position = c(0.15, 0.08),
     panel.grid.major = element_blank(), 
@@
                 top = 0.38, ignore_tag = TRUE)
-
+#pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
 plot_grid(p1.2, p2.2, labels = c("A", "B"), label_size = 16,
   label_fontface = "bold", rel_widths=c(1,1))
+#dev.off()
+} else {
+  plot(image_read_pdf("./output/figures/Figure2.pdf"))
+}
 

### 2021-04-28 ms.Rmd
Added some stats methods. Still need details on visual modelling. Removed section numbering

@@
     toc_depth: 6
-    number_sections: true
+    number_sections: false
     reference_docx: template.docx
@@
 
-Martin J. Whiting $^1,*$, Brenden S. Holland $^2$, J. Scott Keogh$^3$, Daniel W.A.Noble$^3$, Katrina J. Rankin$^4$ and Devi M. Stuart-Fox$^4$
+Martin J. Whiting $^{1,*}$, Brenden S. Holland $^{2}$, J. Scott Keogh$^{3}$, Daniel W.A.Noble$^{3}$, Katrina J. Rankin$^{4}$ and Devi M. Stuart-Fox$^{4}$
 
-$^1$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
+$^{1}$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
 
@@
 In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; Courtship displays: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, table S1).
-Kenyan chameleons were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r coefAIC_display`, 95% CI = `r CI_AC_display[5,1]` to  `r CI_AC_display[5,2]`, df = `r summary(ConspicDisplayAC)$coefficients[3, "df"]`, t =  `r summary(ConspicDisplayAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicDisplayAC)$coefficients[3, "Pr(>|t|)"])`, Courtship: $\beta$ = `r coefAIC_court`, 95% CI = `r CI_AC_court[5,1]` to  `r CI_AC_court[5,2]`, df = `r summary(ConspicCourtshipAC)$coefficients[3, "df"]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicCourtshipAC)$coefficients[3, "Pr(>|t|)"])`, see figure 1a & b).
+Kenyan chameleons from populations where predation is high were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r coefAIC_display`, 95% CI = `r CI_AC_display[5,1]` to  `r CI_AC_display[5,2]`, df = `r summary(ConspicDisplayAC)$coefficients[3, "df"]`, t =  `r summary(ConspicDisplayAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicDisplayAC)$coefficients[3, "Pr(>|t|)"])`, Courtship: $\beta$ = `r coefAIC_court`, 95% CI = `r CI_AC_court[5,1]` to  `r CI_AC_court[5,2]`, df = `r summary(ConspicCourtshipAC)$coefficients[3, "df"]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicCourtshipAC)$coefficients[3, "Pr(>|t|)"])`, see figure 1a & b).
 
@@
 ```
+
+# Materials and Methods
+
+### *Statistical Analysis*
+
+Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package ().
+Across the different contexts (i.e., social contexts, predation and background environments), chromatic and acromatic contrasts were analysed in seperate models as response variables.
+For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects.
+For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or snake), population of origin and their interaction as fixed effects.
+Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii.
+Model interactions were first tesed using a Type III sum of squares ANOVA with a Saitterwaite degrees of freedom correction, and if found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
+For all individuals, we measured signals for multiple body regions, each of which were contratsed between multiple background measurements (i.e., 2 random leaves, and 2 random stems).
+For our main analyses we average estimates across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results).
+Deriving multiple contrats for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely indpeendent of each other.
+To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) on our final model to correct standard errors for within-individual non-independence.
+In all cases, our results and conclusions were not impacted (see supplmentary code).
+
+# References

### 2021-04-28 ms.Rmd
Adding code to implement RVE across the board including pairwise comparisons

@@
   if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
-  if(x >= 0.01) {tmp =x}
+  if(x >= 0.01) {tmp = round(x, digits =2)}
   return(tmp)
@@
 Model interactions were first tesed using a Type III sum of squares ANOVA with a Saitterwaite degrees of freedom correction, and if found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
-For all individuals, we measured signals for multiple body regions, each of which were contratsed between multiple background measurements (i.e., 2 random leaves, and 2 random stems).
+For all individuals, we measured signals for multiple body regions, each of which were contratsed between multiple background measurements (i.e., top and bottoms of two leaves and two stems).
 For our main analyses we average estimates across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results).

### 2021-04-28 supp.Rmd
Adding code to implement RVE across the board including pairwise comparisons

@@
 #########################################################
-#Load data and clean a bit
+  # Load data and clean a bit
   Display_1 <- read.table("./data/Display_1.csv", 
@@
 
-#exclude bottom flank.
+  # Exclude bottom flank.
   Display_1 <- Display_1 %>% 
@@
 
-# create as factor
+  # Create as factor
       Display_1$BodyRegion <- factor(Display_1$BodyRegion)
@@
 
+##########
+## Achromatic Contrast - dL
+##########
+
     # Fit the model that esatimtes changes across body regions
-ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Background 
-                         + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
+    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
+    summary(ConspicDisplayAC)
+
+    # Robust variance estimators (RVE). Mostly all consistent
+    Display_subset <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
+    robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, cluster = Display_subset$ID, vcov="CR2")
+
+    # ANOVA. 
+        anova(Display_subset)
 
@@
            Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
-ConspicDisplayAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Display_1)
+            ConspicDisplayAC2 <- lmer(dL ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Display_1)
+            summary(ConspicDisplayAC2)
+    # Fit RVE
+     robust_ConspicDisplayAC2 <- coef_test(ConspicDisplayAC2, cluster = Display_subset$ID, vcov="CR2")
 
+     # Contrasts hypothesis tests, but use RVE 
+    result1 <-Wald_test(ConspicDisplayAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+    result1$p_val_bonfer <- sapply(p.adjust(result1$p_val, "bonferroni"), function(x) p_value(x))
     result1 = glht(ConspicDisplayAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
@@
 
-## Display - Chromatic contrast (CC or dS)
-## just this next section has Drew's code, which averages the backround (Leaf_B1, Leaf_B2, Leaf_T1, Leaf_T2, Stem1, Stem2) for each lizard body region measurement.
-## This is the full model; however, we do not need background because the way the data is set up R is taking an average. See next model (background dropped)
+##########
+## Chromatic Contrast - dL
+##########
 
 # Fit the model that esatimtes changes across body regions
-  ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Background 
-                         + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
+  ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
 Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)

### 2021-04-29 supp.Rmd
updating the supplement

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, pander, kableExtra, flextable, jpeg)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, pander, kableExtra, flextable, jpeg, clubSandwich)
+
+p_value <- function(x){
+  if(x <= 0.0001) {tmp = "< 0.0001"}
+  if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
+  if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
+  if(x >= 0.01) {tmp = round(x, digits =2)}
+  return(tmp)
+}
 ```
@@
 
-    # Fit the model that esatimtes changes across body regions
+    # Fit the model that estimates changes across body regions
     ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
@@
 
-    # ANOVA. 
-        anova(Display_subset)
-
     # Fit model for linear hypothesis tests
@@
             summary(ConspicDisplayAC2)
+
      # Fit RVE
@@
      # Contrasts hypothesis tests, but use RVE 
-    result1 <-Wald_test(ConspicDisplayAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+    result1 <- Wald_test(ConspicDisplayAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
     result1$p_val_bonfer <- sapply(p.adjust(result1$p_val, "bonferroni"), function(x) p_value(x))
-    result1 = glht(ConspicDisplayAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
-                                                                "Hawaii.midflank - Kenya.midflank = 0",     
-                                                                "Hawaii.topflank - Kenya.topflank = 0", 
-                                                                 "Hawaii.tailbase - Kenya.tailbase = 0" )))
 
@@
   ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
-Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
+
+ # Robust variance estimators (RVE). Mostly all consistent
+             Display_subset <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
+    robust_ConspicDisplayCC <- coef_test(ConspicDisplayCC, cluster = Display_subset$ID, vcov="CR2")
 
   # Fit model for linear hypothesis tests
-ConspicDisplayCC2 <- lmer(dS ~ PopBRInt + (1|ID), data=Display_1)
-summary(glht(ConspicDisplayCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
-                                                          "Hawaii.midflank - Kenya.midflank = 0", 
-                                                          "Hawaii.topflank - Kenya.topflank = 0", 
-                                                          "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+    Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
+     ConspicDisplayCC2 <- lmer(dS ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Display_1)
+     summary(ConspicDisplayCC2)
+
+ # Contrasts hypothesis tests, but use RVE 
+                 result2 <- Wald_test(ConspicDisplayCC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+    result2$p_val_bonfer <- sapply(p.adjust(result1$p_val, "bonferroni"), function(x) p_value(x))
 

### 2021-04-29 supp.Rmd
Up to predators for model checking

@@
                  result2 <- Wald_test(ConspicDisplayCC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
-    result2$p_val_bonfer <- sapply(p.adjust(result1$p_val, "bonferroni"), function(x) p_value(x))
+    result2$p_val_bonfer <- sapply(p.adjust(result2$p_val, "bonferroni"), function(x) p_value(x))
 
@@
 
+##########
+## Achromatic Contrast - dL
+##########
   # Fit model
-ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population + BodyRegion +  
-                             + Population:BodyRegion + (1|ID), data=Courtship)
-anova(ConspicCourtshipAC)
-
-ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + (1|ID), data=Courtship)
+    ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Courtship)
     summary(ConspicCourtshipAC)
@@
 
-# Courtship - Chromatic dl
+  # Robust variance estimators (RVE). Mostly all consistent
+             Courtship_subset <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
+    robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, cluster = Courtship_subset$ID, vcov="CR2")
 
+  # Refit model for contrasts
      Courtship$PopBRInt <- interaction(Courtship$Population,Courtship$BodyRegion)
-ConspicCourtshipAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Courtship)
-summary(glht(ConspicCourtshipAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
-                                                            "Hawaii.midflank - Kenya.midflank = 0", 
-                                                            "Hawaii.topflank - Kenya.topflank = 0", 
-                                                            "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+    ConspicCourtshipAC2 <- lmer(dL ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Courtship)
 
-# Courtship - Chromatic dS
-ConspicCourtshipCC <- lmer(dS ~ SVL + Population + BodyRegion + Background 
-                           + Population:Background + Population:BodyRegion + (1|ID), data=Courtship)
+  # Contrasts hypothesis tests, but use RVE 
+                 result3 <- Wald_test(ConspicCourtshipAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+    result3$p_val_bonfer <- sapply(p.adjust(result3$p_val, "bonferroni"), function(x) p_value(x))
+
+##########
+## Chromatic Contrast - dL
+#########
+
+    # Fit model
+        ConspicCourtshipCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Courtship)
+
+    # Robust variance estimators (RVE). Mostly all consistent
+             Courtship_subset <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
+    robust_ConspicCourtshipCC <- coef_test(ConspicCourtshipCC, cluster = Courtship_subset$ID, vcov="CR2")
+
+    # Refit model to get the contrasts between coefficienst we want
      Courtship$PopBRInt <- interaction(Courtship$Population,Courtship$BodyRegion)
-ConspicCourtshipCC2 <- lmer(dS ~ PopBRInt + (1|ID), data=Courtship)
-summary(glht(ConspicCourtshipCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
-                                                            "Hawaii.midflank - Kenya.midflank = 0", 
-                                                            "Hawaii.topflank - Kenya.topflank = 0", 
-                                                            "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+    ConspicCourtshipCC2 <- lmer(dS ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Courtship)
+
+    # Contrasts hypothesis tests, but use RVE 
+                 result4 <- Wald_test(ConspicCourtshipCC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+    result4$p_val_bonfer <- sapply(p.adjust(result4$p_val, "bonferroni"), function(x) p_value(x))
 
@@
 
-# Transformation, just in case residuals are non-normal they may help.
-Remember, models assume normality of residuals not normality of the data.
-None-normality could be driven by effects you wish to model.
-So, you need to account for these first.
-Predator <- Predator %>% mutate(log_dS = log(dS),
-                                log_dL = log(dL),
-                                sqrt_dS = sqrt(dS),
-                                sqrt_dL = sqrt(dL))
+##########
+## Achromatic Contrast - dL
+##########
 
-## Predator Model - Achromatic contrast (AC or dL)
-## just this next section has Drew's code, which averages the backround (Leaf_B1, Leaf_B2, Leaf_T1, Leaf_T2, Stem1, Stem2) for each lizard body region measurement.
-## This is the full model; however, we do not need background because the way the data is set up R is taking an average. See next model (background dropped)
-PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion + Background 
-                         + Predator:Background + Predator:BodyRegion + (1|ID), data=Predator)
+   # Fit model
+      PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion + Predator:BodyRegion + Predator:Population + (1|ID), data=Predator)
       anova(PredatorAC)
 
+   # Robust variance estimators (RVE). Mostly all consistent
+             Predator_subset <- Predator %>% filter(!is.na(Population), !is.na(BodyRegion))
+    robust_PredatorAC <- coef_test(PredatorAC, cluster = Predator_subset$ID, vcov="CR2")
+
+  # Fit model for linear hypothesis tests
+    Predator$PopBRInt <- interaction(Predator$Population,Predator$Predator, Predator$BodyRegion)
+     ConspicDisplayCC2 <- lmer(dS ~ -1 + PopBRInt + (1|ID), data=Display_1)
+     summary(ConspicDisplayCC2)
+
+ # Contrasts hypothesis tests, but use RVE 
+                 result2 <- Wald_test(ConspicDisplayCC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+    result2$p_val_bonfer <- sapply(p.adjust(result2$p_val, "bonferroni"), function(x) p_value(x))
 # Posthoc tests

### 2021-04-29 supp.Rmd
Supplement update

@@
       PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion + Predator:BodyRegion + Predator:Population + (1|ID), data=Predator)
+      summary(PredatorAC)
       anova(PredatorAC)
@@
     Predator$PopBRInt <- interaction(Predator$Population,Predator$Predator, Predator$BodyRegion)
-     ConspicDisplayCC2 <- lmer(dS ~ -1 + PopBRInt + (1|ID), data=Display_1)
-     summary(ConspicDisplayCC2)
+     PredatorAC2 <- lmer(dS ~ -1 + PopBRInt + (1|ID), data=Predator)
+     summary(PredatorAC2)
 
  # Contrasts hypothesis tests, but use RVE 
-                 result2 <- Wald_test(ConspicDisplayCC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
-    result2$p_val_bonfer <- sapply(p.adjust(result2$p_val, "bonferroni"), function(x) p_value(x))
-# Posthoc tests
-Predator$PopBRInt <- interaction(Predator$Population,Predator$BodyRegion)
-PredatorAC2 <- lmer(sqrt_dL ~ PopBRInt + (1|ID), data=Predator)
-summary(glht(PredatorAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
-                                                    "Hawaii.midflank - Kenya.midflank = 0", 
-                                                    "Hawaii.topflank - Kenya.topflank = 0", 
-                                                    "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+                 result5 <- Wald_test(PredatorAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+    result5$p_val_bonfer <- sapply(p.adjust(result5$p_val, "bonferroni"), function(x) p_value(x))
+
 ## dropping background from model (it is still averaged)

### 2021-04-30 supp.Rmd
Cleaned supplement

@@
 
-## dropping background from model (it is still averaged)
-# anova gives us significance of fixed effects
-# ranova tests random effects (note: report just LRT and df)
-# need the SD of residuals and random effects from summary (just report random effects)
+##########
+## Chromatic Contrast - dL
+########
 
-PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion +  
-                           + Population:BodyRegion + (1|ID), data= Predator) # we have gularbot for Kenya but not Hawaii, so exclude it
+  # Fit Model
+        PredatorCC <- lmer(dS ~ Predator + Population + BodyRegion + Predator:BodyRegion + Predator:Population + (1|ID), data=Predator)
+        summary(PredatorCC)
           anova(PredatorAC)
-ranova(PredatorAC)
 
-#for random effects, use ranova (is variance > 0?) report just LRT and df and need the SD of residuals and random effects from summary (just report random effects)
+  # Robust variance estimators (RVE). Mostly all consistent
+             Predator_subset <- Predator %>% filter(!is.na(Population), !is.na(BodyRegion))
+           robust_PredatorCC <- coef_test(PredatorCC, cluster = Predator_subset$ID, vcov="CR2")
 
-Predator$PopBRInt <- interaction(Predator$Population,Predator$BodyRegion)
-PredatorAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Predator)
-summary(glht(PredatorAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
-                                                    "Hawaii.midflank - Kenya.midflank = 0", 
-                                                    "Hawaii.topflank - Kenya.topflank = 0", 
-                                                    "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+  # Fit model for linear hypothesis tests
+    Predator$PopBRInt <- interaction(Predator$Population,Predator$Predator, Predator$BodyRegion)
+          PredatorCC2 <- lmer(dS ~ -1 + PopBRInt + (1|ID), data=Predator)
+          summary(PredatorCC2)
 
-# Chromatic Contast - dS
-
-## Predator Model - Chromatic contrast (CC or dS)
-## just this next section has Drew's code, which averages the backround (Leaf_B1, Leaf_B2, Leaf_T1, Leaf_T2, Stem1, Stem2) for each lizard body region measurement.
-## This is the full model; however, we do not need background because the way the data is set up R is taking an average. See next model (background dropped)
-PredatorCC <- lmer(sqrt_dS ~ Predator + Population + BodyRegion + Background 
-                   + Predator:Background + Predator:BodyRegion + (1|ID), data=Predator)
-Predator$PopBRInt <- interaction(Predator$Population,Predator$BodyRegion)
-PredatorCC2 <- lmer(sqrt_dS ~ PopBRInt + (1|ID), data=Predator)
-summary(glht(PredatorCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
-                                                    "Hawaii.tailbase - Kenya.tailbase = 0" ))))
-
-## dropping background from model (it is still averaged)
-# anova gives us significance of fixed effects
-# ranova tests random effects (note: report just LRT and df)
-# need the SD of residuals and random effects from summary (just report random effects)
-
-PredatorCC <- lmer(sqrt_dS ~ Predator + Population + BodyRegion +  
-                     + Population:BodyRegion + (1|ID), data=Predator[Predator$BodyRegion!='gularbot',]) # we have gularbot for Kenya but not Hawaii, so exclude it
-anova(PredatorCC)
-ranova(PredatorCC)
-
-### This is old code for when in the above model, predator was nonsignificant, so we can drop it from the model (below).
-But after sqrt it is significant, so don't use this code.
-Just keeping it in case.
-PredatorCC <- lmer(sqrt_dS ~ Population + BodyRegion +  
-                     + Population:BodyRegion + (1|ID), data=Predator[Predator$BodyRegion!='gularbot',]) # we have gularbot for Kenya but not Hawaii, so exclude it
-anova(PredatorCC)
-ranova(PredatorCC)
-###
-
-#for random effects, use ranova (is variance > 0?) report just LRT and df and need the SD of residuals and random effects from summary (just report random effects)
-
-Predator$PopBRInt <- interaction(Predator$Population,Predator$BodyRegion)
-PredatorCC2 <- lmer(sqrt_dS ~ PopBRInt + (1|ID), data=Predator[Predator$BodyRegion!='gularbot',])
-summary(glht(PredatorCC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0","Hawaii.midflank - Kenya.midflank = 0", "Hawaii.topflank - Kenya.topflank = 0", 
-                                                    "Hawaii.tailbase - Kenya.tailbase = 0" ))))
+ # Contrasts hypothesis tests, but use RVE 
+                 result6 <- Wald_test(PredatorCC2, constraints = constrain_pairwise(c(2:9), with_zero = TRUE), vcov="CR2", tidy = TRUE)
+    result6$p_val_bonfer <- sapply(p.adjust(result6$p_val, "bonferroni"), function(x) p_value(x))
 
@@
   summarise( Mean = mean(dS, na.rm = T), 
-               SE = sd(dS)/sqrt(length(unique(ID))),
+               SE = sd(dS)/sqrt(length(unique(ID))-1),
             Upper = Mean + 1.96*SE,
@@
   summarise(Mean = mean(dL, na.rm = T), 
-            SE = sd(dL)/sqrt(length(unique(ID))),
+            SE = sd(dL)/sqrt(length(unique(ID))-1),
             Upper = Mean + 1.96*SE,

### 2021-04-30 ms.Rmd
Adding in CIs with RVE methods.

@@
        robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, cluster = Display_1_IDs$ID, vcov = "CR2")  
+    robust_ConspicDisplayAC_CI <- conf_int(ConspicDisplayAC, cluster = Display_1_IDs$ID, vcov = "CR2")
 
@@
     robust_ConspicDisplayCC <- coef_test(ConspicDisplayCC, cluster = Display_1_IDs$ID, vcov = "CR2")
+ robust_ConspicDisplayCC_CI <- conf_int(ConspicDisplayCC, cluster = Display_1_IDs$ID, vcov = "CR2")
 
@@
     robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, cluster = Courtship_IDs$ID, vcov = "CR2")
+ robust_ConspicCourtshipAC_CI <- conf_int(ConspicCourtshipAC, cluster = Courtship_IDs$ID, vcov = "CR2")
 
@@
        robust_ConspicCourtshipCC <- coef_test(ConspicCourtshipCC, cluster = Courtship_IDs$ID, vcov = "CR2")
+    robust_ConspicCourtshipCC_CI <- conf_int(ConspicCourtshipCC, cluster = Courtship_IDs$ID, vcov = "CR2")
 

### 2021-04-30 ms.Rmd
Fixed all the main MS so that all contrasts, p-values etc are using robust variance esatimatos to corrects SEs and p-values, also using Bonferroni corrections as a super conservative path

@@
        robust_PredatorAC <- coef_test(PredatorAC, cluster = Predator_IDs$ID, vcov = "CR2")
+    robust_PredatorAC_CI <-  conf_int(PredatorAC, cluster = Predator_IDs$ID, vcov = "CR2")
 
@@
       Predator$pred_pop <- with(Predator, interaction(Predator, Population))
-    PredatorCC_pred_pop <- lmer(dS ~ pred_pop + (1|ID), data = Predator) 
+    PredatorCC_pred_pop <- lmer(dS ~ -1 + pred_pop + (1|ID), data = Predator) 
     summary(PredatorCC_pred_pop)
 
-    # Conduct hypothesis tests of interest
-    hypthesis_test <- glht(PredatorCC_pred_pop, df = 101-1, # df to get t-distribution, use number of lizards not data points
-             linfct = mcp(pred_pop = c("snake.Hawaii - snake.Kenya = 0",
-                                       "bird.Hawaii  - bird.Kenya = 0",
-                                       "snake.Hawaii - bird.Hawaii = 0",
-                                    "snake.Kenya  - bird.Kenya = 0")))
-    # Get the contrasts, and statistical data for these plannned comparions.
-      summary(hypthesis_test)
-      CI_hyp_test <- as.data.frame(confint(hypthesis_test)$confint)
+    # Fit RVE, we want contrast effects because annoyingly RVE doesn't provide them so will relevel to get them in the context of RVE
 
-    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-    Predator_IDs <- Predator %>% filter(!is.na(Predator), !is.na(Population)) # filter out misssing data
-    robust_PredatorCC <- coef_test(PredatorCC, cluster = Predator_IDs$ID, vcov = "CR2")
+    # Model allows us to get bird in Hwaii compared to bird in Kenya (contrast 3)
+      PredatorCC_pred_pop1 <- lmer(dS ~  pred_pop + (1|ID), data = Predator) 
+      summary(PredatorCC_pred_pop1)
+      PredatorCC_pred_pop2 <- lmer(dS ~  relevel(pred_pop, ref = "snake.Hawaii") + (1|ID), data = Predator) 
+      summary(PredatorCC_pred_pop2)
+
+      # Fit the models so we get the estimate and corrected robust standard error for the contrasts we want to make
+      robust_PredatorCC_pred_pop1 <- conf_int(PredatorCC_pred_pop1, cluster = Predator$ID, vcov="CR2")
+      robust_PredatorCC_pred_pop2 <- conf_int(PredatorCC_pred_pop2, cluster = Predator$ID, vcov="CR2")
+
+     # Contrasts hypothesis tests, but use RVE 
+    result_pred <- Wald_test(PredatorCC_pred_pop, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+    result_pred$p_val_bonfer <- sapply(p.adjust(result_pred$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Get the contrasts and Wald stats for comparions
+     CI_hyp_test <- data.frame(
+              comparison = c("Snake (Hawaii-Kenya)", "Birds (Hawaii-Kenya)"), 
+                Estimate = c(robust_PredatorCC_pred_pop2[4,1], robust_PredatorCC_pred_pop1[3, 1]),
+                lwr = c(robust_PredatorCC_pred_pop2[4,4], robust_PredatorCC_pred_pop1[3, 4]),
+                upr = c(robust_PredatorCC_pred_pop2[4,5], robust_PredatorCC_pred_pop1[3, 5]),
+              Fstat = c(result_pred[5,3], result_pred[2,3]),
+              df = c(result_pred[5,5], result_pred[2,5]),
+              p_val = c(result_pred[5,7], result_pred[2,7]),
+              p_bonf = c(result_pred[5,8], result_pred[2,8]))
 
@@
   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-  robust_localAdapt <- clubSandwich::coef_test(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+  robust_localAdapt <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
@@
   DisplayBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = DisplayBgrd) 
+  summary(DisplayBgrd_pop_back)
 
-  # df to get t-distribution, use number of lizards not data points
-  hypthesis_test_preppop <- glht(DisplayBgrd_pop_back, df = 69-1, 
-               linfct = mcp(pop_back = c("Hawaii.Kenya  -Hawaii.Hawaii = 0")))
+  robust_DisplayBgrd_pop_back <- clubSandwich::conf_int(DisplayBgrd_pop_back, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
-  # Grab contrast stats for paper
-  summary(hypthesis_test_preppop)
-  p_displ_AC <- as.numeric(summary(hypthesis_test_preppop)$test$pvalues)
-   CI_dis_AC <- as.data.frame(confint(hypthesis_test_preppop)$confint)
+  # Contrasts hypothesis tests, but use RVE 
+    result_DisplayBgrd_pop_back <- Wald_test(DisplayBgrd_pop_back, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+    result_DisplayBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
+
+  # Grab contrasts that we need from the robust esatimtes
+  p_displ_AC <- as.numeric(result_DisplayBgrd_pop_back[2,8])
+   CI_dis_AC <- data.frame(Estimate=robust_DisplayBgrd_pop_back$beta[3],
+                           lwr = robust_DisplayBgrd_pop_back$CI_L[3],
+                           upr = robust_DisplayBgrd_pop_back$CI_U[3])
+
 ##########
@@
 
-    # Conduct hypothesis test, df to get t-distribution, use number of lizards not data points
-      hypthesis_test_preppop_CC <- glht(DisplayBgrd_pop_backCC, df = 69-1, 
-             linfct = mcp(pop_back = c("Hawaii.Kenya  -Hawaii.Hawaii = 0")))
+    # Fit Robust model
+    robust_DisplayBgrd_pop_backCC <- clubSandwich::conf_int(DisplayBgrd_pop_backCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
-    # Grab the statistics for the paper
-      summary(hypthesis_test_preppop_CC)
-      p_displ_CC <- as.numeric(summary(hypthesis_test_preppop_CC)$test$pvalues)
-      CI_Disp_CC <- as.data.frame(confint(hypthesis_test_preppop_CC)$confint)
+    # Contrasts hypothesis tests, but use RVE 
+      result_DisplayBgrd_pop_backCC <- Wald_test(DisplayBgrd_pop_backCC, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+      result_DisplayBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Grab contrasts that we need from the robust esatimtes
+      p_displ_CC <- result_DisplayBgrd_pop_backCC[2,8]
+       CI_Disp_CC <- data.frame(Estimate=robust_DisplayBgrd_pop_backCC$beta[3],
+                               lwr = robust_DisplayBgrd_pop_backCC$CI_L[3],
+                               upr = robust_DisplayBgrd_pop_backCC$CI_U[3])
 
@@
 
-  # Hypothesis tests
-      hypthesis_test_preppopCourtship <- glht(CourtshipBgrd_pop_back, df = 69-1, 
-                                        # df to get t-distribution, use number of lizards not data points
-             linfct = mcp(pop_back = c("Hawaii.Kenya  -Hawaii.Hawaii = 0")))
+   # Fit Robust model
+    robust_CourtshipBgrd_pop_back <- clubSandwich::conf_int(CourtshipBgrd_pop_back, cluster = CourtshipBgrd$ID, vcov =  "CR2")
 
-      summary(hypthesis_test_preppopCourtship)
+    # Contrasts hypothesis tests, but use RVE 
+      result_CourtshipBgrd_pop_back <- Wald_test(CourtshipBgrd_pop_back, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+      result_CourtshipBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
 
-  # Grab the stats for paper
-      p_courtship_AC <- as.numeric(summary(hypthesis_test_preppopCourtship)$test$pvalues)
-CI_AC_Courtship <- data.frame(confint(hypthesis_test_preppopCourtship)$confint)
+    # Grab contrasts that we need from the robust esatimtes
+      p_courtship_AC <- result_CourtshipBgrd_pop_back[2,8]
+       CI_AC_Courtship <- data.frame(Estimate=robust_CourtshipBgrd_pop_back$beta[3],
+                               lwr = robust_CourtshipBgrd_pop_back$CI_L[3],
+                               upr = robust_CourtshipBgrd_pop_back$CI_U[3])
 
@@
 
-  # Specific planned comparisons, # df to get t-distribution, use number of lizards not data points
-      hypthesis_test_preppopCourtshipCC <- glht(CourtshipBgrd_pop_backCC, df = 69-1, 
-             linfct = mcp(pop_back = c("Hawaii.Kenya - Hawaii.Hawaii = 0")))
+     # Fit Robust model
+    robust_CourtshipBgrd_pop_backCC <- clubSandwich::conf_int(CourtshipBgrd_pop_backCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
 
-  # Grab summary stats for paper
-    summary(hypthesis_test_preppopCourtshipCC)
-    p_courtship_CC <- as.numeric(summary(hypthesis_test_preppopCourtshipCC)$test$pvalues)
-    CI_CC_Courtship <- data.frame(confint(hypthesis_test_preppopCourtshipCC)$confint)
+    # Contrasts hypothesis tests, but use RVE 
+      result_CourtshipBgrd_pop_backCC <- Wald_test(CourtshipBgrd_pop_backCC, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+      result_CourtshipBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_backCC$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Grab contrasts that we need from the robust esatimtes
+      p_courtship_CC <- result_CourtshipBgrd_pop_backCC[2,8]
+       CI_CC_Courtship <- data.frame(Estimate=robust_CourtshipBgrd_pop_backCC$beta[3],
+                               lwr = robust_CourtshipBgrd_pop_backCC$CI_L[3],
+                               upr = robust_CourtshipBgrd_pop_backCC$CI_U[3])
+
 ```
@@
 Differences in chromatic contrast between populations, however,  depended on the predator (significant predator by population interaction: $F_{`r anova_predators_interCC$NumDF[3]`,`r anova_predators_interCC$DenDF[3]`}$ = `r anova_predators_interCC[5][3,]`, p = `r p_value(anova_predators_interCC[6][3,])`).
-Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,1]`, 95% CI = `r CI_hyp_test[1,2]` to `r CI_hyp_test[1,3]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,1]`, 95% CI = `r CI_hyp_test[2,2]` to `r CI_hyp_test[2,3]`, p = < 0.001).
+Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).
 

### 2021-05-04 supp.Rmd
Updated supplement with new fancy tables. Took a bit, but should be now easy to modify styles to suite what we need for all the contrast tables

@@
     Predator$PopBRInt <- interaction(Predator$Population,Predator$Predator, Predator$BodyRegion)
-     PredatorAC2 <- lmer(dS ~ -1 + PopBRInt + (1|ID), data=Predator)
+     PredatorAC2 <- lmer(dL ~ -1 + PopBRInt + (1|ID), data=Predator)
      summary(PredatorAC2)
@@
  # Contrasts hypothesis tests, but use RVE 
-                 result5 <- Wald_test(PredatorAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+                 result5 <- Wald_test(PredatorAC2, constraints = constrain_pairwise(c(1:16)), vcov="CR2", tidy = TRUE)
+ string <- c("PopBRIntKenya.snake.midflank - PopBRIntHawaii.snake.midflank|PopBRIntKenya.snake.gular - PopBRIntHawaii.snake.gular|PopBRIntKenya.snake.tailbase - PopBRIntHawaii.snake.tailbase|PopBRIntKenya.snake.topflank - PopBRIntHawaii.snake.topflank|PopBRIntKenya.bird.midflank - PopBRIntHawaii.bird.midflank|PopBRIntKenya.bird.gular - PopBRIntHawaii.bird.gular|PopBRIntKenya.bird.tailbase - PopBRIntHawaii.bird.tailbase|PopBRIntKenya.bird.topflank - PopBRIntHawaii.bird.topflank")
+    result5 <- result5[grep(string, result5$hypothesis),]
     result5$p_val_bonfer <- sapply(p.adjust(result5$p_val, "bonferroni"), function(x) p_value(x))
@@
 ##########
-## Chromatic Contrast - dL
+## Chromatic Contrast - dS
 ########
@@
  # Contrasts hypothesis tests, but use RVE 
-                 result6 <- Wald_test(PredatorCC2, constraints = constrain_pairwise(c(2:9), with_zero = TRUE), vcov="CR2", tidy = TRUE)
+                 result6 <- Wald_test(PredatorCC2, constraints = constrain_pairwise(c(1:16), with_zero = TRUE), vcov="CR2", tidy = TRUE)
+                   string <- c("PopBRIntKenya.snake.midflank - PopBRIntHawaii.snake.midflank|PopBRIntKenya.snake.gular - PopBRIntHawaii.snake.gular|PopBRIntKenya.snake.tailbase - PopBRIntHawaii.snake.tailbase|PopBRIntKenya.snake.topflank - PopBRIntHawaii.snake.topflank|PopBRIntKenya.bird.midflank - PopBRIntHawaii.bird.midflank|PopBRIntKenya.bird.gular - PopBRIntHawaii.bird.gular|PopBRIntKenya.bird.tailbase - PopBRIntHawaii.bird.tailbase|PopBRIntKenya.bird.topflank - PopBRIntHawaii.bird.topflank")
+    result6 <- result6[grep(string, result6$hypothesis),]
     result6$p_val_bonfer <- sapply(p.adjust(result6$p_val, "bonferroni"), function(x) p_value(x))
@@
 
-```{r tableS1}
-summary(result1)
+### Changes in visual signals across body regions show similar patterns
+We took measurements of spectral reflectance for different body regions of each chamaleon.
+In our primary analysis, we averaged across these different body regions because we were interested only in whether visual signals were different in Kenya and Hawaii overall.
+This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figure \@ref(fig:FigS1) and Table \@ref(tab:tableS1)).
+
+```{r tableS1, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
+Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
+Wald test uses a robust variance correction on the standard errors."}
+
+# Split snakes and birds
+result5_snake <-  result5[grep("snake", result5$hypothesis), ] %>% mutate(Species = c("Snake", rep("",3)))
+ result5_bird <- result5[grep("bird", result5$hypothesis), ] %>% mutate(Species = c("Bird", rep("",3)))
+
+ # Bind together in order
+  final_r5 <- rbind(result5_snake,result5_bird)
+
+  # Clean
+  final_r5 <- final_r5 %>% 
+              select(-test, -delta, -df_num) %>%
+              mutate(hypothesis = gsub("PopBRInt", "", hypothesis)) %>%
+              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
+
+table_style <- function(table){
+    table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
+             align(align = "center", part = "header") %>% 
+              align(i = 1:8, j = 1:6, align = "center", part = "body") %>% 
+            compose(part = "header", j = 1, value = as_paragraph(as_b("Predator"))) %>% 
+             compose( part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>% 
+            compose(part = "header", j = 3, value = as_paragraph(as_b("F"))) %>%
+            compose(part = "header", j = 4, value = as_paragraph(as_b("DF"))) %>% 
+            compose(part = "header", j = 5, value = as_paragraph(as_b("p"))) %>% 
+            compose(part = "header", j = 6, value = as_paragraph(as_b("p"), as_sub(as_b("Bonferroni")))) %>%
+            compose(part = "body", i = 1, j = 2, 
+                    value = as_paragraph("Kenya", as_sub("gular"), " - ", "Hawaii", as_sub("gular")))  %>% 
+            compose(part = "body", i = 2, j = 2, 
+                    value = as_paragraph("Kenya", as_sub("midflank"), " - ", "Hawaii", as_sub("midflank"))) %>% 
+            compose(part = "body", i = 3, j = 2, 
+                    value = as_paragraph("Kenya", as_sub("tailbase"), " - ", "Hawaii", as_sub("tailbase"))) %>% 
+            compose(part = "body", i = 4, j = 2, 
+                    value = as_paragraph("Kenya", as_sub("topflank"), " - ", "Hawaii", as_sub("topflank"))) %>% 
+            compose(part = "body", i = 5, j = 2, 
+                    value = as_paragraph("Kenya", as_sub("gular"), " - ", "Hawaii", as_sub("gular")))  %>% 
+            compose(part = "body", i = 6, j = 2, 
+                    value = as_paragraph("Kenya", as_sub("midflank"), " - ", "Hawaii", as_sub("midflank"))) %>% 
+            compose(part = "body", i = 7, j = 2, 
+                    value = as_paragraph("Kenya", as_sub("tailbase"), " - ", "Hawaii", as_sub("tailbase"))) %>% 
+            compose(part = "body", i = 8, j = 2, 
+                    value = as_paragraph("Kenya", as_sub("topflank"), " - ", "Hawaii", as_sub("topflank")))
+ return(table)
+}
+
+ft5 <- flextable(final_r5) %>%
+        table_style(.) %>% 
+        add_header_row(values = "Achromatic Contrast", colwidths = 6)
+
+ ft5
+
+```
+
+```{r tableS2, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
+Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
+Wald test uses a robust variance correction on the standard errors."}
+
+# Split snakes and birds
+
+result6_snake <- result6[grep("snake", result6$hypothesis), ] %>% mutate(Species = c("Snake", rep("",3)))
+ result6_bird <- result6[grep("bird", result6$hypothesis), ] %>% mutate(Species = c("Bird", rep("",3)))
+
+# Bind together in order
+ final_r6 <- rbind(result6_snake,result6_bird)
+
+# Clean
+ final_r6 <- final_r6 %>% 
+              select(-test, -delta, -df_num) %>%
+              mutate(hypothesis = gsub("PopBRInt", "", hypothesis)) %>%
+              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
+
+ # Table
+ ft6 <- flextable(final_r6) %>%
+        table_style(.) %>% 
+        add_header_row(values = "Chromatic Contrast", colwidths = 6)
+
+ ft6
 ```

### 2021-05-04 ms.Rmd
minor header change

@@
 
-  # Get model stats for ppaer
+  # Get model stats for paper
        coefAIC_court <- fixef(ConspicCourtshipAC)[3]

### 2021-05-04 supp.Rmd
Updated p_value function to properly work for whole vector now, cleans up table

@@
               select(-test, -delta, -df_num) %>%
-              mutate(hypothesis = gsub("PopBRInt", "", hypothesis)) %>%
+              mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
+                     p_val = sapply(p_val, function(x) p_value(x))) %>%
               select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)

### 2021-05-05 ms.Rmd
Corrected MS with wald tests

@@
 pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich)
-
-p_value <- function(x){
-  if(x <= 0.0001) {tmp = "< 0.0001"}
-  if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
-  if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
-  if(x >= 0.01) {tmp = round(x, digits =2)}
-  return(tmp)
-}
+source("./R/func.R")
 
@@
 ```{r Models, echo = FALSE, results = "hide"}
+source("./R/func.R")
 #########################################################
@@
 
-    # Get the coefficients, and ANOVA statistics for the paper
-          coefAIC_display <- fixef(ConspicDisplayAC)[3]
-            CI_AC_display <- data.frame(confint(ConspicDisplayAC))
-    anovaConspicDisplayAC <- anova(ConspicDisplayAC)
-
     # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
     Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
-       robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, cluster = Display_1_IDs$ID, vcov = "CR2")  
-    robust_ConspicDisplayAC_CI <- conf_int(ConspicDisplayAC, cluster = Display_1_IDs$ID, vcov = "CR2")
+       robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, 
+                                            cluster = Display_1_IDs$ID, vcov = "CR2")  
+    robust_ConspicDisplayAC_CI <- conf_int(ConspicDisplayAC, 
+                                           cluster = Display_1_IDs$ID, vcov = "CR2")
+
+     # Wald test
+            Population_ConspicDisplayAC <- Wald_test(ConspicDisplayAC, 
+                                             constraints = constrain_zero(3),
+                                            cluster = Display_1_IDs$ID, vcov = "CR2")
 
@@
 
-    # Get ANOVA statistics for model for the paper
-         anovaConspicDisplayCC <- anova(ConspicDisplayCC)
-
     # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
@@
 
+  # Wald test
+            Population_ConspicDisplayCC <- Wald_test(ConspicDisplayCC, 
+                                             constraints = constrain_zero(3),
+                                            cluster = Display_1_IDs$ID, vcov = "CR2")
 #########################################################
@@
 
-  # Get model stats for paper
-       coefAIC_court <- fixef(ConspicCourtshipAC)[3]
-         CI_AC_court <- data.frame(confint(ConspicCourtshipAC))
-    courship_anovaAC <- anova(ConspicCourtshipAC)
-
  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
     Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
-    robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, cluster = Courtship_IDs$ID, vcov = "CR2")
- robust_ConspicCourtshipAC_CI <- conf_int(ConspicCourtshipAC, cluster = Courtship_IDs$ID, vcov = "CR2")
+    robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, 
+                                           cluster = Courtship_IDs$ID, vcov = "CR2")
+ robust_ConspicCourtshipAC_CI <- conf_int(ConspicCourtshipAC, 
+                                          cluster = Courtship_IDs$ID, vcov = "CR2")
+
+    # Wald test
+            Population_ConspicCourtshipAC <- Wald_test(ConspicCourtshipAC, 
+                                             constraints = constrain_zero(3),
+                                            cluster = Courtship_IDs$ID, vcov = "CR2")
 
@@
 
-  # Get the ANOVA stats for paper
-    courship_anovaCC <- anova(ConspicCourtshipCC)
-
   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
@@
     robust_ConspicCourtshipCC_CI <- conf_int(ConspicCourtshipCC, cluster = Courtship_IDs$ID, vcov = "CR2")
+
+    # Wald test
+            Population_ConspicCourtshipCC <- Wald_test(ConspicCourtshipCC, 
+                                             constraints = constrain_zero(3),
+                                            cluster = Courtship_IDs$ID, vcov = "CR2")
 
@@
   # Fit model testing this idea above
-  PredatorAC_inter <- lmer(dL ~ Predator + Population + Predator:Population + (1|ID), data = Predator) 
+  PredatorAC_inter <- lmer(dL ~ Predator + Population + Predator:Population + (1|ID), 
+                           data = Predator) 
   summary(PredatorAC_inter)
@@
 
-  # Fit main effects model
+  #Refit main effects so we can test main effects on own.
     PredatorAC <- lmer(dL ~ Predator + Population + (1|ID), data = Predator) 
@@
     anova_predators <- anova(PredatorAC)
+
+     # Wald tests
+            back_pred_AC_Pred <- Wald_test(PredatorAC_inter, constraints = constrain_zero(4), 
+                                   cluster = Predator$ID, vcov = "CR2")
+
+             Predator_AC_Pred <- Wald_test(PredatorAC, constraints = constrain_zero(2), 
+                                  cluster = Predator$ID, vcov = "CR2")
+
+           Population_AC_Pred <- Wald_test(PredatorAC, constraints = constrain_zero(3), 
+                                  cluster = Predator$ID, vcov = "CR2")
 
@@
     # Fit the interaction model.
-    PredatorCC <- lmer(dS ~ Predator + Population + Predator:Population + (1|ID), data = Predator) 
+    PredatorCC <- lmer(dS ~ Predator + Population + Predator:Population + (1|ID), 
+                       data = Predator) 
     summary(PredatorCC)
 
-    # Get ANOVA statistics for the paper
-    anova_predators_interCC <- anova(PredatorCC)
+    #Refit main effects so we can test main effects on own.
+      PredatorCC_main <- lmer(dS ~ Predator + Population + (1|ID), data = Predator)
+
+     # Wald tests
+            back_pred_CC_Pred <- Wald_test(PredatorCC, constraints = constrain_zero(4), 
+                                   cluster = Predator$ID, vcov = "CR2")
+
+             Predator_CC_Pred <- Wald_test(PredatorCC_main, constraints = constrain_zero(2), 
+                                  cluster = Predator$ID, vcov = "CR2")
+
+           Population_CC_Pred <- Wald_test(PredatorCC_main, constraints = constrain_zero(3), 
+                                  cluster = Predator$ID, vcov = "CR2")
 
@@
 
+  # Add Wald test to test signifiacnce of each factor, background and population using roust variance
+  background_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(3), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  Population_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(2), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
   # Planned contrast. Fit the model in a slightly different way
@@
   # Grab contrasts that we need from the robust esatimtes
-  p_displ_AC <- as.numeric(result_DisplayBgrd_pop_back[2,8])
+  p_displ_AC <- result_DisplayBgrd_pop_back[2,8]
    CI_dis_AC <- data.frame(Estimate=robust_DisplayBgrd_pop_back$beta[3],
@@
       robust_DispBgrdCC <- clubSandwich::coef_test(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+    # Check significance of each factor
+        background_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(3), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  Population_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(2), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
 
@@
 
+   # Wald tests
+           background_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(3), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+           Population_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(2), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
   # Planned contrasts / hypothesis tests. Need to fit model a little differently
@@
   robust_CourtshipCC <- clubSandwich::coef_test(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+   # Wald tests
+           background_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(3), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+           Population_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(2), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
 
@@
 Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals whereas greater conspicuousness of sexual signals should be selected in social contexts.
-In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; Courtship displays: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, table S1).
-Kenyan chameleons from populations where predation is high were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r coefAIC_display`, 95% CI = `r CI_AC_display[5,1]` to  `r CI_AC_display[5,2]`, df = `r summary(ConspicDisplayAC)$coefficients[3, "df"]`, t =  `r summary(ConspicDisplayAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicDisplayAC)$coefficients[3, "Pr(>|t|)"])`, Courtship: $\beta$ = `r coefAIC_court`, 95% CI = `r CI_AC_court[5,1]` to  `r CI_AC_court[5,2]`, df = `r summary(ConspicCourtshipAC)$coefficients[3, "df"]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicCourtshipAC)$coefficients[3, "Pr(>|t|)"])`, see figure 1a & b).
+In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; Courtship displays: Acromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was  true across all body regions measured (See Supplemental Materials, table S1).
+Kenyan chameleons from populations where predation is high were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, Courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see figure 1a & b).
 
@@
 Differential predation pressure resulting from snake and bird predation varying would also result in populations varying in their conspicuousness to different predators.
-In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c; Achromatic Contrast: ($dl$); $F_{`r anova_predators$NumDF[2]`,`r anova_predators$DenDF[2]`}$ = `r anova_predators[5][2,]`, p = `r p_value(anova_predators[6][2,])`).
-Differences in chromatic contrast between populations, however,  depended on the predator (significant predator by population interaction: $F_{`r anova_predators_interCC$NumDF[3]`,`r anova_predators_interCC$DenDF[3]`}$ = `r anova_predators_interCC[5][3,]`, p = `r p_value(anova_predators_interCC[6][3,])`).
+In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c; Achromatic Contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`).
+Differences in chromatic contrast between populations, however,  depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
 Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).
@@
 Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment.
-While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chamleeons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_value(p_displ_AC)`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_value(p_displ_CC)`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_value(p_courtship_AC)`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_value(p_courtship_CC)`).
+While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chamleeons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
 

### 2021-05-05 R/func.R
Added p_value function to file to clean up a bit

@@ -1,0 +1,7 @@
+p_value <- function(x){
+  if(x <= 0.0001) {tmp = "< 0.0001"}
+  if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
+  if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
+  if(x >= 0.01) {tmp = round(x, digits =2)}
+  return(tmp)
+}

### 2021-05-05 ms.Rmd
Update to methods

@@
 Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii.
-Model interactions were first tesed using a Type III sum of squares ANOVA with a Saitterwaite degrees of freedom correction, and if found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
+Model interactions were first tested using Wald tests using robust variance estimators and a Saitterwaite degrees of freedom correction.
+If found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
+We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Saitterwaite degrees of freedom correction.
 For all individuals, we measured signals for multiple body regions, each of which were contratsed between multiple background measurements (i.e., top and bottoms of two leaves and two stems).

### 2021-05-05 supp.Rmd
Added local adaptation figure to suplement with body regions

@@
 
-```{r FigS2}
+```{r FigS3, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Local adaptation.
+All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
+Mean JNDs are provided along with 95% confidence intervals.
+A & B represent male-male contest displays whereas C & D represent courtship displays"}
+
+##########
+# Data
+##########
+courtship_summary <- CourtshipBgrd %>% 
+                    group_by(Population, Background_pop, BodyRegion) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>% filter(Population == "Hawaii") %>% as.data.frame() %>% mutate(Social_Context = "Courtship")
+
+display_summary <- DisplayBgrd %>%
+                group_by(Population, Background_pop, BodyRegion) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>% filter(Population == "Hawaii") %>% as.data.frame() %>% mutate(Social_Context = "Male-male contest")
+
+##########
+# Plots
+##########
+#Grab photos
+
+#plots
+p1.2 <- ggplot(courtship_summary, aes(x = Background_pop, y = Mean_dS, color = BodyRegion, group = BodyRegion)) +
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Background Environment") + 
+  ylab("Chromatic contrast (JND)") +
+  # annotate("segment", x = 0.85, xend = 1.85, y = 6.1, yend = 6.1, colour = "black") +
+  # annotate("segment", x = 1.12, xend = 2.12, y = 5.7, yend = 5.7, colour = "black") + 
+  # annotate("text", x = 0.85+(1.85-0.85)/2,  y = 6.2, label = "NS",  size = 6) + 
+  # annotate("text", x = 1.12+(2.12-1.12)/2, y = 5.8, label = "NS", size = 6) + 
+  theme_bw() +
+  labs(colour = "Body Region") +
+  theme(legend.position = c(0.15, 0.08),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+p2.2 <- ggplot(display_summary, aes(x = Background_pop, y = Mean_dL, group = BodyRegion, color = BodyRegion)) +
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Background Environment") + 
+  ylab("Achromatic contrast (JND)") +
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+#pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
+plot_grid(p1.2, p2.2, labels = c("A", "B"), label_size = 16,
+  label_fontface = "bold", rel_widths=c(1,1))
+
+```
+
+```{r FigS3}
 

### 2021-05-05 supp.Rmd
Updated functions. Added data for local adpation plots.

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, pander, kableExtra, flextable, jpeg, clubSandwich)
-
-p_value <- function(x){
-  if(x <= 0.0001) {tmp = "< 0.0001"}
-  if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
-  if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
-  if(x >= 0.01) {tmp = round(x, digits =2)}
-  return(tmp)
-}
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich)
+source("./R/func.R")
 ```
@@
 We took measurements of spectral reflectance for different body regions of each chamaleon.
-In our primary analysis, we averaged across these different body regions because we were interested only in whether visual signals were different in Kenya and Hawaii overall.
-This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figure \@ref(fig:FigS1) and Table \@ref(tab:tableS1)).
+In our primary analysis, we averaged across different body regions because we were interested only in whether visual signals were different in Kenya and Hawaii overall.
+This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figure \@ref(fig:FigS1) and Table \@ref(tab:tableS1), \@ref(tab:tableS2)).
 
@@
 
-```{r FigS3, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Local adaptation.
+```{r FigS2.1, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Local adaptation.
 All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
 A & B represent male-male contest displays whereas C & D represent courtship displays"}
+
+```
+
+```{r FigS2, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Local adaptation.
+All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
+Mean JNDs are provided along with 95% confidence intervals.
+A & B represent male-male contest displays whereas C & D represent courtship displays"}
+# Courship data
+
+CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+CourtshipBgrd <- CourtshipBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
+
+# Display data
+DisplayBgrd <- read.table("./data/Display_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+DisplayBgrd <- DisplayBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
 
@@
 ```
+
 # References

### 2021-05-05 supp.Rmd
Added preddator figure across body regions

@@
 In our primary analysis, we averaged across different body regions because we were interested only in whether visual signals were different in Kenya and Hawaii overall.
-This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figure \@ref(fig:FigS1) and Table \@ref(tab:tableS1), \@ref(tab:tableS2)).
+This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS1), \@ref(fig:FigS2), \@ref(fig:FigS3)) and Tables \@ref(tab:tableS1), \@ref(tab:tableS2)).
 
@@
 
-```{r FigS2.1, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Local adaptation.
-All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
+```{r FigS2, fig.width = 12, fig.height = 12, fig.cap = "Predators All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
 Mean JNDs are provided along with 95% confidence intervals.
@@
 
+# Load the data
+Predator <- read.table("./data/Predator_2_3.csv", 
+                       header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Clean up
+ Predator <- Predator %>% filter(!BodyRegion == "botflank" & !BodyRegion =='gularbot') %>% as.data.frame() #exclude bottom flank.
+Predator$BodyRegion <- factor(Predator$BodyRegion) #updates levels for the variable BodyRegion
+
+# Calculate summary data
+summary_data_pred <- Predator %>%
+                group_by(Predator, Population, BodyRegion) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>%
+                mutate(Predator = if_else(Predator == "bird", "Birds", "Snakes"))
+
+# PLot 
+p3 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dL, group = BodyRegion, color = BodyRegion)) + facet_wrap(~Predator) +
+  ylim(8, 30) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Achromatic contrast (JND)") +
+  theme_bw() +
+  labs(color = "Body Region") + 
+  theme(legend.position = c(0.85, 0.85), 
+        strip.background = element_blank(),
+    strip.text.x = element_text(size = 24, face = "bold"),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+p4 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dS, group = BodyRegion, color = BodyRegion)) + facet_wrap(~Predator) +
+  ylim(1, 9) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Chromatic contrast (JND)") +
+  labs(colour = "Body Region") + 
+  theme_bw() +
+  theme(legend.position = c(0.85, 0.85),
+    strip.background = element_blank(),
+    strip.text.x = element_text(size = 24, face = "bold"),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+plot_grid(p4, p3, nrow = 2, ncol = 1, labels = c("A", "B"), label_size = 16,
+  label_fontface = "bold", rel_widths=c(1,1))
+
 ```
 
-```{r FigS2, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Local adaptation.
+```{r FigS3, fig.width = 13, fig.height = 8, fig.cap = "Local adaptation.
 All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
 
-```{r FigS3}
+```{r FigS4, fig.cap = "Examples of chameleon colour signal change in response to different social and predatory contexts"}
 

### 2021-05-05 supp.Rmd
Fixes to figure legends. They should be ready to go now

--- 2021-05-05 supp.Rmd
@@
 
-```{r FigS1, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Chormatic (A, C) and Achromatic (B, D) contrast in just noticible differences (JND) for Hawaiian and Kenyan lizards relative to their environments.
+```{r FigS1, fig.width = 13, fig.height = 12, fig.cap = "Chormatic (A, C) and Achromatic (B, D) contrast in just noticible differences (JND) for Hawaiian and Kenyan lizards relative to their environments.
 All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
   labs(color = "Body Region") + 
-  theme(#legend.position = "none",
+  theme(legend.position = c(0.85, 0.85),
     panel.grid.major = element_blank(), 
@@
   theme_bw() +
-  theme(legend.position = "none",
+  theme(legend.position = c(0.85, 0.85),
     panel.grid.major = element_blank(), 
@@
 
-p1 + p2 + p3 + p4 + plot_annotation(tag_levels = c('A', 'B', 'C', 'D')) & theme(plot.tag = element_text(face = 'bold', size=20))
+p1 + p2 + p3 + p4 + plot_annotation(tag_levels = c('A', 'B', 'C', 'D')) & theme(plot.tag = element_text(face = 'bold', size=30))
 
@@
 
+# Load the data
+Predator <- read.table("./data/Predator_2_3.csv", 
+                       header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Clean up
+ Predator <- Predator %>% filter(!BodyRegion == "botflank" & !BodyRegion =='gularbot') %>% as.data.frame() #exclude bottom flank.
+Predator$BodyRegion <- factor(Predator$BodyRegion) #updates levels for the variable BodyRegion
+
+# Calculate summary data
+summary_data_pred <- Predator %>%
+                group_by(Predator, Population, BodyRegion) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>%
+                mutate(Predator = if_else(Predator == "bird", "Birds", "Snakes"))
+
+# PLot 
+p3 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dL, group = BodyRegion, color = BodyRegion)) + facet_wrap(~Predator) +
+  ylim(8, 30) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Achromatic contrast (JND)") +
+  theme_bw() +
+  labs(color = "Body Region") + 
+  theme(legend.position = c(0.90, 0.85), 
+        strip.background = element_blank(),
+    strip.text.x = element_text(size = 24, face = "bold"),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+p4 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dS, group = BodyRegion, color = BodyRegion)) + facet_wrap(~Predator) +
+  ylim(1, 9) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Chromatic contrast (JND)") +
+  labs(colour = "Body Region") + 
+  theme_bw() +
+  theme(legend.position = c(0.90, 0.85),
+    strip.background = element_blank(),
+    strip.text.x = element_text(size = 24, face = "bold"),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+plot_grid(p4, p3, nrow = 2, ncol = 1, labels = c("A", "B"), label_size = 30,
+  label_fontface = "bold", rel_widths=c(1,1))
+
 ```
 
-```{r FigS2, fig.width = 14.33333, fig.height = 11.00000, fig.cap = "Local adaptation.
+```{r FigS3, fig.width = 13, fig.height = 7.5, fig.cap = "Local adaptation.
 All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
   labs(colour = "Body Region") +
-  theme(legend.position = c(0.15, 0.08),
+  theme(legend.position = "none",
     panel.grid.major = element_blank(), 
@@
   theme_bw() +
-  theme(legend.position = "none",
+  theme(legend.position = c(0.80, 0.85),
     panel.grid.major = element_blank(), 
@@
 #pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
-plot_grid(p1.2, p2.2, labels = c("A", "B"), label_size = 16,
+plot_grid(p1.2, p2.2, labels = c("A", "B"), label_size = 30,
   label_fontface = "bold", rel_widths=c(1,1))

### 2021-05-06 R/func.R
Updated supp to fix dataframe for table output

@@ -5,3 +5,33 @@ p_value <- function(x){
   if(x >= 0.01) {tmp = round(x, digits =2)}
   return(tmp)
 }
+
+
+wald_table <- function(model_main, model_inter){
+  
+       inter <- Wald_test(model_inter, 
+                                        constraints = constrain_zero(7:9),
+                                        vcov="CR2", tidy = TRUE)
+  BodyRegion <- Wald_test(model_main, 
+                                             constraints = constrain_zero(4:6),
+                                             vcov="CR2", tidy = TRUE)
+  SVL <- Wald_test(model_main, 
+                                      constraints = constrain_zero(2),
+                                      vcov="CR2", tidy = TRUE)
+  Pop <- Wald_test(model_main, 
+                                      constraints = constrain_zero(3),
+                                      vcov="CR2", tidy = TRUE)
+  
+  WaldTab <- rbind(SVL,
+                                      Pop, 
+                                      BodyRegion, 
+                                      inter)
+  WaldTab <- WaldTab %>% 
+    mutate(Variable = c("SVL", 
+                        "Population", 
+                        "Body Region", 
+                        "Population:Body Region")) %>% 
+    select(Variable, Fstat, df_num, df_denom, p_val) %>% as.data.frame()
+  
+  return(WaldTab)
+}

### 2021-05-06 supp.Rmd
Added new functions for creating Wald Tables, and added all the tables.

@@
     summary(ConspicDisplayAC)
+    ConspicDisplayAC_main <- lmer(dL ~ sc_SVL + Population + BodyRegion +  (1|ID), data=Display_1)
+
+    # Wald Table - Test for interaction, plus refit model and test sig of each factor. Uses RVEs
+    wald_table_ConspicDisplayAC <-  wald_table(model_main =  ConspicDisplayAC_main , model_inter = ConspicDisplayAC)
 
@@
     ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
+    ConspicDisplayCC_main <- lmer(dS ~ sc_SVL + Population + BodyRegion + (1|ID), data=Display_1)
+
+# Wald Table - Test for interaction, plus refit model and test sig of each factor. Uses RVEs
+    wald_table_ConspicDisplayCC <-  wald_table(model_main =  ConspicDisplayCC_main, model_inter = ConspicDisplayCC)
 
@@
     summary(ConspicCourtshipAC)
-      anova(ConspicCourtshipAC)
 
+  # Main effects Model
+    ConspicCourtshipAC_main <- lmer(dL ~ sc_SVL + Population + BodyRegion + (1|ID), data=Courtship)
+
+   # Wald Table - Test for interaction, plus refit model and test sig of each factor. Uses RVEs
+    wald_table_ConspicCourtshipAC <-  wald_table(model_main =  ConspicCourtshipAC_main, model_inter = ConspicCourtshipAC)  
   # Robust variance estimators (RVE). Mostly all consistent
@@
 
-    # Fit model
+    # Fit model, interaction
         ConspicCourtshipCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Courtship)
+
+    # Fit main effects
+    ConspicCourtshipCC_main <- lmer(dS ~ sc_SVL + Population + BodyRegion + (1|ID), data=Courtship)
+
+    # Wald Table - Test for interaction, plus refit model and test sig of each factor. Uses RVEs
+    wald_table_ConspicCourtshipCC <-  wald_table(model_main =  ConspicCourtshipCC_main, model_inter = ConspicCourtshipCC)
 
@@
     # Contrasts hypothesis tests, but use RVE 
-                 result4 <- Wald_test(ConspicCourtshipCC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
+            result4 <- Wald_test(ConspicCourtshipCC2, 
+                                 constraints = constrain_pairwise(c(2:9)), 
+                                 vcov="CR2", tidy = TRUE)
+      string <- c("PopBRIntKenya.gular - PopBRIntHawaii.gular|PopBRIntKenya.midflank - PopBRIntHawaii.midflank|PopBRIntKenya.tailbase - PopBRIntHawaii.tailbase|PopBRIntKenya.topflank - PopBRIntHawaii.topflank")
+      result4 <- result4[grep(string, result4$hypothesis),]
     result4$p_val_bonfer <- sapply(p.adjust(result4$p_val, "bonferroni"), function(x) p_value(x))

### 2021-05-06 R/func.R
New functins added for flextable styles

@@ -7,7 +7,7 @@ p_value <- function(x){
 }
 
 
-wald_table <- function(model_main, model_inter){
+  wald_table <- function(model_main, model_inter){
   
        inter <- Wald_test(model_inter, 
                                         constraints = constrain_zero(7:9),
@@ -34,4 +34,85 @@ wald_table <- function(model_main, model_inter){
     select(Variable, Fstat, df_num, df_denom, p_val) %>% as.data.frame()
   
   return(WaldTab)
+  }
+  
+  
+  wald_table_p <- function(model_main, model_inter){
+    
+    inter_BR_Pred <- Wald_test(model_inter, 
+                       constraints = constrain_zero(7:9),
+                       vcov="CR2", tidy = TRUE)
+    
+    inter_Pop_Pred <- Wald_test(model_inter, 
+                               constraints = constrain_zero(10),
+                               vcov="CR2", tidy = TRUE)
+    
+    BodyRegion <- Wald_test(model_main, 
+                            constraints = constrain_zero(4:6),
+                            vcov="CR2", tidy = TRUE)
+    Pred <- Wald_test(model_main, 
+                     constraints = constrain_zero(2),
+                     vcov="CR2", tidy = TRUE)
+    Pop <- Wald_test(model_main, 
+                     constraints = constrain_zero(3),
+                     vcov="CR2", tidy = TRUE)
+    
+    WaldTab <- rbind(BodyRegion,
+                     Pop, 
+                     Pred, 
+                     inter_BR_Pred,
+                     inter_Pop_Pred)
+    WaldTab <- WaldTab %>% 
+      mutate(Variable = c("Body Region", 
+                          "Population", 
+                          "Predator", 
+                          "Predator:Body Region",
+                          "Predator:Population")) %>% 
+      select(Variable, Fstat, df_num, df_denom, p_val) %>% as.data.frame()
+    
+    return(WaldTab)
+  }
+
+ table_style <- function(table){
+  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
+    align(align = "center", part = "header") %>% 
+    align(i = 1:8, j = 1:6, align = "center", part = "body") %>% 
[…]

### 2021-05-06 R/func.R
Function 3 style

@@ -116,3 +116,16 @@ table_style2 <- function(table){
     compose(part = "header", j = 7, value = as_paragraph(as_b("p"))) 
   return(table)
 }
+
+table_style3 <- function(table){
+  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
+    align(align = "center", part = "header") %>% 
+    align(i = 1:8, j = 1:6, align = "center", part = "body") %>% 
+    compose(part = "header", j = 1, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
+    compose(part = "header", j = 2, value = as_paragraph(as_b("Variable"))) %>% 
+    compose( part = "header", j = 3, value = as_paragraph(as_b("F"))) %>% 
+    compose(part = "header", j = 4, value = as_paragraph(as_b("df"), as_sub(as_b("num")))) %>%
+    compose(part = "header", j = 5, value = as_paragraph(as_b("df"), as_sub(as_b("denom")))) %>% 
+    compose(part = "header", j = 6, value = as_paragraph(as_b("p"))) 
+  return(table)
+}

### 2021-05-06 supp.Rmd
Functions and supplemental tables for Wald tests done

@@
 ```{r display, results = "hide"}
-
+source("./R/func.R")
 #########################################################
@@
    # Fit model
-      PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion + Predator:BodyRegion + Predator:Population + (1|ID), data=Predator)
+      PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion + Predator:BodyRegion + Predator:Population + 
+                           (1|ID), data=Predator)
       summary(PredatorAC)
-      anova(PredatorAC)
+
+  PredatorAC_main <- lmer(dL ~ Predator + Population + BodyRegion  + (1|ID), data=Predator)
+
+  # Wald table RVE
+
+    Wald_table_PredatorAC <- wald_table_p(PredatorAC_main, PredatorAC)
 
@@
           anova(PredatorAC)
+
+    PredatorCC_main <- lmer(dS ~ Predator + Population + BodyRegion  + (1|ID), data=Predator)
+
+  # Wald table RVE
+
+    Wald_table_PredatorCC <- wald_table_p(PredatorCC_main, PredatorCC)
 
@@
 
-table_style <- function(table){
-    table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
-             align(align = "center", part = "header") %>% 
-              align(i = 1:8, j = 1:6, align = "center", part = "body") %>% 
-            compose(part = "header", j = 1, value = as_paragraph(as_b("Predator"))) %>% 
-             compose( part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>% 
-            compose(part = "header", j = 3, value = as_paragraph(as_b("F"))) %>%
-            compose(part = "header", j = 4, value = as_paragraph(as_b("DF"))) %>% 
-            compose(part = "header", j = 5, value = as_paragraph(as_b("p"))) %>% 
-            compose(part = "header", j = 6, value = as_paragraph(as_b("p"), as_sub(as_b("Bonferroni")))) %>%
-            compose(part = "body", i = 1, j = 2, 
-                    value = as_paragraph("Kenya", as_sub("gular"), " - ", "Hawaii", as_sub("gular")))  %>% 
-            compose(part = "body", i = 2, j = 2, 
-                    value = as_paragraph("Kenya", as_sub("midflank"), " - ", "Hawaii", as_sub("midflank"))) %>% 
-            compose(part = "body", i = 3, j = 2, 
-                    value = as_paragraph("Kenya", as_sub("tailbase"), " - ", "Hawaii", as_sub("tailbase"))) %>% 
-            compose(part = "body", i = 4, j = 2, 
-                    value = as_paragraph("Kenya", as_sub("topflank"), " - ", "Hawaii", as_sub("topflank"))) %>% 
-            compose(part = "body", i = 5, j = 2, 
-                    value = as_paragraph("Kenya", as_sub("gular"), " - ", "Hawaii", as_sub("gular")))  %>% 
-            compose(part = "body", i = 6, j = 2, 
-                    value = as_paragraph("Kenya", as_sub("midflank"), " - ", "Hawaii", as_sub("midflank"))) %>% 
-            compose(part = "body", i = 7, j = 2, 
-                    value = as_paragraph("Kenya", as_sub("tailbase"), " - ", "Hawaii", as_sub("tailbase"))) %>% 
-            compose(part = "body", i = 8, j = 2, 
-                    value = as_paragraph("Kenya", as_sub("topflank"), " - ", "Hawaii", as_sub("topflank")))
- return(table)
-}
-
 ft5 <- flextable(final_r5) %>%
@@
  ft6
+```
+
+```{r tableS3, tab.cap = "Wald Tests comparing significance of model parameters.
+Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
+Interactions were first tested and models refit to test the signifiacnce of main effects"}
+fullTable <- rbind(wald_table_ConspicDisplayAC, 
+                   wald_table_ConspicDisplayCC, 
+                   wald_table_ConspicCourtshipAC,
+                   wald_table_ConspicCourtshipCC)
+
+fullTable <- fullTable %>% 
+              mutate(Context = c("Male-Male Display", rep("", 7), "Courtship Display", rep("", 7)),
+                     JND = c("Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3), "Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3)),
+                     p_val = sapply(p_val, function(x) p_value(x))) %>% select(Context, JND, Variable, Fstat, df_num, df_denom, p_val)
+
+S3 <- flextable(fullTable) %>%
+        table_style2(.)
+S3
+```
+
+```{r table4, tab.cap = "Wald Tests comparing significance of model parameters for predators.
+Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
+Interactions were first tested and models refit to test the signifiacnce of main effects"}
+
+fulltableP <- rbind(Wald_table_PredatorAC,
+                    Wald_table_PredatorCC)
+
+fulltableP <- fulltableP %>% mutate(JND = c("Achromatic Contrast", rep("",4), "Chromatc Contrast", rep("",4)),
+                                            p_val = sapply(p_val, function(x) p_value(x))) %>% 
+                                      select(JND, Variable, Fstat, df_num, df_denom, p_val)
+
+S4 <- flextable(fulltableP) %>%
+        table_style3(.)
+S4
 ```

### 2021-05-06 R/func.R

@@ -104,7 +104,7 @@ p_value <- function(x){
 
 
 table_style2 <- function(table){
-  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
+  table <- table %>% width(j = 1, width = 1) %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
     align(align = "center", part = "header") %>% 
     align(i = 1:8, j = 1:6, align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Social Context"))) %>% 

### 2021-05-06 supp.Rmd
Updated style tables and supplement. Corrected a few errors

@@
     # Fit the model that estimates changes across body regions
-    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
+    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + 
+                               Population:BodyRegion + (1|ID), data=Display_1)
     summary(ConspicDisplayAC)
@@
   # Fit model
-    ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Courtship)
+    ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + 
+                                 Population:BodyRegion + (1|ID), data=Courtship)
     summary(ConspicCourtshipAC)
@@
 S4
+```
+
+```{r tableS5, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction."}
+robust_sc_table <- rbind(robust_ConspicDisplayAC,
+                          robust_ConspicDisplayCC,
+                          robust_ConspicCourtshipAC,
+                          robust_ConspicCourtshipCC) %>% 
+                    mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)),
+                           SocialContext = c("Male-Male Display", rep("", 17), 
+                                             "Courtship Display", rep("", 17)),
+                           JND = c("Achromatic Contrast", rep("", 8), 
+                                   "Chromatic Contrast", rep("", 8), 
+                                   "Achromatic Contrast", rep("", 8), 
+                                   "Chromatic Contrast", rep("", 8)))
+
+robust_sc_table$parameter <- rownames(robust_sc_table)
+
+robust_sc_table <- robust_sc_table %>% select(SocialContext, JND, parameter, beta, SE, tstat, df, p_Satt)
+
+tableS5 <- flextable(robust_sc_table) %>%  table_style4()
+tableS5
 ```

### 2021-05-06 R/func.R

@@ -104,9 +104,9 @@ p_value <- function(x){
 
 
 table_style2 <- function(table){
-  table <- table %>% width(j = 1, width = 1) %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
+  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 1) %>% width(j = 2, width = 2.5) %>% 
     align(align = "center", part = "header") %>% 
-    align(i = 1:8, j = 1:6, align = "center", part = "body") %>% 
+    align(i = 1:16, j = 1:7, align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Social Context"))) %>% 
     compose(part = "header", j = 2, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
     compose(part = "header", j = 3, value = as_paragraph(as_b("Variable"))) %>% 
@@ -120,12 +120,40 @@ table_style2 <- function(table){
 table_style3 <- function(table){
   table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
     align(align = "center", part = "header") %>% 
-    align(i = 1:8, j = 1:6, align = "center", part = "body") %>% 
+    align(i = 1:10, j = 1:6, align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
     compose(part = "header", j = 2, value = as_paragraph(as_b("Variable"))) %>% 
     compose( part = "header", j = 3, value = as_paragraph(as_b("F"))) %>% 
     compose(part = "header", j = 4, value = as_paragraph(as_b("df"), as_sub(as_b("num")))) %>%
     compose(part = "header", j = 5, value = as_paragraph(as_b("df"), as_sub(as_b("denom")))) %>% 
     compose(part = "header", j = 6, value = as_paragraph(as_b("p"))) 
+  return(table)
+}
+
+
+table_style4 <- function(table){
+  table <- table %>% width(j = 1, width = 1) %>% 
+                    width(j = 2, width = 1) %>% 
+                    width(j = 3, width = 3) %>% 
+    align(align = "center", part = "header") %>% 
+    align(i = 1:36, j = 1:7, align = "center", part = "body") %>% 
[…]

### 2021-05-07 R/func.R
Updated table styles functions. Added annotation

@@ -1,3 +1,5 @@
+
+# Cleans up look of p_values
 p_value <- function(x){
   if(x <= 0.0001) {tmp = "< 0.0001"}
   if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
@@ -6,8 +8,8 @@ p_value <- function(x){
   return(tmp)
 }
 
-
-  wald_table <- function(model_main, model_inter){
+# Create wald tables. Functions using RVE's to look at overall significance of interactions and main effects in models
+wald_table   <- function(model_main, model_inter){
   
        inter <- Wald_test(model_inter, 
                                         constraints = constrain_zero(7:9),
@@ -36,8 +38,7 @@ p_value <- function(x){
   return(WaldTab)
   }
   
-  
-  wald_table_p <- function(model_main, model_inter){
+wald_table_p <- function(model_main, model_inter){
     
     inter_BR_Pred <- Wald_test(model_inter, 
                        constraints = constrain_zero(7:9),
@@ -73,10 +74,12 @@ p_value <- function(x){
     return(WaldTab)
   }
 
- table_style <- function(table){
+# Table styling functions
+
+table_style  <- function(table){
   table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
     align(align = "center", part = "header") %>% 
-    align(i = 1:8, j = 1:6, align = "center", part = "body") %>% 
+    align(align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Predator"))) %>% 
     compose( part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>% 
     compose(part = "header", j = 3, value = as_paragraph(as_b("F"))) %>%
@@ -102,11 +105,10 @@ p_value <- function(x){
   return(table)
 }
 
-
 table_style2 <- function(table){
-  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 1) %>% width(j = 2, width = 2.5) %>% 
+  table <- table %>%  width(j = 3, width = 3) %>% 
     align(align = "center", part = "header") %>% 
-    align(i = 1:16, j = 1:7, align = "center", part = "body") %>% 
+    align( align = "center", part = "body") %>% 
[…]

### 2021-05-07 supp.Rmd
Re-oragnised tables and figures so they flow on from each other. Renamed chuncks

--- 2021-05-06 supp.Rmd
@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
 source("./R/func.R")
@@
 In our primary analysis, we averaged across different body regions because we were interested only in whether visual signals were different in Kenya and Hawaii overall.
-This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS1), \@ref(fig:FigS2), \@ref(fig:FigS3)) and Tables \@ref(tab:tableS1), \@ref(tab:tableS2)).
-
-```{r tableS1, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
-Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
-Wald test uses a robust variance correction on the standard errors."}
-
-# Split snakes and birds
-result5_snake <-  result5[grep("snake", result5$hypothesis), ] %>% mutate(Species = c("Snake", rep("",3)))
- result5_bird <- result5[grep("bird", result5$hypothesis), ] %>% mutate(Species = c("Bird", rep("",3)))
-
- # Bind together in order
-  final_r5 <- rbind(result5_snake,result5_bird)
-
-  # Clean
-  final_r5 <- final_r5 %>% 
-              select(-test, -delta, -df_num) %>%
-              mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
-                     p_val = sapply(p_val, function(x) p_value(x))) %>%
-              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
-
-ft5 <- flextable(final_r5) %>%
-        table_style(.) %>% 
-        add_header_row(values = "Achromatic Contrast", colwidths = 6)
-
- ft5
-
-```
-
-```{r tableS2, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
-Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
-Wald test uses a robust variance correction on the standard errors."}
-
-# Split snakes and birds
-
-result6_snake <- result6[grep("snake", result6$hypothesis), ] %>% mutate(Species = c("Snake", rep("",3)))
- result6_bird <- result6[grep("bird", result6$hypothesis), ] %>% mutate(Species = c("Bird", rep("",3)))
-
-# Bind together in order
- final_r6 <- rbind(result6_snake,result6_bird)
-
-# Clean
- final_r6 <- final_r6 %>% 
-              select(-test, -delta, -df_num) %>%
-              mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
-                     p_val = sapply(p_val, function(x) p_value(x))) %>%
-              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
-
- # Table
- ft6 <- flextable(final_r6) %>%
-        table_style(.) %>% 
-        add_header_row(values = "Chromatic Contrast", colwidths = 6)
-
- ft6
-```
-
-```{r tableS3, tab.cap = "Wald Tests comparing significance of model parameters.
-Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
-Interactions were first tested and models refit to test the signifiacnce of main effects"}
-fullTable <- rbind(wald_table_ConspicDisplayAC, 
-                   wald_table_ConspicDisplayCC, 
-                   wald_table_ConspicCourtshipAC,
-                   wald_table_ConspicCourtshipCC)
-
-fullTable <- fullTable %>% 
-              mutate(Context = c("Male-Male Display", rep("", 7), "Courtship Display", rep("", 7)),
-                     JND = c("Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3), "Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3)),
-                     p_val = sapply(p_val, function(x) p_value(x))) %>% select(Context, JND, Variable, Fstat, df_num, df_denom, p_val)
-
-S3 <- flextable(fullTable) %>%
-        table_style2(.)
-S3
-```
-
-```{r table4, tab.cap = "Wald Tests comparing significance of model parameters for predators.
-Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
-Interactions were first tested and models refit to test the signifiacnce of main effects"}
-
-fulltableP <- rbind(Wald_table_PredatorAC,
-                    Wald_table_PredatorCC)
-
-fulltableP <- fulltableP %>% mutate(JND = c("Achromatic Contrast", rep("",4), "Chromatc Contrast", rep("",4)),
-                                            p_val = sapply(p_val, function(x) p_value(x))) %>% 
-                                      select(JND, Variable, Fstat, df_num, df_denom, p_val)
-
-S4 <- flextable(fulltableP) %>%
-        table_style3(.)
-S4
-```
+This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS1), \@ref(fig:FigS2), \@ref(fig:FigS3)) and Tables \@ref(tab:tableS1), \@ref(tab:tableS2), \@ref(tab:tableS3), \@ref(tab:tableS4).
+\@ref(tab:tableS5)).
 
@@
 
+```{r tableS2, tab.cap = "Wald Tests comparing significance of model parameters.
+Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
+Interactions were first tested and models refit to test the signifiacnce of main effects"}
+fullTable <- rbind(wald_table_ConspicDisplayAC, 
+                   wald_table_ConspicDisplayCC, 
+                   wald_table_ConspicCourtshipAC,
+                   wald_table_ConspicCourtshipCC)
+
+fullTable <- fullTable %>% 
+              mutate(Context = c("Male-Male Display", rep("", 7), "Courtship Display", rep("", 7)),
+                     JND = c("Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3), "Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3)),
+                     p_val = sapply(p_val, function(x) p_value(x))) %>% select(Context, JND, Variable, Fstat, df_num, df_denom, p_val)
+
+S3 <- flextable(fullTable) %>%
+        table_style2(.)
+S3
+```
+
+```{r tableS3, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction.Social Context"}
+robust_sc_table <- rbind(robust_ConspicDisplayAC,
+                          robust_ConspicDisplayCC,
+                          robust_ConspicCourtshipAC,
+                          robust_ConspicCourtshipCC) %>% 
+                    mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)),
+                           SocialContext = c("Male-Male Display", rep("", 17), 
+                                             "Courtship Display", rep("", 17)),
+                           JND = c("Achromatic Contrast", rep("", 8), 
+                                   "Chromatic Contrast", rep("", 8), 
+                                   "Achromatic Contrast", rep("", 8), 
+                                   "Chromatic Contrast", rep("", 8)))
+
+robust_sc_table$parameter <- rownames(robust_sc_table)
+
+robust_sc_table <- robust_sc_table %>% select(SocialContext, JND, parameter, beta, SE, tstat, df, p_Satt)
+
+tableS5 <- flextable(robust_sc_table) %>%  table_style4()
+tableS5
+```
+
 ```{r FigS2, fig.width = 12, fig.height = 12, fig.cap = "Predators All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
 
+```
+
+```{r tableS4, tab.cap = "Wald Tests comparing significance of model parameters for predators.
+Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
+Interactions were first tested and models refit to test the signifiacnce of main effects"}
+
+fulltableP <- rbind(Wald_table_PredatorAC,
+                    Wald_table_PredatorCC)
+
+fulltableP <- fulltableP %>% mutate(JND = c("Achromatic Contrast", rep("",4), "Chromatc Contrast", rep("",4)),
+                                            p_val = sapply(p_val, function(x) p_value(x))) %>% 
+                                      select(JND, Variable, Fstat, df_num, df_denom, p_val)
+
+S4 <- flextable(fulltableP) %>%
+        table_style3(.)
+S4
+```
+
+```{r tableS5, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction.Predators"}
+robust_sc_pred_table <- rbind(robust_PredatorAC,
+                         robust_PredatorCC) %>% 
+                    mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)),
+                           JND = c("Achromatic Contrast", rep("", 9), 
+                                   "Chromatic Contrast", rep("", 9)))
+
+robust_sc_pred_table$parameter <- rownames(robust_sc_pred_table)
+
+robust_sc_pred_table <- robust_sc_pred_table %>% select(JND, parameter, beta, SE, tstat, df, p_Satt)
+
+tableS6 <- flextable(robust_sc_pred_table) %>%  table_style5()
+tableS6
+```
+
+```{r tableS6, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
+Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
+Wald test uses a robust variance correction on the standard errors."}
+
+# Split snakes and birds
+result5_snake <-  result5[grep("snake", result5$hypothesis), ] %>% mutate(Species = c("Snake", rep("",3)))
+ result5_bird <- result5[grep("bird", result5$hypothesis), ] %>% mutate(Species = c("Bird", rep("",3)))
+
+ # Bind together in order
+  final_r5 <- rbind(result5_snake,result5_bird)
+
+  # Clean
+  final_r5 <- final_r5 %>% 
+              select(-test, -delta, -df_num) %>%
+              mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
+                     p_val = sapply(p_val, function(x) p_value(x))) %>%
+              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
+
+ft5 <- flextable(final_r5) %>%
+        table_style(.) %>% 
+        add_header_row(values = "Achromatic Contrast", colwidths = 6)
+
+ ft5
+
+```
+
+```{r tableS7, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
+Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
+Wald test uses a robust variance correction on the standard errors."}
+
+# Split snakes and birds
+
+result6_snake <- result6[grep("snake", result6$hypothesis), ] %>% mutate(Species = c("Snake", rep("",3)))
+ result6_bird <- result6[grep("bird", result6$hypothesis), ] %>% mutate(Species = c("Bird", rep("",3)))
+
+# Bind together in order
+ final_r6 <- rbind(result6_snake,result6_bird)
+
+# Clean
+ final_r6 <- final_r6 %>% 
+              select(-test, -delta, -df_num) %>%
+              mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
+                     p_val = sapply(p_val, function(x) p_value(x))) %>%
+              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
+
+ # Table
+ ft6 <- flextable(final_r6) %>%
+        table_style(.) %>% 
+        add_header_row(values = "Chromatic Contrast", colwidths = 6)
+
+ ft6
 ```

### 2021-05-07 supp.Rmd
reorganised figures and tables so they follow on from each other.

@@
 
+```{r FigS1, fig.cap = "Examples of chameleon colour signal change in response to different social and predatory contexts"}
+
+files <- list.files("./photos/ESM/")
+img1 <- readJPEG(paste0("./photos/ESM/", files[1]))
+img2 <- readJPEG(paste0("./photos/ESM/", files[2]))
+img3 <- readJPEG(paste0("./photos/ESM/", files[3]))
+img4 <- readJPEG(paste0("./photos/ESM/", files[4]))
+img5 <- readJPEG(paste0("./photos/ESM/", files[5]))
+img6 <- readJPEG(paste0("./photos/ESM/", files[6]))
+
+p1 <- cowplot::ggdraw()+
+        cowplot::draw_image(img1)
+p2 <- cowplot::ggdraw()+
+        cowplot::draw_image(img2)
+p3 <- cowplot::ggdraw()+
+        cowplot::draw_image(img3)
+p4 <- cowplot::ggdraw()+
+        cowplot::draw_image(img4)
+p5 <- cowplot::ggdraw()+
+        cowplot::draw_image(img5)
+p6 <- cowplot::ggdraw()+
+        cowplot::draw_image(img6)
+
+(p1 + p3 + p6) / (p2 + p5 + p4) + plot_annotation(tag_levels = c("A", "B", "C", "D", "E", "F"))
+
+```
+
 ## Colour measurements, habitat background and irradiance
@@
 
-### Changes in visual signals across body regions show similar patterns
+## Assessing changes in sexual signals across different body regions
 We took measurements of spectral reflectance for different body regions of each chamaleon.
-In our primary analysis, we averaged across different body regions because we were interested only in whether visual signals were different in Kenya and Hawaii overall.
-This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS1), \@ref(fig:FigS2), \@ref(fig:FigS3)) and Tables \@ref(tab:tableS1), \@ref(tab:tableS2), \@ref(tab:tableS3), \@ref(tab:tableS4).
+In our primary analysis, we averaged across different body regions because we were interested only in whether visual signals overall were different between Kenya and Hawaii.
+This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS2), \@ref(fig:FigS3), \@ref(fig:FigS4) and Tables \@ref(tab:tableS1), \@ref(tab:tableS2), \@ref(tab:tableS3), \@ref(tab:tableS4).
 \@ref(tab:tableS5)).
 
-```{r FigS1, fig.width = 13, fig.height = 12, fig.cap = "Chormatic (A, C) and Achromatic (B, D) contrast in just noticible differences (JND) for Hawaiian and Kenyan lizards relative to their environments.
+#### Body regions in social contexts (Male-male and Courtship Display) show different magnitude of responses, but similar directionality
+
+```{r FigS2, fig.width = 13, fig.height = 12, fig.cap = "Chormatic (A, C) and Achromatic (B, D) contrast in just noticible differences (JND) for Hawaiian and Kenyan lizards relative to their environments.
 All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
 
-```{r tableS2, tab.cap = "Wald Tests comparing significance of model parameters.
+```{r tableS1, tab.cap = "Wald Tests comparing significance of model parameters.
 Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
@@
 
-```{r tableS3, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction.Social Context"}
+```{r tableS2, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction.Social Context"}
 robust_sc_table <- rbind(robust_ConspicDisplayAC,
@@
 
-```{r FigS2, fig.width = 12, fig.height = 12, fig.cap = "Predators All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
+```{r FigS3, fig.width = 12, fig.height = 12, fig.cap = "Predators All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
 Mean JNDs are provided along with 95% confidence intervals.
@@
 
-```{r tableS4, tab.cap = "Wald Tests comparing significance of model parameters for predators.
+```{r tableS3, tab.cap = "Wald Tests comparing significance of model parameters for predators.
 Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
@@
 
-```{r tableS5, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction.Predators"}
+```{r tableS4, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction.Predators"}
 robust_sc_pred_table <- rbind(robust_PredatorAC,
@@
 
-```{r tableS6, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
+```{r tableS5, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
 Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
@@
 
-```{r tableS7, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
+```{r tableS6, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
 Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
@@
 
-```{r FigS3, fig.width = 13, fig.height = 7.5, fig.cap = "Local adaptation.
+```{r FigS4, fig.width = 13, fig.height = 7.5, fig.cap = "Local adaptation.
 All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
 
-```{r FigS4, fig.cap = "Examples of chameleon colour signal change in response to different social and predatory contexts"}
-
-files <- list.files("./photos/ESM/")
-img1 <- readJPEG(paste0("./photos/ESM/", files[1]))
-img2 <- readJPEG(paste0("./photos/ESM/", files[2]))
-img3 <- readJPEG(paste0("./photos/ESM/", files[3]))
-img4 <- readJPEG(paste0("./photos/ESM/", files[4]))
-img5 <- readJPEG(paste0("./photos/ESM/", files[5]))
-img6 <- readJPEG(paste0("./photos/ESM/", files[6]))
-
-p1 <- cowplot::ggdraw()+
-        cowplot::draw_image(img1)
-p2 <- cowplot::ggdraw()+
-        cowplot::draw_image(img2)
-p3 <- cowplot::ggdraw()+
-        cowplot::draw_image(img3)
-p4 <- cowplot::ggdraw()+
-        cowplot::draw_image(img4)
-p5 <- cowplot::ggdraw()+
-        cowplot::draw_image(img5)
-p6 <- cowplot::ggdraw()+
-        cowplot::draw_image(img6)
-
-(p1 + p3 + p6) / (p2 + p5 + p4) + plot_annotation(tag_levels = c("A", "B", "C", "D", "E", "F"))
-
-```
-
 # References

### 2021-05-07 supp.Rmd
Wriyting some results to go along to summarise all tables and figures

@@
 In our primary analysis, we averaged across different body regions because we were interested only in whether visual signals overall were different between Kenya and Hawaii.
-This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS2), \@ref(fig:FigS3), \@ref(fig:FigS4) and Tables \@ref(tab:tableS1), \@ref(tab:tableS2), \@ref(tab:tableS3), \@ref(tab:tableS4).
-\@ref(tab:tableS5)).
+This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS2) - \@ref(fig:FigS4) and Tables \@ref(tab:tableS1)- \@ref(tab:tableS5)).
 
 #### Body regions in social contexts (Male-male and Courtship Display) show different magnitude of responses, but similar directionality
+
+Overall, different body regions explained significant variation in achromatic and chromatic contrast in both social contexts, whereas differences between Kenya and Hawaii only existed in achormatic contrast (Figure \@ref(fig:FigS2) & Table \@ref(tab:tableS1)).
+A significant interaction between body region and population was found for achormatic contrast when males were displaying to each other (Table \@ref(tab:tableS1)).
+However, in all cases Kenya showed a string reduction in in acromatic contrast relative to Hawaii lizards, despite some variation in the magnitude of effect (Figure \@ref(fig:FigS2)B & D; Table \@ref(tab:tableS2)).
 
@@
 
+#### Body regions are perceived differently by predators but exhibited similar effect sizes
+
+Chromatic contrast of colour signals against the background varied by predator type, population and body region (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
+Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small diferences among populations in how snakes perceived colour signals (Figure \@ref(fig:FigS3)A).
+There were no significant interactions between predator type ('bird' vs 'snake') and either body region or population in achromatic contrast against the background, but a significant population and body region effect (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
+Effects did not vary by predator type (Table \@ref(tab:tableS3)).
+Nonetheless, effect sizes across body regions were in the same direction (Figure \@ref(fig:FigS3)B & Table \@ref(tab:tableS4)).
+Overall, there were significant differences between Kenya and Hawaii in achormatic contrast of the gular and tailbase regions, but not for mid- and top-flank (Table \@ref(tab:tableS5)).
+In contrast, no sigificant differences between Kenya and Hawaii existed across any body region, whereas significant differences existed between gular, midflank and tailbase (Table \@ref(tab:tableS6)).
+
 ```{r FigS3, fig.width = 12, fig.height = 12, fig.cap = "Predators All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
 
+#### All body regions exibit similar contrasts between Kenyan and Hawaiian backgrounds suggesting local adaptation of entire visual signal
+
 ```{r tableS6, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.

### 2021-05-07 ms.Rmd
Updated MS and added some refs

@@
 
-Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package ().
+Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package (@Bateslme4) in combination with the package `clubSandwich` (@Pustejovsky21).
+For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems).Deriving multiple contrast for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely independent of each other.
+To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Saitterwaite degrees of freedom correction.
 Across the different contexts (i.e., social contexts, predation and background environments), chromatic and acromatic contrasts were analysed in seperate models as response variables.
 For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects.
-For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or snake), population of origin and their interaction as fixed effects.
+For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or bird), population of origin and their interaction as fixed effects.
 Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii.
-Model interactions were first tested using Wald tests using robust variance estimators and a Saitterwaite degrees of freedom correction.
+Relevant model interactions were first tested using Wald tests.
 If found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
 We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Saitterwaite degrees of freedom correction.
-For all individuals, we measured signals for multiple body regions, each of which were contratsed between multiple background measurements (i.e., top and bottoms of two leaves and two stems).
-For our main analyses we average estimates across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results).
-Deriving multiple contrats for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely indpeendent of each other.
-To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) on our final model to correct standard errors for within-individual non-independence.
+For our main analyses we averaged estimates (marginalised) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results).
 In all cases, our results and conclusions were not impacted (see supplmentary code).

### 2021-05-07 supp.Rmd
Udpated supp with some results

@@
     result6$p_val_bonfer <- sapply(p.adjust(result6$p_val, "bonferroni"), function(x) p_value(x))
+
+#########################################################
+# Testing local adaptation contrasting backgrounds between hawaii and Kenya
+#########################################################
+##################
+# Social Context: Male-male competition
+##################
+DisplayBgrd <- read.table("./data/Display_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+DisplayBgrd <- DisplayBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
+
+##########
+# Achromatic Contrast - dL
+##########
+
+  # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit the interaction model
+  DispBgrdAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC_inter)
+
+  # Fit the main effect model
+  DispBgrdAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC)
+
+  # Doesn't look like an interaction, so test.
+  anova(DispBgrdAC_inter, DispBgrdAC)
+
+  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+  robust_localAdapt <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+  # Add Wald test to test signifiacnce of each factor, background and population using roust variance
+  background_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(3), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  Population_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(2), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  # Planned contrast. Fit the model in a slightly different way
+  DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
+  DisplayBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = DisplayBgrd) 
+  summary(DisplayBgrd_pop_back)
+
+  robust_DisplayBgrd_pop_back <- clubSandwich::conf_int(DisplayBgrd_pop_back, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+  # Contrasts hypothesis tests, but use RVE 
+    result_DisplayBgrd_pop_back <- Wald_test(DisplayBgrd_pop_back, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+    result_DisplayBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
+
+  # Grab contrasts that we need from the robust esatimtes
+  p_displ_AC <- result_DisplayBgrd_pop_back[2,8]
+   CI_dis_AC <- data.frame(Estimate=robust_DisplayBgrd_pop_back$beta[3],
+                           lwr = robust_DisplayBgrd_pop_back$CI_L[3],
+                           upr = robust_DisplayBgrd_pop_back$CI_U[3])
+
+##########
+# Chromatic Contrast - dS
+##########
+
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+   # Fit the interaction model
+    DispBgrdCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
+    summary(DispBgrdCC_inter)
+
+    # Fit the main effect model
+          DispBgrdCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
+          summary(DispBgrdCC)
+
+    # Looks like weak evidence for interaction    
+        anova(DispBgrdCC_inter, DispBgrdCC)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+      robust_DispBgrdCC <- clubSandwich::coef_test(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+    # Check significance of each factor
+        background_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(3), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  Population_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(2), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+    # Planned contrasts. Refit model in slight different form
+      DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
+    DisplayBgrd_pop_backCC <- lmer(dS ~ pop_back + (1|ID), data = DisplayBgrd)
+
+    # Fit Robust model
+    robust_DisplayBgrd_pop_backCC <- clubSandwich::conf_int(DisplayBgrd_pop_backCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+    # Contrasts hypothesis tests, but use RVE 
+      result_DisplayBgrd_pop_backCC <- Wald_test(DisplayBgrd_pop_backCC, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+      result_DisplayBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Grab contrasts that we need from the robust esatimtes
+      p_displ_CC <- result_DisplayBgrd_pop_backCC[2,8]
+       CI_Disp_CC <- data.frame(Estimate=robust_DisplayBgrd_pop_backCC$beta[3],
+                               lwr = robust_DisplayBgrd_pop_backCC$CI_L[3],
+                               upr = robust_DisplayBgrd_pop_backCC$CI_U[3])
+
+###################
+# Social Context: Courtship
+##################
+CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+CourtshipBgrd <- CourtshipBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
+
+##########
+## Achromatic Contrast - dL
+##########
+
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit the interaction model
+    CourtshipAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
+
+  # Fit main effects model
+    CourtshipAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
+    summary(CourtshipAC)
+
+  # Interaction needed? No
+    anova(CourtshipAC_inter, CourtshipAC)
+
+   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+   robust_CourtshipAC <- clubSandwich::coef_test(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+   # Wald tests
+           background_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(3), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+           Population_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(2), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+  # Planned contrasts / hypothesis tests. Need to fit model a little differently
+    CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
+    CourtshipBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = CourtshipBgrd)
+
+   # Fit Robust model
+    robust_CourtshipBgrd_pop_back <- clubSandwich::conf_int(CourtshipBgrd_pop_back, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+    # Contrasts hypothesis tests, but use RVE 
+      result_CourtshipBgrd_pop_back <- Wald_test(CourtshipBgrd_pop_back, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+      result_CourtshipBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Grab contrasts that we need from the robust esatimtes
+      p_courtship_AC <- result_CourtshipBgrd_pop_back[2,8]
+       CI_AC_Courtship <- data.frame(Estimate=robust_CourtshipBgrd_pop_back$beta[3],
+                               lwr = robust_CourtshipBgrd_pop_back$CI_L[3],
+                               upr = robust_CourtshipBgrd_pop_back$CI_U[3])
+
+##########
+# Chromatic - dS
+##########
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit interaction model
+    CourtshipCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
+
+  # Fit main effect model
+      CourtshipCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC)
+      anova(CourtshipCC)
+
+  # Test whether interaction is supported
+    anova(CourtshipCC_inter, CourtshipCC)
+
+  # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+  robust_CourtshipCC <- clubSandwich::coef_test(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+   # Wald tests
+           background_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(3), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+           Population_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(2), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+  # Planned comparions
+    CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
+  CourtshipBgrd_pop_backCC <- lmer(dS ~ pop_back + (1|ID), data = CourtshipBgrd)
+
+     # Fit Robust model
+    robust_CourtshipBgrd_pop_backCC <- clubSandwich::conf_int(CourtshipBgrd_pop_backCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+    # Contrasts hypothesis tests, but use RVE 
+      result_CourtshipBgrd_pop_backCC <- Wald_test(CourtshipBgrd_pop_backCC, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+      result_CourtshipBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_backCC$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Grab contrasts that we need from the robust esatimtes
+      p_courtship_CC <- result_CourtshipBgrd_pop_backCC[2,8]
+       CI_CC_Courtship <- data.frame(Estimate=robust_CourtshipBgrd_pop_backCC$beta[3],
+                               lwr = robust_CourtshipBgrd_pop_backCC$CI_L[3],
+                               upr = robust_CourtshipBgrd_pop_backCC$CI_U[3])
 

### 2021-05-07 ms.Rmd
Need to check contrasts for preddtor models. Something not quite right

@@
     robust_PredatorAC_CI <-  conf_int(PredatorAC, cluster = Predator_IDs$ID, vcov = "CR2")
+     robust_PredatorAC_CI_inter <-  conf_int(PredatorAC_inter, cluster = Predator_IDs$ID, vcov = "CR2")
+
+  # Model allows us to get bird in Hwaii compared to bird in Kenya (contrast 3)
+      Predator$pred_pop <- with(Predator, interaction(Predator, Population))
+
+      PredatorAC_pred_pop1 <- lmer(dL ~  pred_pop + (1|ID), data = Predator) 
+      summary(PredatorAC_pred_pop1)
+      PredatorAC_pred_pop2 <- lmer(dL ~  relevel(pred_pop, ref = "snake.Hawaii") + (1|ID), data = Predator) 
+      summary(PredatorAC_pred_pop2)
+
+      # Fit the models so we get the estimate and corrected robust standard error for the contrasts we want to make
+      robust_PredatorAC_pred_pop1 <- conf_int(PredatorAC_pred_pop1, cluster = Predator$ID, vcov="CR2")
+      robust_PredatorAC_pred_pop2 <- conf_int(PredatorAC_pred_pop2, cluster = Predator$ID, vcov="CR2")
+
+     # Contrasts hypothesis tests, but use RVE 
+    result_pred <- Wald_test(PredatorAC_pred_pop, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+    result_pred$p_val_bonfer <- sapply(p.adjust(result_pred$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Get the contrasts and Wald stats for comparions
+     CI_hyp_testAC <- data.frame(
+              comparison = c("Snake (Hawaii-Kenya)", "Birds (Hawaii-Kenya)"), 
+                Estimate = c(robust_PredatorAC_pred_pop2[4,1], robust_PredatorAC_pred_pop1[3, 1]),
+                lwr = c(robust_PredatorAC_pred_pop2[4,4], robust_PredatorAC_pred_pop1[3, 4]),
+                upr = c(robust_PredatorAC_pred_pop2[4,5], robust_PredatorAC_pred_pop1[3, 5]),
+              Fstat = c(result_pred[5,3], result_pred[2,3]),
+              df = c(result_pred[5,5], result_pred[2,5]),
+              p_val = c(result_pred[5,7], result_pred[2,7]),
+              p_bonf = c(result_pred[5,8], result_pred[2,8]))
 
@@
     summary(PredatorCC_pred_pop)
+
+    # Present full model
+         robust_PredatorCC_CI_inter <-  conf_int(PredatorCC, cluster = Predator_IDs$ID, vcov = "CR2")
 
@@
 
+### Supplementary Tables - Add to Supplement. Easier to generate here given all code is here.
+
+```{r table1, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
+Standard errors are corrected by using robust variance estimators.
+"}
+
+table1 <- as.data.frame(rbind(robust_ConspicDisplayAC_CI,
+                robust_ConspicDisplayCC_CI,
+                robust_ConspicCourtshipAC_CI,
+                robust_ConspicCourtshipCC_CI))
+table1$Parameter <- rownames(table1)
+rownames(table1) <- NULL
+# Clean
+ final_T1 <- table1 %>% 
+             mutate(SocialContext = c("Male-Male Display", rep("", 5), "Courtship Display", rep("", 5)),
+                    VisualRealm = c("Acromatic Contrast", rep("", 2), "Chromatic Contrast", rep("", 2), "Acromatic Contrast", rep("", 2), "Chromatic Contrast", rep("", 2))) %>%
+              select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
+
+ # Table
+table1 <- flextable(final_T1) %>%
+        table_style6()
+
+table1
+```
+
+```{r table2, tab.cap = ""Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
+Full models are provided and standard errors are corrected by using robust variance estimators."}
+
+table2 <- as.data.frame(rbind(robust_PredatorAC_CI_inter,
+                              robust_PredatorCC_CI_inter))
+table2$Parameter <- rownames(table2)
+rownames(table2) <- NULL
+
+# Clean
+ final_T2 <- table2 %>% 
+             mutate(JND = c("Acromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3))) %>%
+              select(JND, Parameter, beta, SE, df, CI_L, CI_U)
+
+ # Table
+table2 <- flextable(final_T2) %>%
+           table_style7()
+
+table2
+```
+
+CI_hyp_test
+
 # References

### 2021-05-07 R/func.R

@@ -187,3 +187,39 @@ table_style5 <- function(table){
   
   return(table)
 }
+
+table_style6 <- function(table){
+  table <- table %>% width(j = 1, width = 1) %>% width(j = 3, width = 3.5) %>% 
+    align(align = "center", part = "header") %>% 
+    align(align = "center", part = "body") %>% 
+    compose(part = "header", j = 1, value = as_paragraph(as_b("Social Context"))) %>% 
+    compose(part = "header", j = 2, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
+    compose(part = "header", j = 3, value = as_paragraph(as_b("Parameter"))) %>% 
+    compose( part = "header", j = 4, value = as_paragraph(as_b("Estimate"))) %>% 
+    compose(part = "header", j = 5, value = as_paragraph(as_b("SE"))) %>%
+    compose(part = "header", j = 6, value = as_paragraph(as_b("df"))) %>% 
+    compose(part = "header", j = 7, value = as_paragraph(as_b("95% CI"), as_sub(as_b("lower")))) %>% 
+    compose(part = "header", j = 8, value = as_paragraph(as_b("95% CI"), as_sub(as_b("upper")))) %>% 
+    compose(part = "body", j = 3, i = c(1,4,7,10), value = as_paragraph("Intercept")) %>%
+    compose(part = "body", j = 3, i = c(1,4,7,10)+1, value = as_paragraph("SVL", as_sub("z-transformed"))) %>%
+    compose(part = "body", j = 3, i = c(1,4,7,10)+2, value = as_paragraph("Population", as_sub("Kenya"))) %>%
+  return(table)
+}
+
+table_style7 <- function(table){
+  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 3.5) %>% 
+    align(align = "center", part = "header") %>% 
+    align(align = "center", part = "body") %>% 
+    compose(part = "header", j = 1, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
+    compose(part = "header", j = 2, value = as_paragraph(as_b("Parameter"))) %>% 
+    compose( part = "header", j = 3, value = as_paragraph(as_b("Estimate"))) %>% 
+    compose(part = "header", j = 4, value = as_paragraph(as_b("SE"))) %>%
+    compose(part = "header", j = 5, value = as_paragraph(as_b("df"))) %>% 
[…]

### 2021-05-07 R/func.R
Update to table styles

@@ -189,7 +189,7 @@ table_style5 <- function(table){
 }
 
 table_style6 <- function(table){
-  table <- table %>% width(j = 1, width = 1) %>% width(j = 3, width = 3.5) %>% 
+  table <- table  %>% 
     align(align = "center", part = "header") %>% 
     align(align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Social Context"))) %>% 
@@ -202,12 +202,12 @@ table_style6 <- function(table){
     compose(part = "header", j = 8, value = as_paragraph(as_b("95% CI"), as_sub(as_b("upper")))) %>% 
     compose(part = "body", j = 3, i = c(1,4,7,10), value = as_paragraph("Intercept")) %>%
     compose(part = "body", j = 3, i = c(1,4,7,10)+1, value = as_paragraph("SVL", as_sub("z-transformed"))) %>%
-    compose(part = "body", j = 3, i = c(1,4,7,10)+2, value = as_paragraph("Population", as_sub("Kenya"))) %>%
+    compose(part = "body", j = 3, i = c(1,4,7,10)+2, value = as_paragraph("Population", as_sub("Kenya")))
   return(table)
 }
 
 table_style7 <- function(table){
-  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 3.5) %>% 
+  table <- table %>% 
     align(align = "center", part = "header") %>% 
     align(align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
@@ -222,4 +222,20 @@ table_style7 <- function(table){
     compose(part = "body", j = 2, i = c(1,5)+2, value = as_paragraph("Population", as_sub("Kenya"))) %>%
     compose(part = "body", j = 2, i = c(1,5)+3, value = as_paragraph("Predator", as_sub("Snake"), "* Population", as_sub("Kenya")))
     return(table)
+}
+
+table_style8 <- function(table){
+  table <- table %>%
+    align(align = "center", part = "header") %>% 
+    align( align = "center", part = "body") %>% 
+    compose(part = "header", j = 1, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
+    compose(part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>% 
[…]

### 2021-05-07 ms.Rmd
Updated results with tables we can move to supplement

@@
 
-  # Get ANOVA stats for reporting in the paper of interaction
-  anova_predators_inter <- anova(PredatorAC_inter)
-
-# No support for an interaction.
-Drop interaction and fit the main effects model.
-Tells us that predation, regardless of whether we are talking about snakes or birds is just low in Hawaii, otherwise, we would predict that signals would evolve to match the predatory system
-
   #Refit main effects so we can test main effects on own.
@@
         summary(PredatorAC)
-
-  # Grab the ANOVA summary stats
-    anova_predators <- anova(PredatorAC)
 
@@
      # Contrasts hypothesis tests, but use RVE
-    result_pred <- Wald_test(PredatorAC_pred_pop, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+
+      PredatorAC_pred_pop <- lmer(dL ~  0+ pred_pop + (1|ID), data = Predator) 
+    result_pred <- Wald_test(PredatorAC_pred_pop, constraints = constrain_pairwise(c(1:4)), 
+                             vcov="CR2", tidy = TRUE)
     result_pred$p_val_bonfer <- sapply(p.adjust(result_pred$p_val, "bonferroni"), function(x) p_value(x))
@@
   # Doesn't look like an interaction, so test.
-  anova(DispBgrdAC_inter, DispBgrdAC)
+  inter_AC_Disp <- Wald_test(DispBgrdAC_inter, constraints = constrain_zero(4), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
 
@@
   # Add Wald test to test signifiacnce of each factor, background and population using roust variance
+
   background_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(3), 
@@
    # Fit the interaction model
-    DispBgrdCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
+    DispBgrdCC_inter <- lmer(dS ~ Population  + Background_pop + 
+                                  Population:Background_pop + (1|ID), data=DisplayBgrd)
     summary(DispBgrdCC_inter)
@@
 
-    # Looks like weak evidence for interaction    
-        anova(DispBgrdCC_inter, DispBgrdCC)
+    # Doesn't look like an interaction, so test.
+          inter_CC_Disp <- Wald_test(DispBgrdCC_inter, constraints = constrain_zero(4), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
 
@@
 
-```{r table1, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
+```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
 Standard errors are corrected by using robust variance estimators.
@@
  # Table
-table1 <- flextable(final_T1) %>%
-        table_style6()
+table1 <- flextable(final_T1) %>% table_style6()
 
@@
 
-```{r table2, tab.cap = ""Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
+```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
 Full models are provided and standard errors are corrected by using robust variance estimators."}
@@
 
-CI_hyp_test
+```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Mean differences between Hawaiian and Kenyan visual signals for snake and bird predators."}
+
+contrast <- rbind(CI_hyp_test,
+                 CI_hyp_testAC)
+
+contrast <- contrast %>% mutate(VisualSpec = c("Achromatic Contrast", "", "Chromatic Contrast", ""),
+                                p_val = sapply(p_val, function(x) p_value(x))) %>% 
+            select(VisualSpec, comparison, Estimate, lwr, upr, Fstat, df, p_val, p_bonf)
+
+table3 <- flextable(contrast) %>% table_style8()
+table3
+```
 

### 2021-05-07 ms.Rmd
Adding Wald tests, but not sure they are working correct

@@
       robust_DispBgrdCC <- clubSandwich::coef_test(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+      robust_DispBgrdCC_inter <- clubSandwich::coef_test(DispBgrdCC_inter, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
@@
 
-  # Interaction needed? No
-    anova(CourtshipAC_inter, CourtshipAC)
+  # Wald test of interaction
+    background_AC_Court <- Wald_test(CourtshipAC_inter, constraints = constrain_zero(4), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
 
@@
   # Fit interaction model
-    CourtshipCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
+    CourtshipCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + 
+                                (1|ID), data=CourtshipBgrd)
 
@@
       summary(CourtshipCC)
-      anova(CourtshipCC)
 
-  # Test whether interaction is supported
-    anova(CourtshipCC_inter, CourtshipCC)
+    # Wald test of interaction
+    inter_AC_Court <- Wald_test(CourtshipCC_inter, constraints = constrain_zero(4), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
 

### 2021-05-07 ms.Rmd
Updates.

@@
 
-  # Fit the interaction model
-  DispBgrdAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
-  summary(DispBgrdAC_inter)
-
   # Fit the main effect model
@@
   summary(DispBgrdAC)
-
-  # Doesn't look like an interaction, so test.
-  inter_AC_Disp <- Wald_test(DispBgrdAC_inter, constraints = constrain_zero(4), 
-                                  cluster = DisplayBgrd$ID, vcov = "CR2")
 
@@
 
-   # Fit the interaction model
-    DispBgrdCC_inter <- lmer(dS ~ Population  + Background_pop + 
-                                  Population:Background_pop + (1|ID), data=DisplayBgrd)
-    summary(DispBgrdCC_inter)
-
     # Fit the main effect model
@@
           summary(DispBgrdCC)
-
-    # Doesn't look like an interaction, so test.
-          inter_CC_Disp <- Wald_test(DispBgrdCC_inter, constraints = constrain_zero(4), 
-                                     cluster = DisplayBgrd$ID, vcov = "CR2")
 
@@
 
-  # Fit the interaction model
-    CourtshipAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
-
   # Fit main effects model
@@
     summary(CourtshipAC)
-
-  # Wald test of interaction
-    background_AC_Court <- Wald_test(CourtshipAC_inter, constraints = constrain_zero(4), 
-                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
 
@@
 
-  # Fit interaction model
-    CourtshipCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + 
-                                (1|ID), data=CourtshipBgrd)
-
   # Fit main effect model
@@
       summary(CourtshipCC)
-
-    # Wald test of interaction
-    inter_AC_Court <- Wald_test(CourtshipCC_inter, constraints = constrain_zero(4), 
-                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
 

### 2021-05-07 R/func.R
Added new style for last table

@@ -239,3 +239,21 @@ table_style8 <- function(table){
     compose(part = "header", j = 9, value = as_paragraph(as_b("p"), as_sub(as_b("Bonferroni")))) %>% 
   return(table)
 }
+
+table_style9 <- function(table){
+  table <- table  %>% 
+    align(align = "center", part = "header") %>% 
+    align(align = "center", part = "body") %>% 
+    compose(part = "header", j = 1, value = as_paragraph(as_b("Social Context"))) %>% 
+    compose(part = "header", j = 2, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
+    compose(part = "header", j = 3, value = as_paragraph(as_b("Parameter"))) %>% 
+    compose( part = "header", j = 4, value = as_paragraph(as_b("Estimate"))) %>% 
+    compose(part = "header", j = 5, value = as_paragraph(as_b("SE"))) %>%
+    compose(part = "header", j = 6, value = as_paragraph(as_b("df"))) %>% 
+    compose(part = "header", j = 7, value = as_paragraph(as_b("95% CI"), as_sub(as_b("lower")))) %>% 
+    compose(part = "header", j = 8, value = as_paragraph(as_b("95% CI"), as_sub(as_b("upper")))) %>% 
+    compose(part = "body", j = 3, i = c(1,4,7,10), value = as_paragraph("Intercept")) %>%
+    compose(part = "body", j = 3, i = c(1,4,7,10)+1, value = as_paragraph("Population", as_sub("Kenya"))) %>%
+    compose(part = "body", j = 3, i = c(1,4,7,10)+2, value = as_paragraph("Background", as_sub("Kenya")))
+  return(table)
+}

### 2021-05-07 ms.Rmd
Created final table for local adaptation

@@
   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-  robust_localAdapt <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+  robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
@@
     # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-      robust_DispBgrdCC <- clubSandwich::coef_test(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
-      robust_DispBgrdCC_inter <- clubSandwich::coef_test(DispBgrdCC_inter, cluster = DisplayBgrd$ID, vcov =  "CR2")
+      robust_DispBgrdCC <- clubSandwich::conf_int(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
@@
     # Fit Robust model
-    robust_DisplayBgrd_pop_backCC <- clubSandwich::conf_int(DisplayBgrd_pop_backCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+    robust_DisplayBgrd_pop_backCC <- clubSandwich::conf_int(DisplayBgrd_pop_backCC, 
+                                                            cluster = DisplayBgrd$ID, vcov =  "CR2")
 
@@
    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-   robust_CourtshipAC <- clubSandwich::coef_test(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+   robust_CourtshipAC <- clubSandwich::conf_int(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
 
@@
   # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
-  robust_CourtshipCC <- clubSandwich::coef_test(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+  robust_CourtshipCC <- clubSandwich::conf_int(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
 
@@
 
+# Social context. We only need models because contarsts are already there
+
 ```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
@@
 
+# Predators - Models AND contrasts. We don't get all contrasts because we have two factors in model
 ```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
@@
 
+# Local Adaptation.
+We have all the contrasts we need from the model because we are only comparing Hawaii-Hawaii(background) to Hawaii-Kenya(background).
+These are therefore in the table output already.
+
+```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Mean differences between Hawaiian lizards on introduced Hawaiian background and native Kenyan background."}
+
+tb4 <- data.frame(rbind(robust_localAdapt_DispBgrdAC,
+             robust_DispBgrdCC,
+             robust_CourtshipAC,
+             robust_CourtshipCC))
+tb4$parameter <- rownames(tb4)
+rownames(tb4) <- NULL
+
+tb4 <- tb4 %>% mutate(SocialContext = c("Male-Male Display", rep("", 5), 
+                                        "Courtship Display", rep("", 5)),
+                    VisualRealm = c("Acromatic Contrast", rep("", 2), "Chromatic Contrast", 
+                                    rep("", 2), "Acromatic Contrast", rep("", 2), 
+                                    "Chromatic Contrast", rep("", 2))) %>%
+              select(SocialContext, VisualRealm, parameter, beta, SE, df, CI_L, CI_U)
+
+table4 <- flextable(tb4) %>% table_style9()
+
+```
+
 # References

### 2021-05-08 ms.Rmd
Fix refs and tables in MS formatted, but move dto seperate doc to make sure they don't get re-formtted each time we render

@@
 
-```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Mean differences between Hawaiian and Kenyan visual signals for snake and bird predators."}
+```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean acromatric and chormatic contrast in Hawaii and Kenya for both predators (snakes and birds).
+Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
+Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).
+We also applied a Bonferroni correction to p-value to control for multiple comparisons."}
 
@@
 
-```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Mean differences between Hawaiian lizards on introduced Hawaiian background and native Kenyan background."}
+```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for comparing acromatric and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (Background = Kenya) in both social contexts (Male-male competition and Courtship).
+Models were fit using robust variance estaimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.
+Significance parameter estimates are those where 95% CI does not include zero."}
 
@@
 table4 <- flextable(tb4) %>% table_style9()
-
+table4
 ```

### 2021-05-08 ms.Rmd
Fixed spelling mistakes, added some spaces for subscripts

@@
              mutate(SocialContext = c("Male-Male Display", rep("", 5), "Courtship Display", rep("", 5)),
-                    VisualRealm = c("Acromatic Contrast", rep("", 2), "Chromatic Contrast", rep("", 2), "Acromatic Contrast", rep("", 2), "Chromatic Contrast", rep("", 2))) %>%
+                    VisualRealm = c("Achromatic Contrast", rep("", 2), "Chromatic Contrast", rep("", 2), "Achromatic Contrast", rep("", 2), "Chromatic Contrast", rep("", 2))) %>%
               select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
@@
  final_T2 <- table2 %>% 
-             mutate(JND = c("Acromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3))) %>%
+             mutate(JND = c("Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3))) %>%
               select(JND, Parameter, beta, SE, df, CI_L, CI_U)
@@
                                         "Courtship Display", rep("", 5)),
-                    VisualRealm = c("Acromatic Contrast", rep("", 2), "Chromatic Contrast", 
-                                    rep("", 2), "Acromatic Contrast", rep("", 2), 
+                    VisualRealm = c("Achromatic Contrast", rep("", 2), "Chromatic Contrast", 
+                                    rep("", 2), "Achromatic Contrast", rep("", 2), 
                                     "Chromatic Contrast", rep("", 2))) %>%

### 2021-05-08 R/func.R

@@ -85,23 +85,23 @@ table_style  <- function(table){
     compose(part = "header", j = 3, value = as_paragraph(as_b("F"))) %>%
     compose(part = "header", j = 4, value = as_paragraph(as_b("DF"))) %>% 
     compose(part = "header", j = 5, value = as_paragraph(as_b("p"))) %>% 
-    compose(part = "header", j = 6, value = as_paragraph(as_b("p"), as_sub(as_b("Bonferroni")))) %>%
+    compose(part = "header", j = 6, value = as_paragraph(as_b("p"), as_sub(as_b(" Bonferroni")))) %>%
     compose(part = "body", i = 1, j = 2, 
-            value = as_paragraph("Kenya", as_sub("gular"), " - ", "Hawaii", as_sub("gular")))  %>% 
+            value = as_paragraph("Kenya", as_sub(" gular"), " - ", "Hawaii", as_sub(" gular")))  %>% 
     compose(part = "body", i = 2, j = 2, 
-            value = as_paragraph("Kenya", as_sub("midflank"), " - ", "Hawaii", as_sub("midflank"))) %>% 
+            value = as_paragraph("Kenya", as_sub(" midflank"), " - ", "Hawaii", as_sub(" midflank"))) %>% 
     compose(part = "body", i = 3, j = 2, 
-            value = as_paragraph("Kenya", as_sub("tailbase"), " - ", "Hawaii", as_sub("tailbase"))) %>% 
+            value = as_paragraph("Kenya", as_sub(" tailbase"), " - ", "Hawaii", as_sub(" tailbase"))) %>% 
     compose(part = "body", i = 4, j = 2, 
-            value = as_paragraph("Kenya", as_sub("topflank"), " - ", "Hawaii", as_sub("topflank"))) %>% 
+            value = as_paragraph("Kenya", as_sub(" topflank"), " - ", "Hawaii", as_sub(" topflank"))) %>% 
     compose(part = "body", i = 5, j = 2, 
-            value = as_paragraph("Kenya", as_sub("gular"), " - ", "Hawaii", as_sub("gular")))  %>% 
+            value = as_paragraph("Kenya", as_sub(" gular"), " - ", "Hawaii", as_sub(" gular")))  %>% 
     compose(part = "body", i = 6, j = 2, 
-            value = as_paragraph("Kenya", as_sub("midflank"), " - ", "Hawaii", as_sub("midflank"))) %>% 
[…]

### 2021-05-08 R/ms.R
Writing code from .Rmd to pure code files to make it easier to look at code altogether withut intermixing . Easy to do using knitr::purl("./supp.Rmd", output = "./R/supp.R") and knitr::purl("./ms.Rmd", output = "./R/ms.R")

@@ -1,0 +1,799 @@
+## ----setup, include=FALSE-------------------------------------------------------------------------------------------------------------
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
+## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
+  ## and rounded to 2 digits      ## and rounded to 2 digits
+  options(digits = 2)
+
+
+## ----getlibrary, echo = FALSE, results = "hide"---------------------------------------------------------------------------------------
+# load libraries
+devtools::install_github("karthik/wesanderson")
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
+source("./R/func.R")
+
+
+
+## ----Models, echo = FALSE, results = "hide"-------------------------------------------------------------------------------------------
+source("./R/func.R")
+#########################################################
+# Social Context: Male-male Competition Displays
+#########################################################
+#Load data and clean a bit
+Display_1 <- read.table("./data/Display_1.csv", 
+                          header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Center SVL to make intercept interpretable
+Display_1$sc_SVL <- with(Display_1, scale(SVL, scale=FALSE))
+
+#exclude bottom flank.
+Display_1 <- Display_1 %>% 
+              filter(!BodyRegion == "botflank") %>% as.data.frame() 
+
+# create as factor
+    Display_1$BodyRegion <- factor(Display_1$BodyRegion)
+
+##########
+## Achromatic Contrast - dL
+##########
+      
+    # Fit the model
+    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + (1|ID), data=Display_1)
+      summary(ConspicDisplayAC)
+     
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+       robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, 
[…]

### 2021-05-08 R/supp.R

@@ -1,0 +1,860 @@
+## ----setup, include=FALSE-------------------------------------------------------------------------------------------------------------
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
+## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
+  ## and rounded to 2 digits      ## and rounded to 2 digits
+  options(digits = 2)
+
+
+## ----getlibrary, echo = FALSE, results = "hide"---------------------------------------------------------------------------------------
+# load libraries
+devtools::install_github("karthik/wesanderson")
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
+source("./R/func.R")
+
+
+## ----display, results = "hide"--------------------------------------------------------------------------------------------------------
+source("./R/func.R")
+#########################################################
+# Male-male Competition Displays
+#########################################################
+  # Load data and clean a bit
+  Display_1 <- read.table("./data/Display_1.csv", 
+                            header=TRUE, sep=",", dec=".", strip.white=TRUE)
+  
+  # Exclude bottom flank.
+  Display_1 <- Display_1 %>% 
+                filter(!BodyRegion == "botflank") %>% as.data.frame() 
+  
+  # Create as factor
+      Display_1$BodyRegion <- factor(Display_1$BodyRegion)
+      
+  # Center SVL to make intercept interpretable
+  Display_1$sc_SVL <- with(Display_1, scale(SVL, scale=FALSE))
+
+
+##########
+## Achromatic Contrast - dL
+##########
+
+    # Fit the model that estimates changes across body regions
+    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + 
+                               Population:BodyRegion + (1|ID), data=Display_1)
+    summary(ConspicDisplayAC)
[…]

### 2021-05-08 supp.Rmd
Adding purl = FALSE to R chunk so that the chunck isn't exported in final R script. Pretty neat

@@
 ---
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, purl = FALSE}
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)

### 2021-05-08 ms.Rmd
Added purl = FALSE

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, purl = FALSE}
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)

### 2021-05-09 supp.Rmd
Adding in and correcting local adaptation..includes body region

@@
 
-  # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
-The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
-
   # Fit the interaction model
-  DispBgrdAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
-  summary(DispBgrdAC_inter)
-
-  # Fit the main effect model
-  DispBgrdAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
+  DispBgrdAC <- lmer(dL ~ Population  + Background_pop + BodyRegion + (1|ID), data=DisplayBgrd)
   summary(DispBgrdAC)
 
-  # Doesn't look like an interaction, so test.
-  anova(DispBgrdAC_inter, DispBgrdAC)
-
   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-  robust_localAdapt <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+  robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
 
@@
 
-  # Planned contrast. Fit the model in a slightly different way
-  DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
-  DisplayBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = DisplayBgrd) 
-  summary(DisplayBgrd_pop_back)
+  BodyRegion_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(4:6), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
 
-  robust_DisplayBgrd_pop_back <- clubSandwich::conf_int(DisplayBgrd_pop_back, cluster = DisplayBgrd$ID, vcov =  "CR2")
+  # Planned contrast. Fit the model in a slightly different way ********************** UP TO HERE
+  DisplayBgrd$pop_back_region <- with(DisplayBgrd, interaction(Population, Background_pop, BodyRegion))
+  DisplayBgrd_pop_back_region <- lmer(dL ~ pop_back_region + (1|ID), data = DisplayBgrd) 
+  summary(DisplayBgrd_pop_back_region)
+
+  robust_DisplayBgrd_pop_back_region <- clubSandwich::conf_int(DisplayBgrd_pop_back_region, cluster = DisplayBgrd$ID, vcov =  "CR2")
 

### 2021-05-09 supp.Rmd
Added in local adaptation contrasts for male display and courtship. THere are some problems iwth male display that need to be sorted. For some reason they are not estimating one paramter for Achromatic. gular.

@@
 
-  # Planned contrast. Fit the model in a slightly different way ********************** UP TO HERE
+  # Planned contrast. Fit the model in a slightly different way ********** PROBLEMS HERE TO WORK OUT WITH CS
   DisplayBgrd$pop_back_region <- with(DisplayBgrd, interaction(Population, Background_pop, BodyRegion))
-  DisplayBgrd_pop_back_region <- lmer(dL ~ pop_back_region + (1|ID), data = DisplayBgrd) 
+  DisplayBgrd_Hawaii <- DisplayBgrd %>% filter(Population == "Hawaii")
+  DisplayBgrd_pop_back_region <- lmer(dL ~ pop_back_region + (1|ID), data = DisplayBgrd_Hawaii) 
   summary(DisplayBgrd_pop_back_region)
 
-  robust_DisplayBgrd_pop_back_region <- clubSandwich::conf_int(DisplayBgrd_pop_back_region, cluster = DisplayBgrd$ID, vcov =  "CR2")
+  robust_DisplayBgrd_pop_back_region <- clubSandwich::conf_int(DisplayBgrd_pop_back_region, 
+                                                               cluster = DisplayBgrd_Hawaii$ID, 
+                                                               vcov =  "CR2")
 
   # Contrasts hypothesis tests, but use RVE 
-    result_DisplayBgrd_pop_back <- Wald_test(DisplayBgrd_pop_back, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
-    result_DisplayBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
-
-  # Grab contrasts that we need from the robust esatimtes
-  p_displ_AC <- result_DisplayBgrd_pop_back[2,8]
-   CI_dis_AC <- data.frame(Estimate=robust_DisplayBgrd_pop_back$beta[3],
-                           lwr = robust_DisplayBgrd_pop_back$CI_L[3],
-                           upr = robust_DisplayBgrd_pop_back$CI_U[3])
+    result_DisplayBgrd_pop_back_region <- Wald_test(DisplayBgrd_pop_back_region, 
+                                             constraints = constrain_pairwise(c(1:16)),
+                                             vcov="CR2", tidy = TRUE)
+    result_DisplayBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"),
+                                                       function(x) p_value(x))
 
@@
 
- # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
-The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
-
-   # Fit the interaction model
-    DispBgrdCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=DisplayBgrd)
-    summary(DispBgrdCC_inter)
-
     # Fit the main effect model
-          DispBgrdCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
+          DispBgrdCC <- lmer(dS ~ Population  + Background_pop  + BodyRegion + (1|ID), data=DisplayBgrd)
           summary(DispBgrdCC)
-
-    # Looks like weak evidence for interaction    
-        anova(DispBgrdCC_inter, DispBgrdCC)
 
@@
 
+        BodyRegion_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(4:6), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
     # Planned contrasts. Refit model in slight different form
-      DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
-    DisplayBgrd_pop_backCC <- lmer(dS ~ pop_back + (1|ID), data = DisplayBgrd)
+    DisplayBgrd_pop_back_regionCC <- lmer(dS ~ -1 + pop_back_region + (1|ID), data = DisplayBgrd_Hawaii)
 
     # Fit Robust model
-    robust_DisplayBgrd_pop_backCC <- clubSandwich::conf_int(DisplayBgrd_pop_backCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+    robust_DisplayBgrd_pop_back_regionCC <- clubSandwich::conf_int(DisplayBgrd_pop_back_regionCC, 
+                                                            cluster = DisplayBgrd_Hawaii$ID, 
+                                                            vcov =  "CR2")
 
     # Contrasts hypothesis tests, but use RVE 
-      result_DisplayBgrd_pop_backCC <- Wald_test(DisplayBgrd_pop_backCC, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
-      result_DisplayBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
+      result_DisplayBgrd_pop_back_regionCC <- Wald_test(DisplayBgrd_pop_back_regionCC, 
+                                                 constraints = constrain_pairwise(c(1:8)), 
+                                                 vcov="CR2", tidy = TRUE)
 
-    # Grab contrasts that we need from the robust esatimtes
-      p_displ_CC <- result_DisplayBgrd_pop_backCC[2,8]
-       CI_Disp_CC <- data.frame(Estimate=robust_DisplayBgrd_pop_backCC$beta[3],
-                               lwr = robust_DisplayBgrd_pop_backCC$CI_L[3],
-                               upr = robust_DisplayBgrd_pop_backCC$CI_U[3])
+      result_DisplayBgrd_pop_back_regionCC$p_val_bonfer <- 
+                                  sapply(p.adjust(result_DisplayBgrd_pop_back_regionCC$p_val, "bonferroni"), 
+                                                                  function(x) p_value(x))
 

### 2021-05-09 supp.Rmd
Implemented all RVEs for models...some problems with convergence on a few achromatix models, odd. Need to check these. Add in tables now is all that is left.

@@
 
- # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
-The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
-
-  # Fit the interaction model
-    CourtshipAC_inter <- lmer(dL ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
-
   # Fit main effects model
-    CourtshipAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
+    CourtshipAC <- lmer(dL ~ Population  + Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd)
     summary(CourtshipAC)
 
-  # Interaction needed? No
-    anova(CourtshipAC_inter, CourtshipAC)
-
    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-   robust_CourtshipAC <- clubSandwich::coef_test(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+   robust_CourtshipAC <- clubSandwich::conf_int(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
 
@@
 
+           BodyRegion_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(4:6), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
   # Planned contrasts / hypothesis tests. Need to fit model a little differently
-    CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
-    CourtshipBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = CourtshipBgrd)
+    CourtshipBgrd$pop_back_BR <- with(CourtshipBgrd, interaction(Population, Background_pop, BodyRegion))
+    CourtshipBgrd_Hawaii <-CourtshipBgrd %>% filter(Population == "Hawaii")
+    CourtshipBgrd_pop_back_BR <- lmer(dL ~ -1 + pop_back_BR + (1|ID), data = CourtshipBgrd_Hawaii)
 
    # Fit Robust model
-    robust_CourtshipBgrd_pop_back <- clubSandwich::conf_int(CourtshipBgrd_pop_back, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+    robust_CourtshipBgrd_pop_back_BR <- clubSandwich::conf_int(CourtshipBgrd_pop_back_BR, 
+                                                            cluster = CourtshipBgrd_Hawaii$ID, 
+                                                            vcov =  "CR2")
 
-    # Contrasts hypothesis tests, but use RVE 
-      result_CourtshipBgrd_pop_back <- Wald_test(CourtshipBgrd_pop_back, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+    # Contrasts hypothesis tests, but use RVE ********** PROBLEM WITH CONVERGENCE NON-POSITIVE DEF MATRICES
+      result_CourtshipBgrd_pop_back_BR <- Wald_test(CourtshipBgrd_pop_back_BR, 
+                                                 constraints = constrain_pairwise(c(1:8)), 
+                                                 vcov="CR2", tidy = TRUE)
+
       result_CourtshipBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
-
-    # Grab contrasts that we need from the robust esatimtes
-      p_courtship_AC <- result_CourtshipBgrd_pop_back[2,8]
-       CI_AC_Courtship <- data.frame(Estimate=robust_CourtshipBgrd_pop_back$beta[3],
-                               lwr = robust_CourtshipBgrd_pop_back$CI_L[3],
-                               upr = robust_CourtshipBgrd_pop_back$CI_U[3])
 
@@
 ##########
- # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
-The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
-
-  # Fit interaction model
-    CourtshipCC_inter <- lmer(dS ~ Population  + Background_pop + Population:Background_pop + (1|ID), data=CourtshipBgrd)
 
   # Fit main effect model
-      CourtshipCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
+      CourtshipCC <- lmer(dS ~ Population  + Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd)
       summary(CourtshipCC)
-      anova(CourtshipCC)
-
-  # Test whether interaction is supported
-    anova(CourtshipCC_inter, CourtshipCC)
 
@@
 
+           BodyRegion_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(4:6), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
   # Planned comparions
-    CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
-  CourtshipBgrd_pop_backCC <- lmer(dS ~ pop_back + (1|ID), data = CourtshipBgrd)
+
+  CourtshipBgrd_pop_back_BR_CC <- lmer(dS ~ -1 + pop_back_BR + (1|ID), data = CourtshipBgrd_Hawaii)
 
      # Fit Robust model
-    robust_CourtshipBgrd_pop_backCC <- clubSandwich::conf_int(CourtshipBgrd_pop_backCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+    robust_CourtshipBgrd_pop_back_BR_CC <- clubSandwich::conf_int(CourtshipBgrd_pop_back_BR_CC, 
+                                                              cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
 
     # Contrasts hypothesis tests, but use RVE 
-      result_CourtshipBgrd_pop_backCC <- Wald_test(CourtshipBgrd_pop_backCC, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+      result_CourtshipBgrd_pop_back_BR_CC <- Wald_test(CourtshipBgrd_pop_back_BR_CC, 
+                                                   constraints = constrain_pairwise(c(1:8)),
+                                                   cluster = CourtshipBgrd_Hawaii$ID,
+                                                   vcov="CR2", tidy = TRUE)
       result_CourtshipBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_backCC$p_val, "bonferroni"), function(x) p_value(x))
@@
        # Grab contrasts that we need from the robust esatimtes
-      p_courtship_CC <- result_CourtshipBgrd_pop_backCC[2,8]
-       CI_CC_Courtship <- data.frame(Estimate=robust_CourtshipBgrd_pop_backCC$beta[3],
-                               lwr = robust_CourtshipBgrd_pop_backCC$CI_L[3],
-                               upr = robust_CourtshipBgrd_pop_backCC$CI_U[3])
+      string <- c("pop_back_BRHawaii.Kenya.gular - pop_back_BRHawaii.Hawaii.gular|pop_back_BRHawaii.Kenya.midflank - pop_back_BRHawaii.Hawaii.midflank|pop_back_BRHawaii.Kenya.tailbase - pop_back_BRHawaii.Hawaii.tailbase|pop_back_BRHawaii.Kenya.topflank - pop_back_BRHawaii.Hawaii.topflank")
+
+         result_CourtshipBgrd_pop_backCC <- result_CourtshipBgrd_pop_back_BR_CC[
+                                            grep(string, result_CourtshipBgrd_pop_back_BR_CC$hypothesis),]
 

### 2021-05-09 supp.Rmd
Dealing with model convergence problems. Will focus on models that converge normally

@@
 
+# We only really want to compar the Hawaiian population, so we can subset that data
+  DisplayBgrd$pop_back_region <- with(DisplayBgrd, interaction(Population, Background_pop, BodyRegion))
+           DisplayBgrd_Hawaii <- DisplayBgrd %>% filter(Population == "Hawaii")
 ##########
@@
   # Fit the interaction model
-  DispBgrdAC <- lmer(dL ~ Population  + Background_pop + BodyRegion + (1|ID), data=DisplayBgrd)
-  summary(DispBgrdAC)
+  DispBgrdAC_inter <- lmer(dL ~  Background_pop + BodyRegion + Background_pop : BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
+  summary(DispBgrdAC_inter)
 
-  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-  robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+  # Main effercts
+   DispBgrdAC_main <- lmer(dL ~  Background_pop + BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
+  summary(DispBgrdAC_main)
 
-  # Add Wald test to test signifiacnce of each factor, background and population using roust variance
-  background_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(3), 
+  # Wald test of interaction; Not important, so we'll simplify and assume just additive effects. This makes sense when we look at the raw data.
+  intearction_AC_Disp <- Wald_test(DispBgrdAC_inter, constraints = constrain_zero(6:8), 
+                                  cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
+
+  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE).
+Given that there is no interaction, we can just present the main effects model with teh figure.
+Plus wald tests
+  robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC_main, cluster = DisplayBgrd_Hawaii$ID, vcov =  "CR2")
+
+  # Add Wald test to test signifiacnce of each factor using roust variance 
+  background_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(2), 
                                   cluster = DisplayBgrd$ID, vcov = "CR2")
 
-  Population_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(2), 
+  BodyRegion_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(3:5), 
                                   cluster = DisplayBgrd$ID, vcov = "CR2")
-
-  BodyRegion_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(4:6), 
-                                  cluster = DisplayBgrd$ID, vcov = "CR2")
-
-  # Planned contrast. Fit the model in a slightly different way ********** PROBLEMS HERE TO WORK OUT WITH CS
-  DisplayBgrd$pop_back_region <- with(DisplayBgrd, interaction(Population, Background_pop, BodyRegion))
-  DisplayBgrd_Hawaii <- DisplayBgrd %>% filter(Population == "Hawaii")
-  DisplayBgrd_pop_back_region <- lmer(dL ~ pop_back_region + (1|ID), data = DisplayBgrd_Hawaii) 
-  summary(DisplayBgrd_pop_back_region)
-
-  robust_DisplayBgrd_pop_back_region <- clubSandwich::conf_int(DisplayBgrd_pop_back_region, 
-                                                               cluster = DisplayBgrd_Hawaii$ID, 
-                                                               vcov =  "CR2")
-
-  # Contrasts hypothesis tests, but use RVE 
-    result_DisplayBgrd_pop_back_region <- Wald_test(DisplayBgrd_pop_back_region, 
-                                             constraints = constrain_pairwise(c(1:16)),
-                                             vcov="CR2", tidy = TRUE)
-    result_DisplayBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"),
-                                                       function(x) p_value(x))
 
@@
     CourtshipBgrd$pop_back_BR <- with(CourtshipBgrd, interaction(Population, Background_pop, BodyRegion))
-    CourtshipBgrd_Hawaii <-CourtshipBgrd %>% filter(Population == "Hawaii")
+         CourtshipBgrd_Hawaii <- CourtshipBgrd %>% filter(Population == "Hawaii")
     CourtshipBgrd_pop_back_BR <- lmer(dL ~ -1 + pop_back_BR + (1|ID), data = CourtshipBgrd_Hawaii)
@@
                                                    vcov="CR2", tidy = TRUE)
-      result_CourtshipBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_backCC$p_val, "bonferroni"), function(x) p_value(x))
+             result_CourtshipBgrd_pop_back_BR_CC$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_back_BR_CC$p_val, "bonferroni"), 
+                                                                    function(x) p_value(x))
 
@@
 
-         result_CourtshipBgrd_pop_backCC <- result_CourtshipBgrd_pop_back_BR_CC[
+         result_CourtshipBgrd_pop_back_BR_CC <- result_CourtshipBgrd_pop_back_BR_CC[
                                             grep(string, result_CourtshipBgrd_pop_back_BR_CC$hypothesis),]

### 2021-05-10 supp.Rmd
Fixed up last figure. Missed male-male context. Also added some new methods to body regions analysis.

@@
 
-  # Fit the interaction model
-  DispBgrdAC_inter <- lmer(dL ~  Background_pop + BodyRegion + Background_pop : BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
-  summary(DispBgrdAC_inter)
-
   # Main effercts
@@
     summary(DispBgrdAC_main)
-
-  # Wald test of interaction; Not important, so we'll simplify and assume just additive effects. This makes sense when we look at the raw data.
-  intearction_AC_Disp <- Wald_test(DispBgrdAC_inter, constraints = constrain_zero(6:8), 
-                                  cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
 
@@
 
+     # Fit the inetarction model
+          DispBgrdCC_inter <- lmer(dS ~ Background_pop  + BodyRegion + Background_pop:BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
+          summary(DispBgrdCC_inter)
+
     # Fit the main effect model
-          DispBgrdCC <- lmer(dS ~ Population  + Background_pop  + BodyRegion + (1|ID), data=DisplayBgrd)
-          summary(DispBgrdCC)
+          DispBgrdCC_main <- lmer(dS ~ Background_pop  + BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
+          summary(DispBgrdCC_main)
+
+    # Wald test to see if interaction is important. Yes, body regions way in how they change from a Hawaiian and Kenyan population. Fig S4, this makes sense. 
+        intearction_CC_Disp <- Wald_test(DispBgrdCC_inter, constraints = constrain_zero(6:8), 
+                                  cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
 
     # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-      robust_DispBgrdCC <- clubSandwich::coef_test(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+      robust_DispBgrdCC <- clubSandwich::coef_test(DispBgrdCC_inter, cluster = DisplayBgrd_Hawaii$ID, vcov =  "CR2")
 
-    # Check significance of each factor
-        background_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(3), 
-                                  cluster = DisplayBgrd$ID, vcov = "CR2")
-
-        Population_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(2), 
-                                  cluster = DisplayBgrd$ID, vcov = "CR2")
-
-        BodyRegion_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(4:6), 
-                                  cluster = DisplayBgrd$ID, vcov = "CR2")
-
-    # Planned contrasts. Refit model in slight different form
+    # Get contrasts to understand the specific changes better.
+Refit model in slight different form.
+This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
+The model above is just a contrast-based model.
     DisplayBgrd_pop_back_regionCC <- lmer(dS ~ -1 + pop_back_region + (1|ID), data = DisplayBgrd_Hawaii)
@@
 
+# Just subset out the Hawaiian data as we are only intersted in Hawaii on different backgrounds
+  CourtshipBgrd$pop_back_BR <- with(CourtshipBgrd, interaction(Population, Background_pop, BodyRegion))
+         CourtshipBgrd_Hawaii <- CourtshipBgrd %>% filter(Population == "Hawaii")
 ##########
@@
   # Fit main effects model
-    CourtshipAC <- lmer(dL ~ Population  + Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd)
-    summary(CourtshipAC)
+    CourtshipAC_main <- lmer(dL ~ Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
+    summary(CourtshipAC_main)
+
+  # Wald test of interaction; Not important, so we'll simplify and assume just additive effects. This makes sense when we look at the raw data.
+  intearction_AC_Court <- Wald_test(CourtshipAC_inter, constraints = constrain_zero(6:8), 
+                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
 
   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-   robust_CourtshipAC <- clubSandwich::conf_int(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+   robust_CourtshipAC <- clubSandwich::conf_int(CourtshipAC_inter, cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
 
@@
   # Planned contrasts / hypothesis tests. Need to fit model a little differently
-    CourtshipBgrd$pop_back_BR <- with(CourtshipBgrd, interaction(Population, Background_pop, BodyRegion))
-         CourtshipBgrd_Hawaii <- CourtshipBgrd %>% filter(Population == "Hawaii")
     CourtshipBgrd_pop_back_BR <- lmer(dL ~ -1 + pop_back_BR + (1|ID), data = CourtshipBgrd_Hawaii)
@@
 
+We tested whether different body regions of Hawaiian chameleons exhibited different contrasts between Hawaiian habitat background compared to their ancestral, Kenyan background.
+We started off testing models comparing the interaction between background enviroment and body region for both chormatic and achromatic contrast.
+However, for achromatic contrast, these models had poor convergence.
+Based on the raw data (Figure \@ref(fig:figS4)B & D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
+As such, we simplied models and only present the main effects of body region and background environment for achromatic contrast.
+However, inspection of the raw data (Figure \@ref(fig:figS4)A & C) suggested that body regions may differ in how chromatically contrasting they are against the background environment of Hawaii and Kenya.
+In addition, our main results suggested that predators may differentially drive chromatic signal conspicuousness, and given differet body regions maybe more exposed to bird compared to snake predators, we fit an interaction and main effects model for chromatic contrast.
+
 ```{r tableS6, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
@@
 ##########
-#Grab photos
 
 #plots
-p1.2 <- ggplot(courtship_summary, aes(x = Background_pop, y = Mean_dS, color = BodyRegion, group = BodyRegion)) +
+pS4.1 <- ggplot(courtship_summary, aes(x = Background_pop, y = Mean_dS, color = BodyRegion, group = BodyRegion)) +
   geom_point(size = 3, position = position_dodge(width=0.5)) +
@@
   ylab("Chromatic contrast (JND)") +
-  # annotate("segment", x = 0.85, xend = 1.85, y = 6.1, yend = 6.1, colour = "black") +
-  # annotate("segment", x = 1.12, xend = 2.12, y = 5.7, yend = 5.7, colour = "black") + 
-  # annotate("text", x = 0.85+(1.85-0.85)/2,  y = 6.2, label = "NS",  size = 6) + 
-  # annotate("text", x = 1.12+(2.12-1.12)/2, y = 5.8, label = "NS", size = 6) + 
   theme_bw() +
@@
 
-p2.2 <- ggplot(display_summary, aes(x = Background_pop, y = Mean_dL, group = BodyRegion, color = BodyRegion)) +
+pS4.2 <- ggplot(courtship_summary, aes(x = Background_pop, y = Mean_dL, color = BodyRegion, group = BodyRegion)) +
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Background Environment") + 
+  ylab("Achromatic contrast (JND)") +
+  theme_bw() +
+  labs(colour = "Body Region") +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+pS4.3 <- ggplot(display_summary, aes(x = Background_pop, y = Mean_dL, group = BodyRegion, color = BodyRegion)) +
   geom_point(size = 3, position = position_dodge(width=0.5)) +
@@
 
+pS4.4 <- ggplot(display_summary, aes(x = Background_pop, y = Mean_dS, group = BodyRegion, color = BodyRegion)) +
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Background Environment") + 
+  ylab("Chromatic contrast (JND)") +
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
 #pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
-plot_grid(p1.2, p2.2, labels = c("A", "B"), label_size = 30,
+plot_grid(pS4.4, pS4.3, pS4.1, pS4.2 , labels = c("A", "B", "C", "D"), label_size = 30,
   label_fontface = "bold", rel_widths=c(1,1))

### 2021-05-10 supp.Rmd
Fixing up code given convergence problems for dL.

@@
 
-  # Wald test of interaction; Not important, so we'll simplify and assume just additive effects. This makes sense when we look at the raw data.
-  intearction_AC_Court <- Wald_test(CourtshipAC_inter, constraints = constrain_zero(6:8), 
-                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
-
   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
@@
 
-           Population_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(2), 
-                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
-
            BodyRegion_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(4:6), 
                                   cluster = CourtshipBgrd$ID, vcov = "CR2")
-
-  # Planned contrasts / hypothesis tests. Need to fit model a little differently
-    CourtshipBgrd_pop_back_BR <- lmer(dL ~ -1 + pop_back_BR + (1|ID), data = CourtshipBgrd_Hawaii)
-
-   # Fit Robust model
-    robust_CourtshipBgrd_pop_back_BR <- clubSandwich::conf_int(CourtshipBgrd_pop_back_BR, 
-                                                            cluster = CourtshipBgrd_Hawaii$ID, 
-                                                            vcov =  "CR2")
-
-    # Contrasts hypothesis tests, but use RVE ********** PROBLEM WITH CONVERGENCE NON-POSITIVE DEF MATRICES
-      result_CourtshipBgrd_pop_back_BR <- Wald_test(CourtshipBgrd_pop_back_BR, 
-                                                 constraints = constrain_pairwise(c(1:8)), 
-                                                 vcov="CR2", tidy = TRUE)
-
-      result_CourtshipBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
 
@@
   # Fit main effect model
-      CourtshipCC <- lmer(dS ~ Population  + Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd)
-      summary(CourtshipCC)
+      CourtshipCC_inter <- lmer(dS ~ Background_pop + BodyRegion + Background_pop:BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
+      summary(CourtshipCC_inter)
+
+  # Fit main effect model
+      CourtshipCC_main <- lmer(dS ~ Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
+      summary(CourtshipCC_main)
 
   # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
-  robust_CourtshipCC <- clubSandwich::coef_test(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+  robust_CourtshipCC <- clubSandwich::coef_test(CourtshipCC_inter, cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
+
+  # Wald test of interaction; Not important, so we'll simplify and assume just additive effects. This makes sense when we look at the raw data.
+  intearction_CC_Court <- Wald_test(CourtshipCC_inter, constraints = constrain_zero(6:8), 
+                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
 
    # Wald tests
-           background_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(3), 
-                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+           background_CC_Court <- Wald_test(CourtshipCC_main, constraints = constrain_zero(2), 
+                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
 
-           Population_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(2), 
-                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+           BodyRegion_CC_Court <- Wald_test(CourtshipCC_main, constraints = constrain_zero(3:5), 
+                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
 
-           BodyRegion_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(4:6), 
-                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
-
-     # Planned comparions
+    # Get contrasts to understand the specific changes better.
+Refit model in slight different form.
+This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
+The model above is just a contrast-based model.
 
@@
 
-#### All body regions exibit similar contrasts between Kenyan and Hawaiian backgrounds suggesting local adaptation of entire visual signal
-
-We tested whether different body regions of Hawaiian chameleons exhibited different contrasts between Hawaiian habitat background compared to their ancestral, Kenyan background.
-We started off testing models comparing the interaction between background enviroment and body region for both chormatic and achromatic contrast.
-However, for achromatic contrast, these models had poor convergence.
-Based on the raw data (Figure \@ref(fig:figS4)B & D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
-As such, we simplied models and only present the main effects of body region and background environment for achromatic contrast.
-However, inspection of the raw data (Figure \@ref(fig:figS4)A & C) suggested that body regions may differ in how chromatically contrasting they are against the background environment of Hawaii and Kenya.
-In addition, our main results suggested that predators may differentially drive chromatic signal conspicuousness, and given differet body regions maybe more exposed to bird compared to snake predators, we fit an interaction and main effects model for chromatic contrast.
-
 ```{r tableS6, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
@@
 
-```{r FigS4, fig.width = 13, fig.height = 7.5, fig.cap = "Local adaptation.
+#### All body regions exibit similar contrasts between Kenyan and Hawaiian backgrounds suggesting local adaptation of entire visual signal
+
+We tested whether different body regions of Hawaiian chameleons exhibited different contrasts between Hawaiian habitat background compared to their ancestral, Kenyan background.
+We started off testing models comparing the interaction between background enviroment and body region for both chormatic and achromatic contrast.
+However, for achromatic contrast, these models had poor convergence.
+Based on the raw data (Figure \@ref(fig:figS4)B & D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
+As such, we simplied models and only present the main effects of body region and background environment for achromatic contrast.
+However, inspection of the raw data (Figure \@ref(fig:figS4)A & C) suggested that body regions may differ in how chromatically contrasting they are against the background environment of Hawaii and Kenya.
+In addition, our main results suggested that predators may differentially drive chromatic signal conspicuousness, and given differet body regions maybe more exposed to bird compared to snake predators, we fit an interaction and main effects model for chromatic contrast.
+
+```{r FigS4, fig.width = 13, fig.height = 12, fig.cap = "Local adaptation.
 All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.

### 2021-05-10 R/func.R
Fixed fonts for table styles so they are all teh same. Some small errors, probably explains why some tables were not being corrected.

@@ -101,7 +101,9 @@ table_style  <- function(table){
     compose(part = "body", i = 7, j = 2, 
             value = as_paragraph("Kenya", as_sub(" tailbase"), " - ", "Hawaii", as_sub(" tailbase"))) %>% 
     compose(part = "body", i = 8, j = 2, 
-            value = as_paragraph("Kenya", as_sub(" topflank"), " - ", "Hawaii", as_sub(" topflank")))
+            value = as_paragraph("Kenya", as_sub(" topflank"), " - ", "Hawaii", as_sub(" topflank")))%>% 
+    font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header")
   return(table)
 }
 
@@ -115,7 +117,9 @@ table_style2 <- function(table){
     compose( part = "header", j = 4, value = as_paragraph(as_b("F"))) %>% 
     compose(part = "header", j = 5, value = as_paragraph(as_b("df"), as_sub(as_b(" num")))) %>%
     compose(part = "header", j = 6, value = as_paragraph(as_b("df"), as_sub(as_b(" denom")))) %>% 
-    compose(part = "header", j = 7, value = as_paragraph(as_b("p"))) 
+    compose(part = "header", j = 7, value = as_paragraph(as_b("p"))) %>% 
+    font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header")
   return(table)
 }
 
@@ -128,7 +132,9 @@ table_style3 <- function(table){
     compose( part = "header", j = 3, value = as_paragraph(as_b("F"))) %>% 
     compose(part = "header", j = 4, value = as_paragraph(as_b("df"), as_sub(as_b(" num")))) %>%
     compose(part = "header", j = 5, value = as_paragraph(as_b("df"), as_sub(as_b(" denom")))) %>% 
-    compose(part = "header", j = 6, value = as_paragraph(as_b("p"))) 
+    compose(part = "header", j = 6, value = as_paragraph(as_b("p"))) %>% 
+    font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header")
   return(table)
 }
 
@@ -154,8 +160,8 @@ table_style4 <- function(table){
     compose(part = "body", j = 3, i = c(1, 10, 19, 28)+6, value = as_paragraph("Population", as_sub(" Kenya"), "* BodyRegion", as_sub(" midflank"))) %>%
[…]

### 2021-05-10 supp.Rmd
Added final tables for local adaptation to the MS

@@
     # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-      robust_DispBgrdCC <- clubSandwich::coef_test(DispBgrdCC_inter, cluster = DisplayBgrd_Hawaii$ID, vcov =  "CR2")
+      robust_DispBgrdCC <- clubSandwich::conf_int(DispBgrdCC_inter, cluster = DisplayBgrd_Hawaii$ID, vcov =  "CR2")
 
@@
                                                                   function(x) p_value(x))
+     # Grab contrasts that we need from the robust esatimtes
+      string <- c("pop_back_regionHawaii.Kenya.gular - pop_back_regionHawaii.Hawaii.gular|pop_back_regionHawaii.Kenya.midflank - pop_back_regionHawaii.Hawaii.midflank|pop_back_regionHawaii.Kenya.tailbase - pop_back_regionHawaii.Hawaii.tailbase|pop_back_regionHawaii.Kenya.topflank - pop_back_regionHawaii.Hawaii.topflank")
+      result_DisplayBgrd_pop_back_regionCC <- result_DisplayBgrd_pop_back_regionCC[grep(string, result_DisplayBgrd_pop_back_regionCC$hypothesis),]
 
@@
   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-   robust_CourtshipAC <- clubSandwich::conf_int(CourtshipAC_inter, cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
+   robust_CourtshipAC <- clubSandwich::conf_int(CourtshipAC_main, cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
 
@@
   # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
-  robust_CourtshipCC <- clubSandwich::coef_test(CourtshipCC_inter, cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
+  robust_CourtshipCC <- clubSandwich::conf_int(CourtshipCC_inter, cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
 
@@
 
+The magnitude of chromatic contrast differences between Hawaiian chamaleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:figS4)A & C & Table \@ref(tab:table7)).
+However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:figS4)A & C), and only some body regions showed a significant decrease (Table \@ref(tab:table8)).
+
+The magnitude of achromatic contrast, showed a significant overall decrease in achromatic contrast when viewed by conspevcifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court$p_val)`, Figure \@ref(fig:figS4)B & D along with & Table \@ref(tab:table7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, p = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, p = `r p_value(BodyRegion_AC_Court$p_val)`, Figure \@ref(fig:figS4)B & D along with & Table \@ref(tab:table7)).
+
 ```{r FigS4, fig.width = 13, fig.height = 12, fig.cap = "Local adaptation.
@@
 
+```{r table7, fig.cap = "Models"}
+table7 <- rbind(robust_localAdapt_DispBgrdAC,
+                robust_DispBgrdCC,
+                robust_CourtshipAC,
+                robust_CourtshipCC) %>% as.data.frame()
+
+table7$Paramater <- rownames(table7)
+rownames(table7) <- NULL
+
+table7 <- table7 %>% 
+              mutate(Context = c("Male-Male Display", rep("", 12), "Courtship Display", rep("", 12)),
+                     JND = c("Achromatic Contrast", rep("", 4), "Chromatic Contrast", rep("", 7), 
+                             "Achromatic Contrast", rep("", 4), "Chromatic Contrast", rep("", 7))) %>% 
+                      select(Context, JND, Paramater, beta, SE, df, CI_L, CI_U)
+
+Table7 <- flextable(table7) %>% table_style10()
+
+```
+
+```{r table8, fig.cap = "contarsts chromatic"}
+tab8 <- rbind(result_DisplayBgrd_pop_back_regionCC,
+              result_CourtshipBgrd_pop_back_BR_CC) %>%  as.data.frame()
+
+tab8 <- tab8 %>% mutate(Context = c("Male-Male Display", rep("",3), "Courtship Display", rep("",3)),
+                        p_val = sapply(p_val, function(x) p_value(x))) %>% 
+                        select(Context,c(1:8), -test, -delta, -df_num)
+
+Table8 <- flextable(tab8) %>% table_style11()
+Table8
+```
+
 # References

### 2021-05-10 ms.Rmd
Added export table from main MS so that all table can quickly be brought into the supp and rendered all together

@@
               select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
+ write.csv(final_T1, file = "./output/tables/Table1_MS.csv", row.names = FALSE)
 
@@
               select(JND, Parameter, beta, SE, df, CI_L, CI_U)
-
+ write.csv(final_TT, file = "./output/tables/Table2_MS.csv", row.names = FALSE)
  # Table
@@
             select(VisualSpec, comparison, Estimate, lwr, upr, Fstat, df, p_val, p_bonf)
+
+write.csv(contrast, file = "./output/tables/Table3_MS.csv", row.names = FALSE)
 
@@
 
+write.csv(tb4, file = "./output/tables/Table4_MS.csv", row.names = FALSE)
 table4 <- flextable(tb4) %>% table_style9()

### 2021-05-10 ms.Rmd
Updated MS to export tables, tracking tables so Supp can grab these and bring them into main supp files. Fixed functions to ensure that all have fonts correct

@@
               select(JND, Parameter, beta, SE, df, CI_L, CI_U)
- write.csv(final_TT, file = "./output/tables/Table2_MS.csv", row.names = FALSE)
+ write.csv(final_T2, file = "./output/tables/Table2_MS.csv", row.names = FALSE)
  # Table

### 2021-05-10 supp.Rmd
Updated MS to export tables, tracking tables so Supp can grab these and bring them into main supp files. Fixed functions to ensure that all have fonts correct

@@
   background_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(2), 
-                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+                                  cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
 
   BodyRegion_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(3:5), 
-                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+                                  cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
 
@@
    # Wald tests
-           background_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(3), 
-                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
-
-           BodyRegion_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(4:6), 
-                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+           background_AC_Court <- Wald_test(CourtshipAC_main, constraints = constrain_zero(3), 
+                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
+
+           BodyRegion_AC_Court <- Wald_test(CourtshipAC_main, constraints = constrain_zero(3:5), 
+                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
 
@@
 
+## Supporting Supplementay Tables
+
+```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
+Standard errors are corrected by using robust variance estimators.
+"}
+
+final_T1 <-  read.csv(file = "./output/tables/Table1_MS.csv")
+
+ # Table
+table1 <- flextable(final_T1) %>% table_style6()
+
+table1
+```
+
+```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
+Full models are provided and standard errors are corrected by using robust variance estimators."}
+
+ final_T2 <- read.csv(file = "./output/tables/Table2_MS.csv")
+ # Table
+table2 <- flextable(final_T2) %>%
+           table_style7()
+
+table2
+```
+
+```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean acromatric and chormatic contrast in Hawaii and Kenya for both predators (snakes and birds).
+Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
+Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).
+We also applied a Bonferroni correction to p-value to control for multiple comparisons."}
+
+contrast <- read.csv(file = "./output/tables/Table3_MS.csv")
+
+table3 <- flextable(contrast) %>% table_style8()
+table3
+```
+
+```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for comparing acromatric and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (Background = Kenya) in both social contexts (Male-male competition and Courtship).
+Models were fit using robust variance estaimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.
+Significance parameter estimates are those where 95% CI does not include zero."}
+
+tb4 <- read.csv(file = "./output/tables/Table4_MS.csv")
+table4 <- flextable(tb4) %>% table_style9()
+table4
+```
+
 ## Assessing changes in sexual signals across different body regions
@@
 A significant interaction between body region and population was found for achormatic contrast when males were displaying to each other (Table \@ref(tab:tableS1)).
-However, in all cases Kenya showed a string reduction in in acromatic contrast relative to Hawaii lizards, despite some variation in the magnitude of effect (Figure \@ref(fig:FigS2)B & D; Table \@ref(tab:tableS2)).
+However, in all cases Kenya showed a strong reduction in acromatic contrast relative to Hawaii lizards, despite some variation in the magnitude of effect (Figure \@ref(fig:FigS2)B & D; Table \@ref(tab:tableS2)).
 
@@
 
-```{r tableS1, tab.cap = "Wald Tests comparing significance of model parameters.
-Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
+```{r tableS1, tab.cap = "Wald F tests comparing overall significance of model parameters.
+Wald F tests were computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
 Interactions were first tested and models refit to test the signifiacnce of main effects"}
@@
 
-```{r tableS2, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction.Social Context"}
+```{r tableS2, tab.cap = "Linear mixed effects model results using robust variance estimators and Saitterwaite degrees of freedom correction.Social Context"}
 robust_sc_table <- rbind(robust_ConspicDisplayAC,
@@
 
-```{r FigS3, fig.width = 12, fig.height = 12, fig.cap = "Predators All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
-Mean JNDs are provided along with 95% confidence intervals.
-A & B represent male-male contest displays whereas C & D represent courtship displays"}
+```{r FigS3, fig.width = 12, fig.height = 12, fig.cap = "All four body regions (gular, mid-flank, tailbase and topflank) and their achromatic (B) and chromatic contrast (A) to Hawaiian and Kenyan backgrounds (each population against their own background) are shown averaged across stems and leaves found in their habitat.
+Just noticable differences (JNDs) are calculated for both bird and snake predators.
+Mean JNDs are provided along with 95% confidence intervals."}
 
@@
 
-```{r tableS3, tab.cap = "Wald Tests comparing significance of model parameters for predators.
-Wald tests computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
+```{r tableS3, tab.cap = "Wald F tests comparing the significance of model parameters for predator models.
+Wald F tests were computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
 Interactions were first tested and models refit to test the signifiacnce of main effects"}
@@
 
-```{r tableS4, tab.cap = "Linear mixed effects model results using sandwich estimators and Saitterwaite degrees of freedom correction.Predators"}
+```{r tableS4, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter from linear mixed effects models.
+Standard errors were corrected using robust variance estimators and Saitterwaite degrees of freedom correction was used for calculated significance"}
 robust_sc_pred_table <- rbind(robust_PredatorAC,
@@
 ft5 <- flextable(final_r5) %>%
-        table_style(.) %>% 
-        add_header_row(values = "Achromatic Contrast", colwidths = 6)
+        table_style(.)
 
@@
  ft6 <- flextable(final_r6) %>%
-        table_style(.) %>% 
-        add_header_row(values = "Chromatic Contrast", colwidths = 6)
+        table_style(.)
 
@@
 However, for achromatic contrast, these models had poor convergence.
-Based on the raw data (Figure \@ref(fig:figS4)B & D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
+Based on the raw data (Figure \@ref(fig:FigS4)B & D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
 As such, we simplied models and only present the main effects of body region and background environment for achromatic contrast.
-However, inspection of the raw data (Figure \@ref(fig:figS4)A & C) suggested that body regions may differ in how chromatically contrasting they are against the background environment of Hawaii and Kenya.
+However, inspection of the raw data (Figure \@ref(fig:FigS4)A & C) suggested that body regions may differ in how chromatically contrasting they are against the background environment of Hawaii and Kenya.
 In addition, our main results suggested that predators may differentially drive chromatic signal conspicuousness, and given differet body regions maybe more exposed to bird compared to snake predators, we fit an interaction and main effects model for chromatic contrast.
 
-The magnitude of chromatic contrast differences between Hawaiian chamaleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:figS4)A & C & Table \@ref(tab:table7)).
-However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:figS4)A & C), and only some body regions showed a significant decrease (Table \@ref(tab:table8)).
+The magnitude of chromatic contrast differences between Hawaiian chamaleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:FigS4)A & C & Table \@ref(tab:table7)).
+However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:FigS4)A & C), and only some body regions showed a significant decrease (Table \@ref(tab:table8)).
 
@@
 
-```{r table7, fig.cap = "Models"}
+```{r table7, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter.
+Main effects and interaction models are presented.
+Only main effect models are presented for achromatic contrast (see text for details), whereas interaction models are presented for chromatic contrast.
+Pairwise contrasts are presented for chromatic contrast models (see Table 8) for ease of interpretation."}
 table7 <- rbind(robust_localAdapt_DispBgrdAC,
@@
 
-```{r table8, fig.cap = "contarsts chromatic"}
+```{r table8, tab.cap = "Pairwise Wald F tests comparing changes in chromatic contrast between different body regions for Hawaiian lizards against both Hawaiian and Kenyan backgrounds during male-male and courtship displats.
+F statistic, degrees of freedom (DF) and p-values are provided (p).
+A bonferroni correction to p-value is also provided to control for multiple testing.
+See Table 7 for model estimates."}
 tab8 <- rbind(result_DisplayBgrd_pop_back_regionCC,

### 2021-05-10 R/func.R

@@ -77,7 +77,7 @@ wald_table_p <- function(model_main, model_inter){
 # Table styling functions
 
 table_style  <- function(table){
-  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 2.5) %>% 
+  table <- table  %>% 
     align(align = "center", part = "header") %>% 
     align(align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Predator"))) %>% 
@@ -101,7 +101,7 @@ table_style  <- function(table){
     compose(part = "body", i = 7, j = 2, 
             value = as_paragraph("Kenya", as_sub(" tailbase"), " - ", "Hawaii", as_sub(" tailbase"))) %>% 
     compose(part = "body", i = 8, j = 2, 
-            value = as_paragraph("Kenya", as_sub(" topflank"), " - ", "Hawaii", as_sub(" topflank")))%>% 
+            value = as_paragraph("Kenya", as_sub(" topflank"), " - ", "Hawaii", as_sub(" topflank"))) %>% 
     font(fontname="Times", part = "body") %>% 
     font(fontname="Times", part = "header")
   return(table)

### 2021-05-10 ms.Rmd
Fixes to MS, but should have 1st draft of everything now

@@
 Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals whereas greater conspicuousness of sexual signals should be selected in social contexts.
-In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; Courtship displays: Acromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was  true across all body regions measured (See Supplemental Materials, table S1).
+In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; Courtship displays: Acromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was  true across all body regions measured (See Supplemental Materials).
 Kenyan chameleons from populations where predation is high were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, Courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see figure 1a & b).
@@
 Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment.
-While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chamleeons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
+While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chameleons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
 

### 2021-05-10 supp.Rmd
Fixes to MS, but should have 1st draft of everything now

@@
 
-The magnitude of chromatic contrast differences between Hawaiian chamaleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:FigS4)A & C & Table \@ref(tab:table7)).
-However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:FigS4)A & C), and only some body regions showed a significant decrease (Table \@ref(tab:table8)).
-
-The magnitude of achromatic contrast, showed a significant overall decrease in achromatic contrast when viewed by conspevcifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court$p_val)`, Figure \@ref(fig:figS4)B & D along with & Table \@ref(tab:table7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, p = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, p = `r p_value(BodyRegion_AC_Court$p_val)`, Figure \@ref(fig:figS4)B & D along with & Table \@ref(tab:table7)).
+The magnitude of chromatic contrast differences between Hawaiian chamaleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:FigS4)A & C & Table \@ref(tab:tableS7)).
+However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:FigS4)A & C), and only some body regions showed a significant decrease (Table \@ref(tab:tableS8)).
+
+The magnitude of achromatic contrast, showed a significant overall decrease in achromatic contrast when viewed by conspevcifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court$p_val)`, Figure \@ref(fig:FigS4)B & D along with Table \@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, p = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, p = `r p_value(BodyRegion_AC_Court$p_val)`, Figure \@ref(fig:FigS4)B & D along with Table \@ref(tab:tableS7)).
 
@@
 
-```{r table7, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter.
+```{r tableS7, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter.
 Main effects and interaction models are presented.
@@
 Table7 <- flextable(table7) %>% table_style10()
-
-```
-
-```{r table8, tab.cap = "Pairwise Wald F tests comparing changes in chromatic contrast between different body regions for Hawaiian lizards against both Hawaiian and Kenyan backgrounds during male-male and courtship displats.
+Table7
+```
+
+```{r tableS8, tab.cap = "Pairwise Wald F tests comparing changes in chromatic contrast between different body regions for Hawaiian lizards against both Hawaiian and Kenyan backgrounds during male-male and courtship displats.
 F statistic, degrees of freedom (DF) and p-values are provided (p).

### 2021-05-10 R/func.R

@@ -81,7 +81,7 @@ table_style  <- function(table){
     align(align = "center", part = "header") %>% 
     align(align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Predator"))) %>% 
-    compose( part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>% 
+    compose(part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>% 
     compose(part = "header", j = 3, value = as_paragraph(as_b("F"))) %>%
     compose(part = "header", j = 4, value = as_paragraph(as_b("DF"))) %>% 
     compose(part = "header", j = 5, value = as_paragraph(as_b("p"))) %>% 
@@ -103,12 +103,12 @@ table_style  <- function(table){
     compose(part = "body", i = 8, j = 2, 
             value = as_paragraph("Kenya", as_sub(" topflank"), " - ", "Hawaii", as_sub(" topflank"))) %>% 
     font(fontname="Times", part = "body") %>% 
-    font(fontname="Times", part = "header")
+    font(fontname="Times", part = "header") 
   return(table)
 }
 
 table_style2 <- function(table){
-  table <- table %>%  width(j = 3, width = 3) %>% 
+  table <- table %>%  width(j = 2, width = 1.5) %>% 
     align(align = "center", part = "header") %>% 
     align( align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Social Context"))) %>% 
@@ -140,7 +140,6 @@ table_style3 <- function(table){
 
 table_style4 <- function(table){
   table <- table %>%
-    width(j = 3, width = 6) %>% 
     align(align = "center", part = "header") %>% 
     align(align = "center", part = "body") %>% 
     compose(part = "header", j = 1, value = as_paragraph(as_b("Social Context"))) %>% 
@@ -160,6 +159,7 @@ table_style4 <- function(table){
     compose(part = "body", j = 3, i = c(1, 10, 19, 28)+6, value = as_paragraph("Population", as_sub(" Kenya"), "* BodyRegion", as_sub(" midflank"))) %>%
     compose(part = "body", j = 3, i = c(1, 10, 19, 28)+7, value = as_paragraph("Population", as_sub(" Kenya"), "* BodyRegion", as_sub(" tailbase"))) %>%
[…]

### 2021-05-10 R/spec_figs.R
Addding spec data and scrips do that releavnt spectral reflectance curves can be added

@@ -1,0 +1,81 @@
+# Create spectral reflectance curves
+  rm(list=ls())
+# Libraries
+  pacman::p_load(pavo, tidyverse)
+  
+  # Function to extract, clean and aggregate background
+  get_background <- function(dir, background = "leaf", label = NULL){
+    ## Import spec data, name wl
+    bkgrd <- read.csv(file = dir)
+    bkgrd <- bkgrd %>% rename("wl" = "wavelength")
+    
+    ## Get leaves
+    bkgrd <- bkgrd[, c("wl",grep(background, colnames(bkgrd), value = TRUE))]
+    
+    ## Smooth specs
+    bkgrd <- procspec(bkgrd, opt = "smooth" , span = 0.2)
+    
+    # Aggregate specs
+    avg_bkgrd <- aggspec(bkgrd)
+    colnames(avg_bkgrd)[2] <- paste0(background, ".",  label)
+   
+     return(avg_bkgrd)
+    
+  }
+  
+  
+  leaves_hawaii <- get_background(dir = "./data/specs/processed/hawaii_background.csv", label = "hawaii")
+   leaves_kenya <- get_background(dir = "./data/specs/processed/kenya_background.csv", label = "kenya")
+  
+  # Function to extract, clean and aggregate body regions of interst. More than one body region at once. Out put is a column for each body region, the boy regions are aggregated across all individuals
+   get_body_regions <- function(dir, body_region = "gular|topflank", label = NULL){
+     ## Import spec data, name wl
+     bkgrd <- read.csv(file = dir)
+     bkgrd <- bkgrd %>% rename("wl" = "wavelength")
+     
+     ## Get body regions of interest
+     bkgrd <- bkgrd[, c("wl",grep(body_region, colnames(bkgrd), value = TRUE))]
+     
+     ## Smooth specs
+     bkgrd <- procspec(bkgrd, opt = "smooth" , span = 0.2)
+     
+     # Aggregate specs
+     body_region_split <- unlist((strsplit(body_region, "|", fixed = TRUE)))
+     avg_bkgrd <- data.frame(wl = bkgrd[,1])
+     
+     for(i in 1:length(body_region_split)){
+            avg_bkgrd[,i+1] <- aggspec(bkgrd[, c("wl", grep(body_region_split[i], colnames(bkgrd), value = TRUE))])[,2]
+            colnames(avg_bkgrd)[i+1] <- paste0(body_region_split[i],".",label)
+     }
+     
+     return(avg_bkgrd)
+     
+   }
[…]

### 2021-05-10 R/spec_figs.R
Created all spec figures

@@ -57,13 +57,21 @@
    
   
 # Merge all together, add some grouping variables
-    bird_spec_data <- cbind(leaves_hawaii, leaves_kenya, bodyregions_hawaii_bird, bodyregions_kenya_bird)[,-c(3,5,8)]
-    bird_spec_data <- pivot_longer(bird_spec_data, cols = c(2:7)) %>% arrange(name) %>% 
-                          mutate(bodyreg = gsub(".hawaii|.kenya", "", name),
-                                 pop = gsub("gular.|topflank.|leaf.", "", name))
+
+data_merge <- function(table1_background, table2_background, table3, table4){
+    spec_data <- cbind(table1_background, table2_background, table3, table4)[,-c(3,5,8)]
+    spec_data <- pivot_longer(spec_data, cols = c(2:7)) %>% 
+                 arrange(name) %>% 
+                 mutate(bodyreg = gsub(".hawaii|.kenya", "", name),
+                            pop = gsub("gular.|topflank.|leaf.", "", name))
+    return(spec_data)
+}
+bird_spec_data <- data_merge(leaves_hawaii, leaves_kenya, bodyregions_hawaii_bird, bodyregions_kenya_bird)
+
 # Plot!
   pdf(file = "./output/figures/spec_birds.pdf", width = 8.588235,  height = 6.955882)  
-      ggplot(bird_spec_data, aes(x = wl, y = value, group = name)) + 
+plot_spec <- function(data){
+        ggplot(data, aes(x = wl, y = value, group = name)) + 
       geom_line(aes(colour = bodyreg, linetype = pop), size = 1.2) + 
       theme_bw() +
       labs(x = "Wavelength (nm)", y = "Reflectance (%)", linetype = "Population", colour = "Body Region / Environment") + 
@@ -75,7 +83,54 @@
             axis.text = element_text(size = 16),
             axis.title = element_text(size = 24),
             legend.text = element_text(size = 16),
-            legend.title = element_text(size = 20))
+            legend.title = element_text(size = 20))}
+      plot_spec(bird_spec_data)
+
   dev.off()  
   
-  
+
+### Now, run through more simplified processing using all the functions, we already have the background, so now it's just getting the contexts
+
+  # Get snakes
[…]

### 2021-05-10 ms.Rmd
Fancy new figure 1 and updated MS.

@@
 JNDs are calculated based on the chameleon visual system.
-NS denotes non-significant contasts, whereas astericks indicates significant differences is < 0.001."}
+NS denotes non-significant contasts, whereas astericks indicates significant differences is < 0.001.
+Insets showing spectral reflectance curves of two representative body regions ('gular' and 'top-flank), along with average signalling background (averaged leaves) are provided for each social and predatory context.
+A) spectral curves of signals during male-male displays; B) spectral curves of signals during courtship displays; C) spectral curves of signals during bird encounters; and D) spectral curves of signals during snake encounters."}
 rerun = FALSE
@@
 
-### Supplementary Tables - Add to Supplement. Easier to generate here given all code is here.
-
-# Social context. We only need models because contarsts are already there
-
 ```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
@@
 
- # Table
-table1 <- flextable(final_T1) %>% table_style6()
-
-table1
 ```
 
-# Predators - Models AND contrasts. We don't get all contrasts because we have two factors in model
 ```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
@@
  write.csv(final_T2, file = "./output/tables/Table2_MS.csv", row.names = FALSE)
- # Table
-table2 <- flextable(final_T2) %>%
-           table_style7()
 
-table2
 ```
@@
 
-table3 <- flextable(contrast) %>% table_style8()
-table3
 ```
-
-# Local Adaptation.
-We have all the contrasts we need from the model because we are only comparing Hawaii-Hawaii(background) to Hawaii-Kenya(background).
-These are therefore in the table output already.
 
@@
 write.csv(tb4, file = "./output/tables/Table4_MS.csv", row.names = FALSE)
-table4 <- flextable(tb4) %>% table_style9()
-table4
+
 ```

### 2021-05-31 R/func.R
Added new table formatting

@@ -322,3 +322,17 @@ table_style11  <- function(table){
   
   return(table)
 }
+
+
+table_style12 <- function(table){
+  table <- table %>%  width(j = 1, width = 1.5) %>% width(j = 2, width = 3) %>% 
+    align(align = "center", part = "header") %>% 
+    compose(part = "header", j = 1, value = as_paragraph(as_b("Hypothesis"))) %>% 
+    compose(part = "header", j = 2, value = as_paragraph(as_b("Prediction"))) %>% 
+    compose(part = "header", j = 3, value = as_paragraph(as_b("Population"))) %>% 
+    compose( part = "header", j = 4, value = as_paragraph(as_b("Visual System"))) %>% 
+    compose( part = "header", j = 5, value = as_paragraph(as_b("Background"))) %>% 
+    font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header")
+  return(table)
+}

### 2021-05-31 ms.Rmd
Transferred over Martin's revisions, fixed up figures

@@
 ---
-title: "Evolution of a conspicuous dynamic visual signal in an introduced chameleon"
+title: "Rapid evolution of a conspicuous dynamic visual signal in an introduced chameleon"
 author: Whiting et al. 
@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, magick, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
 source("./R/func.R")
@@
 
-    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+    # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
+    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out missing data
        robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, 
@@
 
- # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-    Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+ # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
+    Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out missing data
     robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, 
@@
 
-  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+  # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
   robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
@@
 
-Martin J. Whiting $^{1,*}$, Brenden S. Holland $^{2}$, J. Scott Keogh$^{3}$, Daniel W.A.Noble$^{3}$, Katrina J. Rankin$^{4}$ and Devi M. Stuart-Fox$^{4}$
+Martin J. Whiting <sup>1,*<sup>, Brenden S. Holland <sup>2<sup>, J. Scott Keogh <sup>3<sup>, Daniel W.A.Noble <sup>3<sup>, Katrina J. Rankin <sup>4<sup> and Devi M. Stuart-Fox <sup>4<sup>
 
-$^{1}$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
+1 Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
+2 Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
+3 Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
+4 Department of Zoology, The University of Melbourne, Melbourne, Victoria 3010, Australia
 
@@
 Rapid evolution frequently follows a significant change to an organism’s environment, such as during colonization of a novel habitat.
-Vertebrates, including Caribbean Anolis lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales.
-Traits that are sexually selected can be similarly affected, although well-documented examples from wild populations are rare.
-The Jackson’s chameleon (*Chamaeleo x.
-jacksonii*) from East Africa was inadvertently introduced to the Hawaiian island of Oahu in 1972.
-In its native range, Jackson’s chameleons are preyed upon by a wide range of predators, including snakes and birds, which are thought to shape the dynamics of color change in chameleons.
-We took advantage of the Hawaiian introduction, an environment with only a few potential predators and therefore, very low predation pressure, to test for character release of conspicuous social color signals.
-To test this hypothesis, we measured color change in response to multiple social and predator stimuli in wild chameleons in both Hawai’i and Kenya.
-We used visual modelling of colour signals to estimate signal conspicuousness to both the chameleon and predator (bird, snake) visual systems.
-The Hawai’i and Kenya backgrounds were a similar color, but the major difference was in brightness: the Hawai’i background was darker.
-As a result, chameleons from both Kenya and Hawai’i were brighter against the Hawai’i background.
-However, Hawai’i males were more conspicuous (courtship, display) against their own background for some body regions (mainly gular), providing evidence for character release.
-Also, Hawai’i males were more conspicuous (courtship, display) to snake and bird predators against their own background.
-We suggest that relaxing natural selection can result in signal release in systems experiencing strong sexual selection.
+We show that Jackson’s chameleons (*Trioceros j.
+xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) have evolved significantly more conspicuous sexual signals.
+Compared to Kenya, Oahu has few potential predators and therefore, very low predation pressure.
+We used visual modelling to quantify the conspicuousness of wild chameleons from both Hawaii and Kenya to multiple social and predator stimuli in order to test for character release of sexual signals.
+The strongest effect was for achromatic contrast, where Hawaiian chameleons were brighter during male contests and courtship and were more conspicuous to snake and bird predators.
+Chameleon signals also showed evidence of local adaptation.
+Relaxing natural selection can result in signal release in systems experiencing strong sexual selection, including animals with dynamic colour change, and may act differently on the two primary visual channels (chromatic and achromatic).
 
@@
 
-When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods.
-This view understandably persisted for at least another 120 years following the publication of On the Origin of Species (Darwin, 1859).
+When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods (1).
 We now have numerous examples of evolutionary change following natural selection in both the laboratory and wild, and not only in invertebrates with short generation times.
-Vertebrates, including Caribbean Anolis lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales (e.g.
-@losos97).
-Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in phenotype over short time intervals.
+Vertebrates, including Caribbean *Anolis* lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales (e.g.
+(2, 3)).
+Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in behaviour and phenotype over short time intervals (e.g., (4-6)).
 
-Conspicuous colour signals evolve because of the fitness advantage conferred to the signaller, and frequently becomes increasingly elaborate through a combination of male contest competition and female preference (e.g.
-runaway selection).
-However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly.
+Conspicuous colour signals evolve because of the fitness advantage conferred to the signaller.
+These signals frequently become increasingly elaborate through a combination of male contest competition and female preference (e.g.
+runaway selection (7, 8)).
+However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly (9).
 Evolution can respond to this selection pressure in a variety of ways that include shifting conspicuous signals to areas of the body that might be concealed to predators.
-For example, many conspicuous lizards have colourful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards.
-Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-dense streams.
+For example, many conspicuous lizards have colourful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards (10).
+Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-rich streams (6, 9).
 These examples are from animals with static colour signals.
-A parallel solution that reduces the risk of exposure to predators is dynamic colour change because signals can be switched on and off relatively rapidly through neural or endocrine control.
+A parallel solution that reduces the risk of exposure to predators is dynamic colour change because signals can be switched on and off relatively rapidly through neural or endocrine control (11, 12).
 A wide range of invertebrates (e.g.
 cephalopods) and vertebrates (e.g.
-lizards and frogs) have convergently evolved this ability.
+lizards and frogs) have convergently evolved this ability (13-15).
 
@@
 These novel environments typically have a completely different suite of organisms, including predators, and offer a unique opportunity to study signal evolution in founder populations.
-Jackson’s chameleons (*Chamaeleo j.
-xanthopholus*) are just such a case.
+Jackson’s chameleons (*Trioceros j.
+xanthopholus*) are just such a case (Fig.\@ref(fig:figure1) and fig.
+S1).
 In 1972, a shipment of chameleons was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
-The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island.
-On Oahu, the number of potential predators of chameleons is few (see methods), while in Kenya they are preyed upon by a wide range bird, snake and occasionally, mammal, predators.
-This accidental introduction represents a unique opportunity to study the evolution of a dynamic visual signal by comparing native (Kenya) and introduced (Oahu) chameleons.
+The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island (16).
+On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors), while in Kenya they are preyed upon by a wide range bird, snake and occasionally, mammal, predators.
+This accidental introduction represents a unique opportunity to study the evolution of a dynamic visual signal by comparing native (Kenya) and introduced (Oahu) chameleons under low and high levels of natural selection.
+Specifically, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal.
+We tested the following predictions that underpin this hypothesis: 1.) display colouration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; and 2.) display colouration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawaii than Kenya.
+We also asked whether colour signals were consistent with local adaptation.
+Specifically, we predicted that chameleon display colouration in each population would be more conspicuous within its own signaling environment compared to the other population’s signaling environment.
+For example, Hawaiian chameleons were historically found in Kenya but under local adaptation they are predicted to be more conspicuous in their current signaling environment, Hawaii.
+Likewise, Kenyan chameleons have no history in Hawaii and are therefore predicted to be more conspicuous within their current signaling environment, Kenya (table S1).
 
-# Results
+To test for character release and whether color signals show local adaptation, we quantified the local signalling environment (fig.
+S2) and manipulated the social and predatory context chameleons encountered, in both their native Kenyan, and introduced Hawaiian, range (fig.
+S3).
+We presented wild-caught male chameleons with the following treatments: a male, a female, a model bird predator (African cuckoo-hawk, *Aviceda cuculoides*) and a model snake predator (boomslang, *Dispholidus typus*) (Fig.\@ref(fig:figure1) and fig.
+S1,S4).
+We measured the spectral reflectance of males in dominant, subordinate, courtship, and antipredator (bird, snake) display states (Fig.\@ref(fig:figure2) and fig.
+S5).
+We then used visual modelling against a known chameleon visual system for sexual signals, and a bird and snake visual system for predation trials, to quantify chromatic and achromatic contrast against the backgrounds of each environment (supplementary text).
 
-We manipulated both the social and predatory context chameleons encountered, in both their native Kenyan range and their introduced Hawaiian range, to test three hypotheses about colour signal evolution.
-Using spectrophotometry and visual modelling that relied on our understanding of the known chameleon visual system we quntified chromatic and achromatic contrast against the backgrounds in each environment (i.e., leaves and stems in Kenya and Hawaii).
-
-Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals whereas greater conspicuousness of sexual signals should be selected in social contexts.
-In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; Courtship displays: Acromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was  true across all body regions measured (See Supplemental Materials).
-Kenyan chameleons from populations where predation is high were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, Courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see figure 1a & b).
+Under relaxed predation, sexual selection is expected to favour more conspicuous dynamic sexual signals in Hawaii than Kenya.
+In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig XX) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast ( Fig XX, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses).
+Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (acromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Figs XXXX).
 
 As a consequence of signal release, sexual signals are expected to be more easily detected by predators (i.e., snakes and birds) in Hawaii.
-Differential predation pressure resulting from snake and bird predation varying would also result in populations varying in their conspicuousness to different predators.
-In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c; Achromatic Contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`).
+Also, differential predation pressure resulting from differences in predation rates by snake and bird predators would also result in populations varying in their conspicuousness to each predator type.
+In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators with respect to brightness (FigXX; achromatic contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`).
 Differences in chromatic contrast between populations, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
-Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).
+Interestingly, Kenyan and Hawaiian chameleons were equally detectable by snakes, but Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (FIG XX; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).
 
 Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment.
-While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chameleons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
+While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis.
+Hawaiian chameleons were more conspicuous to other chameleons against heir own (Hawaiian) background as opposiued to the Kenyan background, in both social contexts and for both achromatic and chromatic contrast (FIG XX; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
 
-```{r Figure1, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves).
-A & B) Contrast of male chameleons during male-male contests and courtship.
+We have documented a surprisingly strong effect of environment on the evolution of a conspicuous, dynamic sexual signal.
+For a founding population of chameleons, the tropical environment of Hawaii represents a particularly favourable set of conditions with respect to habitat, temperature, and rainfall.
+Furthermore, Jackson’s chameleons are dietary generalists (17) and likely experience low competition for relatively abundant arthropod prey.
+But most importantly, Hawaii has no snakes and no raptors that feed on lizards.
+Domestic and feral cats would be their main threat on Oahu, and cats have very limited colour vision (18).
+In the absence of the usual suite of predators they would experience in their native rage of Kenya—a diverse community of snake and bird predators, the conditions were highly favourable for character release of sexual signals—particularly in a polygynous species in which males have intense male contest competition and elaborate courtship displays (Fig.\@ref(fig:figure1)).
+Not only do males use colorful displays, but they have three horns that they lock and twist into rivals during intense bouts that sometimes result in physical injury.
+Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated—an increasingly bright sexual signal used in both male contests and during courtship.
+
+Evolutionary biologists, particularly those interested in speciation, seldom consider the role of natural selection in shaping sexually selected signals even though sexual signals evolve in an ecological context (19) (but see (5, 20, 21).
+Landmark studies such as those on Trinidadian guppies (6, 9) have shown the important interaction between natural and sexual selection in static signals following experimental manipulation.
+Ours is one of the only studies that show that a dynamic sexual signal—rapid colour change—is likewise constrained by natural selection.
+Our results suggest that dynamic sexual signals are not simply plastic in a way that allows an individual to moderate the extent to which they ‘turn on’ a display but rather, that there is a threshold constrained by natural selection.
+This may in part be explained by the nature of colour change in chameleons.
+Chameleons have two layers of superposed iridiophores that facilitates color change that produces both camouflage and conspicuous sexual signals (22).
+Interestingly, while Hawaiian chameleons have evolved brighter display coloration, other elements of colour including hue and saturation have not experienced similar levels of selection except to become more conspicuous against their own background (local adaptation).
+This suggests that the two visual channels (achromatic vs chromatic) experience differing levels of selection which is perhaps not surprising given that they act independently of one another (23).
+While studying evolution is challenging because it is not possible to go back in time to witness key evolutionary events, we can take advantage of unique opportunities such as an invasion, to better understand how natural and sexual selection interact to shape animal signals.
+
+```{r figure1, fig.cap = "Chameleon colour signal change in response to different social and predatory stimuli.
+Male chameleons experience intense sexual selection.
+During the breeding season they change from a neutral state (green) to a highly conspicuous bright yellow display signal.
+They also readily fight by locking horns and sometimes pierce their rivals skin with their horns.
+(A) A typical dominant male in display coloration.
+(B) A subordinate male that lost a contest and turned from bright yellow to brown.
+(C) Two males fighting, both are in display colouration and at a stage where they are relatively even matched.
+(D) A male courting female in full display color while the female has turned to a contrasting colour, which suggests rejection of the male"}
+
+img1 <- readJPEG("./photos for figs/MS_images/Fig1a.jpg")
+img2 <- readJPEG("./photos for figs/MS_images/Fig1b.jpg")
+img3 <- readJPEG("./photos for figs/MS_images/Fig1c.jpg")
+img4 <- readJPEG("./photos for figs/MS_images/Fig1d.jpg")
+
+p1 <- cowplot::ggdraw()+
+        cowplot::draw_image(img1)
+p2 <- cowplot::ggdraw()+
+        cowplot::draw_image(img2)
+p3 <- cowplot::ggdraw()+
+        cowplot::draw_image(img3)
+p4 <- cowplot::ggdraw()+
+        cowplot::draw_image(img4)
+
+(p1 + p2) / (p3 + p4) + plot_annotation(tag_levels = c("A", "B", "C", "D"))
+
+```
+
+```{r figure2, fig.width = 14, fig.height= 14, fig.cap="Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya.
+The context for measurement was (A) male contest displays; (B) courtship; (C) bird predator; and (D) snake predator.
+More details in online supplementary material and Fig.
+S5"}
+png <- image_read("./output/figures/Figure3.png")
+png
+
+```
+
+```{r figure3, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves).
+(A & B) Contrast of male chameleons during male-male contests and courtship.
 JNDs are calculated based on the chameleon visual system.
-NS denotes non-significant contasts, whereas astericks indicates significant differences is < 0.001.
-Insets showing spectral reflectance curves of two representative body regions ('gular' and 'top-flank), along with average signalling background (averaged leaves) are provided for each social and predatory context.
-A) spectral curves of signals during male-male displays; B) spectral curves of signals during courtship displays; C) spectral curves of signals during bird encounters; and D) spectral curves of signals during snake encounters."}
+NS denotes non-significant contrasts, whereas asterisks indicates significant differences where p is less than 0.001.
+Insets showing spectral reflectance curves of two representative body regions ('gular' and 'top-flank'), along with average signalling background (averaged leaves) are provided for each social and predatory context.
+(A) spectral curves of signals during male-male displays; (B) spectral curves of signals during courtship displays; (C) spectral curves of signals during bird encounters; and (D) spectral curves of signals during snake encounters."}
 rerun = FALSE
@@
 
-```{r Figure2, fig.width = 14, fig.height= 9, fig.cap= "Local adaptation of chromatic and acromatic contrast for Hawaiian chameleons against Hawaiian and Kenyan backgrounds."}
+```{r figure4, fig.width = 14, fig.height= 9, fig.cap= "Evidence for local adaptation of color signals in both (A) chromatic- and (B) achromatic contrast for Hawaiian chameleons.
+For local adaptation, signals are predicted to be more conspicuous against their own background.
+In this case, Hawaiian chameleons were both significantly more chromatic and brighter (achromatic contrast)  against their own background compared to the Kenyan background suggesting that selection has acted on increasing local signal conspicuousness."}
 rerun=FALSE

### 2021-05-31 ms.Rmd
First draft of MS done, new refs, new figures and edit MS

@@
 title: "Rapid evolution of a conspicuous dynamic visual signal in an introduced chameleon"
-author: Whiting et al. 
-date: "`r Sys.Date()`"
-bibliography: refs.bib
+bibliography: ms_refs.bib
 csl: science.csl
@@
 
-Martin J. Whiting <sup>1,*<sup>, Brenden S. Holland <sup>2<sup>, J. Scott Keogh <sup>3<sup>, Daniel W.A.Noble <sup>3<sup>, Katrina J. Rankin <sup>4<sup> and Devi M. Stuart-Fox <sup>4<sup>
+Martin J. Whiting^1*^, Brenden S. Holland^2^, J. Scott Keogh^3^, Daniel W.A.Noble^3^, Katrina J. Rankin^4^ and Devi M. Stuart-Fox^4^
 
-1 Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
-2 Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
-3 Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
-4 Department of Zoology, The University of Melbourne, Melbourne, Victoria 3010, Australia
+^1^ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
+
+^2^ Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
+
+^3^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
+
+^4^ Department of Zoology, The University of Melbourne, Melbourne, Victoria 3010, Australia
 
@@
 
-When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods (1).
+When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods @RN1.
 We now have numerous examples of evolutionary change following natural selection in both the laboratory and wild, and not only in invertebrates with short generation times.
 Vertebrates, including Caribbean *Anolis* lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales (e.g.
-(2, 3)).
-Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in behaviour and phenotype over short time intervals (e.g., (4-6)).
+[@RN2; @RN3]).
+Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in behaviour and phenotype over short time intervals (e.g., [@RN4; @RN5; @RN6]).
 
@@
 These signals frequently become increasingly elaborate through a combination of male contest competition and female preference (e.g.
-runaway selection (7, 8)).
-However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly (9).
+runaway selection [@RN7; @RN8]).
+However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly @RN9.
 Evolution can respond to this selection pressure in a variety of ways that include shifting conspicuous signals to areas of the body that might be concealed to predators.
-For example, many conspicuous lizards have colourful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards (10).
-Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-rich streams (6, 9).
+For example, many conspicuous lizards have colourful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards @RN10.
+Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-rich streams [@RN6; @RN9].
 These examples are from animals with static colour signals.
-A parallel solution that reduces the risk of exposure to predators is dynamic colour change because signals can be switched on and off relatively rapidly through neural or endocrine control (11, 12).
+A parallel solution that reduces the risk of exposure to predators is dynamic colour change because signals can be switched on and off relatively rapidly through neural or endocrine control [@RN11; @RN12].
 A wide range of invertebrates (e.g.
 cephalopods) and vertebrates (e.g.
-lizards and frogs) have convergently evolved this ability (13-15).
+lizards and frogs) have convergently evolved this ability [@RN13; @RN14; @RN15].
 
@@
 In 1972, a shipment of chameleons was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
-The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island (16).
+The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island @RN16.
 On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors), while in Kenya they are preyed upon by a wide range bird, snake and occasionally, mammal, predators.
 This accidental introduction represents a unique opportunity to study the evolution of a dynamic visual signal by comparing native (Kenya) and introduced (Oahu) chameleons under low and high levels of natural selection.
+
 Specifically, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal.
@@
 Under relaxed predation, sexual selection is expected to favour more conspicuous dynamic sexual signals in Hawaii than Kenya.
-In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig XX) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast ( Fig XX, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses).
-Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (acromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Figs XXXX).
+In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig.\@ref(fig:figure2) &  Fig.\@ref(fig:figure3)A) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses).
+Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (acromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Fig.\@ref(fig:figure3)).
 
@@
 Also, differential predation pressure resulting from differences in predation rates by snake and bird predators would also result in populations varying in their conspicuousness to each predator type.
-In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators with respect to brightness (FigXX; achromatic contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`).
+In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators with respect to brightness (Fig.\@ref(fig:figure3)C; achromatic contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`).
 Differences in chromatic contrast between populations, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
-Interestingly, Kenyan and Hawaiian chameleons were equally detectable by snakes, but Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (FIG XX; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).
+Interestingly, Kenyan and Hawaiian chameleons were equally detectable by snakes, but Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).
 
@@
 While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis.
-Hawaiian chameleons were more conspicuous to other chameleons against heir own (Hawaiian) background as opposiued to the Kenyan background, in both social contexts and for both achromatic and chromatic contrast (FIG XX; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
+Hawaiian chameleons were more conspicuous to other chameleons against heir own (Hawaiian) background as opposed to the Kenyan background, in both social contexts and for both achromatic and chromatic contrast (Fig.\@ref(fig:figure4); Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
 
@@
 For a founding population of chameleons, the tropical environment of Hawaii represents a particularly favourable set of conditions with respect to habitat, temperature, and rainfall.
-Furthermore, Jackson’s chameleons are dietary generalists (17) and likely experience low competition for relatively abundant arthropod prey.
+Furthermore, Jackson’s chameleons are dietary generalists @RN17 and likely experience low competition for relatively abundant arthropod prey.
 But most importantly, Hawaii has no snakes and no raptors that feed on lizards.
-Domestic and feral cats would be their main threat on Oahu, and cats have very limited colour vision (18).
-In the absence of the usual suite of predators they would experience in their native rage of Kenya—a diverse community of snake and bird predators, the conditions were highly favourable for character release of sexual signals—particularly in a polygynous species in which males have intense male contest competition and elaborate courtship displays (Fig.\@ref(fig:figure1)).
+Domestic and feral cats would be their main threat on Oahu, and cats have very limited colour vision @RN18.
+In the absence of the usual suite of predators they would experience in their native range of Kenya—a diverse community of snake and bird predators, the conditions were highly favourable for character release of sexual signals—particularly in a polygynous species in which males have intense male contest competition and elaborate courtship displays (Fig.\@ref(fig:figure1)).
 Not only do males use colorful displays, but they have three horns that they lock and twist into rivals during intense bouts that sometimes result in physical injury.
@@
 
-Evolutionary biologists, particularly those interested in speciation, seldom consider the role of natural selection in shaping sexually selected signals even though sexual signals evolve in an ecological context (19) (but see (5, 20, 21).
-Landmark studies such as those on Trinidadian guppies (6, 9) have shown the important interaction between natural and sexual selection in static signals following experimental manipulation.
+Evolutionary biologists, particularly those interested in speciation, seldom consider the role of natural selection in shaping sexually selected signals even though sexual signals evolve in an ecological context @RN19 (but see [@RN5; @RN20, @RN21]).
+Landmark studies such as those on Trinidadian guppies [@RN6; @RN9]  have shown the important interaction between natural and sexual selection in static signals following experimental manipulation.
 Ours is one of the only studies that show that a dynamic sexual signal—rapid colour change—is likewise constrained by natural selection.
@@
 This may in part be explained by the nature of colour change in chameleons.
-Chameleons have two layers of superposed iridiophores that facilitates color change that produces both camouflage and conspicuous sexual signals (22).
+Chameleons have two layers of superposed iridiophores that facilitates color change that produces both camouflage and conspicuous sexual signals @RN22.
 Interestingly, while Hawaiian chameleons have evolved brighter display coloration, other elements of colour including hue and saturation have not experienced similar levels of selection except to become more conspicuous against their own background (local adaptation).
-This suggests that the two visual channels (achromatic vs chromatic) experience differing levels of selection which is perhaps not surprising given that they act independently of one another (23).
+This suggests that the two visual channels (achromatic vs chromatic) experience differing levels of selection which is perhaps not surprising given that they act independently of one another @RN23.
 While studying evolution is challenging because it is not possible to go back in time to witness key evolutionary events, we can take advantage of unique opportunities such as an invasion, to better understand how natural and sexual selection interact to shape animal signals.
 
-```{r figure1, fig.cap = "Chameleon colour signal change in response to different social and predatory stimuli.
+```{r figure1, cache=FALSE, fig.cap = "Chameleon color signal change in response to different social and predatory stimuli.
 Male chameleons experience intense sexual selection.
@@
 (D) A male courting female in full display color while the female has turned to a contrasting colour, which suggests rejection of the male"}
+library(magick)
 
-img1 <- readJPEG("./photos for figs/MS_images/Fig1a.jpg")
-img2 <- readJPEG("./photos for figs/MS_images/Fig1b.jpg")
-img3 <- readJPEG("./photos for figs/MS_images/Fig1c.jpg")
-img4 <- readJPEG("./photos for figs/MS_images/Fig1d.jpg")
-
-p1 <- cowplot::ggdraw()+
-        cowplot::draw_image(img1)
-p2 <- cowplot::ggdraw()+
-        cowplot::draw_image(img2)
-p3 <- cowplot::ggdraw()+
-        cowplot::draw_image(img3)
-p4 <- cowplot::ggdraw()+
-        cowplot::draw_image(img4)
-
-(p1 + p2) / (p3 + p4) + plot_annotation(tag_levels = c("A", "B", "C", "D"))
+img65 <- image_read("./output/figures/fig1.pdf")
+img65
 
@@
 S5"}
-png <- image_read("./output/figures/Figure3.png")
-png
+library(magick)
+img <- image_read_pdf("./output/figures/Figure3.pdf")
+img
 
@@
 
-```{r figure3, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves).
-(A & B) Contrast of male chameleons during male-male contests and courtship.
+```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves).
+(A & B) Acromatic and chromatic contrast of male chameleons during male-male contests and female courtship.
 JNDs are calculated based on the chameleon visual system.
-NS denotes non-significant contrasts, whereas asterisks indicates significant differences where p is less than 0.001.
-Insets showing spectral reflectance curves of two representative body regions ('gular' and 'top-flank'), along with average signalling background (averaged leaves) are provided for each social and predatory context.
-(A) spectral curves of signals during male-male displays; (B) spectral curves of signals during courtship displays; (C) spectral curves of signals during bird encounters; and (D) spectral curves of signals during snake encounters."}
+(B & C) Acromatic and chromatic contrasts of male chameleons during snake and bird predator encounters.
+JNDs are calculated based on the snake and bird visual systems.
+Contrasts between means, 'Beta', that are bold indicate significant effects."}
 rerun = FALSE
@@
 } else {
-  plot(image_read_pdf("./output/figures/Figure1.pdf"))
+
+  img2 <- image_read("./output/figures/Figure1.png")
+  img2
 }
@@
 
-```{r figure4, fig.width = 14, fig.height= 9, fig.cap= "Evidence for local adaptation of color signals in both (A) chromatic- and (B) achromatic contrast for Hawaiian chameleons.
+```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "Evidence for local adaptation of color signals in both (A) chromatic- and (B) achromatic contrast for Hawaiian chameleons.
 For local adaptation, signals are predicted to be more conspicuous against their own background.
@@
 } else {
-  plot(image_read_pdf("./output/figures/Figure2.pdf"))
+  img3 <- image_read("./output/figures/Figure2.png")
+  img3
 }
@@
 
-# Materials and Methods
-
-### *Statistical Analysis*
-
-Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package (@Bateslme4) in combination with the package `clubSandwich` (@Pustejovsky21).
-For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems).Deriving multiple contrast for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely independent of each other.
-To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Saitterwaite degrees of freedom correction.
-Across the different contexts (i.e., social contexts, predation and background environments), chromatic and acromatic contrasts were analysed in seperate models as response variables.
-For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects.
-For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or bird), population of origin and their interaction as fixed effects.
-Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii.
-Relevant model interactions were first tested using Wald tests.
-If found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
-We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Saitterwaite degrees of freedom correction.
-For our main analyses we averaged estimates (marginalised) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results).
-In all cases, our results and conclusions were not impacted (see supplmentary code).
-
-```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
-Standard errors are corrected by using robust variance estimators.
-"}
-
-table1 <- as.data.frame(rbind(robust_ConspicDisplayAC_CI,
-                robust_ConspicDisplayCC_CI,
-                robust_ConspicCourtshipAC_CI,
-                robust_ConspicCourtshipCC_CI))
-table1$Parameter <- rownames(table1)
-rownames(table1) <- NULL
-# Clean
- final_T1 <- table1 %>% 
-             mutate(SocialContext = c("Male-Male Display", rep("", 5), "Courtship Display", rep("", 5)),
-                    VisualRealm = c("Achromatic Contrast", rep("", 2), "Chromatic Contrast", rep("", 2), "Achromatic Contrast", rep("", 2), "Chromatic Contrast", rep("", 2))) %>%
-              select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
- write.csv(final_T1, file = "./output/tables/Table1_MS.csv", row.names = FALSE)
-
-```
-
-```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
-Full models are provided and standard errors are corrected by using robust variance estimators."}
-
-table2 <- as.data.frame(rbind(robust_PredatorAC_CI_inter,
-                              robust_PredatorCC_CI_inter))
-table2$Parameter <- rownames(table2)
-rownames(table2) <- NULL
-
-# Clean
- final_T2 <- table2 %>% 
-             mutate(JND = c("Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3))) %>%
-              select(JND, Parameter, beta, SE, df, CI_L, CI_U)
- write.csv(final_T2, file = "./output/tables/Table2_MS.csv", row.names = FALSE)
-
-```
-
-```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean acromatric and chormatic contrast in Hawaii and Kenya for both predators (snakes and birds).
-Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
-Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).
-We also applied a Bonferroni correction to p-value to control for multiple comparisons."}
-
-contrast <- rbind(CI_hyp_test,
-                 CI_hyp_testAC)
-
-contrast <- contrast %>% mutate(VisualSpec = c("Achromatic Contrast", "", "Chromatic Contrast", ""),
-                                p_val = sapply(p_val, function(x) p_value(x))) %>% 
-            select(VisualSpec, comparison, Estimate, lwr, upr, Fstat, df, p_val, p_bonf)
-
-write.csv(contrast, file = "./output/tables/Table3_MS.csv", row.names = FALSE)
-
-```
-
-```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for comparing acromatric and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (Background = Kenya) in both social contexts (Male-male competition and Courtship).
-Models were fit using robust variance estaimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.
-Significance parameter estimates are those where 95% CI does not include zero."}
-
-tb4 <- data.frame(rbind(robust_localAdapt_DispBgrdAC,
-             robust_DispBgrdCC,
-             robust_CourtshipAC,
-             robust_CourtshipCC))
-tb4$parameter <- rownames(tb4)
-rownames(tb4) <- NULL
-
-tb4 <- tb4 %>% mutate(SocialContext = c("Male-Male Display", rep("", 5), 
-                                        "Courtship Display", rep("", 5)),
-                    VisualRealm = c("Achromatic Contrast", rep("", 2), "Chromatic Contrast", 
-                                    rep("", 2), "Achromatic Contrast", rep("", 2), 
-                                    "Chromatic Contrast", rep("", 2))) %>%
-              select(SocialContext, VisualRealm, parameter, beta, SE, df, CI_L, CI_U)
-
-write.csv(tb4, file = "./output/tables/Table4_MS.csv", row.names = FALSE)
-
-```
-
 # References

### 2021-05-31 supp.Rmd
Ignore update, and small changes to Supp. Supp needs work now

@@
 
-```{r FigS1, fig.cap = "Examples of chameleon colour signal change in response to different social and predatory contexts"}
-
-files <- list.files("./photos/ESM/")
-img1 <- readJPEG(paste0("./photos/ESM/", files[1]))
-img2 <- readJPEG(paste0("./photos/ESM/", files[2]))
-img3 <- readJPEG(paste0("./photos/ESM/", files[3]))
-img4 <- readJPEG(paste0("./photos/ESM/", files[4]))
-img5 <- readJPEG(paste0("./photos/ESM/", files[5]))
-img6 <- readJPEG(paste0("./photos/ESM/", files[6]))
-
-p1 <- cowplot::ggdraw()+
-        cowplot::draw_image(img1)
-p2 <- cowplot::ggdraw()+
-        cowplot::draw_image(img2)
-p3 <- cowplot::ggdraw()+
-        cowplot::draw_image(img3)
-p4 <- cowplot::ggdraw()+
-        cowplot::draw_image(img4)
-p5 <- cowplot::ggdraw()+
-        cowplot::draw_image(img5)
-p6 <- cowplot::ggdraw()+
-        cowplot::draw_image(img6)
-
-(p1 + p3 + p6) / (p2 + p5 + p4) + plot_annotation(tag_levels = c("A", "B", "C", "D", "E", "F"))
-
-```
-
 ## Colour measurements, habitat background and irradiance
@@
 
+# Supplmental Methods
+
+### *Statistical Analysis*
+
+Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package (@Bateslme4) in combination with the package `clubSandwich` (@Pustejovsky21).
+For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems).Deriving multiple contrast for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely independent of each other.
+To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Saitterwaite degrees of freedom correction.
+Across the different contexts (i.e., social contexts, predation and background environments), chromatic and acromatic contrasts were analysed in seperate models as response variables.
+For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects.
+For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or bird), population of origin and their interaction as fixed effects.
+Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii.
+Relevant model interactions were first tested using Wald tests.
+If found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
+We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Saitterwaite degrees of freedom correction.
+For our main analyses we averaged estimates (marginalised) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results).
+In all cases, our results and conclusions were not impacted (see supplmentary code).
+
 # Supplemental Results
-
-## Supporting Supplementay Tables
-
-```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
-Standard errors are corrected by using robust variance estimators.
-"}
-
-final_T1 <-  read.csv(file = "./output/tables/Table1_MS.csv")
-
- # Table
-table1 <- flextable(final_T1) %>% table_style6()
-
-table1
-```
-
-```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
-Full models are provided and standard errors are corrected by using robust variance estimators."}
-
- final_T2 <- read.csv(file = "./output/tables/Table2_MS.csv")
- # Table
-table2 <- flextable(final_T2) %>%
-           table_style7()
-
-table2
-```
-
-```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean acromatric and chormatic contrast in Hawaii and Kenya for both predators (snakes and birds).
-Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
-Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).
-We also applied a Bonferroni correction to p-value to control for multiple comparisons."}
-
-contrast <- read.csv(file = "./output/tables/Table3_MS.csv")
-
-table3 <- flextable(contrast) %>% table_style8()
-table3
-```
-
-```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for comparing acromatric and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (Background = Kenya) in both social contexts (Male-male competition and Courtship).
-Models were fit using robust variance estaimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.
-Significance parameter estimates are those where 95% CI does not include zero."}
-
-tb4 <- read.csv(file = "./output/tables/Table4_MS.csv")
-table4 <- flextable(tb4) %>% table_style9()
-table4
-```
-
-## Assessing changes in sexual signals across different body regions
-We took measurements of spectral reflectance for different body regions of each chamaleon.
-In our primary analysis, we averaged across different body regions because we were interested only in whether visual signals overall were different between Kenya and Hawaii.
-This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS2) - \@ref(fig:FigS4) and Tables \@ref(tab:tableS1)- \@ref(tab:tableS5)).
-
-#### Body regions in social contexts (Male-male and Courtship Display) show different magnitude of responses, but similar directionality
-
-Overall, different body regions explained significant variation in achromatic and chromatic contrast in both social contexts, whereas differences between Kenya and Hawaii only existed in achormatic contrast (Figure \@ref(fig:FigS2) & Table \@ref(tab:tableS1)).
-A significant interaction between body region and population was found for achormatic contrast when males were displaying to each other (Table \@ref(tab:tableS1)).
-However, in all cases Kenya showed a strong reduction in acromatic contrast relative to Hawaii lizards, despite some variation in the magnitude of effect (Figure \@ref(fig:FigS2)B & D; Table \@ref(tab:tableS2)).
-
-```{r FigS2, fig.width = 13, fig.height = 12, fig.cap = "Chormatic (A, C) and Achromatic (B, D) contrast in just noticible differences (JND) for Hawaiian and Kenyan lizards relative to their environments.
-All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
-Mean JNDs are provided along with 95% confidence intervals.
-A & B represent male-male contest displays whereas C & D represent courtship displays"}
-
-#Making a point plot with error bars split by body size and population 
-#Compute raw means and standard errors for body part and population for dS and assigned this data into Display_plot_dS_dat
-Display_plot_dS_dat <- Display_1 %>% #data for dS Chromatic constrast
-  group_by(Population, BodyRegion) %>% 
-  summarise( Mean = mean(dS, na.rm = T), 
-               SE = sd(dS)/sqrt(length(unique(ID))-1),
-            Upper = Mean + 1.96*SE,
-            Lower = Mean - 1.96*SE)
-
-Display_plot_dL_dat <- Display_1 %>% #data for dL Achromatic contrast
-  group_by(Population, BodyRegion) %>% 
-  summarise(Mean = mean(dL, na.rm = T), 
-            SE = sd(dL)/sqrt(length(unique(ID))-1),
-            Upper = Mean + 1.96*SE,
-            Lower = Mean - 1.96*SE)
-
-#Plotting Chromatic contrast (dS)
-p1 <- ggplot(Display_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.2)) +
-  geom_line(position = position_dodge(width=0.2)) +
-  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("") + 
-  ylab("Chromatic contrast (JND)") +
-  theme_bw() +
-  theme(legend.position = "none",
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-#Plotting Achromatic contrast (dL)
-p2 <- ggplot(Display_plot_dL_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.2)) +
-  geom_line(position = position_dodge(width=0.2)) +
-  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("") + 
-  ylab("Achromatic contrast (JND)") +
-  theme_bw() +
-  labs(color = "Body Region") + 
-  theme(legend.position = c(0.85, 0.85),
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-#Making a point plot with error bars split by body size and population 
-#Compute raw means and standard errors for body part and population for dS and assigned this data into Display_plot_dS_dat
-Courtship_plot_dS_dat <- Courtship %>% #data for dS Chromatic constrast
-  group_by(Population, BodyRegion) %>% 
-  summarise(Mean = mean(dS, na.rm = T), 
-            SE = sd(dS)/sqrt(length(unique(ID))),
-            Upper = Mean + 1.96*SE,
-            Lower = Mean - 1.96*SE)
-
-Courtship_plot_dL_dat <- Courtship %>% #data for dL Achromatic contrast
-  group_by(Population, BodyRegion) %>% 
-  summarise(Mean = mean(dL, na.rm = T), 
-            SE = sd(dL)/sqrt(length(unique(ID))),
-            Upper = Mean + 1.96*SE,
-            Lower = Mean - 1.96*SE)
-
-#Plotting Chromatic contrast (dS)
-p3 <- ggplot(Courtship_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.2)) +
-  geom_line(position = position_dodge(width=0.2)) +
-  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("") + 
-  ylab("Chromatic contrast (JND)") +
-  theme_bw() +
-  theme(legend.position = "none",
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-#Plotting Achromatic contrast (dL)
-p4 <- ggplot(Courtship_plot_dL_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.2)) +
-  geom_line(position = position_dodge(width=0.2)) +
-  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("") + 
-  ylab("Achromatic contrast (JND)") +
-  theme_bw() +
-  theme(legend.position = c(0.85, 0.85),
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-p1 + p2 + p3 + p4 + plot_annotation(tag_levels = c('A', 'B', 'C', 'D')) & theme(plot.tag = element_text(face = 'bold', size=30))
-
-```
-
-```{r tableS1, tab.cap = "Wald F tests comparing overall significance of model parameters.
-Wald F tests were computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
-Interactions were first tested and models refit to test the signifiacnce of main effects"}
-fullTable <- rbind(wald_table_ConspicDisplayAC, 
-                   wald_table_ConspicDisplayCC, 
-                   wald_table_ConspicCourtshipAC,
-                   wald_table_ConspicCourtshipCC)
-
-fullTable <- fullTable %>% 
-              mutate(Context = c("Male-Male Display", rep("", 7), "Courtship Display", rep("", 7)),
-                     JND = c("Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3), "Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3)),
-                     p_val = sapply(p_val, function(x) p_value(x))) %>% select(Context, JND, Variable, Fstat, df_num, df_denom, p_val)
-
-S3 <- flextable(fullTable) %>%
-        table_style2(.)
-S3
-```
-
-```{r tableS2, tab.cap = "Linear mixed effects model results using robust variance estimators and Saitterwaite degrees of freedom correction.Social Context"}
-robust_sc_table <- rbind(robust_ConspicDisplayAC,
-                          robust_ConspicDisplayCC,
-                          robust_ConspicCourtshipAC,
-                          robust_ConspicCourtshipCC) %>% 
-                    mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)),
-                           SocialContext = c("Male-Male Display", rep("", 17), 
-                                             "Courtship Display", rep("", 17)),
-                           JND = c("Achromatic Contrast", rep("", 8), 
-                                   "Chromatic Contrast", rep("", 8), 
-                                   "Achromatic Contrast", rep("", 8), 
-                                   "Chromatic Contrast", rep("", 8)))
-
-robust_sc_table$parameter <- rownames(robust_sc_table)
-
-robust_sc_table <- robust_sc_table %>% select(SocialContext, JND, parameter, beta, SE, tstat, df, p_Satt)
-
-tableS5 <- flextable(robust_sc_table) %>%  table_style4()
-tableS5
-```
 

### 2021-06-01 R/func.R
Problems with flextable, and I belive it is because purr has the same function called "compose", so being explicit now on what one we are using.

@@ -80,27 +80,27 @@ table_style  <- function(table){
   table <- table  %>% 
     align(align = "center", part = "header") %>% 
     align(align = "center", part = "body") %>% 
-    compose(part = "header", j = 1, value = as_paragraph(as_b("Predator"))) %>% 
-    compose(part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>% 
-    compose(part = "header", j = 3, value = as_paragraph(as_b("F"))) %>%
-    compose(part = "header", j = 4, value = as_paragraph(as_b("DF"))) %>% 
-    compose(part = "header", j = 5, value = as_paragraph(as_b("p"))) %>% 
-    compose(part = "header", j = 6, value = as_paragraph(as_b("p"), as_sub(as_b(" Bonferroni")))) %>%
-    compose(part = "body", i = 1, j = 2, 
+    flextable::compose(part = "header", j = 1, value = as_paragraph(as_b("Predator"))) %>% 
+    flextable::compose(part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>% 
+    flextable::compose(part = "header", j = 3, value = as_paragraph(as_b("F"))) %>%
+    flextable::compose(part = "header", j = 4, value = as_paragraph(as_b("DF"))) %>% 
+    flextable::compose(part = "header", j = 5, value = as_paragraph(as_b("p"))) %>% 
+    flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("p"), as_sub(as_b(" Bonferroni")))) %>%
+    flextable::compose(part = "body", i = 1, j = 2, 
             value = as_paragraph("Kenya", as_sub(" gular"), " - ", "Hawaii", as_sub(" gular")))  %>% 
-    compose(part = "body", i = 2, j = 2, 
+    flextable::compose(part = "body", i = 2, j = 2, 
             value = as_paragraph("Kenya", as_sub(" midflank"), " - ", "Hawaii", as_sub(" midflank"))) %>% 
-    compose(part = "body", i = 3, j = 2, 
+    flextable::compose(part = "body", i = 3, j = 2, 
             value = as_paragraph("Kenya", as_sub(" tailbase"), " - ", "Hawaii", as_sub(" tailbase"))) %>% 
-    compose(part = "body", i = 4, j = 2, 
+    flextable::compose(part = "body", i = 4, j = 2, 
[…]

### 2021-06-01 R/ms.R
Found small label mixup in code for table 3. Fixed now

@@ -767,10 +767,10 @@ table2
 
 ## ----table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean acromatric and chormatic contrast in Hawaii and Kenya for both predators (snakes and birds). Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper). Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).  We also applied a Bonferroni correction to p-value to control for multiple comparisons."----
 
-contrast <- rbind(CI_hyp_test,
-                 CI_hyp_testAC)
+contrast <- rbind(CI_hyp_testAC,
+                  CI_hyp_test)
 
-contrast <- contrast %>% mutate(VisualSpec = c("Achromatic Contrast", "", "Chromatic Contrast", ""),
+contrast <- contrast %>% mutate(VisualSpec = c("Achromatic Contrast",  "", "Chromatic Contrast", ""),
                                 p_val = sapply(p_val, function(x) p_value(x))) %>% 
             select(VisualSpec, comparison, Estimate, lwr, upr, Fstat, df, p_val, p_bonf)
 

### 2021-06-01 R/func.R
Update to functions. Jusy small reorganisation of the width

@@ -230,7 +230,8 @@ table_style7 <- function(table){
     flextable::compose(part = "body", j = 2, i = c(1,5)+2, value = as_paragraph("Population", as_sub(" Kenya"))) %>%
     flextable::compose(part = "body", j = 2, i = c(1,5)+3, value = as_paragraph("Predator", as_sub(" Snake"), "* Population", as_sub(" Kenya"))) %>% 
     font(fontname="Times", part = "body") %>% 
-    font(fontname="Times", part = "header")
+    font(fontname="Times", part = "header") %>% 
+    width(j = 2, width = 4)
     return(table)
 }
 
@@ -300,7 +301,7 @@ table_style10 <- function(table){
 }
 
 table_style11  <- function(table){
-  table <- table %>% width(j = 1, width = 1) %>% width(j = 2, width = 4) %>% 
+  table <- table %>% 
     align(align = "center", part = "header") %>% 
     align(align = "center", part = "body") %>% 
     flextable::compose(part = "header", j = 1, value = as_paragraph(as_b("Context"))) %>% 
@@ -318,7 +319,8 @@ table_style11  <- function(table){
     flextable::compose(part = "body", i = c(4,8), j = 2, 
             value = as_paragraph("Kenyan Background", as_sub(" topflank"), " - ", "Hawaiian Background", as_sub(" topflank"))) %>% 
     font(fontname="Times", part = "body") %>% 
-    font(fontname="Times", part = "header")
+    font(fontname="Times", part = "header") %>% 
+    width(j = 1, width = 1) %>% width(j = 2, width = 8)
   
   return(table)
 }

### 2021-06-01 supp.Rmd
Finally osrted problems with word rendering. Not sure why by func.R was being cleared. This was not happening with html, but added source() to func.R back into at first table and it worked fine

@@
 date: "`r Sys.Date()`"
-bibliography: refs.bib
+bibliography: supp_refs.bib
 csl: science.csl
@@
     bibliography: refs.bib
-editor_options: 
-  chunk_output_type: console
 ---
 ```{r setup, include=FALSE, purl = FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
 ## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
+pacman::p_load(flextable, lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich)
 source("./R/func.R")
@@
 
-# Materials and Methods
+<!-- \renewcommand{\thefigure}{S\arabic{figure}} -->
+<!-- \setcounter{figure}{0} -->
+<!-- \renewcommand{\thetable}{S\arabic{table}} -->
+<!-- \setcounter{table}{0} -->
+<!-- \renewcommand{\theequation}{S\arabic{equation}} -->
+<!-- \setcounter{equation}{0} -->
 
+# Expanded Hypotheses and Predictions
+
+<!-- ```{r tablehyp_prd, tab.align = 'center', tab.cap = "Hypotheses and associated predictions tested in this paper in relation to the population of interest, the receiver visual system used in visual modelling, and the background against which the contrast of the signal was calculated for chromatic and achromatic contrast "} -->
+<!--  hyp <-  read.csv(file = "./output/tables/Table_hyp_pred.csv") -->
+
+<!--   # Table -->
+<!--  hyp2 <- flextable(hyp) %>% table_style12() -->
+
+<!--  hyp2 -->
+<!-- ``` -->
+
+# Supplemental Materials and Methods
 ## Study area and collection of chameleons
 
-We collected most chameleons at night by spotlighting, and a small proportion of individuals during the day.
+We collected most chameleons (Fig.
+\@ref(fig:Figcolsig)) at night by spotlighting, and a small proportion of individuals during the day.
 When chameleons were beyond reach, we used 4-m extensible poles that chameleons could be encouraged to step on to, before lowering them to the ground.
 We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
-Field work was conducted in Hawai’i during January-February 2006 and in Kenya during 5-13 April 2006.
-In Hawai’i, chameleons were collected from a single population in the Manoa Valley, Oahu, accessible along Tantalus Drive at an elevation of ca.
-x-x.
-The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth.
+Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006.
+In Hawaii, chameleons were collected from a single population in the Manoa Valley, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
+400-600m.
+The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth (Fig.
+\@ref(fig:Figcolsig2)A,B).
 Most of the plants in this area were nonnative, and chameleons also occurred in hedge-grows and stands of bamboo.
-The area experiences high annual rainfall (mean: xx).
+The area experiences high annual rainfall (mean: ca.
+1500 mm) @RN1 and the mean annual temperature is 22-24&deg;C @RN2.
 In Kenya, field work was conducted near the foothills of Mt.
@@
 The area is partially cleared for agriculture, leaving small stands of forest.
-The vegetation was a mix of native trees and shrubs interspersed with exotic species.
+The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
+\@ref(fig:Figcolsig2)C,D).
+Mean annual temperature is 20.2&deg;C (mean max: 24.8&deg;C) and the mean annual rainfall is 121 mm @RN3.
 
-## Behaviour trials
+<!-- ```{r Figcolsig, fig.align = 'center', fig.cap = "Additional examples of chameleon color signals and behavior in response to different social contexts.
+(A) Two different males in conspicuous display color adopted at the start of a contest and retained by the dominant male.
+(B) Two examples of males engaged in horn locking.
+Males in the bottom plane are evenly matched at this stage and both are in full display colour.
+In the top image, the male on the left is starting to change to a subordinate color.
+(C) Conspicuous lemon-yellow male courtship color.
+In this case, the female is using a contrasting stress color consistent with rejection."} -->
+<!-- figs1 <- image_read("./output/figures/FigS1.png") -->
+<!-- figs1 -->
 
-In order to quantify display colouration we staged encounters between conspecifics (male-male and male-female) and model (bird and snake) predators and measured their subsequent display colouration (details below).
+<!-- ``` -->
+
+<!-- ```{r Figcolsig2, fig.align = 'center', fig.cap = "Representative photos of habitat in Hawaii (A, B) and Kenya (C, D).
+In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines."} -->
+<!-- figs2 <- image_read("./output/figures/FigS2.png") -->
+<!-- figs2 -->
+
+<!-- ``` -->
+
+## Behavioral trials
+
+In order to quantify display coloration we staged encounters between conspecifics (male-male and male-female) and model (bird and snake) predators and measured their subsequent display coloration (details below).
 All chameleons were used in both predator and social trials although in some cases we were unable to quantify all social signals for a particular individual.
 We conducted all conspecific-social trials first, in case anti-predator trials unduly influenced their social interactions.
-Behavioural trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
+Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
 We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand.
 This structure was about 1.5 m high and each arm was approximately 60 cm.
-The triangular setup allowed chameleons to display at a comfortable distance and also to escape from aggressive interactions, particularly during male-male trials.
+The triangular setup allowed chameleons to display at a comfortable distance and also to escape from aggressive interactions, particularly during male-male trials (Fig.
+\@ref(fig:Figcolsig4)).
 All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
@@
 
+<!-- ```{r Figcolsig3, fig.align = 'center', fig.cap = "The experimental setup for behavioral trials.
+We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male."} -->
+<!-- figs3 <- image_read("./output/figures/FigS3.png") -->
+<!-- figs3 -->
+
+<!-- ``` -->
+
 We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
-This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing colour and behaviour.
-In male-male trials both individuals would adopt a lemon-yellow display coloration (Fig.
-x) and advance towards each other.
-This would be accompanied by head shakes and would lead to horn-locking if males were not obviously asymmetric.
-Once dominance was settled, the subordinate would turn brown and flee.
-In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking.
-The day following social trials we presented each chameleon with a snake and bird model predator (separately).
+This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing colour and behaviour (Fig.
+\@ref(fig:Figcolsig)A,B).
+In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
+This would be accompanied by head shakes and would lead to horn-locking (Fig.
+\@ref(fig:Figcolsig)B) if males were not obviously asymmetric.
+Once dominance was settled, the subordinate would turn brown (Fig.
+\@ref(fig:Figcolsig)B) and flee.
+In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
+\@ref(fig:Figcolsig)C).
+The day following social trials we presented each chameleon with a snake (Fig.
+\@ref(fig:Figcolsig4)A) and bird (Fig.
+\@ref(fig:Figcolsig4)B) model predator (separately).
 Lizards were drawn at random and the order of presentation (snake or bird first) was random.
@@
 A trial began when a chameleon was removed from its cloth bag and placed on the triangular stand.
-At this point the lizard was in a neutral colour state.
-In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon.
-In the case of the bird trials, the bird was attached to a x cm wooden pole and was ‘flown’ past the chameleon from above and below it.
-Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behaviour by chameleons.
-Once chameleons changed colour, we quantified their spectral reflectance following the same procedure described for social trials.
+At this point the lizard was in a neutral color state.
+In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Fig.
+\@ref(fig:Figcolsig4)A).
+In the case of the bird trials, the bird was attached to an ca.
+2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Fig.
+\@ref(fig:Figcolsig4)B).
+Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
+Once chameleons changed color, we quantified their spectral reflectance following the same procedure described for social trials.
 We also conducted seven trials with a just a branch with no predator attached, as a control for the model predator.
-The snake model was moulded from a dead boomslang (Dispholidus typus) and painted by a professional artist to resemble a typical green form of this species (previously used in Stuart-Fox…).
-The bird was ….
+The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN6;@RN7]).
+The bird was a mounted African cuckoo-hawk (*Aviceda cuculoides*).
 
-## Colour measurements, habitat background and irradiance
+<!-- ```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator.
+(A) A model snake (Dispholidus typus) causing a characteristic contrasting color change and (B) a model bird (Aviceda cuculoides) which was flown over the top of the chameleon."} -->
+<!-- figs4 <- image_read("./output/figures/FigS4.png") -->
+<!-- figs4 -->
 
-During all behaviour trials, we measured the colour of each body region (Fig.
-x) once only, because chameleons change colour rapidly and we had to reduce the handling stress to the animal.
+<!-- ``` -->
+
+## Color measurements, habitat background and irradiance
+
+During all behavior trials, we measured the color of each body region (Fig.
+\@ref(fig:Figcolsig5)) once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
 Patches were measured in random sequence to avoid any bias of order.
-When a chameleon had completely changed colour (to a human observer) and was also performing typical behaviour associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 optic spectrophotometer and a PX2 light source.
-We used a custom-made probe-holder which we placed in contact with the animal at a standard angle (45; @RN2) and distance (5 mm), thereby excluding ambient light.
+When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 optic spectrophotometer and a PX2 light source.
+We used a custom-made probe-holder which we placed in contact with the animal at a standard angle (45&deg; @RN2) and distance (5 mm), thereby excluding ambient light.
 All measurements were relative to a dark, and a certified 99% white reflectance standard.
-When chameleons started to change colour in response to our handling we stopped measuring colour and allowed them to continue a social trial before continuing to measure spectral reflectance.
+When chameleons started to change color in response to our handling we stopped measuring color and allowed them to continue a social trial before continuing to measure spectral reflectance.
 In the case of a predator trial we would once again present them with the predator model until they returned to their anti-predator display.
-This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behaviour following handling.
+This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behavior following handling.
 
@@
 
+```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "Location of body regions we measured for spectral reflectance. See details in text."}
+figs5 <- image_read("./output/figures/FigS5.jpg")
+figs5
+
+```
+
 ## Visual modelling
 
-# Supplmental Methods
+## Statistical Analysis
 
-### *Statistical Analysis*
-
-Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package (@Bateslme4) in combination with the package `clubSandwich` (@Pustejovsky21).
-For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems).Deriving multiple contrast for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely independent of each other.
-To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Saitterwaite degrees of freedom correction.
-Across the different contexts (i.e., social contexts, predation and background environments), chromatic and acromatic contrasts were analysed in seperate models as response variables.
+Chromatic and achromatic contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN4 in combination with the package `clubSandwich` @RN5.
+For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems).
+While we present sensitivity analyses for all body regions (see below - *Additional Supplemental Results*), we were mainly interested in overall signal differences displayed by chameleons.
+As such, for our main analyses we averaged estimates (marginalized) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see *Additional Supplemental Results*).
+In all cases, our results and conclusions were not impacted (see supplementary code).
+Deriving multiple contrast for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other.
+To ensure that non-independence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction.
+Across the different contexts (i.e., social contexts, predation, and background environments), chromatic and achromatic contrasts were analyzed in separate models as response variables.
 For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects.
 For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or bird), population of origin and their interaction as fixed effects.
-Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii.
+Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were observed on a native Kenyan background or their current introduced habitat background in Hawaii.
 Relevant model interactions were first tested using Wald tests.
 If found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
-We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Saitterwaite degrees of freedom correction.
-For our main analyses we averaged estimates (marginalised) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results).
-In all cases, our results and conclusions were not impacted (see supplmentary code).
+We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.
 
 # Supplemental Results
+## Supporting Supplementary Tables for Main Text
 
-#### Body regions are perceived differently by predators but exhibited similar effect sizes
+```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing achromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
+Standard errors are corrected by using robust variance estimators.
+"}
+source("./R/func.R")
+final_T1 <-  read.csv(file = "./output/tables/Table1_MS.csv")
 
-Chromatic contrast of colour signals against the background varied by predator type, population and body region (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
-Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small diferences among populations in how snakes perceived colour signals (Figure \@ref(fig:FigS3)A).
+  # Table
+ table1 <- flextable(final_T1) %>%  table_style6() %>% fit_to_width(max_width = 7)
+ table1
+
+```
+
+```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing achromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
+Full models are provided and standard errors are corrected by using robust variance estimators."}
+  final_T2 <- read.csv(file = "./output/tables/Table2_MS.csv")
+  # Table
+ table2 <- flextable(final_T2) %>%
+            table_style7(.) %>% fit_to_width(max_width = 7)
+
+ table2
+```
+
+```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean achromatic and chormatic contrast in Hawaii and Kenya for both predators (snakes and birds).
+Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
+Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).
+We also applied a Bonferroni correction to p-value to control for multiple comparisons."}
+ contrast <- read.csv(file = "./output/tables/Table3_MS.csv")
+ table3 <- flextable(contrast) %>% table_style8() %>% fit_to_width(max_width = 7)
+ table3
+```
+
+```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for comparing achromatic and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (Background = Kenya) in both social contexts (Male-male competition and Courtship).
+Models were fit using robust variance estaimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.
+Significance parameter estimates are those where 95% CI does not include zero."}
+ tb4 <- read.csv(file = "./output/tables/Table4_MS.csv")
+ table4 <- flextable(tb4) %>% table_style9() %>% fit_to_width(max_width = 7)
+ table4
+```
+
+# Additional Supplemental Results
+## Body regions are perceived differently by predators but exhibited similar effect sizes
+
+Chromatic contrast of color signals against the background varied by predator type, population and body region (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
+Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Figure \@ref(fig:FigS3)A).
 There were no significant interactions between predator type ('bird' vs 'snake') and either body region or population in achromatic contrast against the background, but a significant population and body region effect (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
@@
 Nonetheless, effect sizes across body regions were in the same direction (Figure \@ref(fig:FigS3)B & Table \@ref(tab:tableS4)).
-Overall, there were significant differences between Kenya and Hawaii in achormatic contrast of the gular and tailbase regions, but not for mid- and top-flank (Table \@ref(tab:tableS5)).
-In contrast, no sigificant differences between Kenya and Hawaii existed across any body region, whereas significant differences existed between gular, midflank and tailbase (Table \@ref(tab:tableS6)).
+Overall, there were significant differences between Kenya and Hawaii in achromatic contrast of the gular and tailbase regions, but not for mid- and top-flank (Table \@ref(tab:tableS5)).
+In contrast, no significant differences between Kenya and Hawaii existed across any body region, whereas significant differences existed between gular, midflank and tailbase (Table \@ref(tab:tableS6)).
 
-```{r FigS3, fig.width = 12, fig.height = 12, fig.cap = "All four body regions (gular, mid-flank, tailbase and topflank) and their achromatic (B) and chromatic contrast (A) to Hawaiian and Kenyan backgrounds (each population against their own background) are shown averaged across stems and leaves found in their habitat.
+```{r FigS3, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "All four body regions (gular, mid-flank, tailbase and topflank) and their achromatic (B) and chromatic contrast (A) to Hawaiian and Kenyan backgrounds (each population against their own background) are shown averaged across stems and leaves found in their habitat.
 Just noticable differences (JNDs) are calculated for both bird and snake predators.
@@
 S4 <- flextable(fulltableP) %>%
-        table_style3(.)
+        table_style3(.) %>% fit_to_width(max_width = 7)
 S4
@@
 
-tableS6 <- flextable(robust_sc_pred_table) %>%  table_style5()
+tableS6 <- flextable(robust_sc_pred_table) %>%  table_style5() %>% fit_to_width(max_width = 7)
 tableS6
@@
 ft5 <- flextable(final_r5) %>%
-        table_style(.)
+        table_style(.) %>% fit_to_width(max_width = 7)
 
@@
  ft6 <- flextable(final_r6) %>%
-        table_style(.)
+        table_style(.) %>% fit_to_width(max_width = 7)
 
@@
 
-#### All body regions exibit similar contrasts between Kenyan and Hawaiian backgrounds suggesting local adaptation of entire visual signal
+## All body regions exhibit similar contrasts between Kenyan and Hawaiian backgrounds suggesting local adaptation of entire visual signal
 
 We tested whether different body regions of Hawaiian chameleons exhibited different contrasts between Hawaiian habitat background compared to their ancestral, Kenyan background.
-We started off testing models comparing the interaction between background enviroment and body region for both chormatic and achromatic contrast.
+We started off testing models comparing the interaction between background environment and body region for both chromatic and achromatic contrast.
 However, for achromatic contrast, these models had poor convergence.
 Based on the raw data (Figure \@ref(fig:FigS4)B & D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
-As such, we simplied models and only present the main effects of body region and background environment for achromatic contrast.
+As such, we simplified models and only present the main effects of body region and background environment for achromatic contrast.
 However, inspection of the raw data (Figure \@ref(fig:FigS4)A & C) suggested that body regions may differ in how chromatically contrasting they are against the background environment of Hawaii and Kenya.
-In addition, our main results suggested that predators may differentially drive chromatic signal conspicuousness, and given differet body regions maybe more exposed to bird compared to snake predators, we fit an interaction and main effects model for chromatic contrast.
+In addition, our main results suggested that predators may deferentially drive chromatic signal conspicuousness, and given different body regions maybe more exposed to bird compared to snake predators, we fit an interaction and main effects model for chromatic contrast.
 
-The magnitude of chromatic contrast differences between Hawaiian chamaleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:FigS4)A & C & Table \@ref(tab:tableS7)).
+The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:FigS4)A & C & Table \@ref(tab:tableS7)).
 However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:FigS4)A & C), and only some body regions showed a significant decrease (Table \@ref(tab:tableS8)).
@@
 
-```{r FigS4, fig.width = 13, fig.height = 12, fig.cap = "Local adaptation.
+```{r FigS4, fig.align = 'center', fig.width = 13, fig.height = 12, fig.cap = "Local adaptation.
 All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
@@
 
-Table7 <- flextable(table7) %>% table_style10()
+Table7 <- flextable(table7) %>% table_style10() %>% fit_to_width(max_width = 7)
 Table7
@@
 
-Table8 <- flextable(tab8) %>% table_style11()
+Table8 <- flextable(tab8) %>% table_style11() %>% fit_to_width(max_width = 7)
 Table8

### 2021-06-01 ms.Rmd
Small fixes

@@
 In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig.\@ref(fig:figure2) &  Fig.\@ref(fig:figure3)A) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses).
-Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (acromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Fig.\@ref(fig:figure3)).
+Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (achromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Fig.\@ref(fig:figure3)).
 

### 2021-06-01 supp.Rmd
Supp now ready, just need Fig4

@@
 
-<!-- \renewcommand{\thefigure}{S\arabic{figure}} -->
-<!-- \setcounter{figure}{0} -->
-<!-- \renewcommand{\thetable}{S\arabic{table}} -->
-<!-- \setcounter{table}{0} -->
-<!-- \renewcommand{\theequation}{S\arabic{equation}} -->
-<!-- \setcounter{equation}{0} -->
-
 # Expanded Hypotheses and Predictions
 
-<!-- ```{r tablehyp_prd, tab.align = 'center', tab.cap = "Hypotheses and associated predictions tested in this paper in relation to the population of interest, the receiver visual system used in visual modelling, and the background against which the contrast of the signal was calculated for chromatic and achromatic contrast "} -->
-<!--  hyp <-  read.csv(file = "./output/tables/Table_hyp_pred.csv") -->
+We provided an expanded table of hypotheses and associated predictions. We also identify the data needed to test these predictions.
 
-<!--   # Table -->
-<!--  hyp2 <- flextable(hyp) %>% table_style12() -->
+```{r tablehyp, tab.cap = "Hypotheses and associated predictions tested in this paper in relation to the population of interest, the receiver visual system used in visual modelling, and the background against which the contrast of the signal was calculated for chromatic and achromatic contrast."}
+  source("./R/func.R") 
+  hyp <-  read.csv(file = "./output/tables/Table_hyp_pred.csv")
 
-<!--  hyp2 -->
-<!-- ``` -->
+  # Table
+  hyp2 <- flextable(hyp) %>% table_style12()
+
+  hyp2
+```
 
@@
 
-<!-- ```{r Figcolsig, fig.align = 'center', fig.cap = "Additional examples of chameleon color signals and behavior in response to different social contexts.
+```{r Figcolsig, fig.align = 'center', fig.cap = "Additional examples of chameleon color signals and behavior in response to different social contexts.
 (A) Two different males in conspicuous display color adopted at the start of a contest and retained by the dominant male.
@@
 (C) Conspicuous lemon-yellow male courtship color.
-In this case, the female is using a contrasting stress color consistent with rejection."} -->
-<!-- figs1 <- image_read("./output/figures/FigS1.png") -->
-<!-- figs1 -->
+In this case, the female is using a contrasting stress color consistent with rejection."}
+figs1 <- image_read("./output/figures/FigS1.png")
+figs1
 
-<!-- ``` -->
+```
 
-<!-- ```{r Figcolsig2, fig.align = 'center', fig.cap = "Representative photos of habitat in Hawaii (A, B) and Kenya (C, D).
-In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines."} -->
-<!-- figs2 <- image_read("./output/figures/FigS2.png") -->
-<!-- figs2 -->
+```{r Figcolsig2, fig.align = 'center', fig.cap = "Representative photos of habitat in Hawaii (A, C) and Kenya (B, D).
+In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines."}
+figs2 <- image_read("./output/figures/FigS2.png")
+figs2
 
-<!-- ``` -->
+```
 
@@
 
-<!-- ```{r Figcolsig3, fig.align = 'center', fig.cap = "The experimental setup for behavioral trials.
-We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male."} -->
-<!-- figs3 <- image_read("./output/figures/FigS3.png") -->
-<!-- figs3 -->
+```{r Figcolsig3, fig.align = 'center', fig.cap = "The experimental setup for behavioral trials.
+We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male."}
+figs3 <- image_read("./output/figures/FigS3.jpg")
+figs3
 
-<!-- ``` -->
+```
 

### 2021-06-01 supp.Rmd
Final fig and final fixes to supp

@@
 
-<!-- ```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator.
-(A) A model snake (Dispholidus typus) causing a characteristic contrasting color change and (B) a model bird (Aviceda cuculoides) which was flown over the top of the chameleon."} -->
-<!-- figs4 <- image_read("./output/figures/FigS4.png") -->
-<!-- figs4 -->
+```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator.
+(A) A model snake (Dispholidus typus) causing a characteristic contrasting color change and (B) a model bird (Aviceda cuculoides) which was flown over the top of the chameleon."}
+figs4 <- image_read("./output/figures/FigS4.jpg")
+figs4
 
-<!-- ``` -->
+```
 

### 2021-06-01 ms.Rmd
fixed some spelling to uS

@@
 Chameleon signals also showed evidence of local adaptation.
-Relaxing natural selection can result in signal release in systems experiencing strong sexual selection, including animals with dynamic colour change, and may act differently on the two primary visual channels (chromatic and achromatic).
+Relaxing natural selection can result in signal release in systems experiencing strong sexual selection, including animals with dynamic color change, and may act differently on the two primary visual channels (chromatic and achromatic).
 
@@
 [@RN2; @RN3]).
-Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in behaviour and phenotype over short time intervals (e.g., [@RN4; @RN5; @RN6]).
+Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in behavior and phenotype over short time intervals (e.g., [@RN4; @RN5; @RN6]).
 
-Conspicuous colour signals evolve because of the fitness advantage conferred to the signaller.
+Conspicuous color signals evolve because of the fitness advantage conferred to the signaller.
 These signals frequently become increasingly elaborate through a combination of male contest competition and female preference (e.g.
@@
 Evolution can respond to this selection pressure in a variety of ways that include shifting conspicuous signals to areas of the body that might be concealed to predators.
-For example, many conspicuous lizards have colourful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards @RN10.
+For example, many conspicuous lizards have colorful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards @RN10.
 Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-rich streams [@RN6; @RN9].
-These examples are from animals with static colour signals.
-A parallel solution that reduces the risk of exposure to predators is dynamic colour change because signals can be switched on and off relatively rapidly through neural or endocrine control [@RN11; @RN12].
+These examples are from animals with static color signals.
+A parallel solution that reduces the risk of exposure to predators is dynamic color change because signals can be switched on and off relatively rapidly through neural or endocrine control [@RN11; @RN12].
 A wide range of invertebrates (e.g.
@@
 
-Unlike animals with static colour signals, we have a poor understanding of how animals with dynamic colour change respond to predation pressure and whether relaxing predation pressure will result in an increasingly conspicuous signal.
+Unlike animals with static color signals, we have a poor understanding of how animals with dynamic color change respond to predation pressure and whether relaxing predation pressure will result in an increasingly conspicuous signal.
 Tackling this question in the wild is particularly difficult unless we can take advantage of unplanned changes to an organism’s environment such as during colonization of a novel habitat.
@@
 Specifically, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal.
-We tested the following predictions that underpin this hypothesis: 1.) display colouration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; and 2.) display colouration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawaii than Kenya.
-We also asked whether colour signals were consistent with local adaptation.
-Specifically, we predicted that chameleon display colouration in each population would be more conspicuous within its own signaling environment compared to the other population’s signaling environment.
+We tested the following predictions that underpin this hypothesis: 1.) display coloration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; and 2.) display coloration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawaii than Kenya.
+We also asked whether color signals were consistent with local adaptation.
+Specifically, we predicted that chameleon display coloration in each population would be more conspicuous within its own signaling environment compared to the other population’s signaling environment.
 For example, Hawaiian chameleons were historically found in Kenya but under local adaptation they are predicted to be more conspicuous in their current signaling environment, Hawaii.
@@
 
-Under relaxed predation, sexual selection is expected to favour more conspicuous dynamic sexual signals in Hawaii than Kenya.
-In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig.\@ref(fig:figure2) &  Fig.\@ref(fig:figure3)A) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses).
+Under relaxed predation, sexual selection is expected to favor more conspicuous dynamic sexual signals in Hawaii than Kenya.
+In support of this prediction, male display coloration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig.\@ref(fig:figure2) &  Fig.\@ref(fig:figure3)A) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses).
 Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (achromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Fig.\@ref(fig:figure3)).
@@
 Also, differential predation pressure resulting from differences in predation rates by snake and bird predators would also result in populations varying in their conspicuousness to each predator type.
-In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators with respect to brightness (Fig.\@ref(fig:figure3)C; achromatic contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`).
+In support of these predictions, Hawaiian chameleons were significantly more conspicuous to both snake and bird predators with respect to brightness (Fig.\@ref(fig:figure3)C; achromatic contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`).
 Differences in chromatic contrast between populations, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
@@
 We have documented a surprisingly strong effect of environment on the evolution of a conspicuous, dynamic sexual signal.
-For a founding population of chameleons, the tropical environment of Hawaii represents a particularly favourable set of conditions with respect to habitat, temperature, and rainfall.
+For a founding population of chameleons, the tropical environment of Hawaii represents a particularly favorable set of conditions with respect to habitat, temperature, and rainfall.
 Furthermore, Jackson’s chameleons are dietary generalists @RN17 and likely experience low competition for relatively abundant arthropod prey.
 But most importantly, Hawaii has no snakes and no raptors that feed on lizards.
-Domestic and feral cats would be their main threat on Oahu, and cats have very limited colour vision @RN18.
-In the absence of the usual suite of predators they would experience in their native range of Kenya—a diverse community of snake and bird predators, the conditions were highly favourable for character release of sexual signals—particularly in a polygynous species in which males have intense male contest competition and elaborate courtship displays (Fig.\@ref(fig:figure1)).
+Domestic and feral cats would be their main threat on Oahu, and cats have very limited color vision @RN18.
+In the absence of the usual suite of predators they would experience in their native range of Kenya—a diverse community of snake and bird predators, the conditions were highly favorable for character release of sexual signals—particularly in a polygynous species in which males have intense male contest competition and elaborate courtship displays (Fig.\@ref(fig:figure1)).
 Not only do males use colorful displays, but they have three horns that they lock and twist into rivals during intense bouts that sometimes result in physical injury.
@@
 Landmark studies such as those on Trinidadian guppies [@RN6; @RN9]  have shown the important interaction between natural and sexual selection in static signals following experimental manipulation.
-Ours is one of the only studies that show that a dynamic sexual signal—rapid colour change—is likewise constrained by natural selection.
+Ours is one of the only studies that show that a dynamic sexual signal—rapid color change—is likewise constrained by natural selection.
 Our results suggest that dynamic sexual signals are not simply plastic in a way that allows an individual to moderate the extent to which they ‘turn on’ a display but rather, that there is a threshold constrained by natural selection.
-This may in part be explained by the nature of colour change in chameleons.
+This may in part be explained by the nature of color change in chameleons.
 Chameleons have two layers of superposed iridiophores that facilitates color change that produces both camouflage and conspicuous sexual signals @RN22.
-Interestingly, while Hawaiian chameleons have evolved brighter display coloration, other elements of colour including hue and saturation have not experienced similar levels of selection except to become more conspicuous against their own background (local adaptation).
+Interestingly, while Hawaiian chameleons have evolved brighter display coloration, other elements of color including hue and saturation have not experienced similar levels of selection except to become more conspicuous against their own background (local adaptation).
 This suggests that the two visual channels (achromatic vs chromatic) experience differing levels of selection which is perhaps not surprising given that they act independently of one another @RN23.

### 2021-06-02 supp.Rmd
Added some more details on vs model, fixed up figS2

@@
 (C) Conspicuous lemon-yellow male courtship color.
-In this case, the female is using a contrasting stress color consistent with rejection."}
+In this case, the female is in a neutral state.
+When females reject males they develop a heavily contrasting pattern."}
 figs1 <- image_read("./output/figures/FigS1.png")
@@
 
-```{r Figcolsig2, fig.align = 'center', fig.cap = "Representative photos of habitat in Hawaii (A, C) and Kenya (B, D).
-In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines."}
-figs2 <- image_read("./output/figures/FigS2.png")
+```{r Figcolsig2, fig.align = 'center', fig.cap = "Representative photos of habitat in Kenya (A, B) and Hawaii (C, D).
+In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines"}
+figs2 <- image_read("./output/figures/FigS2.jpg")
 figs2
@@
 ```{r Figcolsig3, fig.align = 'center', fig.cap = "The experimental setup for behavioral trials.
-We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male."}
+We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male.
+The male on the left won this contest and is in full display color while the male on the right has lost this contest and turned brown (subordinate)."}
 figs3 <- image_read("./output/figures/FigS3.jpg")
@@
 
-## Color measurements, habitat background and irradiance
+## Spectroradiometry, habitat background and irradiance
 
@@
 When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 optic spectrophotometer and a PX2 light source.
-We used a custom-made probe-holder which we placed in contact with the animal at a standard angle (45&deg; @RN2) and distance (5 mm), thereby excluding ambient light.
+We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal at a standard angle (45&deg; @RN2) and distance (5 mm), thereby excluding ambient light.
 All measurements were relative to a dark, and a certified 99% white reflectance standard.
@@
 We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signalling environment.
-This included 1.) the upper surface of leaves; 2.) the underside of leaves; and 3.) branches.
-We also measured the habitat light environment by recording irradiance in the chameleons environment.
+This included 1.) the upper surface of leaves; 2.) the underside of leaves; and 3.) stems We also measured the habitat light environment by recording irradiance in the chameleons environment.
+Measurements were taken parallel to the ground (side-welling irradiance) to mimic signaling conditions for chameleons—light illuminating the chameleon’s flank is more likely to play a role in signaling @RN8.
 
@@
 
+We first smoothed all spectral reflectance files by a factor of 0.16 using the plotsmooth function in Pavo (vers 1.0).
+We then lifted negative reflectance spectra by the factor they were negative by, thereby maintaining the shape of the curve.
+We also took the median reflectance for all leaves (top, bottom) and stems (2 measures/stem) for each population (Hawaii, Kenya), and for irradiance.
+
+We used visual modelling to determine how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the Vorobyev-Osario (V-O) receptor noise model @RN9.
+The V-O model calculates perceptual distance (∆S) between two color patches in units of just noticeable differences (JNDs).
+A difference of at least 1 JND is thought to enable a receiver to discriminate two colors @RN10.
+We calculated JNDs for chromatic- and achromatic contrast for all body regions combined (overall conspicuousness) and for individual body regions in order to compare conspicuousness as a function of population (Hawaii, Kenya) or predator (bird, snake) in relation to background (vegetation).
+The receptor noise model incorporates the median reflectance of background color, irradiance, chameleon spectral reflectance and visual pigment absorbance functions for the Jackson’s chameleon (Loew unpubl.
+data), birds, and the garter snake @RN11.
+
 ## Statistical Analysis
@@
 In all cases, our results and conclusions were not impacted (see supplementary code).
-Deriving multiple contrast for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other.
+Deriving multiple contrasts for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other.
 To ensure that non-independence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction.
@@
 Relevant model interactions were first tested using Wald tests.
-If found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model.
+If found to be non-significant (p > 0.1) we dropped the interaction and fitted a main effects model.
 We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.

### 2021-08-15 ms.Rmd
Fixed all text and refs. Need tio add one block of new analysis

@@
                           header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Sample sizes
+N_display <- data.frame(Display_1 %>% group_by(Population) %>% summarise(n = length(unique(ID))))
 
@@
 
+# Sample sizes
+N_Courtship <- data.frame(Courtship %>% group_by(Population) %>% summarise(n = length(unique(ID))))
+
 # Center SVL do intercept is at the mean body size
@@
                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Sample sizes
+N_Pred <- data.frame(Predator %>% group_by(Population, Predator) %>% summarise(n = length(unique(ID)))) %>% arrange(Predator)
 
@@
 
-# Abstract
-Rapid evolution frequently follows a significant change to an organism’s environment, such as during colonization of a novel habitat.
-We show that Jackson’s chameleons (*Trioceros j.
-xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) have evolved significantly more conspicuous sexual signals.
-Compared to Kenya, Oahu has few potential predators and therefore, very low predation pressure.
-We used visual modelling to quantify the conspicuousness of wild chameleons from both Hawaii and Kenya to multiple social and predator stimuli in order to test for character release of sexual signals.
-The strongest effect was for achromatic contrast, where Hawaiian chameleons were brighter during male contests and courtship and were more conspicuous to snake and bird predators.
-Chameleon signals also showed evidence of local adaptation.
-Relaxing natural selection can result in signal release in systems experiencing strong sexual selection, including animals with dynamic color change, and may act differently on the two primary visual channels (chromatic and achromatic).
+# Summary
+Colonization of novel environments can free introduced populations from the constraints imposed by native-range predators and promote rapid evolutionary change.
+We demonstrate such character release for color signals in Jackson’s chameleons (*Trioceros j.
+xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) where there are few lizard predators.
+Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators, and showed greater change between display and anti-predator color states.
+Consistent with local adaptation, the display colors of Hawaiian chameleons were more conspicuous in their local than ancestral habitats.
+These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
 
-# Introduction
+# Main text
 
-When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods @RN1.
-We now have numerous examples of evolutionary change following natural selection in both the laboratory and wild, and not only in invertebrates with short generation times.
-Vertebrates, including Caribbean *Anolis* lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales (e.g.
-[@RN2; @RN3]).
-Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in behavior and phenotype over short time intervals (e.g., [@RN4; @RN5; @RN6]).
+When species encounter novel environments, such as during human-mediated introductions, evolutionary change can occur more rapidly than Darwin ever thought possible [@RN1; @RN2; @RN3; @RN4].
+In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4; @RN5].
+Such character release should be most evident for social and sexual signals; released from constraints on conspicuousness imposed by predation risk, signals could elaborate in many directions [e.g., @RN4; @RN6; @RN7; @RN8].
+However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7; @RN9; @RN10].
+This is especially true for species with dynamic color change, for which we are not aware of any examples.
+Here, we examine character release of a dynamic color signal by taking advantage of an unintended evolutionary ‘experiment’: accidental translocation of chameleons to a novel, low predation environment.
 
-Conspicuous color signals evolve because of the fitness advantage conferred to the signaller.
-These signals frequently become increasingly elaborate through a combination of male contest competition and female preference (e.g.
-runaway selection [@RN7; @RN8]).
-However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly @RN9.
-Evolution can respond to this selection pressure in a variety of ways that include shifting conspicuous signals to areas of the body that might be concealed to predators.
-For example, many conspicuous lizards have colorful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards @RN10.
-Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-rich streams [@RN6; @RN9].
-These examples are from animals with static color signals.
-A parallel solution that reduces the risk of exposure to predators is dynamic color change because signals can be switched on and off relatively rapidly through neural or endocrine control [@RN11; @RN12].
-A wide range of invertebrates (e.g.
-cephalopods) and vertebrates (e.g.
-lizards and frogs) have convergently evolved this ability [@RN13; @RN14; @RN15].
+In 1972, a shipment of Jackson’s chameleons (*Trioceros jacksonii xanthopholus*; (Fig.\@ref(fig:figure1) and fig.
+S1)) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
+The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11].
+Jackson’s chameleons have a high intrinsic population growth rate (9--12 month generation time and live birth of up to 50 young @RN24), enabling their rapid establishment and potentially rapid evolution (approx.
+50 -- 65 generations).
+On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators.
+This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.
 
-Unlike animals with static color signals, we have a poor understanding of how animals with dynamic color change respond to predation pressure and whether relaxing predation pressure will result in an increasingly conspicuous signal.
-Tackling this question in the wild is particularly difficult unless we can take advantage of unplanned changes to an organism’s environment such as during colonization of a novel habitat.
-Living organisms are occasionally translocated great distances by human action to environments far removed from their native range.
-These novel environments typically have a completely different suite of organisms, including predators, and offer a unique opportunity to study signal evolution in founder populations.
-Jackson’s chameleons (*Trioceros j.
-xanthopholus*) are just such a case (Fig.\@ref(fig:figure1) and fig.
-S1).
-In 1972, a shipment of chameleons was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
-The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island @RN16.
-On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors), while in Kenya they are preyed upon by a wide range bird, snake and occasionally, mammal, predators.
-This accidental introduction represents a unique opportunity to study the evolution of a dynamic visual signal by comparing native (Kenya) and introduced (Oahu) chameleons under low and high levels of natural selection.
+Chameleons are famous for their rapid color change [@RN12; @RN13; @RN14].
+When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green to signal dominance or during courtship (Fig.
+1, fig.
+S1, S3).
+At other times, they are a duller green and can darken to brown or become mottled green and brown, particularly in response to perceived predation risk (Fig.\@ref(fig:figure1), fig.
+S4A).
+Their different color states can appear more or less conspicuous depending on the context, background, and viewer (e.g., conspecific or predator).
+If introduced chameleons in Hawaii experience character release from reduced predation, we expect Hawaiian chameleons more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators.
+We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.
 
-Specifically, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal.
-We tested the following predictions that underpin this hypothesis: 1.) display coloration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; and 2.) display coloration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawaii than Kenya.
-We also asked whether color signals were consistent with local adaptation.
-Specifically, we predicted that chameleon display coloration in each population would be more conspicuous within its own signaling environment compared to the other population’s signaling environment.
-For example, Hawaiian chameleons were historically found in Kenya but under local adaptation they are predicted to be more conspicuous in their current signaling environment, Hawaii.
-Likewise, Kenyan chameleons have no history in Hawaii and are therefore predicted to be more conspicuous within their current signaling environment, Kenya (table S1).
+To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Fig.\@ref(fig:figure1), fig.
+S1, S4).
+We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, N = `r N_display[1,2]` Hawaii, N = `r N_display[2,2]` Kenya), courtship (N = `r N_Courtship[1,2]` Hawaii, N = `r N_Courtship[2,2]` Kenya) and in response to bird (N = `r N_Pred[1,3]` Hawaii, N = `r N_Pred[2,3]` Kenya) and snake (N = `r N_Pred[3,3]` Hawaii, N = `r N_Pred[4,3]` Kenya) predators (Fig.\@ref(fig:figure2), fig.
+S5).
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs^@RN25 against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
 
-To test for character release and whether color signals show local adaptation, we quantified the local signalling environment (fig.
-S2) and manipulated the social and predatory context chameleons encountered, in both their native Kenyan, and introduced Hawaiian, range (fig.
-S3).
-We presented wild-caught male chameleons with the following treatments: a male, a female, a model bird predator (African cuckoo-hawk, *Aviceda cuculoides*) and a model snake predator (boomslang, *Dispholidus typus*) (Fig.\@ref(fig:figure1) and fig.
-S1,S4).
-We measured the spectral reflectance of males in dominant, subordinate, courtship, and antipredator (bird, snake) display states (Fig.\@ref(fig:figure2) and fig.
-S5).
-We then used visual modelling against a known chameleon visual system for sexual signals, and a bird and snake visual system for predation trials, to quantify chromatic and achromatic contrast against the backgrounds of each environment (supplementary text).
+In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.\@ref(fig:figure2), Fig.\@ref(fig:figure3)A) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`).
+This result was consistent for all body regions measured (no significant interaction with body region; supplementary text).
+However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`).
 
-Under relaxed predation, sexual selection is expected to favor more conspicuous dynamic sexual signals in Hawaii than Kenya.
-In support of this prediction, male display coloration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig.\@ref(fig:figure2) &  Fig.\@ref(fig:figure3)A) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses).
+The importance of luminance contrast is consistent with the nature and mechanism of rapid color change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.\@ref(fig:figure1)).
+Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15 or changes in the spacing of guanine crystals within iridophores @RN14, both of which strongly affect luminance.
+Accordingly, character release was observed in the overall luminance of signals rather than their hue.
+
+Released from predation, Hawaiian chameleons were less cryptic than Kenyan chameleons when threatened by both bird and snake predators.
+Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.\@ref(fig:figure3)C; luminance contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type).
+Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
+Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D; Birds (Hawaii-Kenya): $\Beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`, JND values > 5).
+By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.\@ref(fig:figure3)D; Snakes (Hawaii-Kenya): $\Beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; JND values ~ 2).
+The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18].
+The poor color discrimination of snakes may also explain why chameleons from Hawaii show little change relative to their Kenyan counterparts in their chromatic response to snakes, despite the absence of snakes on Hawaii.
+
+Character release may reflect rapid evolutionary change, phenotypic plasticity or a combination of these processes [@RN19; @RN20].
+Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
+Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, and for both luminance and chromatic contrast (Fig.\@ref(fig:figure4); Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: luminance ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: luminance ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
+Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].
+
+To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states .........
+
+We have documented a surprisingly strong effect of recent release from predation on a conspicuous, dynamic visual signal.
+Oahu has no snakes or raptors that feed on lizards.
+Domestic and feral cats would be their main threat on Oahu, and cats have very limited color vision [@RN22].
+In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators -- the conditions were highly favorable for character release of social signals.
+Jackson’s chameleons likely experience strong sexual selection, given their polygynous mating system, elaborate courtship displays and intense male contest competition (Fig.\@ref(fig:figure1)) in which males lock and twist their three horns into rivals during intense bouts that sometimes result in physical injury.
+Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
+
+Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12].
+Our results suggest that even for species capable of dynamic color change there is an upper threshold to signal intensity that is constrained by natural selection.
+This novel finding raises the question of whether other animals that have convergently evolved dynamic color signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
+More generally, our study highlights the opportunities provided by translocations and invasions, which are often otherwise detrimental, to study natural and sexual selection in the wild.
+
+## OLD
+
  Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (achromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Fig.\@ref(fig:figure3)).
 
-As a consequence of signal release, sexual signals are expected to be more easily detected by predators (i.e., snakes and birds) in Hawaii.
-Also, differential predation pressure resulting from differences in predation rates by snake and bird predators would also result in populations varying in their conspicuousness to each predator type.
-In support of these predictions, Hawaiian chameleons were significantly more conspicuous to both snake and bird predators with respect to brightness (Fig.\@ref(fig:figure3)C; achromatic contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`).
-Differences in chromatic contrast between populations, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
-Interestingly, Kenyan and Hawaiian chameleons were equally detectable by snakes, but Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).
-
-Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment.
-While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis.
-Hawaiian chameleons were more conspicuous to other chameleons against heir own (Hawaiian) background as opposed to the Kenyan background, in both social contexts and for both achromatic and chromatic contrast (Fig.\@ref(fig:figure4); Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
-
-We have documented a surprisingly strong effect of environment on the evolution of a conspicuous, dynamic sexual signal.
-For a founding population of chameleons, the tropical environment of Hawaii represents a particularly favorable set of conditions with respect to habitat, temperature, and rainfall.
-Furthermore, Jackson’s chameleons are dietary generalists @RN17 and likely experience low competition for relatively abundant arthropod prey.
-But most importantly, Hawaii has no snakes and no raptors that feed on lizards.
-Domestic and feral cats would be their main threat on Oahu, and cats have very limited color vision @RN18.
-In the absence of the usual suite of predators they would experience in their native range of Kenya—a diverse community of snake and bird predators, the conditions were highly favorable for character release of sexual signals—particularly in a polygynous species in which males have intense male contest competition and elaborate courtship displays (Fig.\@ref(fig:figure1)).
-Not only do males use colorful displays, but they have three horns that they lock and twist into rivals during intense bouts that sometimes result in physical injury.
-Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated—an increasingly bright sexual signal used in both male contests and during courtship.
-
-Evolutionary biologists, particularly those interested in speciation, seldom consider the role of natural selection in shaping sexually selected signals even though sexual signals evolve in an ecological context @RN19 (but see [@RN5; @RN20, @RN21]).
-Landmark studies such as those on Trinidadian guppies [@RN6; @RN9]  have shown the important interaction between natural and sexual selection in static signals following experimental manipulation.
-Ours is one of the only studies that show that a dynamic sexual signal—rapid color change—is likewise constrained by natural selection.
-Our results suggest that dynamic sexual signals are not simply plastic in a way that allows an individual to moderate the extent to which they ‘turn on’ a display but rather, that there is a threshold constrained by natural selection.
-This may in part be explained by the nature of color change in chameleons.
-Chameleons have two layers of superposed iridiophores that facilitates color change that produces both camouflage and conspicuous sexual signals @RN22.
-Interestingly, while Hawaiian chameleons have evolved brighter display coloration, other elements of color including hue and saturation have not experienced similar levels of selection except to become more conspicuous against their own background (local adaptation).
-This suggests that the two visual channels (achromatic vs chromatic) experience differing levels of selection which is perhaps not surprising given that they act independently of one another @RN23.
-While studying evolution is challenging because it is not possible to go back in time to witness key evolutionary events, we can take advantage of unique opportunities such as an invasion, to better understand how natural and sexual selection interact to shape animal signals.
-
-```{r figure1, cache=FALSE, fig.cap = "Chameleon color signal change in response to different social and predatory stimuli.
-Male chameleons experience intense sexual selection.
+```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social and predatory stimuli.
+Male chameleons experience intense sexual selection**.
 During the breeding season they change from a neutral state (green) to a highly conspicuous bright yellow display signal.
-They also readily fight by locking horns and sometimes pierce their rivals skin with their horns.
+They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
 (A) A typical dominant male in display coloration.
 (B) A subordinate male that lost a contest and turned from bright yellow to brown.
-(C) Two males fighting, both are in display colouration and at a stage where they are relatively even matched.
-(D) A male courting female in full display color while the female has turned to a contrasting colour, which suggests rejection of the male"}
+(C) Two males fighting, both are in display coloration and at a stage where they are relatively even matched.
+(D) A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male"}
 library(magick)
@@
 
-```{r figure2, fig.width = 14, fig.height= 14, fig.cap="Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya.
+```{r figure2, fig.width = 14, fig.height= 14, fig.cap="**Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya**.
 The context for measurement was (A) male contest displays; (B) courtship; (C) bird predator; and (D) snake predator.
-More details in online supplementary material and Fig.
+More details in online supplementary text and Fig.
 S5"}
@@
 
-```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves).
-(A & B) Acromatic and chromatic contrast of male chameleons during male-male contests and female courtship.
+```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "**Chromatic and luminance contrast in Just Noticeable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves)**.
+(A & B) Luminance and chromatic contrast of male chameleons during male-male contests and female courtship.
 JNDs are calculated based on the chameleon visual system.
-(B & C) Acromatic and chromatic contrasts of male chameleons during snake and bird predator encounters.
+(B & C) Luminance and chromatic contrasts of male chameleons during snake and bird predator encounters.
 JNDs are calculated based on the snake and bird visual systems.
-Contrasts between means, 'Beta', that are bold indicate significant effects."}
+Contrasts between means, ‘Beta’, that are bold indicate significant effects.."}
 rerun = FALSE
@@
 
-```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "Evidence for local adaptation of color signals in both (A) chromatic- and (B) achromatic contrast for Hawaiian chameleons.
+```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "**Evidence for local adaptation of color signals in both (A) chromatic- and (B) luminance contrast for Hawaiian chameleons**.
 For local adaptation, signals are predicted to be more conspicuous against their own background.
-In this case, Hawaiian chameleons were both significantly more chromatic and brighter (achromatic contrast)  against their own background compared to the Kenyan background suggesting that selection has acted on increasing local signal conspicuousness."}
+In this case, Hawaiian chameleons were both significantly more chromatic and brighter (luminance contrast) against their own background compared to the Kenyan background suggesting that selection has acted on increasing local signal conspicuousness."}
 rerun=FALSE

### 2021-08-15 ms.Rmd
Added new analyses. Not integrated yet, will add to supp and explore in more detail BodyRegions

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, magick, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, magick, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable, readxl)
 source("./R/func.R")
@@
 
+#########################################################
+# Comparing displays
+#########################################################
+ # Load data from Devi
+       Display_bird_diff <- read_excel("data/Display_bird_diff.xlsx")
+
+       Diff <- Display_bird_diff
+
+      names(Diff) # header rows R is reading
+
+    Diff$BodyRegion <- factor(Diff$BodyRegion) #updates levels for the variable BodyRegion
+    str(diff.difftime) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
+    Diff$ID <- factor(Diff$ID) #updates levels for the variable ID
+    str(diff.difftime) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
+
+# Center SVL to make intercept interpretable
+    Diff$sc_SVL <- with(Diff, scale(SVL, scale=FALSE))
+
+# Filter missing SVL data
+  Diff <- Diff %>% filter(!is.na(SVL)) # filter out misssing SVL data
+
+# create as factor
+    Diff$BodyRegion <- factor(Diff$BodyRegion)
+
+# Sample sizes
+    N_diff <- data.frame(Diff %>% group_by(Population) %>% summarise(n = length(unique(ID))))
+
+##################
+# Achromatic contrast
+##################
+
+# Fit model testing this idea above
+  Diff_dL <- lmer(dL_diff_abs ~ SVL + Population + (1|ID), data=Diff)
+  summary(Diff_dL)
+
+# Check residuals of model
+  hist(residuals(Diff_dL))
+  plot(Diff_dL) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
+
+# Refit mdoel using robust standard errors
+  robust_Diff_dl <- clubSandwich::conf_int(Diff_dL, cluster = Diff$ID, vcov =  "CR2")
+
+# Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
+   robust_Diff_dl_pop <- Wald_test(Diff_dL, constraints = constrain_zero(3), 
+                                  cluster = Diff$ID, vcov = "CR2")
+   robust_Diff_dl_svl <- Wald_test(Diff_dL, constraints = constrain_zero(2), 
+                                  cluster = Diff$ID, vcov = "CR2")
+
+##################
+# Chromatic contrast
+##################  
+   # Fit model testing this idea above
+  Diff_dS <- lmer(dS_diff_abs ~  SVL + Population + (1|ID), data=Diff)
+  summary(Diff_dS)
+
+# Check residuals of model
+  hist(residuals(Diff_dS))
+  plot(Diff_dS) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
+
+# Refit mdoel using robust standard errors
+  robust_Diff_dS <- clubSandwich::conf_int(Diff_dS, cluster = Diff$ID, vcov =  "CR2")
+
+# Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
+   robust_Diff_dS_pop <- Wald_test(Diff_dS, constraints = constrain_zero(3), 
+                                  cluster = Diff$ID, vcov = "CR2")
+   robust_Diff_dS_svl <- Wald_test(Diff_dS, constraints = constrain_zero(2), 
+                                  cluster = Diff$ID, vcov = "CR2")
+
 ```
 
-Martin J. Whiting^1*^, Brenden S. Holland^2^, J. Scott Keogh^3^, Daniel W.A.Noble^3^, Katrina J. Rankin^4^ and Devi M. Stuart-Fox^4^
+Martin J. Whiting^1^, Brenden S. Holland^2^, J. Scott Keogh^3^, Daniel W.A.Noble^3^, Katrina J. Rankin^4^ and Devi M. Stuart-Fox^4^
 
@@
 
-# Summary
+# Abstract
 Colonization of novel environments can free introduced populations from the constraints imposed by native-range predators and promote rapid evolutionary change.
@@
 S5).
-We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs^@RN25 against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs @RN25 against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
 
@@
 Released from predation, Hawaiian chameleons were less cryptic than Kenyan chameleons when threatened by both bird and snake predators.
-Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.\@ref(fig:figure3)C; luminance contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type).
+Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type).
 Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
-Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D; Birds (Hawaii-Kenya): $\Beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`, JND values > 5).
-By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.\@ref(fig:figure3)D; Snakes (Hawaii-Kenya): $\Beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; JND values ~ 2).
+Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`, JND values > 5).
+By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.\@ref(fig:figure3)D; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; JND values ~ 2).
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18].
@@
 Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
-Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, and for both luminance and chromatic contrast (Fig.\@ref(fig:figure4); Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: luminance ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: luminance ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
+Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, and for both luminance and chromatic contrast (Fig.\@ref(fig:figure4); male-male contests: luminance ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$), $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: luminance ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
 Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].
+# Current Sec
+To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; N = `r N_diff[1,2]` Hawaii, N = `r N_diff[2,2]` Kenya).
 
-To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states .........
+Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
+While it is possible that differences between Hawaiian and Kenyan chameleons could reflect a founder effect, the observed differences are all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the local environment.
 
@@
 
-## OLD
-
- Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (achromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Fig.\@ref(fig:figure3)).
+# Figures
 
@@
 
-# References
+# References and Notes

### 2021-08-15 ms.Rmd
Fixed Luminance in figure legends, also fixed ai and figures

@@
   xlab("") + 
-  ylab("Achromatic contrast (JND)") +
+  ylab("Luminance contrast (JND)") +
    annotate("segment", x = 0.90, xend = 1.90, y = 40, yend = 40, colour = "black") +
@@
   xlab("") + 
-  ylab("Achromatic contrast (JND)") +
+  ylab("Luminance contrast (JND)") +
    annotate("segment", x = 0.90, xend = 1.90, y = 28, yend = 28, colour = "black") +
@@
   xlab("Background Environment") + 
-  ylab("Achromatic contrast (JND)") +
+  ylab("Luminance contrast (JND)") +
    annotate("segment", x = 0.90, xend = 1.90, y = 40, yend = 40, colour = "black") +

### 2021-08-15 supp.Rmd
Updated supp to Additional Supp Results. Some model fixes needed with SVL. Still need to add in Devi's new analysis

@@
 
+# Invasive chameleons released from predation display more conspicuous colors 
+Martin J. Whiting, Brenden S. Holland, J. Scott Keogh, Daniel W.A.Noble, Katrina J. Rankin and Devi M. Stuart-Fox
+
+Correspondance: Martin J. Whiting
+Email: Martin.Whiting@mq.edu.au
+
+The file includes:
+Supplementary text
+Figures S1 -- Sx
+Tables S1 -- Sx
+Supplementary References
+
 # Expanded Hypotheses and Predictions
@@
 
+# Potential bird predators of chameleons on Oahu
+We are not aware of any evidence of bird predation of chameleons on Oahu.
+We examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons @RN1.
+Among the larger, predatory birds, three are owls (*Asio flammeus*, *Bubo scandiacus*, *Tyto alba*) that are highly unlikely to feed on chameleons because they are either mostly or entirely nocturnal, extremely rare on Oahu, or feed mostly on rodents and/or birds @RN2.
+Smaller raptors include the Merlin (*Falco columbarius*), which is considered rare and feeds on small birds in more open habitats while the peregrine falcon (*Falco peregrinus*) is a predator of medium-sized birds and also considered rare @RN2.
+There are also seven species of egret and heron, which typically have generalist diets.
+Of these, six are considered rare or accidental @RN1, are associated with water bodies, and better known for eating fishes @RN3.
+The cattle egret (*Bubulcus ibis*) is a generalist predator that is mostly insectivorous but does eat lizards opportunistically [@RN4;@RN5].
+However, it forages on the ground in more open habitats, often following domestic animals @RN5 and therefore, is highly unlikely to encounter chameleons.
+Among other potential predators, the Hawaiian Hawk or ʻio (*Buteo solitarius*) is restricted to the Big Island while the Hen Harrier (*Circus cyaneus*) is considered an accidental or rare visitor to the Hawaiian islands.
+Finally, among corvids, the Hawaiian crow (*Corvus hawaiiensis*) is extinct in the wild while the common raven (*Corvus corax*) is rare @RN1.
+In summary, we could not identify any bird species that could be regarded as a major predator of chameleons.
+
 # Supplemental Materials and Methods
@@
 
-We collected most chameleons (Fig.
-\@ref(fig:Figcolsig)) at night by spotlighting, and a small proportion of individuals during the day.
-When chameleons were beyond reach, we used 4-m extensible poles that chameleons could be encouraged to step on to, before lowering them to the ground.
+We collected most chameleons at night by spotlighting, and a small proportion of individuals during the day.
+When chameleons were beyond reach, we used 4 m extensible poles that chameleons could be encouraged to step on to, before lowering them to the ground.
 We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
+We measured snout-vent length (SVL), a standard measure of body size in lizards, from the tip of the snout to the posterior edge of the vent using a plastic ruler to the nearest 1 mm.
 Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006.
 In Hawaii, chameleons were collected from a single population in the Manoa Valley, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
-400-600m.
+400-600 m.
 The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Fig.
 \@ref(fig:Figcolsig2)A,B).
-Most of the plants in this area were nonnative, and chameleons also occurred in hedge-grows and stands of bamboo.
+Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
 The area experiences high annual rainfall (mean: ca.
-1500 mm) @RN1 and the mean annual temperature is 22-24&deg;C @RN2.
+1500 mm) @RN6 and the mean annual temperature is 22-24°C @RN7.
 In Kenya, field work was conducted near the foothills of Mt.
@@
 \@ref(fig:Figcolsig2)C,D).
-Mean annual temperature is 20.2&deg;C (mean max: 24.8&deg;C) and the mean annual rainfall is 121 mm @RN3.
+Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.
 
-```{r Figcolsig, fig.align = 'center', fig.cap = "Additional examples of chameleon color signals and behavior in response to different social contexts.
-(A) Two different males in conspicuous display color adopted at the start of a contest and retained by the dominant male.
-(B) Two examples of males engaged in horn locking.
-Males in the bottom plane are evenly matched at this stage and both are in full display colour.
+```{r Figcolsig, fig.align = 'center', fig.cap = "**Additional examples of chameleon color signals and behavior in response to different social contexts**.
+**(A)** Two different males in conspicuous display color adopted at the start of a contest and retained by the dominant male.
+**(B)** Two examples of males engaged in horn locking.
+Males in the bottom plane are evenly matched at this stage and both are in full display color.
 In the top image, the male on the left is starting to change to a subordinate color.
-(C) Conspicuous lemon-yellow male courtship color.
-In this case, the female is in a neutral state.
-When females reject males they develop a heavily contrasting pattern."}
+**(C)** Conspicuous yellow-green male courtship color.
+When females reject males, they develop a heavily contrasting pattern."}
 figs1 <- image_read("./output/figures/FigS1.png")
@@
 ## Behavioral trials
-
-In order to quantify display coloration we staged encounters between conspecifics (male-male and male-female) and model (bird and snake) predators and measured their subsequent display coloration (details below).
+In order to quantify display coloration, we staged encounters between conspecifics (male-male and male-female) and model predators (bird and snake) and measured their subsequent display coloration (details below).
 All chameleons were used in both predator and social trials although in some cases we were unable to quantify all social signals for a particular individual.
@@
 This structure was about 1.5 m high and each arm was approximately 60 cm.
-The triangular setup allowed chameleons to display at a comfortable distance and also to escape from aggressive interactions, particularly during male-male trials (Fig.
+The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Fig.
 \@ref(fig:Figcolsig4)).
 All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
-The signalling environment was therefore the same for all animals.
+The signaling environment was therefore the same for all animals.
 
-```{r Figcolsig3, fig.align = 'center', fig.cap = "The experimental setup for behavioral trials.
+```{r Figcolsig3, fig.align = 'center', fig.cap = "**The experimental setup for behavioral trials**.
 We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male.
@@
 ```
-
 We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
-This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing colour and behaviour (Fig.
+This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Fig.
 \@ref(fig:Figcolsig)A,B).
@@
 Once dominance was settled, the subordinate would turn brown (Fig.
-\@ref(fig:Figcolsig)B) and flee.
+\@ref(fig:Figcolsig)B and Fig.
+3) and flee.
 In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
@@
 A trial began when a chameleon was removed from its cloth bag and placed on the triangular stand.
-At this point the lizard was in a neutral color state.
 In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Fig.
@@
 Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
-Once chameleons changed color, we quantified their spectral reflectance following the same procedure described for social trials.
+Once chameleons changed color, we quantified their spectral reflectance (details below).
 We also conducted seven trials with a just a branch with no predator attached, as a control for the model predator.
-The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN6;@RN7]).
+The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
 The bird was a mounted African cuckoo-hawk (*Aviceda cuculoides*).
 
-```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator.
-(A) A model snake (Dispholidus typus) causing a characteristic contrasting color change and (B) a model bird (Aviceda cuculoides) which was flown over the top of the chameleon."}
+```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator**.
+**(A)** A model snake (*Dispholidus typus*) causing a characteristic contrasting color change and **(B)** a model bird (*Aviceda cuculoides*) which was flown over the top of the chameleon"}
 figs4 <- image_read("./output/figures/FigS4.jpg")
@@
 
-## Spectroradiometry, habitat background and irradiance
+## Color measurement
 
-During all behavior trials, we measured the color of each body region (Fig.
-\@ref(fig:Figcolsig5)) once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
-Patches were measured in random sequence to avoid any bias of order.
-When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 optic spectrophotometer and a PX2 light source.
-We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal at a standard angle (45&deg; @RN2) and distance (5 mm), thereby excluding ambient light.
-All measurements were relative to a dark, and a certified 99% white reflectance standard.
-When chameleons started to change color in response to our handling we stopped measuring color and allowed them to continue a social trial before continuing to measure spectral reflectance.
-In the case of a predator trial we would once again present them with the predator model until they returned to their anti-predator display.
+During all behavior trials, we measured the color of each of three body regions: top flank, mid flank, tail base and gular region (Fig.
+\@ref(fig:Figcolsig5)).
+We measured each body region once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
+The four body regions were measured in random sequence to avoid any bias of order.
+When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
+We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal.
+The probe holder ensured that measurements were taken at a standard angle (45°) and distance (5 mm) for a 3 x 5 mm area and excluded all ambient light.
+All measurements were relative to a dark, and a certified 99% white reflectance standard (Labsphere, North Sutton, NH, U.S.A.).
+If chameleons started to change color in response to our handling, we returned them to the branch and allowed them to continue a social trial before measuring reflectance of remaining body regions.
+In the case of a predator trial, we would once again present them with the predator model until they returned to their anti-predator display.
 This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behavior following handling.
 
-We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signalling environment.
-This included 1.) the upper surface of leaves; 2.) the underside of leaves; and 3.) stems We also measured the habitat light environment by recording irradiance in the chameleons environment.
-Measurements were taken parallel to the ground (side-welling irradiance) to mimic signaling conditions for chameleons—light illuminating the chameleon’s flank is more likely to play a role in signaling @RN8.
+We obtained spectral reflectance measurements of male dominant color state in male-male contests for 34 individuals from Hawaii and 35 from Kenya, courtship color state for 25 individuals from Hawaii and 32 from Kenya; color state in response to the bird for 56 individuals from Hawaii and 23 from Kenya and color state in response to the snake for 59 individuals from Hawaii and 38 from Kenya.
+
+We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signaling environment.
+We took 120 reflectance measures of the background at each location (Hawaii and Kenya), comprising 40 measurements of the top surface of leaves, 40 of the bottom surface of leaves and 40 stems and branches of suitable width for chameleons.
+To reduce the number of backgrounds, we averaged the reflectance of the 20 darker measurements and 20 lighter measurements for each of the top leaf and bottom leaf measurements and averaged brown stems and green stems.
+This resulted in 6 average background spectra for each location, representing the range of background colors.
 
@@
 
-We first smoothed all spectral reflectance files by a factor of 0.16 using the plotsmooth function in Pavo (vers 1.0).
-We then lifted negative reflectance spectra by the factor they were negative by, thereby maintaining the shape of the curve.
+We visualized and quality-checked all spectra, adjusted very minor negative reflectance values caused by electrical noise by lifting the curves by the maximum negative value, then smoothed all spectra with LOESS smoothing of 0.16 using the plotsmooth function in Pavo 1.0 @RN29.
 We also took the median reflectance for all leaves (top, bottom) and stems (2 measures/stem) for each population (Hawaii, Kenya), and for irradiance.
 
-We used visual modelling to determine how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the Vorobyev-Osario (V-O) receptor noise model @RN9.
-The V-O model calculates perceptual distance (∆S) between two color patches in units of just noticeable differences (JNDs).
-A difference of at least 1 JND is thought to enable a receiver to discriminate two colors @RN10.
-We calculated JNDs for chromatic- and achromatic contrast for all body regions combined (overall conspicuousness) and for individual body regions in order to compare conspicuousness as a function of population (Hawaii, Kenya) or predator (bird, snake) in relation to background (vegetation).
-The receptor noise model incorporates the median reflectance of background color, irradiance, chameleon spectral reflectance and visual pigment absorbance functions for the Jackson’s chameleon (Loew unpubl.
-data), birds, and the garter snake @RN11.
+We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package ‘pavo’ @RN29.
+This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
+Color contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of color discrimination for stimuli viewed simultaneously under ideal viewing conditions @RN17.
+Under natural conditions, discrimination thresholds are likely to be >1 and vary depending on the species and conditions [@RN12;@RN13; @RN14].
+Additionally, perceived conspicuousness may not scale linearly with color distance (JNDs) @RN15.
+Nevertheless, the models provide a reasonable approximation of relative conspicuousness that can be compared between populations and trial types (social and predator).
+
+We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region for each color state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
+We first calculated the quantum catch of each photoreceptor under standard daylight illumination (D65; Commission internationale de l’éclairage – CIE) using the ‘vismodel’ function.
+We applied von Kries chromatic adaptation (vonkries = TRUE) and used the average background reflectance of the relevant background (Hawaii or Kenya) as the adapting background.
+The signal in each photoreceptor class was proportional to the natural logarithm of the quantum catch, in accordance with Fechner's law (“fi” in pavo).
+Next, we calculated the chromatic contrast between pairs of spectra using the function ‘coldist’.
+
+The visual models require information on the spectral sensitivities and photoreceptor noise (Weber fraction), which is a function of the relative density of each photoreceptor type within the retina, for the relevant receiver.
+Chameleons have four spectral classes of single cone enabling tetrachromatic vision @RN16.
+We used spectral sensitivities for the congeneric flap-necked chameleon (*Chameleo dilepis*) @RN16 and a standard Weber fraction of 0.1 @RN17 to account for the receptor noise of the LWS photoreceptor.
+We calculated noise for the other photoreceptor classes using the relative photoreceptor ratio of 1: 1.6: 3.3: 3.2 for the four single cones (UVS, SWS, MWS, and LWS, respectively).
+*Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Fig.
+\@ref(fig:Figcolsig6SPec)).
+
+```{r Figcolsig6SPec, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Spectral sensitivities of chameleons (A), snake (B) and bird (C) used for visual models (details in text)**."}
+figs5 <- image_read("./output/figures/FigS5.jpg")
+figs5
+```
+
+For birds, we used an ultraviolet sensitive (UVS) visual system @RN18 characteristic of raptors and other birds of prey @RN19.
+A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement @RN20.
+We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21.
+
+For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake @RN22.
+We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22.
+
+In each case, visual pigment absorbance curves were multiplied by the transmission spectra of ocular media (lens and cornea) and oil droplets associated with the corresponding photoreceptor class and normalized to equal area under the curve to satisfy the assumption of equal stimulation by white light @RN18.
+
+Luminance contrast was calculated using the sensitivity function for the LWS photoreceptor (with transmission of the oil droplet associated with the double cone in chameleons and birds) for each visual system, assuming a Weber fraction of 0.05 because double cones used for luminance perception in birds and chameleons are by far the most abundant photoreceptor type in the retina [@RN16; @RN23].
 
@@
 
-Chromatic and achromatic contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN4 in combination with the package `clubSandwich` @RN5.
-For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems).
+Chromatic and luminance contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN24 in combination with the package `clubSandwich` @RN25.
+For each individual, measurements for each body region (top flank, mid-flank, tail base and gular region) were contrasted against each of six average background spectra (i.e., average top and bottom of lighter leaves, average top and bottom of darker leaves, average green and average brown stems).
 While we present sensitivity analyses for all body regions (see below - *Additional Supplemental Results*), we were mainly interested in overall signal differences displayed by chameleons.
@@
 Deriving multiple contrasts for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other.
-To ensure that non-independence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction.
-Across the different contexts (i.e., social contexts, predation, and background environments), chromatic and achromatic contrasts were analyzed in separate models as response variables.
-For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects.
-For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or bird), population of origin and their interaction as fixed effects.
-Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were observed on a native Kenyan background or their current introduced habitat background in Hawaii.
-Relevant model interactions were first tested using Wald tests.
+To ensure that non-independence of data did not impact our inferences, we included individual ID as a random effect to account for repeated measures on each individual and used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction.
+
+First, we tested whether Hawaiian chameleons were more conspicuous (as perceived by chameleons) in social contexts than Kenyan chameleons.
+We ran separate linear models with chromatic and luminance contrast of male display or courtship coloration against the local background as the response variable.
+We included population of origin and z-transformed snout-vent length (SVL) as fixed effects.
+Next, we tested whether Kenyan chameleons were more cryptic (less conspicuous against the local background) in response to predators.
+For these models, chromatic or luminance contrast against the local background (as perceived by the corresponding predator type) was the response variable and predator type (i.e, snake or bird), population of origin (Hawaii or Kenya) and their interaction were fixed effects.
+Finally, to evaluate evidence for local adaptation, we tested whether the chromatic or luminance contrast of Hawaiian chameleons during social interactions (male-male and male-female) varied depending on whether contrast was against a native Kenyan background or their current introduced habitat background in Hawaii.
+For all models, interaction terms were first tested using Wald tests.
 If found to be non-significant (p > 0.1) we dropped the interaction and fitted a main effects model.
-We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.
+We then tested the significance of the main effects using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.
 
 # Supplemental Results
+## Body Size Comparisons between Hawaii and Kenya
+```{r svl, results = "hide"}
+SVLdat <- read.table("data/SVL2.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+SVLdat$SVL2 <- with(SVLdat, scale(SVL))
+
+# Fit the model
+SVL_pop <- t.test(SVL ~ Population, data=SVLdat)
+summary(SVL_pop)
+
+# Body size
+svl <- data.frame(SVLdat %>% group_by(Population) %>% summarise(mean = mean(SVL), sd = sd(SVL), n = n()))
+
+```
+Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, N = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]`, N = `r svl[2,"n"]` mm; t = `r SVL_pop$statistic`, p <0.001).
+SVL did not significantly predict luminance contrast against the background; thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
+SVL was a weak predictor of chromatic contrast: larger chameleons had slightly higher chromatic contrast against the background.
+However, the populations did not differ in chromatic contrast during male displays (courtship and contests).
+
 ## Supporting Supplementary Tables for Main Text
 
-```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing achromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship).
-Standard errors are corrected by using robust variance estimators.
-"}
+```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "**Results of models comparing mean differences between Kenyan and Hawaiian lizards in luminance and chromatic contrast against the local background in two social contexts (male-male competition and courtship)**.
+Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI).
+Standard errors are corrected by using robust variance estimators "}
 source("./R/func.R")
@@
 
-```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing achromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards.
+```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "**Results of models comparing luminance and chromatic mean differences between Kenyan and Hawaiian lizards in response to snakes and birds**.
+Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI).
 Full models are provided and standard errors are corrected by using robust variance estimators."}
@@
 
-```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean achromatic and chormatic contrast in Hawaii and Kenya for both predators (snakes and birds).
+```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "**Planned contrasts between mean luminance and chromatic contrast in Hawaii and Kenya for both predators (snakes and birds)**.
 Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
@@
 
-```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for comparing achromatic and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (Background = Kenya) in both social contexts (Male-male competition and Courtship).
-Models were fit using robust variance estaimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.
-Significance parameter estimates are those where 95% CI does not include zero."}
+```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "**Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship)**.
+Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for Models were fit using robust variance estimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.
+Significant parameter estimates are those where the 95% CIs do not include zero."}
  tb4 <- read.csv(file = "./output/tables/Table4_MS.csv")
@@
   xlab("") + 
-  ylab("Achromatic contrast (JND)") +
+  ylab("Luminance contrast (JND)") +
   theme_bw() +
@@
 
-fulltableP <- fulltableP %>% mutate(JND = c("Achromatic Contrast", rep("",4), "Chromatc Contrast", rep("",4)),
+fulltableP <- fulltableP %>% mutate(JND = c("Luminance Contrast", rep("",4), "Chromatc Contrast", rep("",4)),
                                             p_val = sapply(p_val, function(x) p_value(x))) %>% 
@@
                     mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)),
-                           JND = c("Achromatic Contrast", rep("", 9), 
+                           JND = c("Luminance Contrast", rep("", 9), 
                                    "Chromatic Contrast", rep("", 9)))
@@
   xlab("Background Environment") + 
-  ylab("Achromatic contrast (JND)") +
+  ylab("Luminance contrast (JND)") +
   theme_bw() +
@@
               mutate(Context = c("Male-Male Display", rep("", 12), "Courtship Display", rep("", 12)),
-                     JND = c("Achromatic Contrast", rep("", 4), "Chromatic Contrast", rep("", 7), 
-                             "Achromatic Contrast", rep("", 4), "Chromatic Contrast", rep("", 7))) %>% 
+                     JND = c("Luminance Contrast", rep("", 4), "Chromatic Contrast", rep("", 7), 
+                             "Luminance Contrast", rep("", 4), "Chromatic Contrast", rep("", 7))) %>% 
                       select(Context, JND, Paramater, beta, SE, df, CI_L, CI_U)
@@
 
+## Hawaiian chameleons show greater individual change between display and bird color states than Kenyan chameleons
 # References

### 2021-08-15 ms.Rmd
Added code chunk for exporting tables back in MS. Somehow got lost down in past revisions

@@
 
+```{r table4, results='hide'}
+#Section will create and write Tables for Supplement. It is not spit out in the main MS. Exported then imported for the supp
+
+# Table S1
+    table1 <- as.data.frame(rbind(robust_ConspicDisplayAC_CI,
+                    robust_ConspicDisplayCC_CI,
+                    robust_ConspicCourtshipAC_CI,
+                    robust_ConspicCourtshipCC_CI))
+    table1$Parameter <- rownames(table1)
+    rownames(table1) <- NULL
+    # Clean
+     final_T1 <- table1 %>% 
+                 mutate(SocialContext = c("Male-Male Display", rep("", 5), "Courtship Display", rep("", 5)),
+                        VisualRealm = c("Luminance  Contrast", rep("", 2), "Chromatic Contrast", rep("", 2), 
+                                        "Luminance Contrast", rep("", 2), "Chromatic Contrast", rep("", 2))) %>%
+                  select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
+     write.csv(final_T1, file = "./output/tables/Table1_MS.csv", row.names = FALSE)
+
+# Table S2
+    table2 <- as.data.frame(rbind(robust_PredatorAC_CI_inter,
+                                  robust_PredatorCC_CI_inter))
+    table2$Parameter <- rownames(table2)
+    rownames(table2) <- NULL
+
+    # Clean
+     final_T2 <- table2 %>% 
+                 mutate(JND = c("Luminance Contrast", rep("", 3), "Chromatic Contrast", rep("", 3))) %>%
+                  select(JND, Parameter, beta, SE, df, CI_L, CI_U)
+     write.csv(final_T2, file = "./output/tables/Table2_MS.csv", row.names = FALSE)
+
+ #Table S3
+    contrast <- rbind(CI_hyp_test,
+                     CI_hyp_testAC)
+
+    contrast <- contrast %>% mutate(VisualSpec = c("Luminance Contrast", "", "Chromatic Contrast", ""),
+                                    p_val = sapply(p_val, function(x) p_value(x))) %>% 
+                select(VisualSpec, comparison, Estimate, lwr, upr, Fstat, df, p_val, p_bonf)
+
+    write.csv(contrast, file = "./output/tables/Table3_MS.csv", row.names = FALSE)
+
+#TableS4
+    tb4 <- data.frame(rbind(robust_localAdapt_DispBgrdAC,
+                 robust_DispBgrdCC,
+                 robust_CourtshipAC,
+                 robust_CourtshipCC))
+    tb4$parameter <- rownames(tb4)
+    rownames(tb4) <- NULL
+
+    tb4 <- tb4 %>% mutate(SocialContext = c("Male-Male Display", rep("", 5), 
+                                            "Courtship Display", rep("", 5)),
+                        VisualRealm = c("Luminance Contrast", rep("", 2), "Chromatic Contrast", 
+                                        rep("", 2), "Luminance Contrast", rep("", 2), 
+                                        "Chromatic Contrast", rep("", 2))) %>%
+                  select(SocialContext, VisualRealm, parameter, beta, SE, df, CI_L, CI_U)
+
+    write.csv(tb4, file = "./output/tables/Table4_MS.csv", row.names = FALSE)
+
+```
+
 # References and Notes

### 2021-08-15 ms.Rmd
Added in a fix to local adaptation dataframes. I filtered out the Kenyan animals. I broat in SVL data and merged this, and expluded any missing information. Given we are not intersted in contrasts between Kenya and Hawaii, we can ditch the pop_back models that interact populations and differences

@@
                         header=TRUE, sep=",", dec=".", strip.white=TRUE)
+svl_dat <- read.csv("./data/SVL2.csv")
 
@@
 
+# Filter out only Hawaiian and add in SVL for IDs
+DisplayBgrd <- DisplayBgrd %>% 
+               filter(Population == "Hawaii") %>% 
+               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
 ##########
@@
   # Fit the main effect model
-  DispBgrdAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
+  DispBgrdAC <- lmer(dL ~ sc_svl  + Background_pop  + (1|ID), data=DisplayBgrd)
   summary(DispBgrdAC)
@@
 
-  Population_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(2), 
+  svl_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(2), 
                                   cluster = DisplayBgrd$ID, vcov = "CR2")
@@
 
+# Filter out only Hawaiian and add in SVL for IDs
+CourtshipBgrd <- CourtshipBgrd %>% 
+               filter(Population == "Hawaii") %>% 
+               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+               filter(!is.na(SVL))
+
 ##########
@@
   # Fit main effects model
-    CourtshipAC <- lmer(dL ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
+    CourtshipAC <- lmer(dL ~ sc_svl  + Background_pop  + (1|ID), data=CourtshipBgrd)
     summary(CourtshipAC)
@@
 
-           Population_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(2), 
+           svl_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(2), 
                                   cluster = CourtshipBgrd$ID, vcov = "CR2")
@@
   # Fit main effect model
-      CourtshipCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=CourtshipBgrd)
+      CourtshipCC <- lmer(dS ~ sc_svl  + Background_pop  + (1|ID), data=CourtshipBgrd)
       summary(CourtshipCC)
@@
 
-           Population_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(2), 
+           svl_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(2), 
                                   cluster = CourtshipBgrd$ID, vcov = "CR2")

### 2021-08-15 ms.Rmd
Fixed last bits. When removing pop_back. Need to remove code in L577.

@@
     # Fit the main effect model
-          DispBgrdCC <- lmer(dS ~ Population  + Background_pop  + (1|ID), data=DisplayBgrd)
+          DispBgrdCC <- lmer(dS ~ sc_svl + Background_pop  + (1|ID), data=DisplayBgrd)
           summary(DispBgrdCC)
@@
 
-        Population_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(2), 
+        svl_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(2), 
                                   cluster = DisplayBgrd$ID, vcov = "CR2")

### 2021-08-18 ms.Rmd
Removed unecessary pop_back models because we only compare Hawaii now. Fixed the main text to have corrected values form Robust variance estaimte models.

@@
 
-  # Planned contrast. Fit the model in a slightly different way
-  DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
-  DisplayBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = DisplayBgrd) 
-  summary(DisplayBgrd_pop_back)
-
-  robust_DisplayBgrd_pop_back <- clubSandwich::conf_int(DisplayBgrd_pop_back, cluster = DisplayBgrd$ID, vcov =  "CR2")
-
-  # Contrasts hypothesis tests, but use RVE 
-    result_DisplayBgrd_pop_back <- Wald_test(DisplayBgrd_pop_back, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
-    result_DisplayBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
-
-  # Grab contrasts that we need from the robust esatimtes
-  p_displ_AC <- result_DisplayBgrd_pop_back[2,8]
-   CI_dis_AC <- data.frame(Estimate=robust_DisplayBgrd_pop_back$beta[3],
-                           lwr = robust_DisplayBgrd_pop_back$CI_L[3],
-                           upr = robust_DisplayBgrd_pop_back$CI_U[3])
-
  ##########
@@
                                   cluster = DisplayBgrd$ID, vcov = "CR2")
-
-    # Planned contrasts. Refit model in slight different form
-      DisplayBgrd$pop_back <- with(DisplayBgrd, interaction(Population, Background_pop))
-    DisplayBgrd_pop_backCC <- lmer(dS ~ pop_back + (1|ID), data = DisplayBgrd)
-
-    # Fit Robust model
-    robust_DisplayBgrd_pop_backCC <- clubSandwich::conf_int(DisplayBgrd_pop_backCC, 
-                                                            cluster = DisplayBgrd$ID, vcov =  "CR2")
-
-    # Contrasts hypothesis tests, but use RVE 
-      result_DisplayBgrd_pop_backCC <- Wald_test(DisplayBgrd_pop_backCC, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
-      result_DisplayBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_DisplayBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
-
-    # Grab contrasts that we need from the robust esatimtes
-      p_displ_CC <- result_DisplayBgrd_pop_backCC[2,8]
-       CI_Disp_CC <- data.frame(Estimate=robust_DisplayBgrd_pop_backCC$beta[3],
-                               lwr = robust_DisplayBgrd_pop_backCC$CI_L[3],
-                               upr = robust_DisplayBgrd_pop_backCC$CI_U[3])
 
@@
 
-  # Planned contrasts / hypothesis tests. Need to fit model a little differently
-    CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
-    CourtshipBgrd_pop_back <- lmer(dL ~ pop_back + (1|ID), data = CourtshipBgrd)
-
-   # Fit Robust model
-    robust_CourtshipBgrd_pop_back <- clubSandwich::conf_int(CourtshipBgrd_pop_back, cluster = CourtshipBgrd$ID, vcov =  "CR2")
-
-    # Contrasts hypothesis tests, but use RVE 
-      result_CourtshipBgrd_pop_back <- Wald_test(CourtshipBgrd_pop_back, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
-      result_CourtshipBgrd_pop_back$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_back$p_val, "bonferroni"), function(x) p_value(x))
-
-    # Grab contrasts that we need from the robust esatimtes
-      p_courtship_AC <- result_CourtshipBgrd_pop_back[2,8]
-       CI_AC_Courtship <- data.frame(Estimate=robust_CourtshipBgrd_pop_back$beta[3],
-                               lwr = robust_CourtshipBgrd_pop_back$CI_L[3],
-                               upr = robust_CourtshipBgrd_pop_back$CI_U[3])
-
 ##########
@@
                                   cluster = CourtshipBgrd$ID, vcov = "CR2")
-
-  # Planned comparions
-    CourtshipBgrd$pop_back <- with(CourtshipBgrd, interaction(Population, Background_pop))
-  CourtshipBgrd_pop_backCC <- lmer(dS ~ pop_back + (1|ID), data = CourtshipBgrd)
-
-     # Fit Robust model
-    robust_CourtshipBgrd_pop_backCC <- clubSandwich::conf_int(CourtshipBgrd_pop_backCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
-
-    # Contrasts hypothesis tests, but use RVE 
-      result_CourtshipBgrd_pop_backCC <- Wald_test(CourtshipBgrd_pop_backCC, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
-      result_CourtshipBgrd_pop_backCC$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_backCC$p_val, "bonferroni"), function(x) p_value(x))
-
-    # Grab contrasts that we need from the robust esatimtes
-      p_courtship_CC <- result_CourtshipBgrd_pop_backCC[2,8]
-       CI_CC_Courtship <- data.frame(Estimate=robust_CourtshipBgrd_pop_backCC$beta[3],
-                               lwr = robust_CourtshipBgrd_pop_backCC$CI_L[3],
-                               upr = robust_CourtshipBgrd_pop_backCC$CI_U[3])
 
@@
 Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
-Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, and for both luminance and chromatic contrast (Fig.\@ref(fig:figure4); male-male contests: luminance ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$), $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: luminance ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
+Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.\@ref(fig:figure4); male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, p = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, p = `r p_value(background_CC_Court[,6]) `).
 Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].
+
 # Current Sec

### 2021-08-18 ms.Rmd
Added new figures for Devi's new analysis. abs JND by body region and population for chromatic and luminance

@@
 
-# Fit model testing this idea above
-  Diff_dL <- lmer(dL_diff_abs ~ SVL + Population + (1|ID), data=Diff)
+# Fit model testing this idea above. First just test for body region interaction. But, needs RVE
+    Diff_dL_BR_int <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
+    summary(Diff_dL_BR_int)
+
+    # Significance of interaction
+      robust_Diff_dl_int <- Wald_test(Diff_dL_BR_int, constraints = constrain_zero(7:9), 
+                                    cluster = Diff$ID, vcov = "CR2")
+# Main effects BR
+    Diff_dL_BR_me <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
+    summary(Diff_dL_BR_me)
+
+    # Sig of Body Region
+        robust_Diff_dL_BR_me <- Wald_test(Diff_dL_BR_me, constraints = constrain_zero(4:6), 
+                                    cluster = Diff$ID, vcov = "CR2")
+
+# Then fit marginalised model so that body regions are averaged
+  Diff_dL <- lmer(dL_diff_abs ~ sc_SVL + Population + (1|ID), data=Diff)
   summary(Diff_dL)
@@
 # Refit mdoel using robust standard errors
-  robust_Diff_dl <- clubSandwich::conf_int(Diff_dL, cluster = Diff$ID, vcov =  "CR2")
+  robust_Diff_dl_BR <- clubSandwich::conf_int(Diff_dL, cluster = Diff$ID, vcov =  "CR2")
 
 # Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
+
   robust_Diff_dl_pop <- Wald_test(Diff_dL, constraints = constrain_zero(3), 
@@
 ##################
+
+  # Fit model testing this idea above. First just test for body region interaction. But, needs RVE
+    Diff_dS_BR_int <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
+    summary(Diff_dS_BR_int)
+
+    # Significance of interaction
+      robust_Diff_dS_int <- Wald_test(Diff_dS_BR_int, constraints = constrain_zero(7:9), 
+                                    cluster = Diff$ID, vcov = "CR2")
+  # Main effects BR
+    Diff_dS_BR_me <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
+    summary(Diff_dS_BR_me)
+
+    # Sig of Body Region
+      robust_Diff_dS_BR_me <- Wald_test(Diff_dS_BR_me, constraints = constrain_zero(4:6), 
+                                    cluster = Diff$ID, vcov = "CR2")
+
 # Fit model testing this idea above
@@
 
-# Current Sec
 To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; N = `r N_diff[1,2]` Hawaii, N = `r N_diff[2,2]` Kenya).
-
+While body regions varied in there overall effects (significant body region effect: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, p = `r p_value(robust_Diff_dS_BR_me[,6])`; lunminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, p = `r p_value(robust_Diff_dL_BR_me[,6])`, but no population by bodyregion interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, p = `r p_value(robust_Diff_dS_int[,6])`; lunminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, p = `r p_value(robust_Diff_dl_int[,6])`) all exhibited positive absolute JND's between display and antipreadtor color states.
+Overall, Hawaiian chameleons showed greater chromatic but not luminance change (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, p = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, p = `r p_value(robust_Diff_dl_pop[,6])`).
 Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
@@
 
+```{r}
+# Courship data
+
+##########
+# Data
+##########
+abs_display_pred <- Diff %>%
+                group_by(Population, BodyRegion) %>%
+                summarise( Mean_dS_abs = mean(dS_diff_abs, na.rm = T), 
+                             SE_dS_abs = sd(dS_diff_abs)/sqrt(length(unique(ID))),
+                             Upper_dS_abs = Mean_dS_abs + 1.96*SE_dS_abs,
+                             Lower_dS_abs = Mean_dS_abs - 1.96*SE_dS_abs, 
+                           Mean_dL_abs = mean(dL_diff_abs, na.rm = T), 
+                             SE_dL_abs = sd(dL_diff_abs)/sqrt(length(unique(ID))),
+                             Upper_dL_abs = Mean_dL_abs + 1.96*SE_dL_abs,
+                             Lower_dL_abs = Mean_dL_abs - 1.96*SE_dL_abs) %>% as.data.frame()
+
+##########
+# Plots
+##########
+
+#plots
+p1 <- ggplot(abs_display_pred, aes(x = Population, y = Mean_dS_abs, color = BodyRegion, group = BodyRegion)) +
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS_abs, ymax = Upper_dS_abs), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Population") + 
+  ylab("Chromatic contrast (JND)") +
+  theme_bw() +
+  labs(colour = "Body Region") +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+p2 <- ggplot(abs_display_pred, aes(x = Population, y = Mean_dL_abs, color = BodyRegion, group = BodyRegion)) +
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL_abs, ymax = Upper_dL_abs), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Population") + 
+  ylab("Luminance contrast (JND)") +
+  theme_bw() +
+  labs(colour = "Body Region") +
+  theme(legend.position = "right",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+#pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
+plot_grid(p1, p2, labels = c("A", "B"), label_size = 30,
+  label_fontface = "bold", rel_widths=c(1,1))
+```
+
 ```{r table4, results='hide'}

### 2021-08-18 supp.Rmd
Supp now had added new analysis plus figure by bodyrergion for chromatic and luminance change

@@
 ## Hawaiian chameleons show greater individual change between display and bird color states than Kenyan chameleons
+
+```{r}
+#########################################################
+# Comparing displays
+#########################################################
+ # Load data from Devi
+       Display_bird_diff <- read_excel("data/Display_bird_diff.xlsx")
+
+       Diff <- Display_bird_diff
+
+      names(Diff) # header rows R is reading
+
+    Diff$BodyRegion <- factor(Diff$BodyRegion) #updates levels for the variable BodyRegion
+    str(diff.difftime) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
+    Diff$ID <- factor(Diff$ID) #updates levels for the variable ID
+    str(diff.difftime) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
+
+# Center SVL to make intercept interpretable
+    Diff$sc_SVL <- with(Diff, scale(SVL, scale=FALSE))
+
+# Filter missing SVL data
+  Diff <- Diff %>% filter(!is.na(SVL)) # filter out misssing SVL data
+
+# create as factor
+    Diff$BodyRegion <- factor(Diff$BodyRegion)
+
+# Sample sizes
+    N_diff <- data.frame(Diff %>% group_by(Population) %>% summarise(n = length(unique(ID))))
+
+##################
+# Achromatic contrast
+##################
+
+# Fit model testing this idea above. First just test for body region interaction. But, needs RVE
+    Diff_dL_BR_int <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
+    summary(Diff_dL_BR_int)
+
+    # Significance of interaction
+      robust_Diff_dl_int <- Wald_test(Diff_dL_BR_int, constraints = constrain_zero(7:9), 
+                                    cluster = Diff$ID, vcov = "CR2")
+# Main effects BR
+    Diff_dL_BR_me <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
+    summary(Diff_dL_BR_me)
+
+    # Sig of Body Region
+        robust_Diff_dL_BR_me <- Wald_test(Diff_dL_BR_me, constraints = constrain_zero(4:6), 
+                                    cluster = Diff$ID, vcov = "CR2")
+
+# Then fit marginalised model so that body regions are averaged
+  Diff_dL <- lmer(dL_diff_abs ~ sc_SVL + Population + (1|ID), data=Diff)
+  summary(Diff_dL)
+
+# Check residuals of model
+ # hist(residuals(Diff_dL))
+ # plot(Diff_dL) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
+
+# Refit mdoel using robust standard errors
+  robust_Diff_dl_BR <- clubSandwich::conf_int(Diff_dL, cluster = Diff$ID, vcov =  "CR2")
+
+# Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
+
+  robust_Diff_dl_pop <- Wald_test(Diff_dL, constraints = constrain_zero(3), 
+                                  cluster = Diff$ID, vcov = "CR2")
+   robust_Diff_dl_svl <- Wald_test(Diff_dL, constraints = constrain_zero(2), 
+                                  cluster = Diff$ID, vcov = "CR2")
+
+##################
+# Chromatic contrast
+##################
+
+  # Fit model testing this idea above. First just test for body region interaction. But, needs RVE
+    Diff_dS_BR_int <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
+    summary(Diff_dS_BR_int)
+
+    # Significance of interaction
+      robust_Diff_dS_int <- Wald_test(Diff_dS_BR_int, constraints = constrain_zero(7:9), 
+                                    cluster = Diff$ID, vcov = "CR2")
+  # Main effects BR
+    Diff_dS_BR_me <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
+    summary(Diff_dS_BR_me)
+
+    # Sig of Body Region
+      robust_Diff_dS_BR_me <- Wald_test(Diff_dS_BR_me, constraints = constrain_zero(4:6), 
+                                    cluster = Diff$ID, vcov = "CR2")
+
+# Fit model testing this idea above
+  Diff_dS <- lmer(dS_diff_abs ~  SVL + Population + (1|ID), data=Diff)
+  summary(Diff_dS)
+
+# Check residuals of model
+  #hist(residuals(Diff_dS))
+  #plot(Diff_dS) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
+
+# Refit mdoel using robust standard errors
+  robust_Diff_dS <- clubSandwich::conf_int(Diff_dS, cluster = Diff$ID, vcov =  "CR2")
+
+# Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
+   robust_Diff_dS_pop <- Wald_test(Diff_dS, constraints = constrain_zero(3), 
+                                  cluster = Diff$ID, vcov = "CR2")
+   robust_Diff_dS_svl <- Wald_test(Diff_dS, constraints = constrain_zero(2), 
+                                  cluster = Diff$ID, vcov = "CR2")
+
+```
+To gauge the extent of individual color change, we calculated the absolute difference in JNDs between for each individual that had both display (male-male contest) and bird response data (N = `r N_diff[1,2]` individuals from Hawaii, N = `r N_diff[2,2]` Kenya).
+We chose these two color states to maximize sample size and because chameleons and birds have similar tetrachromatic visual systems.
+We tested whether the absolute difference in luminance or chromatic contrast between these two color states differed between populations and body regions.
+We included SVL, Population, Body region and the interaction between population and body region as fixed effects.
+ID was included as a random effect due to measures for more than one body region per individual.
+
+Body regions varied in there overall effects but mainly showed a decrease in absolute JND difference between display and antipredator color states in Kenya compared to Hawaii (Fig.
+\@ref(fig:figDisplayPred)).
+While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, p = `r p_value(robust_Diff_dS_int[,6])`; lunminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, p = `r p_value(robust_Diff_dl_int[,6])`) (Fig.
+\@ref(fig:figDisplayPred)).
+However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, p = `r p_value(robust_Diff_dS_BR_me[,6])`; lunminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, p = `r p_value(robust_Diff_dL_BR_me[,6])`) (Fig.
+\@ref(fig:figDisplayPred)).
+
+```{r, figDisplayPred, fig.align = 'center', fig.width = 13, fig.height = 12, fig.cap = " Extent of color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (N = 30) and Kenyan (N = 13) chameleons.
+**(A)** Average chromatic and **(B)** luminance contrast are provided for each bory region"}
+##########
+# Data
+##########
+abs_display_pred <- Diff %>%
+                group_by(Population, BodyRegion) %>%
+                summarise( Mean_dS_abs = mean(dS_diff_abs, na.rm = T), 
+                             SE_dS_abs = sd(dS_diff_abs)/sqrt(length(unique(ID))),
+                             Upper_dS_abs = Mean_dS_abs + 1.96*SE_dS_abs,
+                             Lower_dS_abs = Mean_dS_abs - 1.96*SE_dS_abs, 
+                           Mean_dL_abs = mean(dL_diff_abs, na.rm = T), 
+                             SE_dL_abs = sd(dL_diff_abs)/sqrt(length(unique(ID))),
+                             Upper_dL_abs = Mean_dL_abs + 1.96*SE_dL_abs,
+                             Lower_dL_abs = Mean_dL_abs - 1.96*SE_dL_abs) %>% as.data.frame()
+##########
+# Plots
+##########
+
+#plots
+p1 <- ggplot(abs_display_pred, aes(x = Population, y = Mean_dS_abs, color = BodyRegion, group = BodyRegion)) +
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS_abs, ymax = Upper_dS_abs), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Population") + 
+  ylab("Chromatic Change (absolute JND)") +
+  theme_bw() +
+  labs(colour = "Body Region") +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+p2 <- ggplot(abs_display_pred, aes(x = Population, y = Mean_dL_abs, color = BodyRegion, group = BodyRegion)) +
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL_abs, ymax = Upper_dL_abs), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Population") + 
+  ylab("Luminance Change (absolute JND)") +
+  theme_bw() +
+  labs(colour = "Body Region") +
+  theme(legend.position = c(0.80, 0.85),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold"))
+
+#pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
+plot_grid(p1, p2, labels = c("A", "B"), label_size = 30,
+  label_fontface = "bold", rel_widths=c(1,1))
+```
+
 # References

### 2021-08-18 ms.Rmd
Removed body region anlaysis from ms, added to supp and Added a new tab5 to be exported for supplement

--- 2021-08-18 ms.Rmd
@@
 
-# Fit model testing this idea above
-  Diff_dL <- lmer(dL_diff_abs ~ SVL + Population + (1|ID), data=Diff)
+# Fit model testing this idea above. First just test for body region interaction. But, needs RVE
+    Diff_dL_BR_int <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
+    summary(Diff_dL_BR_int)
+
+    # Refit mdoel using robust standard errors
+  robust_Diff_dL_BR_int <- clubSandwich::conf_int(Diff_dL_BR_int, cluster = Diff$ID, vcov =  "CR2")
+
+    # Significance of interaction
+      robust_Diff_dl_int <- Wald_test(Diff_dL_BR_int, constraints = constrain_zero(7:9), 
+                                    cluster = Diff$ID, vcov = "CR2")
+# Main effects BR
+    Diff_dL_BR_me <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
+    summary(Diff_dL_BR_me)
+
+    # Sig of Body Region
+        robust_Diff_dL_BR_me <- Wald_test(Diff_dL_BR_me, constraints = constrain_zero(4:6), 
+                                    cluster = Diff$ID, vcov = "CR2")
+
+# Then fit marginalised model so that body regions are averaged
+  Diff_dL <- lmer(dL_diff_abs ~ sc_SVL + Population + (1|ID), data=Diff)
   summary(Diff_dL)
@@
 # Check residuals of model
-  hist(residuals(Diff_dL))
-  plot(Diff_dL) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
+ # hist(residuals(Diff_dL))
+ # plot(Diff_dL) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
 
 # Refit mdoel using robust standard errors
-  robust_Diff_dl <- clubSandwich::conf_int(Diff_dL, cluster = Diff$ID, vcov =  "CR2")
+  robust_Diff_dl_BR <- clubSandwich::conf_int(Diff_dL_BR_me, cluster = Diff$ID, vcov =  "CR2")
 
 # Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
+
   robust_Diff_dl_pop <- Wald_test(Diff_dL, constraints = constrain_zero(3), 
@@
 ##################
+
+  # Fit model testing this idea above. First just test for body region interaction. But, needs RVE
+    Diff_dS_BR_int <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
+    summary(Diff_dS_BR_int)
+
+        # Refit mdoel using robust standard errors
+     robust_Diff_dS_BR_int <- clubSandwich::conf_int(Diff_dS_BR_int, cluster = Diff$ID, vcov =  "CR2")
+
+     # Significance of interaction
+      robust_Diff_dS_int <- Wald_test(Diff_dS_BR_int, constraints = constrain_zero(7:9), 
+                                    cluster = Diff$ID, vcov = "CR2")
+  # Main effects BR
+    Diff_dS_BR_me <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
+    summary(Diff_dS_BR_me)
+
+    # Sig of Body Region
+      robust_Diff_dS_BR_me <- Wald_test(Diff_dS_BR_me, constraints = constrain_zero(4:6), 
+                                    cluster = Diff$ID, vcov = "CR2")
+
 # Fit model testing this idea above
@@
 # Check residuals of model
-  hist(residuals(Diff_dS))
-  plot(Diff_dS) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
+  #hist(residuals(Diff_dS))
+  #plot(Diff_dS) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
 
 # Refit mdoel using robust standard errors
-  robust_Diff_dS <- clubSandwich::conf_int(Diff_dS, cluster = Diff$ID, vcov =  "CR2")
+  robust_Diff_dS <- clubSandwich::conf_int(Diff_dS_BR_me, cluster = Diff$ID, vcov =  "CR2")
 
@@
 
-# Current Sec
 To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; N = `r N_diff[1,2]` Hawaii, N = `r N_diff[2,2]` Kenya).
-
+While there were slight differences between body regions (see supplementary text), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, p = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, p = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (Fig.
+SX).
 Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
@@
 
+# Table S5
+    tb5 <- data.frame(rbind(robust_Diff_dl_BR,
+                      robust_Diff_dS))
+    tb5$parameter <- rownames(tb5)
+    rownames(tb5) <- NULL
+
+  tb5 <- tb5 %>% mutate(VisualRealm = c("Luminance Contrast", rep("", 5), "Chromatic Contrast", 
+                                        rep("", 5))) %>%
+                  select(VisualRealm, parameter, beta, SE, df, CI_L, CI_U)
+     write.csv(tb5, file = "./output/tables/Table5_MS.csv", row.names = FALSE)
+
 ```

### 2021-08-18 supp.Rmd
Added new Table style. INteraction models are provided in new supplementary table

@@
 
+```{r tableS9, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter.
+Full interaction models are presented for chromatic and luminance change (absolute JND)."}
+tab9 <- read.csv("./output/tables/Table5_MS.csv")
+
+Table9 <- flextable(tab9) %>% table_style13() %>% fit_to_width(max_width = 7)
+Table9
+```
+
 # References

### 2021-08-18 R/func.R

@@ -338,3 +338,31 @@ table_style12 <- function(table){
     font(fontname="Times", part = "header")
   return(table)
 }
+
+
+table_style13 <- function(table){
+  table <- table %>%
+    width(j = 3, width = 6) %>% 
+    align(align = "center", part = "header") %>% 
+    align(align = "center", part = "body") %>% 
+    flextable::compose(part = "header", j = 1, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
+    flextable::compose(part = "header", j = 2, value = as_paragraph(as_b("Parameter"))) %>% 
+    flextable::compose(part = "header", j = 3, value = as_paragraph(as_b("Estimate"))) %>% 
+    flextable::compose(part = "header", j = 4, value = as_paragraph(as_b("SE"))) %>%
+    flextable::compose(part = "header", j = 5, value = as_paragraph(as_b("df"))) %>% 
+    flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("95% CI"), as_sub(as_b("lower")))) %>% 
+    flextable::compose(part = "header", j = 7, value = as_paragraph(as_b("95% CI"), as_sub(as_b("upper")))) %>% 
+    flextable::compose(part = "body", j = 2, i = c(1, 10), value = as_paragraph("Intercept")) %>%
+    flextable::compose(part = "body", j = 2, i = c(1, 10)+1, value = as_paragraph("Snout-vent Length")) %>%
+    flextable::compose(part = "body", j = 2, i = c(1, 10)+2, value = as_paragraph("Population", as_sub(" Kenya"))) %>%
+    flextable::compose(part = "body", j = 2, i = c(1, 10)+3, value = as_paragraph("Body Region", as_sub(" midflank"))) %>%
+    flextable::compose(part = "body", j = 2, i = c(1, 10)+4, value = as_paragraph("Body Region", as_sub(" tailbase"))) %>%
+    flextable::compose(part = "body", j = 2, i = c(1, 10)+5, value = as_paragraph("Body Region", as_sub(" topflank"))) %>%
+    flextable::compose(part = "body", j = 2, i = c(1, 10)+6, value = as_paragraph("Population", as_sub(" Kenya"), "* BodyRegion", as_sub(" midflank"))) %>%
[…]

### 2021-08-18 ms.Rmd
Printing out tb4 which goes to supplement. Added to main MS

@@
 # Table S5
-    tb5 <- data.frame(rbind(robust_Diff_dl_BR,
-                      robust_Diff_dS))
+    tb5 <- data.frame(rbind(robust_Diff_dL_BR_int,
+                            robust_Diff_dS_BR_int))
     tb5$parameter <- rownames(tb5)
@@
 
-  tb5 <- tb5 %>% mutate(VisualRealm = c("Luminance Contrast", rep("", 5), "Chromatic Contrast", 
-                                        rep("", 5))) %>%
+  tb5 <- tb5 %>% mutate(VisualRealm = c("Luminance Contrast", rep("", 8), "Chromatic Contrast", 
+                                        rep("", 8))) %>%
                   select(VisualRealm, parameter, beta, SE, df, CI_L, CI_U)

### 2021-08-18 supp.Rmd
Added new spec sensitity fig 6. Hid some results to avoid spitting output out and just modified text of Devi's new analysis

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(flextable, lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich)
+pacman::p_load(flextable, lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, readxl, pavo)
 source("./R/func.R")
@@
 
-```{r Figcolsig6SPec, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Spectral sensitivities of chameleons (A), snake (B) and bird (C) used for visual models (details in text)**."}
-figs5 <- image_read("./output/figures/FigS5.jpg")
-figs5
+```{r Figcolsig6SPec, fig.align = 'center', fig.width = 12, fig.height = 10, fig.cap = "**Spectral sensitivities of chameleons (A), bird (B) and snake (C) used for visual models (details in text)**."}
+sens_cham_c_dilepis <- read_excel("./data/Spectral sensitivities.xls", sheet = "C_dilepis")[,1:6]
+        sens_birdUV <- read_excel("./data/Spectral sensitivities.xls", sheet = "birdUV")
+         sens_snake <- read_excel("./data/Spectral sensitivities.xls", sheet = "snake")[,1:4]
+
+cham <- sens_cham_c_dilepis %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
+bird <- sens_birdUV %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
+snake <- sens_snake %>% pivot_longer(cols = c(2:4)) %>% select(wavelength, name, value) %>%  data.frame()
+
+plot_spec <- function(data, tri = FALSE){
+        data %>% ggplot(aes(x = wavelength, y = value)) + 
+      geom_line(aes(color = name), size = 1.2) + 
+      theme_bw() +
+    scale_colour_manual(values = rev(c("#8D25CC", "#3770C9", "#93C572", "#FFA500", "#DBD8D6"))) +
+      labs(x = "Wavelength (nm)", y = "Absorbance", color = "") + 
+        theme(legend.position  = c(0.90, 0.65),
+            panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.text = element_text(size = 16),
+            axis.title = element_text(size = 24),
+            legend.text = element_text(size = 16),
+            legend.title = element_text(size = 20))}
+
+p1 <- plot_spec(cham) + theme(legend.position  = "none")
+p2 <- plot_spec(bird) 
+p3 <- plot_spec(snake) + scale_colour_manual(values = rev(c("#8D25CC", "#93C572", "#FFA500"))) + theme(legend.position  = "none")
+
+plot_grid(p1, p2,p3,nrow = 3, ncol = 1, labels = c("A", "B", "C"), label_size = 30,
+  label_fontface = "bold", rel_widths=c(1,1))
+
 ```
@@
 
-```{r}
+```{r, results='hide'}
 #########################################################
@@
 \@ref(fig:figDisplayPred)).
-While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, p = `r p_value(robust_Diff_dS_int[,6])`; lunminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, p = `r p_value(robust_Diff_dl_int[,6])`) (Fig.
-\@ref(fig:figDisplayPred)).
-However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, p = `r p_value(robust_Diff_dS_BR_me[,6])`; lunminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, p = `r p_value(robust_Diff_dL_BR_me[,6])`) (Fig.
-\@ref(fig:figDisplayPred)).
-
-```{r, figDisplayPred, fig.align = 'center', fig.width = 13, fig.height = 12, fig.cap = " Extent of color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (N = 30) and Kenyan (N = 13) chameleons.
+While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, p = `r p_value(robust_Diff_dS_int[,6])`; luminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, p = `r p_value(robust_Diff_dl_int[,6])`) (Fig.
+\@ref(fig:figDisplayPred) & Table \@ref(tab:tableS9)).
+However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, p = `r p_value(robust_Diff_dS_BR_me[,6])`; luminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, p = `r p_value(robust_Diff_dL_BR_me[,6])`) (Fig.
+\@ref(fig:figDisplayPred); Table \@ref(tab:tableS9)).
+
+```{r, figDisplayPred, fig.align = 'center', fig.width = 13, fig.height = 7, fig.cap = " Extent of color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (N = 30) and Kenyan (N = 13) chameleons.
 **(A)** Average chromatic and **(B)** luminance contrast are provided for each bory region"}

### 2021-08-18 R/func.R
Fixed name of table S5 to make sure this is SVL and not population Kenya.

@@ -266,7 +266,7 @@ table_style9 <- function(table){
     flextable::compose(part = "header", j = 7, value = as_paragraph(as_b("95% CI"), as_sub(as_b(" lower")))) %>% 
     flextable::compose(part = "header", j = 8, value = as_paragraph(as_b("95% CI"), as_sub(as_b(" upper")))) %>% 
     flextable::compose(part = "body", j = 3, i = c(1,4,7,10), value = as_paragraph("Intercept")) %>%
-    flextable::compose(part = "body", j = 3, i = c(1,4,7,10)+1, value = as_paragraph("Population", as_sub(" Kenya"))) %>%
+    flextable::compose(part = "body", j = 3, i = c(1,4,7,10)+1, value = as_paragraph("Snout-vent length (SVL)")) %>%
     flextable::compose(part = "body", j = 3, i = c(1,4,7,10)+2, value = as_paragraph("Background", as_sub(" Kenya"))) %>% 
     font(fontname="Times", part = "body") %>% 
     font(fontname="Times", part = "header")

### 2021-08-18 supp.Rmd
Added in SVL, chromatic still has contrasts table

@@
 
-# We only really want to compar the Hawaiian population, so we can subset that data
-  DisplayBgrd$pop_back_region <- with(DisplayBgrd, interaction(Population, Background_pop, BodyRegion))
-           DisplayBgrd_Hawaii <- DisplayBgrd %>% filter(Population == "Hawaii")
+# Filter out only Hawaiian and add in SVL for IDs
+DisplayBgrd$pop_back_region <- with(DisplayBgrd, interaction(Background_pop, BodyRegion))
+DisplayBgrd_Hawaii <- DisplayBgrd %>% 
+               filter(Population == "Hawaii") %>% 
+               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
 ##########
@@
   # Main effercts
-   DispBgrdAC_main <- lmer(dL ~  Background_pop + BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
+   DispBgrdAC_main <- lmer(dL ~  sc_svl + Background_pop + BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
     summary(DispBgrdAC_main)
@@
   # Add Wald test to test signifiacnce of each factor using roust variance 
-  background_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(2), 
+  background_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(3), 
                                   cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
 
-  BodyRegion_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(3:5), 
+  BodyRegion_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(4:6), 
                                   cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
@@
      # Fit the inetarction model
-          DispBgrdCC_inter <- lmer(dS ~ Background_pop  + BodyRegion + Background_pop:BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
+          DispBgrdCC_inter <- lmer(dS ~ sc_svl + Background_pop  + BodyRegion + Background_pop:BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
           summary(DispBgrdCC_inter)
@@
     # Fit the main effect model
-          DispBgrdCC_main <- lmer(dS ~ Background_pop  + BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
+          DispBgrdCC_main <- lmer(dS ~ sc_svl + Background_pop  + BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
           summary(DispBgrdCC_main)
@@
     # Wald test to see if interaction is important. Yes, body regions way in how they change from a Hawaiian and Kenyan population. Fig S4, this makes sense. 
-        intearction_CC_Disp <- Wald_test(DispBgrdCC_inter, constraints = constrain_zero(6:8), 
+        intearction_CC_Disp <- Wald_test(DispBgrdCC_inter, constraints = constrain_zero(7:9), 
                                   cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
@@
       robust_DispBgrdCC <- clubSandwich::conf_int(DispBgrdCC_inter, cluster = DisplayBgrd_Hawaii$ID, vcov =  "CR2")
-
-    # Get contrasts to understand the specific changes better.
-Refit model in slight different form.
-This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
-The model above is just a contrast-based model.
-    DisplayBgrd_pop_back_regionCC <- lmer(dS ~ -1 + pop_back_region + (1|ID), data = DisplayBgrd_Hawaii)
-
-    # Fit Robust model
-    robust_DisplayBgrd_pop_back_regionCC <- clubSandwich::conf_int(DisplayBgrd_pop_back_regionCC, 
-                                                            cluster = DisplayBgrd_Hawaii$ID, 
-                                                            vcov =  "CR2")
-
-    # Contrasts hypothesis tests, but use RVE 
-      result_DisplayBgrd_pop_back_regionCC <- Wald_test(DisplayBgrd_pop_back_regionCC, 
-                                                 constraints = constrain_pairwise(c(1:8)), 
-                                                 vcov="CR2", tidy = TRUE)
-
-      result_DisplayBgrd_pop_back_regionCC$p_val_bonfer <- 
-                                  sapply(p.adjust(result_DisplayBgrd_pop_back_regionCC$p_val, "bonferroni"), 
-                                                                  function(x) p_value(x))
-     # Grab contrasts that we need from the robust esatimtes
-      string <- c("pop_back_regionHawaii.Kenya.gular - pop_back_regionHawaii.Hawaii.gular|pop_back_regionHawaii.Kenya.midflank - pop_back_regionHawaii.Hawaii.midflank|pop_back_regionHawaii.Kenya.tailbase - pop_back_regionHawaii.Hawaii.tailbase|pop_back_regionHawaii.Kenya.topflank - pop_back_regionHawaii.Hawaii.topflank")
-      result_DisplayBgrd_pop_back_regionCC <- result_DisplayBgrd_pop_back_regionCC[grep(string, result_DisplayBgrd_pop_back_regionCC$hypothesis),]
 
@@
 
-# Just subset out the Hawaiian data as we are only intersted in Hawaii on different backgrounds
-  CourtshipBgrd$pop_back_BR <- with(CourtshipBgrd, interaction(Population, Background_pop, BodyRegion))
-         CourtshipBgrd_Hawaii <- CourtshipBgrd %>% filter(Population == "Hawaii")
+# Filter out only Hawaiian and add in SVL for IDs
+CourtshipBgrd$pop_back_region <- with(CourtshipBgrd, interaction(Background_pop, BodyRegion))
+CourtshipBgrd_Hawaii <- CourtshipBgrd %>% 
+               filter(Population == "Hawaii") %>% 
+               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+               filter(!is.na(SVL))
+
 ##########
@@
   # Fit main effects model
-    CourtshipAC_main <- lmer(dL ~ Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
+    CourtshipAC_main <- lmer(dL ~ sc_svl + Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
     summary(CourtshipAC_main)
@@
 
-           BodyRegion_AC_Court <- Wald_test(CourtshipAC_main, constraints = constrain_zero(3:5), 
+           BodyRegion_AC_Court <- Wald_test(CourtshipAC_main, constraints = constrain_zero(4:6), 
                                   cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
@@
   # Fit main effect model
-      CourtshipCC_inter <- lmer(dS ~ Background_pop + BodyRegion + Background_pop:BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
+      CourtshipCC_inter <- lmer(dS ~ sc_svl + Background_pop + BodyRegion + Background_pop:BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
       summary(CourtshipCC_inter)
@@
   # Fit main effect model
-      CourtshipCC_main <- lmer(dS ~ Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
+      CourtshipCC_main <- lmer(dS ~ sc_svl + Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
       summary(CourtshipCC_main)
@@
   # Wald test of interaction; Not important, so we'll simplify and assume just additive effects. This makes sense when we look at the raw data.
-  intearction_CC_Court <- Wald_test(CourtshipCC_inter, constraints = constrain_zero(6:8), 
+  intearction_CC_Court <- Wald_test(CourtshipCC_inter, constraints = constrain_zero(7:9), 
                                   cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
@@
    # Wald tests
-           background_CC_Court <- Wald_test(CourtshipCC_main, constraints = constrain_zero(2), 
+           background_CC_Court <- Wald_test(CourtshipCC_main, constraints = constrain_zero(3), 
                                   cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
 
-           BodyRegion_CC_Court <- Wald_test(CourtshipCC_main, constraints = constrain_zero(3:5), 
+           BodyRegion_CC_Court <- Wald_test(CourtshipCC_main, constraints = constrain_zero(4:6), 
                                   cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
@@
 
-          CourtshipBgrd_pop_back_BR_CC <- lmer(dS ~ -1 + pop_back_BR + (1|ID), data = CourtshipBgrd_Hawaii)
+          CourtshipBgrd_pop_back_BR_CC <- lmer(dS ~ -1 +  sc_svl + pop_back_BR + (1|ID), data = CourtshipBgrd_Hawaii)
 

### 2021-08-18 supp.Rmd
Fixed up contrasts for local adaptation. Included SVL, dropped population. Only Hawaii now.

@@
 
+ # Get contrasts to understand the specific changes better.
+Refit model in slight different form.
+This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
+The model above is just a contrast-based model.
+    DisplayBgrd_pop_back_regionCC <- lmer(dS ~ -1 + sc_svl + pop_back_region + (1|ID), data = DisplayBgrd_Hawaii) 
+    # Fit Robust model
+    robust_DisplayBgrd_pop_back_regionCC <- clubSandwich::conf_int(DisplayBgrd_pop_back_regionCC, 
+                                                            cluster = DisplayBgrd_Hawaii$ID, 
+                                                            vcov =  "CR2")
+
+    # Contrasts hypothesis tests, but use RVE 
+      result_DisplayBgrd_pop_back_regionCC <- Wald_test(DisplayBgrd_pop_back_regionCC, 
+                                                 constraints = constrain_pairwise(c(1:9)), 
+                                                 vcov="CR2", tidy = TRUE)
+
+      result_DisplayBgrd_pop_back_regionCC$p_val_bonfer <- 
+                                  sapply(p.adjust(result_DisplayBgrd_pop_back_regionCC$p_val, "bonferroni"), 
+                                                                  function(x) p_value(x))
+     # Grab contrasts that we need from the robust esatimtes
+      string <- c("pop_back_regionKenya.gular - pop_back_regionHawaii.gular|pop_back_regionKenya.midflank - pop_back_regionHawaii.midflank|pop_back_regionKenya.tailbase - pop_back_regionHawaii.tailbase|pop_back_regionKenya.topflank - pop_back_regionHawaii.topflank")
+      result_DisplayBgrd_pop_back_regionCC <- result_DisplayBgrd_pop_back_regionCC[grep(string, result_DisplayBgrd_pop_back_regionCC$hypothesis),]
 ###################
@@
           result_CourtshipBgrd_pop_back_BR_CC <- Wald_test(CourtshipBgrd_pop_back_BR_CC, 
-                                                   constraints = constrain_pairwise(c(1:8)),
+                                                   constraints = constrain_pairwise(c(1:9)),
                                                    cluster = CourtshipBgrd_Hawaii$ID,
@@
        # Grab contrasts that we need from the robust esatimtes
-      string <- c("pop_back_BRHawaii.Kenya.gular - pop_back_BRHawaii.Hawaii.gular|pop_back_BRHawaii.Kenya.midflank - pop_back_BRHawaii.Hawaii.midflank|pop_back_BRHawaii.Kenya.tailbase - pop_back_BRHawaii.Hawaii.tailbase|pop_back_BRHawaii.Kenya.topflank - pop_back_BRHawaii.Hawaii.topflank")
+      string <- c("pop_back_regionKenya.gular - pop_back_regionHawaii.gular|pop_back_regionKenya.midflank - pop_back_regionHawaii.midflank|pop_back_regionKenya.tailbase - pop_back_regionHawaii.tailbase|pop_back_regionKenya.topflank - pop_back_regionHawaii.topflank")
 

### 2021-08-19 R/func.R
Fixed Table S6 so that * is used for interaction

@@ -133,6 +133,8 @@ table_style3 <- function(table){
     flextable::compose(part = "header", j = 4, value = as_paragraph(as_b("df"), as_sub(as_b(" num")))) %>%
     flextable::compose(part = "header", j = 5, value = as_paragraph(as_b("df"), as_sub(as_b(" denom")))) %>% 
     flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("p"))) %>% 
+    flextable::compose(part = "body", j = 2, i = c(4,9), value = as_paragraph("Predator", "* BodyRegion")) %>% 
+    flextable::compose(part = "body", j = 2, i = c(4,9)+1, value = as_paragraph("Predator", "* Population")) %>% 
     font(fontname="Times", part = "body") %>% 
     font(fontname="Times", part = "header")
   return(table)

### 2021-08-19 supp.Rmd
Added in contrast results for bodyregion comparions for Devi's analysis

@@
 # Center SVL to make intercept interpretable
-    Diff$sc_SVL <- with(Diff, scale(SVL, scale=FALSE))
+    Diff$sc_SVL <- with(Diff, as.vector(scale(SVL, scale=FALSE)))
 
@@
     Diff$BodyRegion <- factor(Diff$BodyRegion)
+# Add interaction for contrasts
+    Diff$Pop_BodReg <- with(Diff, interaction(Population, BodyRegion))
 
@@
 
+ # Get contrasts to understand the specific changes better.
+Refit model in slight different form.
+This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
+The model above is just a contrast-based model.
+
+          Diff_pop_back_BR_AC <- lmer(dL_diff_abs ~ -1 +  sc_SVL + Pop_BodReg + (1|ID), data = Diff)
+
+     # Fit Robust model
+          robust_Diff_pop_back_BR_AC <- clubSandwich::conf_int(Diff_pop_back_BR_AC, 
+                                                              cluster = Diff$ID, vcov =  "CR2")
+
+     # Contrasts hypothesis tests, but use RVE 
+          result_robust_Diff_pop_back_BR_AC <- Wald_test(Diff_pop_back_BR_AC, 
+                                                   constraints = constrain_pairwise(c(2:9)),
+                                                   cluster = Diff$ID,
+                                                   vcov="CR2", tidy = TRUE)
+             result_robust_Diff_pop_back_BR_AC$p_val_bonfer <- sapply(p.adjust(result_robust_Diff_pop_back_BR_AC$p_val, "bonferroni"), 
+                                                                    function(x) p_value(x))
+
+       # Grab contrasts that we need from the robust esatimtes
+      string <- c("Pop_BodRegKenya.gular - Pop_BodRegHawaii.gular|Pop_BodRegKenya.midflank - Pop_BodRegHawaii.midflank|Pop_BodRegKenya.tailbase - Pop_BodRegHawaii.tailbase|Pop_BodRegKenya.topflank - Pop_BodRegHawaii.topflank")
+
+         result_robust_Diff_pop_back_BR_AC <- result_robust_Diff_pop_back_BR_AC[
+                                            grep(string, result_robust_Diff_pop_back_BR_AC$hypothesis),]
+
 ##################
@@
                                   cluster = Diff$ID, vcov = "CR2")
+
+   # Get contrasts to understand the specific changes better.
+Refit model in slight different form.
+This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
+The model above is just a contrast-based model.
+
+          Diff_pop_back_BR_CC <- lmer(dS_diff_abs ~ -1 +  sc_SVL + Pop_BodReg + (1|ID), data = Diff)
+
+     # Fit Robust model
+          robust_Diff_pop_back_BR_CC <- clubSandwich::conf_int(Diff_pop_back_BR_CC, 
+                                                              cluster = Diff$ID, vcov =  "CR2")
+
+     # Contrasts hypothesis tests, but use RVE 
+          result_robust_Diff_pop_back_BR_CC <- Wald_test(Diff_pop_back_BR_CC, 
+                                                   constraints = constrain_pairwise(c(2:9)),
+                                                   cluster = Diff$ID,
+                                                   vcov="CR2", tidy = TRUE)
+             result_robust_Diff_pop_back_BR_CC$p_val_bonfer <- sapply(p.adjust(result_robust_Diff_pop_back_BR_CC$p_val, "bonferroni"), 
+                                                                    function(x) p_value(x))
+
+       # Grab contrasts that we need from the robust esatimtes
+      string <- c("Pop_BodRegKenya.gular - Pop_BodRegHawaii.gular|Pop_BodRegKenya.midflank - Pop_BodRegHawaii.midflank|Pop_BodRegKenya.tailbase - Pop_BodRegHawaii.tailbase|Pop_BodRegKenya.topflank - Pop_BodRegHawaii.topflank")
+
+         result_robust_Diff_pop_back_BR_CC <- result_robust_Diff_pop_back_BR_CC[
+                                            grep(string, result_robust_Diff_pop_back_BR_CC$hypothesis),]
 

### 2021-08-19 R/func.R
Added new table formatting function

@@ -368,3 +368,29 @@ table_style13 <- function(table){
   
   return(table)
 }
+
+
+table_style14  <- function(table){
+  table <- table %>% 
+    align(align = "center", part = "header") %>% 
+    align(align = "center", part = "body") %>% 
+    flextable::compose(part = "header", j = 1, value = as_paragraph(as_b("Visual Spectrum"))) %>%
+    flextable::compose(part = "header", j = 2, value = as_paragraph(as_b("Contrast"))) %>%
+    flextable::compose(part = "header", j = 3, value = as_paragraph(as_b("F"))) %>%
+    flextable::compose(part = "header", j = 4, value = as_paragraph(as_b("DF"))) %>% 
+    flextable::compose(part = "header", j = 5, value = as_paragraph(as_b("p"))) %>% 
+    flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("p"), as_sub(as_b(" Bonferroni")))) %>%
+    flextable::compose(part = "body", i = c(1,5), j = 2, 
+                       value = as_paragraph("Kenyan", as_sub(" gular"), " - ", "Hawaiian", as_sub(" gular")))  %>% 
+    flextable::compose(part = "body", i = c(2,6), j = 2, 
+                       value = as_paragraph("Kenyan", as_sub(" midflank"), " - ", "Hawaiian", as_sub(" midflank"))) %>% 
+    flextable::compose(part = "body", i = c(3,7), j = 2, 
+                       value = as_paragraph("Kenyan", as_sub(" tailbase"), " - ", "Hawaiian", as_sub(" tailbase"))) %>% 
+    flextable::compose(part = "body", i = c(4,8), j = 2, 
+                       value = as_paragraph("Kenyan", as_sub(" topflank"), " - ", "Hawaiian", as_sub(" topflank"))) %>% 
+    font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header") %>% 
+    width(j = 1, width = 1) %>% width(j = 2, width = 8)
+  
+  return(table)
+}

### 2021-08-19 supp.Rmd
Added new contrast table for devi's analysis

@@
 However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, p = `r p_value(robust_Diff_dS_BR_me[,6])`; luminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, p = `r p_value(robust_Diff_dL_BR_me[,6])`) (Fig.
-\@ref(fig:figDisplayPred); Table \@ref(tab:tableS9)).
+\@ref(fig:figDisplayPred); Table \@ref(tab:tableS9) & Table \@ref(tab:tableS10)).
 
@@
 
+```{r tableS10, tab.cap = "Pairwise Wald F tests for chromatic and luminance change between different body regions for color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (N = 30) and Kenyan (N = 13) chameleons.
+F statistic, degrees of freedom (DF) and p-values are provided (p).
+A bonferroni correction to p-value is also provided to control for multiple testing."}
+
+tabS10 <- rbind(result_robust_Diff_pop_back_BR_AC, result_robust_Diff_pop_back_BR_CC)
+
+tabS10 <- tabS10 %>% mutate(VisRealm = c("Luminance Change", rep("",3), "Chromatic Change", rep("",3))) %>% 
+                        select(VisRealm, hypothesis, Fstat, df_denom, p_val, p_val_bonfer, -test, -delta, -df_num)
+
+Tables10 <- flextable(tabS10) %>% table_style14() %>% fit_to_width(max_width = 7)
+Tables10
+```
+
 # References

### 2021-08-20 ms.Rmd
Changed to read PNG
Checekd all figures from Supp match, and added Supp tables.

@@
 To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Fig.\@ref(fig:figure1), fig.
-S1, S4).
+S4).
 We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, N = `r N_display[1,2]` Hawaii, N = `r N_display[2,2]` Kenya), courtship (N = `r N_Courtship[1,2]` Hawaii, N = `r N_Courtship[2,2]` Kenya) and in response to bird (N = `r N_Pred[1,3]` Hawaii, N = `r N_Pred[2,3]` Kenya) and snake (N = `r N_Pred[3,3]` Hawaii, N = `r N_Pred[4,3]` Kenya) predators (Fig.\@ref(fig:figure2), fig.
@@
 
-In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.\@ref(fig:figure2), Fig.\@ref(fig:figure3)A) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`).
-This result was consistent for all body regions measured (no significant interaction with body region; supplementary text).
-However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`).
+In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.\@ref(fig:figure2), Fig.\@ref(fig:figure3)A, table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`).
+This result was consistent for all body regions measured (supplementary text).
+However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.\@ref(fig:figure3)B & table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`).
 
@@
 Released from predation, Hawaiian chameleons were less cryptic than Kenyan chameleons when threatened by both bird and snake predators.
-Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type).
-Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
-Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`, JND values > 5).
-By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.\@ref(fig:figure3)D; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; JND values ~ 2).
+Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, table S3).
+Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, table S3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
+Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D & table S4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`, JND values > 5).
+By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.\@ref(fig:figure3)D & table S4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; JND values ~ 2).
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18].
@@
 Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
-Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.\@ref(fig:figure4); male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, p = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, p = `r p_value(background_CC_Court[,6]) `).
+Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for chromatic were much larger than for luminance contrast (Fig.\@ref(fig:figure4) & table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, p = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, p = `r p_value(background_CC_Court[,6]) `).
 Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].
@@
 To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; N = `r N_diff[1,2]` Hawaii, N = `r N_diff[2,2]` Kenya).
-While there were slight differences between body regions (see supplementary text), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, p = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, p = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (Fig.
-SX).
+While there was evidence for differences between body regions (see supplementary text, fig.S9, table S12,13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, p = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, p = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (fig.
+S9).
 Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
@@
 library(magick)
-img <- image_read_pdf("./output/figures/Figure3.pdf")
+img <- image_read("./output/figures/Figure3.png")
 img
@@
 
-    contrast <- contrast %>% mutate(VisualSpec = c("Luminance Contrast", "", "Chromatic Contrast", ""),
+    contrast <- contrast %>% mutate(VisualSpec = c("Chromatic Contrast", "", "Luminance Contrast", ""),
                                     p_val = sapply(p_val, function(x) p_value(x))) %>% 

### 2021-08-20 supp.Rmd
Updated supp

@@
 ---
-title: "Supplemental Materials, Methods and Results"
-author: Whiting et al. 
+title: "Invasive chameleons released from predation display more conspicuous colors"
 date: "`r Sys.Date()`"
@@
 
-          CourtshipBgrd_pop_back_BR_CC <- lmer(dS ~ -1 +  sc_svl + pop_back_BR + (1|ID), data = CourtshipBgrd_Hawaii)
+          CourtshipBgrd_pop_back_BR_CC <- lmer(dS ~ -1 +  sc_svl + pop_back_region + (1|ID), data = CourtshipBgrd_Hawaii)
 
@@
 
-# Invasive chameleons released from predation display more conspicuous colors 
 Martin J. Whiting, Brenden S. Holland, J. Scott Keogh, Daniel W.A.Noble, Katrina J. Rankin and Devi M. Stuart-Fox
@@
 
-The file includes:
+**The file includes**:
+
 Supplementary text  
-Figures S1 -- Sx
-Tables S1 -- Sx
+Figures S1 -- S9  
+Tables S1 -- S13  
 Supplementary References
@@
 
-We provided an expanded table of hypotheses and associated predictions. We also identify the data needed to test these predictions.
+We provide an expanded table of hypotheses and associated predictions. We also identify the data needed to test these predictions.
 
@@
 \@ref(fig:Figcolsig)B and Fig.
-3) and flee.
+\@ref(fig:Figcolsig3)) and flee.
 In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
@@
 
-```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "**Results of models comparing mean differences between Kenyan and Hawaiian lizards in luminance and chromatic contrast against the local background in two social contexts (male-male competition and courtship)**.
+```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing mean differences between Kenyan and Hawaiian lizards in luminance and chromatic contrast against the local background in two social contexts (male-male competition and courtship).
 Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI).
@@
 
-```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "**Results of models comparing luminance and chromatic mean differences between Kenyan and Hawaiian lizards in response to snakes and birds**.
+```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic mean differences between Kenyan and Hawaiian lizards in response to snakes and birds.
 Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI).
@@
 
-```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "**Planned contrasts between mean luminance and chromatic contrast in Hawaii and Kenya for both predators (snakes and birds)**.
+```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean luminance and chromatic contrast in Hawaii and Kenya for both predators (snakes and birds).
 Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
@@
 
-```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "**Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship)**.
+```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
 Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for Models were fit using robust variance estimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.

### 2021-08-20 ms.Rmd
Fixed few errors in main MS

@@
 Their different color states can appear more or less conspicuous depending on the context, background, and viewer (e.g., conspecific or predator).
-If introduced chameleons in Hawaii experience character release from reduced predation, we expect Hawaiian chameleons more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators.
+If introduced chameleons in Hawaii experience character release from reduced predation, we expect Hawaiian chameleons to have more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators.
 We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.
 
-To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Fig.\@ref(fig:figure1), fig.
+To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (fig.
 S4).
@@
 Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
-Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for chromatic were much larger than for luminance contrast (Fig.\@ref(fig:figure4) & table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, p = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, p = `r p_value(background_CC_Court[,6]) `).
+Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.\@ref(fig:figure4) & table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, p = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, p = `r p_value(background_CC_Court[,6]) `).
 Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].

### 2021-08-20 supp.Rmd
Fixed a bunch of supp errors. Mainly spelling and grammar. But a few mixups in the referencing to figure blocks.

@@
 The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Fig.
-\@ref(fig:Figcolsig2)A,B).
+\@ref(fig:Figcolsig2)C,D).
 Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
@@
 The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
-\@ref(fig:Figcolsig2)C,D).
+\@ref(fig:Figcolsig2)A,B).
 Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.
@@
 Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
-We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand.
+We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand (Fig.
+\@ref(fig:Figcolsig3)).
 This structure was about 1.5 m high and each arm was approximately 60 cm.
 The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Fig.
-\@ref(fig:Figcolsig4)).
+\@ref(fig:Figcolsig3)).
 All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
@@
 \@ref(fig:Figcolsig)C).
+
 The day following social trials we presented each chameleon with a snake (Fig.
@@
 
-We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package ‘pavo’ @RN29.
+We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package ‘pavo’ @RN29.
 This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
@@
 
-We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region for each color state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
+We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region and color state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
 We first calculated the quantum catch of each photoreceptor under standard daylight illumination (D65; Commission internationale de l’éclairage – CIE) using the ‘vismodel’ function.
@@
 A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement @RN20.
-We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21.
+We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21 (Fig.
+\@ref(fig:Figcolsig6SPec)B).
 
 For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake @RN22.
-We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22.
+We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22 (Fig.
+\@ref(fig:Figcolsig6SPec)C).
 
@@
 ```
-Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, N = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]`, N = `r svl[2,"n"]` mm; t = `r SVL_pop$statistic`, p <0.001).
-SVL did not significantly predict luminance contrast against the background; thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
-SVL was a weak predictor of chromatic contrast: larger chameleons had slightly higher chromatic contrast against the background.
+Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, N = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]` mm, N = `r svl[2,"n"]`; t = `r SVL_pop$statistic`, p <0.001).
+SVL did not significantly predict luminance contrast against the background (Table \@ref(tab:table1)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
+SVL was a weak predictor of chromatic contrast (Table \@ref(tab:table1)): larger chameleons had slightly higher chromatic contrast against the background.
 However, the populations did not differ in chromatic contrast during male displays (courtship and contests).
@@
 
-Chromatic contrast of color signals against the background varied by predator type, population and body region (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
+Chromatic contrast of color signals against the background varied by predator type, population and body region (Figure \@ref(fig:FigS3)A; Table \@ref(tab:tableS3)).
 Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Figure \@ref(fig:FigS3)A).
-There were no significant interactions between predator type ('bird' vs 'snake') and either body region or population in achromatic contrast against the background, but a significant population and body region effect (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
+There was no significant interaction between predator type ('bird' vs 'snake') and population in luminance contrast against the background, but a marginally non-significant population and body region effect (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
 Effects did not vary by predator type (Table \@ref(tab:tableS3)).
-Nonetheless, effect sizes across body regions were in the same direction (Figure \@ref(fig:FigS3)B & Table \@ref(tab:tableS4)).
-Overall, there were significant differences between Kenya and Hawaii in achromatic contrast of the gular and tailbase regions, but not for mid- and top-flank (Table \@ref(tab:tableS5)).
-In contrast, no significant differences between Kenya and Hawaii existed across any body region, whereas significant differences existed between gular, midflank and tailbase (Table \@ref(tab:tableS6)).
+Nonetheless, effect sizes across body regions were generally in the same direction (Figure \@ref(fig:FigS3)B & Table \@ref(tab:tableS4)).
+Overall, there were significant differences between Kenya and Hawaii in luminance contrast of the gular and tailbase regions, but not for mid- and top-flank (Table \@ref(tab:tableS5)).
+In contrast, no significant differences between Kenya and Hawaii existed across any body region for snakes, whereas significant differences existed between gular, midflank and tailbase for birds (Table \@ref(tab:tableS6)).
 
@@
 
-```{r tableS5, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
+```{r tableS5, tab.cap = "Pairwise comparisons of luminance contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
 Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
@@
 
-```{r tableS6, tab.cap = "Pairwise comparisons of achormatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
+```{r tableS6, tab.cap = "Pairwise comparisons of chromatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
 Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
@@
 We tested whether different body regions of Hawaiian chameleons exhibited different contrasts between Hawaiian habitat background compared to their ancestral, Kenyan background.
-We started off testing models comparing the interaction between background environment and body region for both chromatic and achromatic contrast.
-However, for achromatic contrast, these models had poor convergence.
+We started off testing models comparing the interaction between background environment and body region for both chromatic and luminance contrast.
+However, for luminance contrast, these models had poor convergence.
 Based on the raw data (Figure \@ref(fig:FigS4)B & D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
-As such, we simplified models and only present the main effects of body region and background environment for achromatic contrast.
-However, inspection of the raw data (Figure \@ref(fig:FigS4)A & C) suggested that body regions may differ in how chromatically contrasting they are against the background environment of Hawaii and Kenya.
+As such, we simplified models and only present the main effects of body region and background environment for luminance contrast.
+However, inspection of the raw data (Figure \@ref(fig:FigS4)A & C) suggested that body regions may differ in how chromatically contrasting they were against the background environment of Hawaii and Kenya.
 In addition, our main results suggested that predators may deferentially drive chromatic signal conspicuousness, and given different body regions maybe more exposed to bird compared to snake predators, we fit an interaction and main effects model for chromatic contrast.
@@
 The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:FigS4)A & C & Table \@ref(tab:tableS7)).
-However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:FigS4)A & C), and only some body regions showed a significant decrease (Table \@ref(tab:tableS8)).
-
-The magnitude of achromatic contrast, showed a significant overall decrease in achromatic contrast when viewed by conspevcifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court$p_val)`, Figure \@ref(fig:FigS4)B & D along with Table \@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, p = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, p = `r p_value(BodyRegion_AC_Court$p_val)`, Figure \@ref(fig:FigS4)B & D along with Table \@ref(tab:tableS7)).
+However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:FigS4)A & C).
+Tailbase was the only body region showed a non-significant change (Table \@ref(tab:tableS8)).
+
+The magnitude of luminance contrast, showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court$p_val)`, Figure \@ref(fig:FigS4)B & D along with Table \@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, p = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, p = `r p_value(BodyRegion_AC_Court$p_val)`, Figure \@ref(fig:FigS4)B & D along with Table \@ref(tab:tableS7)).
 
@@
 Mean JNDs are provided along with 95% confidence intervals.
-A & B represent male-male contest displays whereas C & D represent courtship displays"}
+**(A & B)** represent male-male contest displays whereas **(C & D)** represent courtship displays"}
 # Courship data
@@
 
-```{r tableS8, tab.cap = "Pairwise Wald F tests comparing changes in chromatic contrast between different body regions for Hawaiian lizards against both Hawaiian and Kenyan backgrounds during male-male and courtship displats.
+```{r tableS8, tab.cap = "Pairwise Wald F tests comparing changes in chromatic contrast between different body regions for Hawaiian lizards against both Hawaiian and Kenyan backgrounds during male-male and courtship displays.
 F statistic, degrees of freedom (DF) and p-values are provided (p).
@@
 ```{r, figDisplayPred, fig.align = 'center', fig.width = 13, fig.height = 7, fig.cap = " Extent of color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (N = 30) and Kenyan (N = 13) chameleons.
-**(A)** Average chromatic and **(B)** luminance contrast are provided for each bory region"}
+**(A)** Average chromatic and **(B)** luminance contrast are provided for each body region"}
 ##########

### 2021-08-21 ms.Rmd
Fix to title, affiliations

@@
 ---
-title: "Rapid evolution of a conspicuous dynamic visual signal in an introduced chameleon"
+title: "Invasive chameleons released from predation display more conspicuous colors"
 bibliography: ms_refs.bib
@@
 
-Martin J. Whiting^1^, Brenden S. Holland^2^, J. Scott Keogh^3^, Daniel W.A.Noble^3^, Katrina J. Rankin^4^ and Devi M. Stuart-Fox^4^
+Martin J. Whiting^1^, Brenden S. Holland^2^, J. Scott Keogh^3^, Daniel W.A.Noble^3^, Katrina J. Rankin^4^ and Devi Stuart-Fox^4^
 
@@
 
-^4^ Department of Zoology, The University of Melbourne, Melbourne, Victoria 3010, Australia
+^4^ School of BioSciences, The University of Melbourne, Parkville, Victoria 3010, Australia
 

### 2021-08-23 supp.Rmd
Fix to title, affiliations
Fixed a few residual achormaticx

@@
 
-Martin J. Whiting, Brenden S. Holland, J. Scott Keogh, Daniel W.A.Noble, Katrina J. Rankin and Devi M. Stuart-Fox
+Martin J. Whiting, Brenden S. Holland, J. Scott Keogh, Daniel W.A.Noble, Katrina J. Rankin and Devi Stuart-Fox
 
@@
 
-```{r tablehyp, tab.cap = "Hypotheses and associated predictions tested in this paper in relation to the population of interest, the receiver visual system used in visual modelling, and the background against which the contrast of the signal was calculated for chromatic and achromatic contrast."}
+```{r tablehyp, tab.cap = "Hypotheses and associated predictions tested in this paper in relation to the population of interest, the receiver visual system used in visual modelling, and the background against which the contrast of the signal was calculated for chromatic and luminance contrast."}
   source("./R/func.R") 
@@
 
-```{r FigS3, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "All four body regions (gular, mid-flank, tailbase and topflank) and their achromatic (B) and chromatic contrast (A) to Hawaiian and Kenyan backgrounds (each population against their own background) are shown averaged across stems and leaves found in their habitat.
+```{r FigS3, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "All four body regions (gular, mid-flank, tailbase and topflank) and their luminance (B) and chromatic contrast (A) to Hawaiian and Kenyan backgrounds (each population against their own background) are shown averaged across stems and leaves found in their habitat.
 Just noticable differences (JNDs) are calculated for both bird and snake predators.
@@
 Main effects and interaction models are presented.
-Only main effect models are presented for achromatic contrast (see text for details), whereas interaction models are presented for chromatic contrast.
+Only main effect models are presented for luminance contrast (see text for details), whereas interaction models are presented for chromatic contrast.
 Pairwise contrasts are presented for chromatic contrast models (see Table 8) for ease of interpretation."}

### 2021-08-23 ms.Rmd
Fixed wording and some small formatting issues

@@
 # Abstract
-Colonization of novel environments can free introduced populations from the constraints imposed by native-range predators and promote rapid evolutionary change.
+Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid evolutionary change.
 We demonstrate such character release for color signals in Jackson’s chameleons (*Trioceros j.
 xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) where there are few lizard predators.
-Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators, and showed greater change between display and anti-predator color states.
+Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator color states.
 Consistent with local adaptation, the display colors of Hawaiian chameleons were more conspicuous in their local than ancestral habitats.
 These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
-
-# Main text
 
@@
 
-In 1972, a shipment of Jackson’s chameleons (*Trioceros jacksonii xanthopholus*; (Fig.\@ref(fig:figure1) and fig.
-S1)) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
+In 1972, a shipment of Jackson’s chameleons (*Trioceros jacksonii xanthopholus*; Fig.
+\@ref(fig:figure1) and fig.
+S1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
 The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11].
@@
 50 -- 65 generations).
-On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators.
+On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014.
 This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.
@@
 Chameleons are famous for their rapid color change [@RN12; @RN13; @RN14].
-When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green to signal dominance or during courtship (Fig.
-1, fig.
-S1, S3).
-At other times, they are a duller green and can darken to brown or become mottled green and brown, particularly in response to perceived predation risk (Fig.\@ref(fig:figure1), fig.
+When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green during courtship or to signal dominance (Fig.
+1, figs.
+S1 and S3).
+At other times, they are a duller green and can darken to brown or become mottled green and brown, particularly in response to perceived predation risk (Fig.
+\@ref(fig:figure1) and fig.
 S4A).
@@
 S4).
-We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, N = `r N_display[1,2]` Hawaii, N = `r N_display[2,2]` Kenya), courtship (N = `r N_Courtship[1,2]` Hawaii, N = `r N_Courtship[2,2]` Kenya) and in response to bird (N = `r N_Pred[1,3]` Hawaii, N = `r N_Pred[2,3]` Kenya) and snake (N = `r N_Pred[3,3]` Hawaii, N = `r N_Pred[4,3]` Kenya) predators (Fig.\@ref(fig:figure2), fig.
+We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, n = `r N_display[1,2]` Hawaii, n = `r N_display[2,2]` Kenya), courtship (n = `r N_Courtship[1,2]` Hawaii, n = `r N_Courtship[2,2]` Kenya) and in response to bird (n = `r N_Pred[1,3]` Hawaii, n = `r N_Pred[2,3]` Kenya) and snake (n = `r N_Pred[3,3]` Hawaii, n = `r N_Pred[4,3]` Kenya) predators (Fig.
+\@ref(fig:figure2) and fig.
 S5).
-We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs @RN25 against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
-
-In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.\@ref(fig:figure2), Fig.\@ref(fig:figure3)A, table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`).
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25) against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
+
+In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.
+\@ref(fig:figure2), Fig.
+\@ref(fig:figure3)A and table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, P = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, P = `r p_value(Population_ConspicCourtshipAC$p_val)`).
 This result was consistent for all body regions measured (supplementary text).
-However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.\@ref(fig:figure3)B & table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`).
-
-The importance of luminance contrast is consistent with the nature and mechanism of rapid color change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.\@ref(fig:figure1)).
+However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.
+\@ref(fig:figure3)B and table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, P = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, P = `r p_value(Population_ConspicCourtshipCC$p_val)`).
+
+The importance of luminance contrast is consistent with the nature and mechanism of rapid color change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.
+\@ref(fig:figure1)).
 Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15 or changes in the spacing of guanine crystals within iridophores @RN14, both of which strongly affect luminance.
@@
 Released from predation, Hawaiian chameleons were less cryptic than Kenyan chameleons when threatened by both bird and snake predators.
-Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, table S3).
-Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, table S3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`).
-Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D & table S4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`, JND values > 5).
-By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.\@ref(fig:figure3)D & table S4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; JND values ~ 2).
+Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.
+\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, P = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, table S3).
+Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, table S3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, P = `r p_value(back_pred_CC_Pred$p_val)`).
+Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.
+\@ref(fig:figure3)D and table S4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, P = `r CI_hyp_test[2,8]`, JND values > 5).
+By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.
+\@ref(fig:figure3)D and table S4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, P = 1; JND values ~ 2).
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18].
@@
 Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
-Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.\@ref(fig:figure4) & table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, p = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, p = `r p_value(background_CC_Court[,6]) `).
+Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.
+\@ref(fig:figure4) and table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, P = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, P = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, P = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, P = `r p_value(background_CC_Court[,6]) `).
 Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].
@@
 To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; N = `r N_diff[1,2]` Hawaii, N = `r N_diff[2,2]` Kenya).
-While there was evidence for differences between body regions (see supplementary text, fig.S9, table S12,13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, p = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, p = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (fig.
+While there was evidence for differences between body regions (see supplementary text, fig.
+S9, tables S12 and S13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, P = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, P = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (fig.
 S9).
@@
 Oahu has no snakes or raptors that feed on lizards.
-Domestic and feral cats would be their main threat on Oahu, and cats have very limited color vision [@RN22].
-In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators -- the conditions were highly favorable for character release of social signals.
-Jackson’s chameleons likely experience strong sexual selection, given their polygynous mating system, elaborate courtship displays and intense male contest competition (Fig.\@ref(fig:figure1)) in which males lock and twist their three horns into rivals during intense bouts that sometimes result in physical injury.
+The main threat on for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22].
+In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014 -- the conditions were highly favorable for character release of social signals.
+Jackson’s chameleons likely experience strong sexual selection given their polygynous mating system, elaborate courtship displays and intense male contest competition (Fig.
+\@ref(fig:figure1)) in which males lock and twist their three horns into rivals during intense bouts that sometimes result in physical injury.
 Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
+To the best of our knowledge, this is also the first example of character release in a dynamic color signal.
 
@@
 They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
-(A) A typical dominant male in display coloration.
-(B) A subordinate male that lost a contest and turned from bright yellow to brown.
-(C) Two males fighting, both are in display coloration and at a stage where they are relatively even matched.
-(D) A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male"}
+**(A)** A typical dominant male in display coloration.
+**(B)** A subordinate male that lost a contest and turned from bright yellow to brown.
+**(C)** Two males fighting, both are in display coloration and at a stage where they are relatively even matched.
+**(D)** A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male."}
 library(magick)
@@
 ```{r figure2, fig.width = 14, fig.height= 14, fig.cap="**Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya**.
-The context for measurement was (A) male contest displays; (B) courtship; (C) bird predator; and (D) snake predator.
+The context for measurement was **(A)** male contest displays; **(B)** courtship; **(C)** bird predator; and **(D)** snake predator.
 More details in online supplementary text and Fig.
-S5"}
+S5."}
 library(magick)
@@
 ```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "**Chromatic and luminance contrast in Just Noticeable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves)**.
-(A & B) Luminance and chromatic contrast of male chameleons during male-male contests and female courtship.
+Luminance and chromatic contrast of male chameleons during male-male contests **(A)** and female courtship **(B)**.
 JNDs are calculated based on the chameleon visual system.
-(B & C) Luminance and chromatic contrasts of male chameleons during snake and bird predator encounters.
+Luminance and chromatic contrasts of male chameleons during snake **(C)** and bird **(D)** predator encounters.
 JNDs are calculated based on the snake and bird visual systems.
-Contrasts between means, ‘Beta’, that are bold indicate significant effects.."}
+Contrasts between means, ‘Beta’, that are bold indicate significant effects."}
 rerun = FALSE

### 2021-08-23 ms.Rmd
Fixed a few n

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, magick, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable, readxl)
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable, readxl)
 source("./R/func.R")
@@
 
-To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; N = `r N_diff[1,2]` Hawaii, N = `r N_diff[2,2]` Kenya).
+To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; n = `r N_diff[1,2]` Hawaii, n = `r N_diff[2,2]` Kenya).
 While there was evidence for differences between body regions (see supplementary text, fig.

### 2021-08-23 ms.Rmd
P val italicised

@@
 S4).
-We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, n = `r N_display[1,2]` Hawaii, n = `r N_display[2,2]` Kenya), courtship (n = `r N_Courtship[1,2]` Hawaii, n = `r N_Courtship[2,2]` Kenya) and in response to bird (n = `r N_Pred[1,3]` Hawaii, n = `r N_Pred[2,3]` Kenya) and snake (n = `r N_Pred[3,3]` Hawaii, n = `r N_Pred[4,3]` Kenya) predators (Fig.
+We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
 \@ref(fig:figure2) and fig.
@@
 \@ref(fig:figure2), Fig.
-\@ref(fig:figure3)A and table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, P = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, P = `r p_value(Population_ConspicCourtshipAC$p_val)`).
+\@ref(fig:figure3)A and table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
 This result was consistent for all body regions measured (supplementary text).
 However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.
-\@ref(fig:figure3)B and table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, P = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, P = `r p_value(Population_ConspicCourtshipCC$p_val)`).
+\@ref(fig:figure3)B and table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
 
@@
 Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.
-\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, P = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, table S3).
-Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, table S3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, P = `r p_value(back_pred_CC_Pred$p_val)`).
+\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, *P* = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, table S3).
+Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, table S3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, *P* = `r p_value(back_pred_CC_Pred$p_val)`).
 Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.
-\@ref(fig:figure3)D and table S4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, P = `r CI_hyp_test[2,8]`, JND values > 5).
+\@ref(fig:figure3)D and table S4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`, JND values > 5).
 By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.
-\@ref(fig:figure3)D and table S4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, P = 1; JND values ~ 2).
+\@ref(fig:figure3)D and table S4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1; JND values ~ 2).
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18].
@@
 Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.
-\@ref(fig:figure4) and table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, P = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, P = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, P = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, P = `r p_value(background_CC_Court[,6]) `).
+\@ref(fig:figure4) and table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
 Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].
 
-To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; n = `r N_diff[1,2]` Hawaii, n = `r N_diff[2,2]` Kenya).
+To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
 While there was evidence for differences between body regions (see supplementary text, fig.
-S9, tables S12 and S13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, P = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, P = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (fig.
+S9, tables S12 and S13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (fig.
 S9).
@@
 Oahu has no snakes or raptors that feed on lizards.
-The main threat on for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22].
+The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22].
 In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014 -- the conditions were highly favorable for character release of social signals.

### 2021-08-24 supp.Rmd
Making supp fixes, checking the table S1 rendering given it has been updated

@@
 Correspondance: Martin J. Whiting  
-Email: Martin.Whiting@mq.edu.au
+Email: martin.whiting@mq.edu.au
 
@@
 
-We provide an expanded table of hypotheses and associated predictions. We also identify the data needed to test these predictions.
-
-```{r tablehyp, tab.cap = "Hypotheses and associated predictions tested in this paper in relation to the population of interest, the receiver visual system used in visual modelling, and the background against which the contrast of the signal was calculated for chromatic and luminance contrast."}
+We provide an expanded table of the key hypothesis and associated predictions. We also identify the data needed to test these predictions.
+
+```{r tablehyp, tab.cap = "Hypothesis and associated predictions tested in this paper in relation to the population of interest, the receiver visual system used in visual modelling, and the background against which the contrast of the signal was calculated for chromatic and luminance contrast."}
   source("./R/func.R") 

### 2021-08-24 R/func.R
Fix export of sleect.

@@ -33,11 +33,11 @@ wald_table   <- function(model_main, model_inter){
                         "Population", 
                         "Body Region", 
                         "Population:Body Region")) %>% 
-    select(Variable, Fstat, df_num, df_denom, p_val) %>% as.data.frame()
+    dplyr::select(Variable, Fstat, df_num, df_denom, p_val) %>% as.data.frame()
   
   return(WaldTab)
   }
-  
+
 wald_table_p <- function(model_main, model_inter){
     
     inter_BR_Pred <- Wald_test(model_inter, 
@@ -69,7 +69,7 @@ wald_table_p <- function(model_main, model_inter){
                           "Predator", 
                           "Predator:Body Region",
                           "Predator:Population")) %>% 
-      select(Variable, Fstat, df_num, df_denom, p_val) %>% as.data.frame()
+      dplyr::select(Variable, Fstat, df_num, df_denom, p_val) %>% as.data.frame()
     
     return(WaldTab)
   }

### 2021-08-24 R/func.R
Table10 style needed updating as SVL was now added to models

@@ -288,14 +288,15 @@ table_style10 <- function(table){
     flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("df"))) %>% 
     flextable::compose(part = "header", j = 7, value = as_paragraph(as_b("95% CI"), as_sub(as_b("lower")))) %>% 
     flextable::compose(part = "header", j = 8, value = as_paragraph(as_b("95% CI"), as_sub(as_b("upper")))) %>% 
-    flextable::compose(part = "body", j = 3, i = c(1, 6, 14, 19), value = as_paragraph("Intercept")) %>%
-    flextable::compose(part = "body", j = 3, i = c(1, 6, 14, 19)+1, value = as_paragraph("Background Population", as_sub(" Kenya"))) %>%
-    flextable::compose(part = "body", j = 3, i = c(1, 6, 14, 19)+2, value = as_paragraph("Body Region", as_sub(" midflank"))) %>%
-    flextable::compose(part = "body", j = 3, i = c(1, 6, 14, 19)+3, value = as_paragraph("Body Region", as_sub(" tailbase"))) %>%
-    flextable::compose(part = "body", j = 3, i = c(1, 6, 14, 19)+4, value = as_paragraph("Body Region", as_sub(" topflank"))) %>%
-    flextable::compose(part = "body", j = 3, i = c(11,24), value = as_paragraph("Background Population", as_sub(" Kenya"), "* BodyRegion", as_sub(" midflank"))) %>%
-    flextable::compose(part = "body", j = 3, i = c(11,24)+1, value = as_paragraph("Background Population", as_sub(" Kenya"), "* BodyRegion", as_sub(" tailbase"))) %>%
-    flextable::compose(part = "body", j = 3, i = c(11,24)+2, value = as_paragraph("Background Population", as_sub(" Kenya"), "* BodyRegion", as_sub(" topflank"))) %>% 
+    flextable::compose(part = "body", j = 3, i = c(1, 7, 16, 22), value = as_paragraph("Intercept")) %>%
+    flextable::compose(part = "body", j = 3, i = c(1, 7, 16, 22)+1, value = as_paragraph("SVL", as_sub(" z"))) %>%
+    flextable::compose(part = "body", j = 3, i = c(1, 7, 16, 22)+2, value = as_paragraph("Background Population", as_sub(" Kenya"))) %>%
[…]

### 2021-08-24 R/func.R
Fixing SVL so it's consistent across tables

@@ -153,7 +153,7 @@ table_style4 <- function(table){
     flextable::compose(part = "header", j = 7, value = as_paragraph(as_b("df"))) %>% 
     flextable::compose(part = "header", j = 8, value = as_paragraph(as_b("p"))) %>% 
     flextable::compose(part = "body", j = 3, i = c(1, 10, 19, 28), value = as_paragraph("Intercept")) %>%
-    flextable::compose(part = "body", j = 3, i = c(1, 10, 19, 28)+1, value = as_paragraph("SVL")) %>%
+    flextable::compose(part = "body", j = 3, i = c(1, 10, 19, 28)+1, value = as_paragraph("Snout-vent Length (SVL)")) %>%
     flextable::compose(part = "body", j = 3, i = c(1, 10, 19, 28)+2, value = as_paragraph("Population", as_sub(" Kenya"))) %>%
     flextable::compose(part = "body", j = 3, i = c(1, 10, 19, 28)+3, value = as_paragraph("Body Region", as_sub(" midflank"))) %>%
     flextable::compose(part = "body", j = 3, i = c(1, 10, 19, 28)+4, value = as_paragraph("Body Region", as_sub(" tailbase"))) %>%
@@ -209,7 +209,7 @@ table_style6 <- function(table){
     flextable::compose(part = "header", j = 7, value = as_paragraph(as_b("95% CI"), as_sub(as_b(" lower")))) %>% 
     flextable::compose(part = "header", j = 8, value = as_paragraph(as_b("95% CI"), as_sub(as_b(" upper")))) %>% 
     flextable::compose(part = "body", j = 3, i = c(1,4,7,10), value = as_paragraph("Intercept")) %>%
-    flextable::compose(part = "body", j = 3, i = c(1,4,7,10)+1, value = as_paragraph("SVL", as_sub(" z-transformed"))) %>%
+    flextable::compose(part = "body", j = 3, i = c(1,4,7,10)+1, value = as_paragraph("Snout-vent Length (SVL)")) %>%
     flextable::compose(part = "body", j = 3, i = c(1,4,7,10)+2, value = as_paragraph("Population", as_sub(" Kenya"))) %>% 
     font(fontname="Times", part = "body") %>% 
     font(fontname="Times", part = "header")
@@ -289,7 +289,7 @@ table_style10 <- function(table){
[…]

### 2021-08-24 supp.Rmd
Some odd formatting issues. Need to re-source .R func file for various locations as formatting seems to be getting cleared across R chunks, which is odd

@@
 Interactions were first tested and models refit to test the signifiacnce of main effects"}
-
+source("./R/func.R")
 fulltableP <- rbind(Wald_table_PredatorAC,
@@
                                             p_val = sapply(p_val, function(x) p_value(x))) %>% 
-                                      select(JND, Variable, Fstat, df_num, df_denom, p_val)
+                                      dplyr::select(JND, Variable, Fstat, df_num, df_denom, p_val)
 
@@
 Pairwise contrasts are presented for chromatic contrast models (see Table 8) for ease of interpretation."}
+source("./R/func.R")
 table7 <- rbind(robust_localAdapt_DispBgrdAC,
@@
 table7 <- table7 %>% 
-              mutate(Context = c("Male-Male Display", rep("", 12), "Courtship Display", rep("", 12)),
-                     JND = c("Luminance Contrast", rep("", 4), "Chromatic Contrast", rep("", 7), 
-                             "Luminance Contrast", rep("", 4), "Chromatic Contrast", rep("", 7))) %>% 
-                      select(Context, JND, Paramater, beta, SE, df, CI_L, CI_U)
+              mutate(Context = c("Male-Male Display", rep("", 14), "Courtship Display", rep("", 14)),
+                     JND = c("Luminance Contrast", rep("", 5), "Chromatic Contrast", rep("", 8), 
+                             "Luminance Contrast", rep("", 5), "Chromatic Contrast", rep("", 8))) %>% 
+                      dplyr::select(Context, JND, Paramater, beta, SE, df, CI_L, CI_U)
 

### 2021-08-24 R/func.R
Added in SVL abbreviation

@@ -356,7 +356,7 @@ table_style13 <- function(table){
     flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("95% CI"), as_sub(as_b("lower")))) %>% 
     flextable::compose(part = "header", j = 7, value = as_paragraph(as_b("95% CI"), as_sub(as_b("upper")))) %>% 
     flextable::compose(part = "body", j = 2, i = c(1, 10), value = as_paragraph("Intercept")) %>%
-    flextable::compose(part = "body", j = 2, i = c(1, 10)+1, value = as_paragraph("Snout-vent Length")) %>%
+    flextable::compose(part = "body", j = 2, i = c(1, 10)+1, value = as_paragraph("Snout-vent Length (SVL)")) %>%
     flextable::compose(part = "body", j = 2, i = c(1, 10)+2, value = as_paragraph("Population", as_sub(" Kenya"))) %>%
     flextable::compose(part = "body", j = 2, i = c(1, 10)+3, value = as_paragraph("Body Region", as_sub(" midflank"))) %>%
     flextable::compose(part = "body", j = 2, i = c(1, 10)+4, value = as_paragraph("Body Region", as_sub(" tailbase"))) %>%

### 2021-08-24 supp.Rmd
Added refs from main MS in front so we can get contonuous line numbers. Also added in source for func.R to Tabe chunks to ensure that everything works

@@
 
+[@RN15ms;@RN7ms;@RN11ms;@RN6ms;@RN22ms;@RN2ms;@RN25ms;@RN8ms;@RN3ms;@RN1ms;@RN13ms;@RN18ms;@RN9ms;@RN12ms;@RN26ms;@RN24ms;@RN19ms;@RN5ms;@Measey2014ms;@RN20ms;@RN14ms;@RN4ms;@RN10ms]
+
 # Expanded Hypotheses and Predictions
@@
 Standard errors were corrected using robust variance estimators and Saitterwaite degrees of freedom correction was used for calculated significance"}
+source("./R/func.R")
 robust_sc_pred_table <- rbind(robust_PredatorAC,
@@
 Wald test uses a robust variance correction on the standard errors."}
-
+source("./R/func.R")
 # Split snakes and birds
@@
 See Table 7 for model estimates."}
+
+source("./R/func.R")
 tab8 <- rbind(result_DisplayBgrd_pop_back_regionCC,
@@
 #########################################################
+source("./R/func.R")
  # Load data from Devi
@@
 Full interaction models are presented for chromatic and luminance change (absolute JND)."}
+source("./R/func.R")
 tab9 <- read.csv("./output/tables/Table5_MS.csv")
@@
 A bonferroni correction to p-value is also provided to control for multiple testing."}
-
+source("./R/func.R")
 tabS10 <- rbind(result_robust_Diff_pop_back_BR_AC, result_robust_Diff_pop_back_BR_CC)

### 2021-08-24 supp.Rmd
Chunk of Supp done

@@
 We are not aware of any evidence of bird predation of chameleons on Oahu.
-We examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons @RN1.
-Among the larger, predatory birds, three are owls (*Asio flammeus*, *Bubo scandiacus*, *Tyto alba*) that are highly unlikely to feed on chameleons because they are either mostly or entirely nocturnal, extremely rare on Oahu, or feed mostly on rodents and/or birds @RN2.
-Smaller raptors include the Merlin (*Falco columbarius*), which is considered rare and feeds on small birds in more open habitats while the peregrine falcon (*Falco peregrinus*) is a predator of medium-sized birds and also considered rare @RN2.
-There are also seven species of egret and heron, which typically have generalist diets.
-Of these, six are considered rare or accidental @RN1, are associated with water bodies, and better known for eating fishes @RN3.
+We consulted with several ornithologists in Hawaii regarding feeding behavior of the few predatory bird species that consume vertebrate prey, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons @RN1.
+Among the larger, predatory birds, two are owls (*Asio flammeus sandwichensis*, *Tyto alba*), which are extremely rare on Oahu.
+Both species are highly unlikely to feed on chameleons because they feed out in the open rather than in forest canopy where chameleons occur, and typically feed on rodents, birds and insects @RN2.
+Smaller to medium raptors include the Merlin (*Falco columbarius*) and the peregrine falcon (*Falco peregrinus*) which are both considered rare non-breeding vagrants in the Hawaiian Islands, and feed on small-medium birds in more open habitats @RN2.
+There are also eight species of egret and heron, which typically have generalist diets.
+Of these, six are considered rare or accidental @RN1, are associated with aquatic habitats, and are better known for eating fishes @RN3.
 The cattle egret (*Bubulcus ibis*) is a generalist predator that is mostly insectivorous but does eat lizards opportunistically [@RN4;@RN5].
-However, it forages on the ground in more open habitats, often following domestic animals @RN5 and therefore, is highly unlikely to encounter chameleons.
-Among other potential predators, the Hawaiian Hawk or ʻio (*Buteo solitarius*) is restricted to the Big Island while the Hen Harrier (*Circus cyaneus*) is considered an accidental or rare visitor to the Hawaiian islands.
-Finally, among corvids, the Hawaiian crow (*Corvus hawaiiensis*) is extinct in the wild while the common raven (*Corvus corax*) is rare @RN1.
+However, it forages on the ground in more open habitats, often following domestic animals and agricultural or landscaping machinery @RN5 and therefore, is not found in forested habitats and is highly unlikely to encounter chameleons.
+Among other potential predators, the Hawaiian Hawk or ʻio (*Buteo solitarius*) is restricted to the Big Island, while the Hen Harrier (*Circus cyaneus*) is considered an accidental or rare visitor to the Hawaiian islands.
+Finally, among corvids, the Hawaiian Crow (*Corvus hawaiiensis*) is extinct in the wild while the Common Raven (*Corvus corax*) is extremely rare @RN1.
 In summary, we could not identify any bird species that could be regarded as a major predator of chameleons.
@@
 
-We collected most chameleons at night by spotlighting, and a small proportion of individuals during the day.
-When chameleons were beyond reach, we used 4 m extensible poles that chameleons could be encouraged to step on to, before lowering them to the ground.
+We collected most chameleons (Fig.
+S\@ref(fig:Figcolsig2)) at night by spotlighting, and a small proportion of individuals during the day.
+When chameleons were beyond reach in the forest canopy, we used lightweight 6-m extensible graphite composite poles (South Bend Kwik Stix™ 20' Telescopic Fishing Pole), that chameleons could be encouraged to step on to, before lowering them to the ground.
 We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
@@
 Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006.
-In Hawaii, chameleons were collected from a single population in the Manoa Valley, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
+In Hawaii, chameleons were collected from a single population in the forested Ko’olau Mountains north of Honolulu, between upper Makiki Valley and the summit of Tantalus Mountain, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
 400-600 m.
 The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Fig.
-\@ref(fig:Figcolsig2)C,D).
+S\@ref(fig:Figcolsig2)C,D).
 Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
@@
 The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
-\@ref(fig:Figcolsig2)A,B).
+S\@ref(fig:Figcolsig2)A,B).
 Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.
@@
 **(B)** Two examples of males engaged in horn locking.
-Males in the bottom plane are evenly matched at this stage and both are in full display color.
+Males in the bottom panel are evenly matched at this stage and both are in full display color.
 In the top image, the male on the left is starting to change to a subordinate color.
@@
 
-```{r Figcolsig2, fig.align = 'center', fig.cap = "Representative photos of habitat in Kenya (A, B) and Hawaii (C, D).
-In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines"}
+```{r Figcolsig2, fig.align = 'center', fig.cap = "**Representative photos of habitat in Kenya (A, B) and Hawaii (C, D)**.
+In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines."}
 figs2 <- image_read("./output/figures/FigS2.jpg")
@@
 We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand (Fig.
-\@ref(fig:Figcolsig3)).
+S\@ref(fig:Figcolsig3)).
 This structure was about 1.5 m high and each arm was approximately 60 cm.
 The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Fig.
-\@ref(fig:Figcolsig3)).
+S\@ref(fig:Figcolsig3)).
 All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
@@
 ```
+
 We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
 This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Fig.
-\@ref(fig:Figcolsig)A,B).
+S\@ref(fig:Figcolsig)A,B).
 In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
-This would be accompanied by head shakes and would lead to horn-locking (Fig.
-\@ref(fig:Figcolsig)B) if males were not obviously asymmetric.
+This would be accompanied by head shakes and gapping, leading to horn-locking (Fig.
+S\@ref(fig:Figcolsig)B) if males were not obviously asymmetric.
 Once dominance was settled, the subordinate would turn brown (Fig.
-\@ref(fig:Figcolsig)B and Fig.
-\@ref(fig:Figcolsig3)) and flee.
+S\@ref(fig:Figcolsig)B and Fig.
+S\@ref(fig:Figcolsig3)) and flee.
 In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
-\@ref(fig:Figcolsig)C).
+S\@ref(fig:Figcolsig)C).
 
 The day following social trials we presented each chameleon with a snake (Fig.
-\@ref(fig:Figcolsig4)A) and bird (Fig.
-\@ref(fig:Figcolsig4)B) model predator (separately).
+S\@ref(fig:Figcolsig4)A) and bird (Fig.
+S\@ref(fig:Figcolsig4)B) model predator model (separately).
 Lizards were drawn at random and the order of presentation (snake or bird first) was random.
@@
 In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Fig.
-\@ref(fig:Figcolsig4)A).
+S\@ref(fig:Figcolsig4)A).
 In the case of the bird trials, the bird was attached to an ca.
 2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Fig.
-\@ref(fig:Figcolsig4)B).
+S\@ref(fig:Figcolsig4)B).
 Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
@@
 The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
-The bird was a mounted African cuckoo-hawk (*Aviceda cuculoides*).
+The bird was a mounted African Cuckoo-hawk (*Aviceda cuculoides*).
 
 ```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator**.
-**(A)** A model snake (*Dispholidus typus*) causing a characteristic contrasting color change and **(B)** a model bird (*Aviceda cuculoides*) which was flown over the top of the chameleon"}
+**(A)** A model snake (*Dispholidus typus*) causing a characteristic contrasting color change and **(B)** a model bird (*Aviceda cuculoides*) which was flown over the top of the chameleon."}
 figs4 <- image_read("./output/figures/FigS4.jpg")
@@
 During all behavior trials, we measured the color of each of three body regions: top flank, mid flank, tail base and gular region (Fig.
-\@ref(fig:Figcolsig5)).
+S\@ref(fig:Figcolsig5)).
 We measured each body region once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
@@
 When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
-We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal.
+We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal during measurement.
 The probe holder ensured that measurements were taken at a standard angle (45°) and distance (5 mm) for a 3 x 5 mm area and excluded all ambient light.
@@
 
-```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "Location of body regions we measured for spectral reflectance. See details in text."}
+```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Location of body regions we measured for spectral reflectance**. See details in text."}
 figs5 <- image_read("./output/figures/FigS5.jpg")
@@
 
-## Visual modelling
-
-We visualized and quality-checked all spectra, adjusted very minor negative reflectance values caused by electrical noise by lifting the curves by the maximum negative value, then smoothed all spectra with LOESS smoothing of 0.16 using the plotsmooth function in Pavo 1.0 @RN29.
-We also took the median reflectance for all leaves (top, bottom) and stems (2 measures/stem) for each population (Hawaii, Kenya), and for irradiance.
-
-We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package ‘pavo’ @RN29.
+## Visual modeling
+
+We visualized and quality-checked all spectra, adjusted very minor negative reflectance values caused by electrical noise by lifting the curves by the maximum negative value, then smoothed all spectra with LOESS smoothing of 0.16 using the plotsmooth function in `Pavo` vers.
+1.0 [@RN29].
+We also took the median reflectance for all leaves (top, bottom) and stems (2 measures/stem) for each population (Hawaii, Kenya).
+
+We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` @RN29.
 This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
@@
 We first calculated the quantum catch of each photoreceptor under standard daylight illumination (D65; Commission internationale de l’éclairage – CIE) using the ‘vismodel’ function.
-We applied von Kries chromatic adaptation (vonkries = TRUE) and used the average background reflectance of the relevant background (Hawaii or Kenya) as the adapting background.
+We applied the von Kries chromatic adaptation (vonkries = TRUE) and used the average background reflectance of the relevant background (Hawaii or Kenya) as the adapting background.
 The signal in each photoreceptor class was proportional to the natural logarithm of the quantum catch, in accordance with Fechner's law (“fi” in pavo).
@@
 *Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Fig.
-\@ref(fig:Figcolsig6SPec)).
+S\@ref(fig:Figcolsig6SPec)).
 
@@
 We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21 (Fig.
-\@ref(fig:Figcolsig6SPec)B).
+S\@ref(fig:Figcolsig6SPec)B).
 
@@
 We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22 (Fig.
-\@ref(fig:Figcolsig6SPec)C).
+S\@ref(fig:Figcolsig6SPec)C).
 
@@
 ```
-Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, N = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]` mm, N = `r svl[2,"n"]`; t = `r SVL_pop$statistic`, p <0.001).
+Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, *n* = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]` mm, *n* = `r svl[2,"n"]`; t = `r SVL_pop$statistic`, *P* <0.001).
 SVL did not significantly predict luminance contrast against the background (Table \@ref(tab:table1)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
-SVL was a weak predictor of chromatic contrast (Table \@ref(tab:table1)): larger chameleons had slightly higher chromatic contrast against the background.
+Snout-vent length was a weak predictor of chromatic contrast (Table S\@ref(tab:table1)): larger chameleons had slightly higher chromatic contrast against the background.
 However, the populations did not differ in chromatic contrast during male displays (courtship and contests).
@@
 Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI).
-Standard errors are corrected by using robust variance estimators "}
+Standard errors are corrected by using robust variance estimators."}
 source("./R/func.R")
@@
 ```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
-Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for Models were fit using robust variance estimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom.
+Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for Models were fit using robust variance estimators to correct standard errors for non-independence, and we used a Satterthwaite degrees of freedom.
 Significant parameter estimates are those where the 95% CIs do not include zero."}
@@
 ```{r tableS3, tab.cap = "Wald F tests comparing the significance of model parameters for predator models.
-Wald F tests were computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom.
+Wald F tests were computed using robust variance estimators with a Satterthwaite correction to the degrees of freedom.
 Interactions were first tested and models refit to test the signifiacnce of main effects"}
@@
 ```{r tableS4, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter from linear mixed effects models.
-Standard errors were corrected using robust variance estimators and Saitterwaite degrees of freedom correction was used for calculated significance"}
+Standard errors were corrected using robust variance estimators and Satterthwaite degrees of freedom correction was used for calculated significance"}
 source("./R/func.R")

### 2021-08-25 supp.Rmd
Up to Table S5

@@
 Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).
-We also applied a Bonferroni correction to p-value to control for multiple comparisons."}
+We also applied a Bonferroni correction to p-values to control for multiple comparisons."}
  contrast <- read.csv(file = "./output/tables/Table3_MS.csv")
@@
 ```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
-Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for Models were fit using robust variance estimators to correct standard errors for non-independence, and we used a Satterthwaite degrees of freedom.
+Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for models were fitted using robust variance estimators to correct standard errors for non-independence, and we used a Satterthwaite degrees of freedom.
 Significant parameter estimates are those where the 95% CIs do not include zero."}

### 2021-08-25 supp.Rmd
Fixed up Supp

@@
 ```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing mean differences between Kenyan and Hawaiian lizards in luminance and chromatic contrast against the local background in two social contexts (male-male competition and courtship).
-Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI).
+Model estimates, standard errors (SE), Satterthwaite's degrees of freedom (df) and 95% confidence internals (95% CI).
 Standard errors are corrected by using robust variance estimators."}
@@
 ```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic mean differences between Kenyan and Hawaiian lizards in response to snakes and birds.
-Model estimates, standard errors (SE), Satterthwaite degrees of freedom (df) and 95% confidence internals (95% CI).
+Model estimates, standard errors (SE), Satterthwaite's degrees of freedom (df) and 95% confidence internals (95% CI).
 Full models are provided and standard errors are corrected by using robust variance estimators."}
@@
 ```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
-Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for models were fitted using robust variance estimators to correct standard errors for non-independence, and we used a Satterthwaite degrees of freedom.
+Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for models were fitted using robust variance estimators to correct standard errors for non-independence, and we used Satterthwaite's degrees of freedom.
 Significant parameter estimates are those where the 95% CIs do not include zero."}
@@
 Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Figure \@ref(fig:FigS3)A).
-There was no significant interaction between predator type ('bird' vs 'snake') and population in luminance contrast against the background, but a marginally non-significant population and body region effect (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
+There was no significant interaction between predator type ('bird' vs 'snake') and population was not significant for luminance contrast against the background, but there was a marginally non-significant population and body region effect (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
 Effects did not vary by predator type (Table \@ref(tab:tableS3)).
@@
 ```{r FigS3, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "All four body regions (gular, mid-flank, tailbase and topflank) and their luminance (B) and chromatic contrast (A) to Hawaiian and Kenyan backgrounds (each population against their own background) are shown averaged across stems and leaves found in their habitat.
-Just noticable differences (JNDs) are calculated for both bird and snake predators.
+Just noticable differences (JNDs) were calculated for both bird and snake predators.
 Mean JNDs are provided along with 95% confidence intervals."}
@@
 Wald F tests were computed using robust variance estimators with a Satterthwaite correction to the degrees of freedom.
-Interactions were first tested and models refit to test the signifiacnce of main effects"}
+Interactions were first tested, and models refitted, to test the significance of main effects."}
 source("./R/func.R")
@@
 ```{r tableS4, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter from linear mixed effects models.
-Standard errors were corrected using robust variance estimators and Satterthwaite degrees of freedom correction was used for calculated significance"}
+Standard errors were corrected using robust variance estimators and Satterthwaite's degrees of freedom correction was used for calculated significance."}
 source("./R/func.R")
@@
 ```{r tableS5, tab.cap = "Pairwise comparisons of luminance contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
-Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
-Wald test uses a robust variance correction on the standard errors."}
+Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald tests for contrast significance.
+The Wald test uses a robust variance correction on the standard errors."}
 source("./R/func.R")
@@
 ```{r tableS6, tab.cap = "Pairwise comparisons of chromatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
-Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald test for contrast significance.
-Wald test uses a robust variance correction on the standard errors."}
+Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with a pairwise Wald test for contrast significance.
+The Wald test uses a robust variance correction on the standard errors."}
 source("./R/func.R")
@@
 However, for luminance contrast, these models had poor convergence.
-Based on the raw data (Figure \@ref(fig:FigS4)B & D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
+Based on the raw data (Fig.
+S\@ref(fig:FigS4)B and S\@ref(fig:FigS4)D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
 As such, we simplified models and only present the main effects of body region and background environment for luminance contrast.
-However, inspection of the raw data (Figure \@ref(fig:FigS4)A & C) suggested that body regions may differ in how chromatically contrasting they were against the background environment of Hawaii and Kenya.
-In addition, our main results suggested that predators may deferentially drive chromatic signal conspicuousness, and given different body regions maybe more exposed to bird compared to snake predators, we fit an interaction and main effects model for chromatic contrast.
-
-The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, p = `r p_value(intearction_CC_Court$p_val)`, Figure \@ref(fig:FigS4)A & C & Table \@ref(tab:tableS7)).
-However, in all cases there was a tendeny for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Figure \@ref(fig:FigS4)A & C).
-Tailbase was the only body region showed a non-significant change (Table \@ref(tab:tableS8)).
-
-The magnitude of luminance contrast, showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court$p_val)`, Figure \@ref(fig:FigS4)B & D along with Table \@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, p = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, p = `r p_value(BodyRegion_AC_Court$p_val)`, Figure \@ref(fig:FigS4)B & D along with Table \@ref(tab:tableS7)).
+However, inspection of the raw data (Fig.
+S\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C) suggested that body regions may differ in how chromatically contrasting they were against the background environment of Hawaii and Kenya.
+In addition, our main results suggested that predators may deferentially drive chromatic signal conspicuousness, and given different body regions maybe more exposed to bird compared to snake, predators, we fitted an interaction and main effects model for chromatic contrast.
+
+The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, *P* = `r p_value(intearction_CC_Court$p_val)`, Fig.
+S\@ref(fig:FigS4)A, S\@ref(fig:FigS4)C and Table S\@ref(tab:tableS7)).
+However, in all cases there was a tendency for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Fig.
+S\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C).
+Tailbase was the only body region that showed a non-significant change (Table S\@ref(tab:tableS8)).
+
+The magnitude of luminance contrast showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court$p_val)`, Fig.
+S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, p = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Fig.
+S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)).
 
@@
 Mean JNDs are provided along with 95% confidence intervals.
-**(A & B)** represent male-male contest displays whereas **(C & D)** represent courtship displays"}
+**(A & B)** represent male-male contest displays whereas **(C & D)** represent courtship displays."}
 # Courship data
@@
 
-```{r tableS7, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter.
+```{r tableS7, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df), and 95% confidence intervals for each parameter.
 Main effects and interaction models are presented.
 Only main effect models are presented for luminance contrast (see text for details), whereas interaction models are presented for chromatic contrast.
-Pairwise contrasts are presented for chromatic contrast models (see Table 8) for ease of interpretation."}
+Pairwise contrasts are presented for chromatic contrast models for ease of interpretation."}
 source("./R/func.R")
@@
 F statistic, degrees of freedom (DF) and p-values are provided (p).
-A bonferroni correction to p-value is also provided to control for multiple testing.
-See Table 7 for model estimates."}
+A Bonferroni correction to p-values is also provided to control for multiple testing."}
 
@@
 ```
-To gauge the extent of individual color change, we calculated the absolute difference in JNDs between for each individual that had both display (male-male contest) and bird response data (N = `r N_diff[1,2]` individuals from Hawaii, N = `r N_diff[2,2]` Kenya).
+To gauge the extent of individual color change, we calculated the absolute difference in JNDs between for each individual that had both display (male-male contest) and bird response data (*n* = `r N_diff[1,2]` individuals from Hawaii, *n* = `r N_diff[2,2]` Kenya).
 We chose these two color states to maximize sample size and because chameleons and birds have similar tetrachromatic visual systems.
 We tested whether the absolute difference in luminance or chromatic contrast between these two color states differed between populations and body regions.
-We included SVL, Population, Body region and the interaction between population and body region as fixed effects.
-ID was included as a random effect due to measures for more than one body region per individual.
-
-Body regions varied in there overall effects but mainly showed a decrease in absolute JND difference between display and antipredator color states in Kenya compared to Hawaii (Fig.
-\@ref(fig:figDisplayPred)).
-While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, p = `r p_value(robust_Diff_dS_int[,6])`; luminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, p = `r p_value(robust_Diff_dl_int[,6])`) (Fig.
-\@ref(fig:figDisplayPred) & Table \@ref(tab:tableS9)).
-However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, p = `r p_value(robust_Diff_dS_BR_me[,6])`; luminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, p = `r p_value(robust_Diff_dL_BR_me[,6])`) (Fig.
-\@ref(fig:figDisplayPred); Table \@ref(tab:tableS9) & Table \@ref(tab:tableS10)).
-
-```{r, figDisplayPred, fig.align = 'center', fig.width = 13, fig.height = 7, fig.cap = " Extent of color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (N = 30) and Kenyan (N = 13) chameleons.
+We included SVL, population, body region and the interaction between population and body region as fixed effects.
+ID was included as a random effect becasue we had measures for more than one body region per individual.
+
+Body regions varied in their overall effects but mainly showed a decrease in absolute JND difference between display and antipredator color states in Kenya compared to Hawaii (Fig.
+S\@ref(fig:figDisplayPred)).
+While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, *P* = `r p_value(robust_Diff_dS_int[,6])`; luminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, *P* = `r p_value(robust_Diff_dl_int[,6])`) (Fig.
+S\@ref(fig:figDisplayPred) and Table S\@ref(tab:tableS9)).
+However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dS_BR_me[,6])`; luminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dL_BR_me[,6])`) (Fig.
+S\@ref(fig:figDisplayPred); Table S\@ref(tab:tableS9) and Table S\@ref(tab:tableS10)).
+
+```{r, figDisplayPred, fig.align = 'center', fig.width = 13, fig.height = 7, fig.cap = " Extent of color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (*n* = 30) and Kenyan (*n* = 13) chameleons.
 **(A)** Average chromatic and **(B)** luminance contrast are provided for each body region"}
@@
 
-```{r tableS10, tab.cap = "Pairwise Wald F tests for chromatic and luminance change between different body regions for color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (N = 30) and Kenyan (N = 13) chameleons.
+```{r tableS10, tab.cap = "Pairwise Wald F tests for chromatic and luminance change between different body regions for color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (*n* = 30) and Kenyan (*n* = 13) chameleons.
 F statistic, degrees of freedom (DF) and p-values are provided (p).
-A bonferroni correction to p-value is also provided to control for multiple testing."}
+A Bonferroni correction to p-values is also provided to control for multiple testing."}
 source("./R/func.R")

### 2021-08-25 supp.Rmd
Supp corrected. P and n values fixes. Missed a paragraoh with S for Fig and Tables. Fixed

@@
 Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, *n* = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]` mm, *n* = `r svl[2,"n"]`; t = `r SVL_pop$statistic`, *P* <0.001).
-SVL did not significantly predict luminance contrast against the background (Table \@ref(tab:table1)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
+SVL did not significantly predict luminance contrast against the background (Table S\@ref(tab:table1)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
 Snout-vent length was a weak predictor of chromatic contrast (Table S\@ref(tab:table1)): larger chameleons had slightly higher chromatic contrast against the background.
@@
 
-Chromatic contrast of color signals against the background varied by predator type, population and body region (Figure \@ref(fig:FigS3)A; Table \@ref(tab:tableS3)).
-Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Figure \@ref(fig:FigS3)A).
-There was no significant interaction between predator type ('bird' vs 'snake') and population was not significant for luminance contrast against the background, but there was a marginally non-significant population and body region effect (Figure \@ref(fig:FigS3)B; Table \@ref(tab:tableS3)).
-Effects did not vary by predator type (Table \@ref(tab:tableS3)).
-Nonetheless, effect sizes across body regions were generally in the same direction (Figure \@ref(fig:FigS3)B & Table \@ref(tab:tableS4)).
-Overall, there were significant differences between Kenya and Hawaii in luminance contrast of the gular and tailbase regions, but not for mid- and top-flank (Table \@ref(tab:tableS5)).
-In contrast, no significant differences between Kenya and Hawaii existed across any body region for snakes, whereas significant differences existed between gular, midflank and tailbase for birds (Table \@ref(tab:tableS6)).
+Chromatic contrast of color signals against the background varied by predator type, population and body region (Fig.
+S\@ref(fig:FigS3)A and Table S\@ref(tab:tableS3)).
+Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Fig.
+S\@ref(fig:FigS3)A).
+There was no significant interaction between predator type ('bird' vs 'snake') and population was not significant for luminance contrast against the background, but there was a marginally non-significant population and body region effect (Fig.
+S\@ref(fig:FigS3)B and Table S\@ref(tab:tableS3)).
+Effects did not vary by predator type (Table S\@ref(tab:tableS3)).
+Nonetheless, effect sizes across body regions were generally in the same direction (Fig.
+S\@ref(fig:FigS3)B and Table S\@ref(tab:tableS4)).
+Overall, there were significant differences between Kenya and Hawaii in luminance contrast of the gular and tailbase regions, but not for mid- and top-flank (Table S\@ref(tab:tableS5)).
+In contrast, no significant differences between Kenya and Hawaii existed across any body region for snakes, whereas significant differences existed between gular, midflank and tailbase for birds (Table S\@ref(tab:tableS6)).
 
@@
 
-The magnitude of luminance contrast showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, p = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, p = `r p_value(background_AC_Court$p_val)`, Fig.
-S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, p = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Fig.
+The magnitude of luminance contrast showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court$p_val)`, Fig.
+S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, *P* = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Fig.
 S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)).

### 2021-08-25 biblio.Rmd
Adding a merged file becuase of stupid Sciences idiocy of having all refs in MS and supp together.

--- 2021-08-25 supp.Rmd
@@
 title: "Invasive chameleons released from predation display more conspicuous colors"
-date: "`r Sys.Date()`"
 bibliography: supp_refs.bib
@@
 ---
+
+```{r setup, include=FALSE, purl = FALSE}
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
+## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
+  ## and rounded to 2 digits      ## and rounded to 2 digits
+  options(digits = 2)
+```
+
+```{r getlibrary, echo = FALSE, results = "hide"}
+# load libraries
+devtools::install_github("karthik/wesanderson")
+pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, flextable, readxl)
+source("./R/func.R")
+
+```
+
+```{r Models, echo = FALSE, results = "hide"}
+source("./R/func.R")
+#########################################################
+# Social Context: Male-male Competition Displays
+#########################################################
+#Load data and clean a bit
+Display_1 <- read.table("./data/Display_1.csv", 
+                          header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Sample sizes
+N_display <- data.frame(Display_1 %>% group_by(Population) %>% summarise(n = length(unique(ID))))
+
+# Center SVL to make intercept interpretable
+Display_1$sc_SVL <- with(Display_1, scale(SVL, scale=FALSE))
+
+#exclude bottom flank.
+Display_1 <- Display_1 %>% 
+              filter(!BodyRegion == "botflank") %>% as.data.frame()
+
+# create as factor
+    Display_1$BodyRegion <- factor(Display_1$BodyRegion)
+
+##########
+## Achromatic Contrast - dL
+##########
+
+    # Fit the model
+    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + (1|ID), data=Display_1)
+      summary(ConspicDisplayAC)
+
+    # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
+    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out missing data
+       robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, 
+                                            cluster = Display_1_IDs$ID, vcov = "CR2")  
+    robust_ConspicDisplayAC_CI <- conf_int(ConspicDisplayAC, 
+                                           cluster = Display_1_IDs$ID, vcov = "CR2")
+
+     # Wald test
+            Population_ConspicDisplayAC <- Wald_test(ConspicDisplayAC, 
+                                             constraints = constrain_zero(3),
+                                            cluster = Display_1_IDs$ID, vcov = "CR2")
+
+##########
+# Chromatic contrast - dS
+##########
+    #Fit the model
+         ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + (1|ID), data=Display_1)
+         summary(ConspicDisplayCC)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Display_1_IDs <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+    robust_ConspicDisplayCC <- coef_test(ConspicDisplayCC, cluster = Display_1_IDs$ID, vcov = "CR2")
+ robust_ConspicDisplayCC_CI <- conf_int(ConspicDisplayCC, cluster = Display_1_IDs$ID, vcov = "CR2")
+
+  # Wald test
+            Population_ConspicDisplayCC <- Wald_test(ConspicDisplayCC, 
+                                             constraints = constrain_zero(3),
+                                            cluster = Display_1_IDs$ID, vcov = "CR2")
+#########################################################
+# Social Context: Courtship Displays
+#########################################################
+
+# Load data
+Courtship <- read.table("./data/Courtship_1.2.csv", 
+                          header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Sample sizes
+N_Courtship <- data.frame(Courtship %>% group_by(Population) %>% summarise(n = length(unique(ID))))
+
+# Center SVL do intercept is at the mean body size
+Courtship$sc_SVL <- with(Courtship, scale(SVL, scale=FALSE))
+
+##########
+## Achromatic Contrast - dL
+##########
+
+  # Fit the model
+    ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population  + (1|ID), data=Courtship)
+    summary(ConspicCourtshipAC)
+
+ # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
+    Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out missing data
+    robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, 
+                                           cluster = Courtship_IDs$ID, vcov = "CR2")
+ robust_ConspicCourtshipAC_CI <- conf_int(ConspicCourtshipAC, 
+                                          cluster = Courtship_IDs$ID, vcov = "CR2")
+
+    # Wald test
+            Population_ConspicCourtshipAC <- Wald_test(ConspicCourtshipAC, 
+                                             constraints = constrain_zero(3),
+                                            cluster = Courtship_IDs$ID, vcov = "CR2")
+
+##########
+# Chromatic contrast - dS
+##########
+
+  # Fit the model
+  ConspicCourtshipCC <- lmer(dS ~ sc_SVL + Population  + (1|ID), data=Courtship)
+    summary(ConspicCourtshipCC)
+
+  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Courtship_IDs <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population)) # filter out misssing data
+       robust_ConspicCourtshipCC <- coef_test(ConspicCourtshipCC, cluster = Courtship_IDs$ID, vcov = "CR2")
+    robust_ConspicCourtshipCC_CI <- conf_int(ConspicCourtshipCC, cluster = Courtship_IDs$ID, vcov = "CR2")
+
+    # Wald test
+            Population_ConspicCourtshipCC <- Wald_test(ConspicCourtshipCC, 
+                                             constraints = constrain_zero(3),
+                                            cluster = Courtship_IDs$ID, vcov = "CR2")
+
+ #########################################################
+# Predator Context: Birds and snakes
+#########################################################
+ Predator <- read.table("./data/Predator_2_3.csv", 
+                       header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+# Sample sizes
+N_Pred <- data.frame(Predator %>% group_by(Population, Predator) %>% summarise(n = length(unique(ID)))) %>% arrange(Predator)
+
+# Clean up
+ Predator <- Predator %>% filter(!BodyRegion == "botflank" & !BodyRegion =='gularbot') %>% as.data.frame() #exclude bottom flank.
+Predator$BodyRegion <- factor(Predator$BodyRegion) #updates levels for the variable BodyRegion
+
+##########
+## Achromatic Contrast - dL
+##########
+
+# First, if predator type varies across populations then we would predcit that whether snake and bird actually are different would depend on the population.
+For example, if, say, in Hawaii and Kenya birds were equally likely as predators then dL would not vary for birds across populations, but would so for snakes.
+This is an interaction.
+
+  # Fit model testing this idea above
+    PredatorAC_inter <- lmer(dL ~ Predator + Population + Predator:Population + (1|ID), 
+                             data = Predator) 
+    summary(PredatorAC_inter)
+
+  #Refit main effects so we can test main effects on own.
+         PredatorAC <- lmer(dL ~ Predator + Population + (1|ID), data = Predator) 
+        summary(PredatorAC)
+
+     # Wald tests
+            back_pred_AC_Pred <- Wald_test(PredatorAC_inter, constraints = constrain_zero(4), 
+                                   cluster = Predator$ID, vcov = "CR2")
+
+             Predator_AC_Pred <- Wald_test(PredatorAC, constraints = constrain_zero(2), 
+                                  cluster = Predator$ID, vcov = "CR2")
+
+           Population_AC_Pred <- Wald_test(PredatorAC, constraints = constrain_zero(3), 
+                                  cluster = Predator$ID, vcov = "CR2")
+
+  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+    Predator_IDs <- Predator %>% filter(!is.na(Predator), !is.na(Population)) # filter out misssing data
+              robust_PredatorAC <- coef_test(PredatorAC, cluster = Predator_IDs$ID, vcov = "CR2")
+           robust_PredatorAC_CI <-  conf_int(PredatorAC, cluster = Predator_IDs$ID, vcov = "CR2")
+     robust_PredatorAC_CI_inter <-  conf_int(PredatorAC_inter, cluster = Predator_IDs$ID, vcov = "CR2")
+
+  # Model allows us to get bird in Hwaii compared to bird in Kenya (contrast 3)
+      Predator$pred_pop <- with(Predator, interaction(Predator, Population))
+
+      PredatorAC_pred_pop1 <- lmer(dL ~  pred_pop + (1|ID), data = Predator) 
+      summary(PredatorAC_pred_pop1)
+      PredatorAC_pred_pop2 <- lmer(dL ~  relevel(pred_pop, ref = "snake.Hawaii") + (1|ID), data = Predator) 
+      summary(PredatorAC_pred_pop2)
+
+      # Fit the models so we get the estimate and corrected robust standard error for the contrasts we want to make
+      robust_PredatorAC_pred_pop1 <- conf_int(PredatorAC_pred_pop1, cluster = Predator$ID, vcov="CR2")
+      robust_PredatorAC_pred_pop2 <- conf_int(PredatorAC_pred_pop2, cluster = Predator$ID, vcov="CR2")
+
+     # Contrasts hypothesis tests, but use RVE
+
+      PredatorAC_pred_pop <- lmer(dL ~  0+ pred_pop + (1|ID), data = Predator) 
+    result_pred <- Wald_test(PredatorAC_pred_pop, constraints = constrain_pairwise(c(1:4)), 
+                             vcov="CR2", tidy = TRUE)
+    result_pred$p_val_bonfer <- sapply(p.adjust(result_pred$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Get the contrasts and Wald stats for comparions
+     CI_hyp_testAC <- data.frame(
+              comparison = c("Snake (Hawaii-Kenya)", "Birds (Hawaii-Kenya)"), 
+                Estimate = c(robust_PredatorAC_pred_pop2[4,1], robust_PredatorAC_pred_pop1[3, 1]),
+                lwr = c(robust_PredatorAC_pred_pop2[4,4], robust_PredatorAC_pred_pop1[3, 4]),
+                upr = c(robust_PredatorAC_pred_pop2[4,5], robust_PredatorAC_pred_pop1[3, 5]),
+              Fstat = c(result_pred[5,3], result_pred[2,3]),
+              df = c(result_pred[5,5], result_pred[2,5]),
+              p_val = c(result_pred[5,7], result_pred[2,7]),
+              p_bonf = c(result_pred[5,8], result_pred[2,8]))
+
+##########
+# Chromatic contrast - dS
+##########
+
+    # First, if predator type varies across populations then we would predcit that whether snake and bird actually are different would depend on the population.
+For example, if, say, in Hawaii and Kenya birds were equally likely as predators then dL would not vary for birds across populations, but would so for snakes.
+This is an interaction.
+
+    # Fit the interaction model.
+    PredatorCC <- lmer(dS ~ Predator + Population + Predator:Population + (1|ID), 
+                       data = Predator) 
+    summary(PredatorCC)
+
+    #Refit main effects so we can test main effects on own.
+      PredatorCC_main <- lmer(dS ~ Predator + Population + (1|ID), data = Predator)
+
+     # Wald tests
+            back_pred_CC_Pred <- Wald_test(PredatorCC, constraints = constrain_zero(4), 
+                                   cluster = Predator$ID, vcov = "CR2")
+
+             Predator_CC_Pred <- Wald_test(PredatorCC_main, constraints = constrain_zero(2), 
+                                  cluster = Predator$ID, vcov = "CR2")
+
+           Population_CC_Pred <- Wald_test(PredatorCC_main, constraints = constrain_zero(3), 
+                                  cluster = Predator$ID, vcov = "CR2")
+
+    # Support for an interaction between predator type and population in the chromatic realm. Planned comparisons. First re-fit model
+      Predator$pred_pop <- with(Predator, interaction(Predator, Population))
+    PredatorCC_pred_pop <- lmer(dS ~ -1 + pred_pop + (1|ID), data = Predator) 
+    summary(PredatorCC_pred_pop)
+
+    # Present full model
+         robust_PredatorCC_CI_inter <-  conf_int(PredatorCC, cluster = Predator_IDs$ID, vcov = "CR2")
+
+    # Fit RVE, we want contrast effects because annoyingly RVE doesn't provide them so will relevel to get them in the context of RVE
+
+    # Model allows us to get bird in Hwaii compared to bird in Kenya (contrast 3)
+      PredatorCC_pred_pop1 <- lmer(dS ~  pred_pop + (1|ID), data = Predator) 
+      summary(PredatorCC_pred_pop1)
+      PredatorCC_pred_pop2 <- lmer(dS ~  relevel(pred_pop, ref = "snake.Hawaii") + (1|ID), data = Predator) 
+      summary(PredatorCC_pred_pop2)
+
+      # Fit the models so we get the estimate and corrected robust standard error for the contrasts we want to make
+      robust_PredatorCC_pred_pop1 <- conf_int(PredatorCC_pred_pop1, cluster = Predator$ID, vcov="CR2")
+      robust_PredatorCC_pred_pop2 <- conf_int(PredatorCC_pred_pop2, cluster = Predator$ID, vcov="CR2")
+
+     # Contrasts hypothesis tests, but use RVE 
+    result_pred <- Wald_test(PredatorCC_pred_pop, constraints = constrain_pairwise(c(1:4)), vcov="CR2", tidy = TRUE)
+    result_pred$p_val_bonfer <- sapply(p.adjust(result_pred$p_val, "bonferroni"), function(x) p_value(x))
+
+    # Get the contrasts and Wald stats for comparions
+     CI_hyp_test <- data.frame(
+              comparison = c("Snake (Hawaii-Kenya)", "Birds (Hawaii-Kenya)"), 
+                Estimate = c(robust_PredatorCC_pred_pop2[4,1], robust_PredatorCC_pred_pop1[3, 1]),
+                lwr = c(robust_PredatorCC_pred_pop2[4,4], robust_PredatorCC_pred_pop1[3, 4]),
+                upr = c(robust_PredatorCC_pred_pop2[4,5], robust_PredatorCC_pred_pop1[3, 5]),
+              Fstat = c(result_pred[5,3], result_pred[2,3]),
+              df = c(result_pred[5,5], result_pred[2,5]),
+              p_val = c(result_pred[5,7], result_pred[2,7]),
+              p_bonf = c(result_pred[5,8], result_pred[2,8]))
+
+#########################################################
+# Local Adaptation: Testing for local adaptation
+#########################################################
+##################
+# Social Context: Male-male competition
+##################
+DisplayBgrd <- read.table("./data/Display_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+svl_dat <- read.csv("./data/SVL2.csv")
+
+DisplayBgrd <- DisplayBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
+
+# Filter out only Hawaiian and add in SVL for IDs
+DisplayBgrd <- DisplayBgrd %>% 
+               filter(Population == "Hawaii") %>% 
+               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
+##########
+# Achromatic Contrast - dL
+##########
+
+  # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit the main effect model
+  DispBgrdAC <- lmer(dL ~ sc_svl  + Background_pop  + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC)
+
+  # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
+  robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+  # Add Wald test to test signifiacnce of each factor, background and population using roust variance
+
+  background_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(3), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  svl_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(2), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+ ##########
+# Chromatic Contrast - dS
+##########
+
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+    # Fit the main effect model
+          DispBgrdCC <- lmer(dS ~ sc_svl + Background_pop  + (1|ID), data=DisplayBgrd)
+          summary(DispBgrdCC)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+      robust_DispBgrdCC <- clubSandwich::conf_int(DispBgrdCC, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+    # Check significance of each factor
+        background_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(3), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+        svl_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(2), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+###################
+# Social Context: Courtship
+##################
+CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+CourtshipBgrd <- CourtshipBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
+
+# Filter out only Hawaiian and add in SVL for IDs
+CourtshipBgrd <- CourtshipBgrd %>% 
+               filter(Population == "Hawaii") %>% 
+               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+               filter(!is.na(SVL))
+
+##########
+## Achromatic Contrast - dL
+##########
+
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit main effects model
+    CourtshipAC <- lmer(dL ~ sc_svl  + Background_pop  + (1|ID), data=CourtshipBgrd)
+    summary(CourtshipAC)
+
+   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+   robust_CourtshipAC <- clubSandwich::conf_int(CourtshipAC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+   # Wald tests
+           background_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(3), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+           svl_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(2), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+##########
+# Chromatic - dS
+##########
+ # Here, we want to test if the conspicuousness of lizard signals is locally adapted to the specific background of the Kenyan and Hawaiian population.
+The predcition is simply that, if we were to place a Hawaiian animal against the Kenyan background that it would be far more conspicuous compared to if we place a Kenyan animal in a Hawaiian background.
+
+  # Fit main effect model
+      CourtshipCC <- lmer(dS ~ sc_svl  + Background_pop  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC)
+
+  # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+  robust_CourtshipCC <- clubSandwich::conf_int(CourtshipCC, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+   # Wald tests
+           background_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(3), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+           svl_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(2), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+#########################################################
+# Comparing displays
+#########################################################
+ # Load data from Devi
+       Display_bird_diff <- read_excel("data/Display_bird_diff.xlsx")
+
+       Diff <- Display_bird_diff
+
+      names(Diff) # header rows R is reading
+
+    Diff$BodyRegion <- factor(Diff$BodyRegion) #updates levels for the variable BodyRegion
+    str(diff.difftime) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
+    Diff$ID <- factor(Diff$ID) #updates levels for the variable ID
+    str(diff.difftime) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
+
+# Center SVL to make intercept interpretable
+    Diff$sc_SVL <- with(Diff, scale(SVL, scale=FALSE))
+
+# Filter missing SVL data
+  Diff <- Diff %>% filter(!is.na(SVL)) # filter out misssing SVL data
+
+# create as factor
+    Diff$BodyRegion <- factor(Diff$BodyRegion)
+
+# Sample sizes
+    N_diff <- data.frame(Diff %>% group_by(Population) %>% summarise(n = length(unique(ID))))
+
+##################
+# Achromatic contrast
+##################
+
+# Fit model testing this idea above. First just test for body region interaction. But, needs RVE
+    Diff_dL_BR_int <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
+    summary(Diff_dL_BR_int)
+
+    # Refit mdoel using robust standard errors
+  robust_Diff_dL_BR_int <- clubSandwich::conf_int(Diff_dL_BR_int, cluster = Diff$ID, vcov =  "CR2")
+
+    # Significance of interaction
+      robust_Diff_dl_int <- Wald_test(Diff_dL_BR_int, constraints = constrain_zero(7:9), 
+                                    cluster = Diff$ID, vcov = "CR2")
+# Main effects BR
+    Diff_dL_BR_me <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
+    summary(Diff_dL_BR_me)
+
+    # Sig of Body Region
+        robust_Diff_dL_BR_me <- Wald_test(Diff_dL_BR_me, constraints = constrain_zero(4:6), 
+                                    cluster = Diff$ID, vcov = "CR2")
+
+# Then fit marginalised model so that body regions are averaged
+  Diff_dL <- lmer(dL_diff_abs ~ sc_SVL + Population + (1|ID), data=Diff)
+  summary(Diff_dL)
+
+# Check residuals of model
+ # hist(residuals(Diff_dL))
+ # plot(Diff_dL) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
+
+# Refit mdoel using robust standard errors
+  robust_Diff_dl_BR <- clubSandwich::conf_int(Diff_dL_BR_me, cluster = Diff$ID, vcov =  "CR2")
+
+# Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
+
+  robust_Diff_dl_pop <- Wald_test(Diff_dL, constraints = constrain_zero(3), 
+                                  cluster = Diff$ID, vcov = "CR2")
+   robust_Diff_dl_svl <- Wald_test(Diff_dL, constraints = constrain_zero(2), 
+                                  cluster = Diff$ID, vcov = "CR2")
+
+##################
+# Chromatic contrast
+##################
+
+  # Fit model testing this idea above. First just test for body region interaction. But, needs RVE
+    Diff_dS_BR_int <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
+    summary(Diff_dS_BR_int)
+
+        # Refit mdoel using robust standard errors
+     robust_Diff_dS_BR_int <- clubSandwich::conf_int(Diff_dS_BR_int, cluster = Diff$ID, vcov =  "CR2")
+
+     # Significance of interaction
+      robust_Diff_dS_int <- Wald_test(Diff_dS_BR_int, constraints = constrain_zero(7:9), 
+                                    cluster = Diff$ID, vcov = "CR2")
+  # Main effects BR
+    Diff_dS_BR_me <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
+    summary(Diff_dS_BR_me)
+
+    # Sig of Body Region
+      robust_Diff_dS_BR_me <- Wald_test(Diff_dS_BR_me, constraints = constrain_zero(4:6), 
+                                    cluster = Diff$ID, vcov = "CR2")
+
+# Fit model testing this idea above
+  Diff_dS <- lmer(dS_diff_abs ~  SVL + Population + (1|ID), data=Diff)
+  summary(Diff_dS)
+
+# Check residuals of model
+  #hist(residuals(Diff_dS))
+  #plot(Diff_dS) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
+
+# Refit mdoel using robust standard errors
+  robust_Diff_dS <- clubSandwich::conf_int(Diff_dS_BR_me, cluster = Diff$ID, vcov =  "CR2")
+
+# Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
+   robust_Diff_dS_pop <- Wald_test(Diff_dS, constraints = constrain_zero(3), 
+                                  cluster = Diff$ID, vcov = "CR2")
+   robust_Diff_dS_svl <- Wald_test(Diff_dS, constraints = constrain_zero(2), 
+                                  cluster = Diff$ID, vcov = "CR2")
+
+```
+
+Martin J. Whiting^1^, Brenden S. Holland^2^, J. Scott Keogh^3^, Daniel W.A.Noble^3^, Katrina J. Rankin^4^ and Devi Stuart-Fox^4^
+
+^1^ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
+
+^2^ Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
+
+^3^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
+
+^4^ School of BioSciences, The University of Melbourne, Parkville, Victoria 3010, Australia
+
+# Abstract
+Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid evolutionary change.
+We demonstrate such character release for color signals in Jackson’s chameleons (*Trioceros j.
+xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) where there are few lizard predators.
+Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator color states.
+Consistent with local adaptation, the display colors of Hawaiian chameleons were more conspicuous in their local than ancestral habitats.
+These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
+
+When species encounter novel environments, such as during human-mediated introductions, evolutionary change can occur more rapidly than Darwin ever thought possible [@RN1ms; @RN2ms; @RN3ms; @RN4ms].
+In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4ms; @RN5ms].
+Such character release should be most evident for social and sexual signals; released from constraints on conspicuousness imposed by predation risk, signals could elaborate in many directions [e.g., @RN4ms; @RN6ms; @RN7ms; @RN8ms].
+However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7ms; @RN9ms; @RN10ms].
+This is especially true for species with dynamic color change, for which we are not aware of any examples.
+Here, we examine character release of a dynamic color signal by taking advantage of an unintended evolutionary ‘experiment’: accidental translocation of chameleons to a novel, low predation environment.
+
+In 1972, a shipment of Jackson’s chameleons (*Trioceros jacksonii xanthopholus*; Fig.
+\@ref(fig:figure1) and fig.
+S1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
+The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11ms].
+Jackson’s chameleons have a high intrinsic population growth rate (9--12 month generation time and live birth of up to 50 young @RN24ms), enabling their rapid establishment and potentially rapid evolution (approx.
+50 -- 65 generations).
+On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014ms.
+This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.
+
+Chameleons are famous for their rapid color change [@RN12ms; @RN13ms; @RN14ms].
+When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green during courtship or to signal dominance (Fig.
+1, figs.
+S1 and S3).
+At other times, they are a duller green and can darken to brown or become mottled green and brown, particularly in response to perceived predation risk (Fig.
+\@ref(fig:figure1) and fig.
+S4A).
+Their different color states can appear more or less conspicuous depending on the context, background, and viewer (e.g., conspecific or predator).
+If introduced chameleons in Hawaii experience character release from reduced predation, we expect Hawaiian chameleons to have more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators.
+We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.
+
+To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (fig.
+S4).
+We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
+\@ref(fig:figure2) and fig.
+S5).
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
+
+In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.
+\@ref(fig:figure2), Fig.
+\@ref(fig:figure3)A and table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
+This result was consistent for all body regions measured (supplementary text).
+However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.
+\@ref(fig:figure3)B and table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
+
+The importance of luminance contrast is consistent with the nature and mechanism of rapid color change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.
+\@ref(fig:figure1)).
+Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15ms or changes in the spacing of guanine crystals within iridophores @RN14, both of which strongly affect luminance.
+Accordingly, character release was observed in the overall luminance of signals rather than their hue.
+
+Released from predation, Hawaiian chameleons were less cryptic than Kenyan chameleons when threatened by both bird and snake predators.
+Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.
+\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, *P* = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, table S3).
+Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, table S3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, *P* = `r p_value(back_pred_CC_Pred$p_val)`).
+Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.
+\@ref(fig:figure3)D and table S4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`, JND values > 5).
+By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.
+\@ref(fig:figure3)D and table S4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1; JND values ~ 2).
+The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18ms].
+The poor color discrimination of snakes may also explain why chameleons from Hawaii show little change relative to their Kenyan counterparts in their chromatic response to snakes, despite the absence of snakes on Hawaii.
+
+Character release may reflect rapid evolutionary change, phenotypic plasticity or a combination of these processes [@RN19ms; @RN20ms].
+Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
+Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.
+\@ref(fig:figure4) and table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
+Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26ms].
+
+To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
+While there was evidence for differences between body regions (see supplementary text, fig.
+S9, tables S12 and S13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (fig.
+S9).
+Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
+While it is possible that differences between Hawaiian and Kenyan chameleons could reflect a founder effect, the observed differences are all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the local environment.
+
+We have documented a surprisingly strong effect of recent release from predation on a conspicuous, dynamic visual signal.
+Oahu has no snakes or raptors that feed on lizards.
+The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22ms].
+In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014ms -- the conditions were highly favorable for character release of social signals.
+Jackson’s chameleons likely experience strong sexual selection given their polygynous mating system, elaborate courtship displays and intense male contest competition (Fig.
+\@ref(fig:figure1)) in which males lock and twist their three horns into rivals during intense bouts that sometimes result in physical injury.
+Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
+To the best of our knowledge, this is also the first example of character release in a dynamic color signal.
+
+Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12ms].
+Our results suggest that even for species capable of dynamic color change there is an upper threshold to signal intensity that is constrained by natural selection.
+This novel finding raises the question of whether other animals that have convergently evolved dynamic color signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
+More generally, our study highlights the opportunities provided by translocations and invasions, which are often otherwise detrimental, to study natural and sexual selection in the wild.
+
+# Figures
+
+```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social and predatory stimuli.
+Male chameleons experience intense sexual selection**.
+During the breeding season they change from a neutral state (green) to a highly conspicuous bright yellow display signal.
+They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
+**(A)** A typical dominant male in display coloration.
+**(B)** A subordinate male that lost a contest and turned from bright yellow to brown.
+**(C)** Two males fighting, both are in display coloration and at a stage where they are relatively even matched.
+**(D)** A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male."}
+library(magick)
+
+img65 <- image_read("./output/figures/fig1.pdf")
+img65
+
+```
+
+```{r figure2, fig.width = 14, fig.height= 14, fig.cap="**Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya**.
+The context for measurement was **(A)** male contest displays; **(B)** courtship; **(C)** bird predator; and **(D)** snake predator.
+More details in online supplementary text and Fig.
+S5."}
+library(magick)
+img <- image_read("./output/figures/Figure3.png")
+img
+
+```
+
+```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "**Chromatic and luminance contrast in Just Noticeable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves)**.
+Luminance and chromatic contrast of male chameleons during male-male contests **(A)** and female courtship **(B)**.
+JNDs are calculated based on the chameleon visual system.
+Luminance and chromatic contrasts of male chameleons during snake **(C)** and bird **(D)** predator encounters.
+JNDs are calculated based on the snake and bird visual systems.
+Contrasts between means, ‘Beta’, that are bold indicate significant effects."}
+rerun = FALSE
+if(rerun){
+## Make figures that average across body regions, but display Social Context
+
+courtship_data <- Courtship %>% 
+                  select(ID, Population, Colour, SVL, BodyRegion, Background, dS, dL)
+  display_data <- Display_1 %>%
+                  select(ID, Population, Colour, SVL, BodyRegion, Background, dS, dL)
+combined_data <- rbind(courtship_data, display_data)
+
+# Now calculate the averages for each population and Colour (i.e., context) / predation
+
+summary_data_pred <- Predator %>%
+                group_by(Predator, Population) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL)
+
+summary_data <- combined_data %>%
+                group_by(Population, Colour) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>%
+                mutate(Colour = if_else(Colour == "Courtship", "Courtship", "Male-male contest"))
+
+#Grab photos
+  males <- readPNG("./photos/male_fight.png", native = TRUE)
+females <- readPNG("./photos/courtship.png",  native = TRUE)
+  snake <- readPNG("./photos/snake_cham.png", native = TRUE)
+   bird <- readPNG("./photos/bird_cham.png",  native = TRUE)
+
+#plots
+p1 <- ggplot(summary_data, aes(x = Population, y = Mean_dS, group = Colour, color = Colour)) +
+    ylim(2.5, 6.5) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Chromatic contrast (JND)") +
+  annotate("segment", x = 0.85, xend = 1.85, y = 6.1, yend = 6.1, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 5.7, yend = 5.7, colour = "black") + 
+  annotate("text", x = 0.85+(1.85-0.85)/2,  y = 6.2, label = "NS",  size = 6) + 
+  annotate("text", x = 1.12+(2.12-1.12)/2, y = 5.8, label = "NS", size = 6) + 
+  theme_bw() +
+  labs(colour = "Social Context") +
+  theme(legend.position = c(0.18, 0.08),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+   inset_element(males,
+                left = 0.45,
+                right = 0.90,
+                bottom = 0.58,
+                top = 0.75, ignore_tag = TRUE) + 
+    inset_element(females,
+                left = 0.30,
+                right = 0.75,
+                bottom = 0.15,
+                top = 0.28, ignore_tag = TRUE) 
+      #annotation_raster(males, xmin = 1.2, xmax = 1.9, ymin = 4.8, ymax = 5.3) + 
+      #annotation_raster(females, xmin = 1, xmax = 1.7, ymin = 3.5, ymax = 4)
+
+p2 <- ggplot(summary_data, aes(x = Population, y = Mean_dL, group = Colour, color = Colour)) +
+  ylim(15, 45) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Luminance contrast (JND)") +
+   annotate("segment", x = 0.90, xend = 1.90, y = 40, yend = 40, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 42, yend = 42, colour = "black") + 
+  annotate("text", x = 0.9+((1.90 - 0.90)/2), y = 41, label = "**",  size = 6) + 
+  annotate("text", x = 1.12+((2.12 - 1.12)/ 2), y = 43, label = "**", size = 6) + 
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+    inset_element(males,
+                left = 0.45,
+                right = 0.90,
+                bottom = 0.60,
+                top = 0.75, ignore_tag = TRUE) + 
+    inset_element(females,
+                left = 0.05,
+                right = 0.50,
+                bottom = 0.25,
+                top = 0.38, ignore_tag = TRUE)
+
+p3 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dL, group = Predator, color = Predator)) +
+  ylim(10, 30) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Luminance contrast (JND)") +
+   annotate("segment", x = 0.90, xend = 1.90, y = 28, yend = 28, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 26, yend = 26, colour = "black") + 
+  annotate("text", x = 0.9+((1.90 - 0.90)/2), y = 28.5, label = "**",  size = 6) + 
+  annotate("text", x = 1.12+((2.12 - 1.12)/ 2), y = 26.5, label = "**", size = 6) + 
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+    inset_element(snake,
+                left = 0.45,
+                right = 0.75,
+                bottom = 0.60,
+                top = 0.70, ignore_tag = TRUE) + 
+    inset_element(bird,
+                left = 0.05,
+                right = 0.50,
+                bottom = 0.25,
+                top = 0.38, ignore_tag = TRUE)
+
+p4 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dS, group = Predator, color = Predator)) +
+  ylim(1, 9) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("") + 
+  ylab("Chromatic contrast (JND)") +
+   annotate("segment", x = 0.88, xend = 1.88, y = 8.8, yend = 8.8, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y =2.5, yend = 2.5, colour = "black") + 
+  annotate("text", x = 0.9+((1.90 - 0.90)/2), y = 8.9, label = "**",  size = 6) + 
+  annotate("text", x = 1.12+((2.12 - 1.12)/ 2), y = 2.7, label = "NS", size = 6) + 
+  theme_bw() +
+  theme(legend.position = c(0.10, 0.08),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+    inset_element(bird,
+                left = 0.30,
+                right = 0.60,
+                bottom = 0.45,
+                top = 0.60, ignore_tag = TRUE) + 
+    inset_element(snake,
+                left = 0.02,
+                right = 0.32,
+                bottom = 0.20,
+                top = 0.35, ignore_tag = TRUE)
+
+#pdf(file = "./output/figures/Figure1.1.pdf", width = 14, height = 14)
+plot_grid(p1, p2, p4, p3, labels = c("A", "B", "C", "D"), label_size = 16,
+  label_fontface = "bold", rel_widths=c(1,1))
+#dev.off()
+} else {
+
+  img2 <- image_read("./output/figures/Figure1.png")
+  img2
+}
+```
+
+```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "**Evidence for local adaptation of color signals in both (A) chromatic- and (B) luminance contrast for Hawaiian chameleons**.
+For local adaptation, signals are predicted to be more conspicuous against their own background.
+In this case, Hawaiian chameleons were both significantly more chromatic and brighter (luminance contrast) against their own background compared to the Kenyan background suggesting that selection has acted on increasing local signal conspicuousness."}
+rerun=FALSE
+
+if(rerun){
+##########
+# Data
+##########
+courtship_summary <- CourtshipBgrd %>% 
+                    group_by(Population, Background_pop) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>% filter(Population == "Hawaii") %>% as.data.frame() %>% mutate(Social_Context = "Courtship")
+
+display_summary <- DisplayBgrd %>%
+                group_by(Population, Background_pop) %>%
+                summarise( Mean_dS = mean(dS, na.rm = T), 
+                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
+                             Upper_dS = Mean_dS + 1.96*SE_dS,
+                             Lower_dS = Mean_dS - 1.96*SE_dS, 
+                           Mean_dL = mean(dL, na.rm = T), 
+                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
+                             Upper_dL = Mean_dL + 1.96*SE_dL,
+                             Lower_dL = Mean_dL - 1.96*SE_dL) %>% filter(Population == "Hawaii") %>% as.data.frame() %>% mutate(Social_Context = "Male-male contest")
+
+summary_data_localAdap <- rbind(courtship_summary, display_summary)
+
+##########
+# Plots
+##########
+#Grab photos
+  males <- readPNG("./photos/male_fight.png", native = TRUE)
+females <- readPNG("./photos/courtship.png",  native = TRUE)
+
+#plots
+p1.2 <- ggplot(summary_data_localAdap, aes(x = Background_pop, y = Mean_dS, group = Social_Context, color = Social_Context)) +
+    ylim(2.5, 6.5) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Background Environment") + 
+  ylab("Chromatic contrast (JND)") +
+  annotate("segment", x = 0.85, xend = 1.85, y = 6.1, yend = 6.1, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 5.7, yend = 5.7, colour = "black") + 
+  annotate("text", x = 0.85+(1.85-0.85)/2,  y = 6.2, label = "NS",  size = 6) + 
+  annotate("text", x = 1.12+(2.12-1.12)/2, y = 5.8, label = "NS", size = 6) + 
+  theme_bw() +
+  labs(colour = "Social Context") +
+  theme(legend.position = c(0.15, 0.08),
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+   inset_element(males,
+                left = 0.45,
+                right = 0.90,
+                bottom = 0.58,
+                top = 0.75, ignore_tag = TRUE) + 
+    inset_element(females,
+                left = 0.30,
+                right = 0.75,
+                bottom = 0.15,
+                top = 0.28, ignore_tag = TRUE)
+
+p2.2 <- ggplot(summary_data_localAdap, aes(x = Background_pop, y = Mean_dL, group = Social_Context, color = Social_Context)) +
+  ylim(15, 45) + 
+  geom_point(size = 3, position = position_dodge(width=0.5)) +
+  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
+  geom_line(position = position_dodge(width=0.5)) + 
+  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
+  xlab("Background Environment") + 
+  ylab("Luminance contrast (JND)") +
+   annotate("segment", x = 0.90, xend = 1.90, y = 40, yend = 40, colour = "black") +
+  annotate("segment", x = 1.12, xend = 2.12, y = 42, yend = 42, colour = "black") + 
+  annotate("text", x = 0.9+((1.90 - 0.90)/2), y = 41, label = "**",  size = 6) + 
+  annotate("text", x = 1.12+((2.12 - 1.12)/ 2), y = 43, label = "**", size = 6) + 
+  theme_bw() +
+  theme(legend.position = "none",
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(),
+    axis.text=element_text(size=18),
+    axis.title=element_text(size=20,face="bold"),
+    legend.text =element_text(size = 12),
+    legend.title =element_text(size = 12, face = "bold")) +
+    inset_element(males,
+                left = 0.45,
+                right = 0.90,
+                bottom = 0.60,
+                top = 0.75, ignore_tag = TRUE) + 
+    inset_element(females,
+                left = 0.05,
+                right = 0.50,
+                bottom = 0.25,
+                top = 0.38, ignore_tag = TRUE)
+#pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
+plot_grid(p1.2, p2.2, labels = c("A", "B"), label_size = 16,
+  label_fontface = "bold", rel_widths=c(1,1))
+#dev.off()
+} else {
+  img3 <- image_read("./output/figures/Figure2.png")
+  img3
+}
+
+```
+
+```{r table4, results='hide'}
+#Section will create and write Tables for Supplement. It is not spit out in the main MS. Exported then imported for the supp
+
+# Table S1
+    table1 <- as.data.frame(rbind(robust_ConspicDisplayAC_CI,
+                    robust_ConspicDisplayCC_CI,
+                    robust_ConspicCourtshipAC_CI,
+                    robust_ConspicCourtshipCC_CI))
+    table1$Parameter <- rownames(table1)
+    rownames(table1) <- NULL
+    # Clean
+     final_T1 <- table1 %>% 
+                 mutate(SocialContext = c("Male-Male Display", rep("", 5), "Courtship Display", rep("", 5)),
+                        VisualRealm = c("Luminance  Contrast", rep("", 2), "Chromatic Contrast", rep("", 2), 
+                                        "Luminance Contrast", rep("", 2), "Chromatic Contrast", rep("", 2))) %>%
+                  select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
+     write.csv(final_T1, file = "./output/tables/Table1_MS.csv", row.names = FALSE)
+
+# Table S2
+    table2 <- as.data.frame(rbind(robust_PredatorAC_CI_inter,
+                                  robust_PredatorCC_CI_inter))
+    table2$Parameter <- rownames(table2)
+    rownames(table2) <- NULL
+
+    # Clean
+     final_T2 <- table2 %>% 
+                 mutate(JND = c("Luminance Contrast", rep("", 3), "Chromatic Contrast", rep("", 3))) %>%
+                  select(JND, Parameter, beta, SE, df, CI_L, CI_U)
+     write.csv(final_T2, file = "./output/tables/Table2_MS.csv", row.names = FALSE)
+
+ #Table S3
+    contrast <- rbind(CI_hyp_test,
+                     CI_hyp_testAC)
+
+    contrast <- contrast %>% mutate(VisualSpec = c("Chromatic Contrast", "", "Luminance Contrast", ""),
+                                    p_val = sapply(p_val, function(x) p_value(x))) %>% 
+                select(VisualSpec, comparison, Estimate, lwr, upr, Fstat, df, p_val, p_bonf)
+
+    write.csv(contrast, file = "./output/tables/Table3_MS.csv", row.names = FALSE)
+
+#TableS4
+    tb4 <- data.frame(rbind(robust_localAdapt_DispBgrdAC,
+                 robust_DispBgrdCC,
+                 robust_CourtshipAC,
+                 robust_CourtshipCC))
+    tb4$parameter <- rownames(tb4)
+    rownames(tb4) <- NULL
+
+    tb4 <- tb4 %>% mutate(SocialContext = c("Male-Male Display", rep("", 5), 
+                                            "Courtship Display", rep("", 5)),
+                        VisualRealm = c("Luminance Contrast", rep("", 2), "Chromatic Contrast", 
+                                        rep("", 2), "Luminance Contrast", rep("", 2), 
+                                        "Chromatic Contrast", rep("", 2))) %>%
+                  select(SocialContext, VisualRealm, parameter, beta, SE, df, CI_L, CI_U)
+
+    write.csv(tb4, file = "./output/tables/Table4_MS.csv", row.names = FALSE)
+
+# Table S5
+    tb5 <- data.frame(rbind(robust_Diff_dL_BR_int,
+                            robust_Diff_dS_BR_int))
+    tb5$parameter <- rownames(tb5)
+    rownames(tb5) <- NULL
+
+  tb5 <- tb5 %>% mutate(VisualRealm = c("Luminance Contrast", rep("", 8), "Chromatic Contrast", 
+                                        rep("", 8))) %>%
+                  select(VisualRealm, parameter, beta, SE, df, CI_L, CI_U)
+     write.csv(tb5, file = "./output/tables/Table5_MS.csv", row.names = FALSE)
+
+```
+
 ```{r setup, include=FALSE, purl = FALSE}
@@
 
-```{r display, results = "hide"}
+```{r display_supp, results = "hide"}
 source("./R/func.R")
@@
 Supplementary References
-
-[@RN15ms;@RN7ms;@RN11ms;@RN6ms;@RN22ms;@RN2ms;@RN25ms;@RN8ms;@RN3ms;@RN1ms;@RN13ms;@RN18ms;@RN9ms;@RN12ms;@RN26ms;@RN24ms;@RN19ms;@RN5ms;@Measey2014ms;@RN20ms;@RN14ms;@RN4ms;@RN10ms]
 

### 2021-08-25 biblio.Rmd
Added new ms/supp merged file beacuse of the continuous ref formatting for Science. Some refs are the same from MS and Supp so numbers need to be the same.

@@
 
-```{r setup, include=FALSE, purl = FALSE}
+```{r setup1, include=FALSE, purl = FALSE}
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
@@
 
-```{r getlibrary, echo = FALSE, results = "hide"}
+```{r getlibrary1, echo = FALSE, results = "hide"}
 # load libraries
@@
 
-```{r table4, results='hide'}
-#Section will create and write Tables for Supplement. It is not spit out in the main MS. Exported then imported for the supp
-
-# Table S1
-    table1 <- as.data.frame(rbind(robust_ConspicDisplayAC_CI,
-                    robust_ConspicDisplayCC_CI,
-                    robust_ConspicCourtshipAC_CI,
-                    robust_ConspicCourtshipCC_CI))
-    table1$Parameter <- rownames(table1)
-    rownames(table1) <- NULL
-    # Clean
-     final_T1 <- table1 %>% 
-                 mutate(SocialContext = c("Male-Male Display", rep("", 5), "Courtship Display", rep("", 5)),
-                        VisualRealm = c("Luminance  Contrast", rep("", 2), "Chromatic Contrast", rep("", 2), 
-                                        "Luminance Contrast", rep("", 2), "Chromatic Contrast", rep("", 2))) %>%
-                  select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
-     write.csv(final_T1, file = "./output/tables/Table1_MS.csv", row.names = FALSE)
-
-# Table S2
-    table2 <- as.data.frame(rbind(robust_PredatorAC_CI_inter,
-                                  robust_PredatorCC_CI_inter))
-    table2$Parameter <- rownames(table2)
-    rownames(table2) <- NULL
-
-    # Clean
-     final_T2 <- table2 %>% 
-                 mutate(JND = c("Luminance Contrast", rep("", 3), "Chromatic Contrast", rep("", 3))) %>%
-                  select(JND, Parameter, beta, SE, df, CI_L, CI_U)
-     write.csv(final_T2, file = "./output/tables/Table2_MS.csv", row.names = FALSE)
-
- #Table S3
-    contrast <- rbind(CI_hyp_test,
-                     CI_hyp_testAC)
-
-    contrast <- contrast %>% mutate(VisualSpec = c("Chromatic Contrast", "", "Luminance Contrast", ""),
-                                    p_val = sapply(p_val, function(x) p_value(x))) %>% 
-                select(VisualSpec, comparison, Estimate, lwr, upr, Fstat, df, p_val, p_bonf)
-
-    write.csv(contrast, file = "./output/tables/Table3_MS.csv", row.names = FALSE)
-
-#TableS4
-    tb4 <- data.frame(rbind(robust_localAdapt_DispBgrdAC,
-                 robust_DispBgrdCC,
-                 robust_CourtshipAC,
-                 robust_CourtshipCC))
-    tb4$parameter <- rownames(tb4)
-    rownames(tb4) <- NULL
-
-    tb4 <- tb4 %>% mutate(SocialContext = c("Male-Male Display", rep("", 5), 
-                                            "Courtship Display", rep("", 5)),
-                        VisualRealm = c("Luminance Contrast", rep("", 2), "Chromatic Contrast", 
-                                        rep("", 2), "Luminance Contrast", rep("", 2), 
-                                        "Chromatic Contrast", rep("", 2))) %>%
-                  select(SocialContext, VisualRealm, parameter, beta, SE, df, CI_L, CI_U)
-
-    write.csv(tb4, file = "./output/tables/Table4_MS.csv", row.names = FALSE)
-
-# Table S5
-    tb5 <- data.frame(rbind(robust_Diff_dL_BR_int,
-                            robust_Diff_dS_BR_int))
-    tb5$parameter <- rownames(tb5)
-    rownames(tb5) <- NULL
-
-  tb5 <- tb5 %>% mutate(VisualRealm = c("Luminance Contrast", rep("", 8), "Chromatic Contrast", 
-                                        rep("", 8))) %>%
-                  select(VisualRealm, parameter, beta, SE, df, CI_L, CI_U)
-     write.csv(tb5, file = "./output/tables/Table5_MS.csv", row.names = FALSE)
-
-```
-
 ```{r setup, include=FALSE, purl = FALSE}
@@
 We also conducted seven trials with a just a branch with no predator attached, as a control for the model predator.
-The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
+The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN18ms;@RN10]).
 The bird was a mounted African Cuckoo-hawk (*Aviceda cuculoides*).
@@
 
-We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` @RN29.
+We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN25ms; @RN28] implemented in the R package `pavo` @RN29.
 This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
@@
          sens_snake <- read_excel("./data/Spectral sensitivities.xls", sheet = "snake")[,1:4]
-
+library(tidyverse)
 cham <- sens_cham_c_dilepis %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()

### 2021-08-25 biblio.Rmd
Missed an "ms" for one of ain ms refs and so added one that didn't exist

@@
 \@ref(fig:figure1)).
-Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15ms or changes in the spacing of guanine crystals within iridophores @RN14, both of which strongly affect luminance.
+Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15ms or changes in the spacing of guanine crystals within iridophores @RN14ms, both of which strongly affect luminance.
 Accordingly, character release was observed in the overall luminance of signals rather than their hue.

### 2021-09-11 ms.Rmd
Added the final MS changes. Pretty minor.

@@
 
-Martin J. Whiting^1^, Brenden S. Holland^2^, J. Scott Keogh^3^, Daniel W.A.Noble^3^, Katrina J. Rankin^4^ and Devi Stuart-Fox^4^
+Martin J. Whiting^1,2^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel W.A.Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
 
@@
 
-^2^ Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
-
-^3^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
-
-^4^ School of BioSciences, The University of Melbourne, Parkville, Victoria 3010, Australia
+^2^    School of Animal, Plant and Environmental Sciences, University of the Witwatersrand, Wits 2050, South Africa
+
+^3^ Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
+
+^4^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
+
+^5^ School of BioSciences, The University of Melbourne, Parkville, Victoria 3010, Australia
 
@@
 Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator color states.
-Consistent with local adaptation, the display colors of Hawaiian chameleons were more conspicuous in their local than ancestral habitats.
+Hawaiian chameleon display colors were more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
 These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
+
+**One sentence summary**: Relaxing predation results in character release and local adaptation in a dynamic social signal in a chameleon.
 
@@
 50 -- 65 generations).
-On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014.
+On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014.
 This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.
@@
 S4A).
-Their different color states can appear more or less conspicuous depending on the context, background, and viewer (e.g., conspecific or predator).
-If introduced chameleons in Hawaii experience character release from reduced predation, we expect Hawaiian chameleons to have more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators.
+Their different color states can appear either more, or less, conspicuous depending on the context, background, and viewer (e.g., conspecific or predator).
+If introduced chameleons in Hawaii experience character release from reduced predation, we expect them to have more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators.
 We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.
@@
 To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (fig.
-S4).
+S4), or a control (stick).
 We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
@@
 Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, table S3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, *P* = `r p_value(back_pred_CC_Pred$p_val)`).
-Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.
+Hawaiian chameleons were significantly more conspicuous to birds than were Kenyan conspecifics (Fig.
 \@ref(fig:figure3)D and table S4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`, JND values > 5).
@@
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18].
-The poor color discrimination of snakes may also explain why chameleons from Hawaii show little change relative to their Kenyan counterparts in their chromatic response to snakes, despite the absence of snakes on Hawaii.
+The poor color discrimination of snakes may also explain why Hawaiian and Kenyan chameleons differ little their chromatic response to snakes, despite the absence of snakes on Hawaii.
 
@@
 \@ref(fig:figure4) and table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
-Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].
+These results are consistent with local adaptation, in which phenotypic plasticity, which can also evolve rapidly [@RN26], may also play a role.
 
@@
 Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
-While it is possible that differences between Hawaiian and Kenyan chameleons could reflect a founder effect, the observed differences are all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the local environment.
+Although differences between Hawaiian and Kenyan chameleons might reflect a founder effect, the traits involved and observed differences are all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the environment.
 
@@
 This novel finding raises the question of whether other animals that have convergently evolved dynamic color signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
-More generally, our study highlights the opportunities provided by translocations and invasions, which are often otherwise detrimental, to study natural and sexual selection in the wild.
+More generally, our study highlights the opportunities provided by translocations and invasions to study natural and sexual selection in the wild.
 
@@
 
-```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social and predatory stimuli.
+```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social stimuli.
 Male chameleons experience intense sexual selection**.
@@
 **(B)** A subordinate male that lost a contest and turned from bright yellow to brown.
-**(C)** Two males fighting, both are in display coloration and at a stage where they are relatively even matched.
-**(D)** A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male."}
+**(C)** Two males fighting, both are in display coloration and at a stage when they are relatively even matched.
+**(D)** A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male.
+See supplementary text for additional photos, including in response to a snake."}
 library(magick)
@@
 ```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "**Chromatic and luminance contrast in Just Noticeable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves)**.
-Luminance and chromatic contrast of male chameleons during male-male contests **(A)** and female courtship **(B)**.
+Luminance **(A)** and chromatic **(B)** contrast of male chameleons during male-male contests and female courtship.
 JNDs are calculated based on the chameleon visual system.
-Luminance and chromatic contrasts of male chameleons during snake **(C)** and bird **(D)** predator encounters.
+Luminance **(C)** and chromatic **(D)** contrasts of male chameleons during snake  and birdpredator encounters.
 JNDs are calculated based on the snake and bird visual systems.
@@
 
-```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "**Evidence for local adaptation of color signals in both (A) chromatic- and (B) luminance contrast for Hawaiian chameleons**.
+```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "**Evidence for local adaptation of social color signals in both (A) chromatic- and (B) luminance contrast for Hawaiian chameleons**.
 For local adaptation, signals are predicted to be more conspicuous against their own background.
@@
 
+**Acknowledgments**: The research was approved by the Kenya Wildlife Service (R.
+Bagine, KWS 5001) and the Ministry of Science & Technology (Research Permit No.
+MOS&T 13/36C 208).
+We thank Idle Farah and Patrick Malonza of the National Museums of Kenya for facilitating the research, Joseph Gathua for field support and Jan Stipala for use of his field vehicle.
+We thank the Ditsong National Museum of Natural History (formerly Transvaal Museum) for the stuffed African Cuckoo Hawk.
+Ben Phillips, Rick Shine and Tobias Uller provided feedback on earlier drafts of this manuscript.
+We thank Fonti Kar for help curating data and Drew Allen for statistical advice.
+Funding: This work was supported by the University of the Witwatersrand.
+Authors contributions: MJW conceived the study.
+MJW and DSF designed the study with help from JSK and BH.
+MJW, BSH, JSK and DSF collected the data.
+MJW, DSF, KJR and DWAN performed data curation, visualisation and exploratory analyses.
+DWAN performed final analyses.
+MJW wrote the first draft, with help from DWAN and DSF.
+All authors contributed substantially to subsequent versions.
+**Competing interests**: we (the authors) declare no competing interests.
+**Data and materials availability**: the data and code are available at OSF.
+
+**List of Supplementary materials**
+Materials and Methods
+Figures S1 – S9
+Tables S1 – S13
+References (24-48)
+
 # References and Notes

### 2021-09-11 ms.Rmd
Final changes to get to the word count

@@
 S1 and S3).
-At other times, they are a duller green and can darken to brown or become mottled green and brown, particularly in response to perceived predation risk (Fig.
+At other times, they are a duller green, brown or become mottled green-brown, particularly in response to predators (Fig.
 \@ref(fig:figure1) and fig.
 S4A).
-Their different color states can appear either more, or less, conspicuous depending on the context, background, and viewer (e.g., conspecific or predator).
-If introduced chameleons in Hawaii experience character release from reduced predation, we expect them to have more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators.
+Their different color states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
+If introduced chameleons in Hawaii experience character release from reduced predation, we expect males to have more conspicuous display colors in response to conspecifics and to be less cryptic in response to bird and snake predators.
 We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.
@@
 S5).
-We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25) against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25) against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
 
@@
 \@ref(fig:figure3)A and table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
-This result was consistent for all body regions measured (supplementary text).
-However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.
+This result was consistent for all body regions (supplementary text).
+However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of color signals (Fig.
 \@ref(fig:figure3)B and table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
@@
 Character release may reflect rapid evolutionary change, phenotypic plasticity or a combination of these processes [@RN19; @RN20].
-Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
+Rapid evolutionary change results in local adaptation, whereby signals become more conspicuous to conspecifics in their local environment than in ancestral environments.
 Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.
@@
 To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
-While there was evidence for differences between body regions (see supplementary text, fig.
+While there was evidence for differences between body regions (supplementary text, fig.
 S9, tables S12 and S13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (fig.
@@
 This novel finding raises the question of whether other animals that have convergently evolved dynamic color signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
-More generally, our study highlights the opportunities provided by translocations and invasions to study natural and sexual selection in the wild.
+More generally, our study highlights the opportunities provided by invasions to study natural and sexual selection in the wild.
 
@@
 Male chameleons experience intense sexual selection**.
-During the breeding season they change from a neutral state (green) to a highly conspicuous bright yellow display signal.
+During the breeding season they change from dull green to a highly conspicuous bright yellow display signal.
 They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
-**(A)** A typical dominant male in display coloration.
+**(A)** A dominant male in display coloration.
 **(B)** A subordinate male that lost a contest and turned from bright yellow to brown.
-**(C)** Two males fighting, both are in display coloration and at a stage when they are relatively even matched.
-**(D)** A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male.
+**(C)** Two males fighting, both are in display coloration and relatively even matched.
+**(D)** A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
 See supplementary text for additional photos, including in response to a snake."}
@@
 
-**Acknowledgments**: The research was approved by the Kenya Wildlife Service (R.
-Bagine, KWS 5001) and the Ministry of Science & Technology (Research Permit No.
+**Acknowledgments**: This research was approved by the Kenya Wildlife Service (R.
+Bagine, KWS 5001) and Ministry of Science & Technology (Research Permit No.
 MOS&T 13/36C 208).
-We thank Idle Farah and Patrick Malonza of the National Museums of Kenya for facilitating the research, Joseph Gathua for field support and Jan Stipala for use of his field vehicle.
-We thank the Ditsong National Museum of Natural History (formerly Transvaal Museum) for the stuffed African Cuckoo Hawk.
-Ben Phillips, Rick Shine and Tobias Uller provided feedback on earlier drafts of this manuscript.
-We thank Fonti Kar for help curating data and Drew Allen for statistical advice.
-Funding: This work was supported by the University of the Witwatersrand.
-Authors contributions: MJW conceived the study.
+**Funding**: This work was supported by the University of the Witwatersrand.
+**Authors contributions**: MJW conceived the study.
 MJW and DSF designed the study with help from JSK and BH.

### 2021-09-11 supp.Rmd
Supp comments integrated

@@
 
-Correspondance: Martin J. Whiting  
+Correspondence: Martin J. Whiting  
 Email: martin.whiting@mq.edu.au
@@
 We are not aware of any evidence of bird predation of chameleons on Oahu.
-We consulted with several ornithologists in Hawaii regarding feeding behavior of the few predatory bird species that consume vertebrate prey, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons @RN1.
+We consulted with several ornithologists in Hawaii regarding feeding behavior of the few predatory bird species that consume vertebrate prey, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons @RN1.
 Among the larger, predatory birds, two are owls (*Asio flammeus sandwichensis*, *Tyto alba*), which are extremely rare on Oahu.
@@
 The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Fig.
-S\@ref(fig:Figcolsig2)C,D).
+S\@ref(fig:Figcolsig2)C, D).
 Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
@@
 The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
-S\@ref(fig:Figcolsig2)A,B).
+S\@ref(fig:Figcolsig2)A, B).
 Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.
@@
 We conducted all conspecific-social trials first, in case anti-predator trials unduly influenced their social interactions.
+Our analyses focused on comparing color responses of chameleons to conspecifics and predators to ensure that the color state corresponded to specific stimuli.
+We did not, and could not, measure a ‘neutral’ color state because it is impossible to meaningfully gauge a neutral state in a color-changing organism such as a chameleon.
+However, for predator trials, we ensured that chameleons were specifically responding to the predators by conducting seven trials with a just a branch with no predator attached, as a control.
+Chameleons did not respond to the branch (see also @RN18ms).
+Furthermore, during trials with predators, chameleons focused on the predator itself, and exhibited classic anti-predator behavior (e.g., contrasting stress color, body inflation, open-mouth threat).
+
 Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
@@
 This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Fig.
-S\@ref(fig:Figcolsig)A,B).
+S\@ref(fig:Figcolsig)A, B).
 In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
@@
 Once chameleons changed color, we quantified their spectral reflectance (details below).
-We also conducted seven trials with a just a branch with no predator attached, as a control for the model predator.
 The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
@@
 ```
+# Extended acknowledgments
+We thank Idle Farah and Patrick Malonza of the National Museums of Kenya for facilitating the research, Joseph Gathua for field support and Jan Stipala for use of his field vehicle.
+We thank the Ditsong National Museum of Natural History (formerly Transvaal Museum) for the stuffed African Cuckoo Hawk.
+Ben Phillips, Rick Shine and Tobias Uller provided feedback on earlier drafts of this manuscript.
+We thank Fonti Kar for help curating data and Drew Allen for statistical advice.
 

### 2021-09-14 ms.Rmd
Changes 'perfect storm' to unqiue set

@@
 \@ref(fig:figure1)) in which males lock and twist their three horns into rivals during intense bouts that sometimes result in physical injury.
-Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
+Consequently, a unique set of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
 To the best of our knowledge, this is also the first example of character release in a dynamic color signal.

### 2021-12-02 ms.Rmd
Added new analyses to address reviewer comments

@@
                                   cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+#########################################################***** NEW
+# Background Analysis - Founder           
+#########################################################***** NEW
+DisplayBgrd <- read.table("./data/Display_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+svl_dat <- read.csv("./data/SVL2.csv")
+
+DisplayBgrd <- DisplayBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, 
+                                               if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>%
+                 as.data.frame() #exclude bottom flank.
+
+DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
+
+DisplayBgrd <- DisplayBgrd %>% 
+                mutate(pop_background = interaction(Population, Background_pop)) %>%  
+                left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+                mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
+#unique(DisplayBgrd$pop_background)
+###################################
+# Male-male display
+###################################
+##########
+# Achromatic Contrast - dL
+##########
+  # Fit the main effect model
+  DispBgrdAC_s6 <- lmer(dL ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC_s6)
+
+  # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
+  robust_localAdapt_DispBgrdAC_s6 <- clubSandwich::conf_int(DispBgrdAC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+##########
+# Chromatic Contrast - dS
+##########
+    # Fit the main effect model
+          DispBgrdCC_s6 <- lmer(dS ~ sc_svl + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=DisplayBgrd)
+          summary(DispBgrdCC_s6)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+      robust_DispBgrdCC_s6 <- clubSandwich::conf_int(DispBgrdCC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+###################
+# Social Context: Courtship
+##################
+CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+CourtshipBgrd <- CourtshipBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
+
+# Filter out only Hawaiian and add in SVL for IDs
+CourtshipBgrd <- CourtshipBgrd %>% 
+               mutate(pop_background = interaction(Population, Background_pop)) %>%  left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
+##########
+## Achromatic Contrast - dL
+##########
+
+  # Fit main effects model
+    CourtshipAC_s6 <- lmer(dL ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=CourtshipBgrd)
+    summary(CourtshipAC_s6)
+
+   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+   robust_CourtshipAC_s6 <- clubSandwich::conf_int(CourtshipAC_s6, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+##########
+# Chromatic - dS
+##########
+
+  # Fit main effect model
+      CourtshipCC_s6 <- lmer(dS ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC_s6)
+
+  # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+  robust_CourtshipCC_s6 <- clubSandwich::conf_int(CourtshipCC_s6, cluster = CourtshipBgrd$ID, vcov =  "CR2")
 

### 2021-12-02 ms.Rmd
Added new table

@@
 CourtshipBgrd <- CourtshipBgrd %>% 
-               mutate(pop_background = interaction(Population, Background_pop)) %>%  left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(pop_background = interaction(Population, Background_pop)) %>%  
+                left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
                mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
@@
 
+# Table S6
+tableS6 <- as.data.frame(rbind(robust_localAdapt_DispBgrdAC_s6,
+                               robust_DispBgrdCC_s6,
+                               robust_CourtshipAC_s6,
+                              robust_CourtshipCC_s6))
+tableS6$Parameter <- rownames(tableS6)
+rownames(tableS6) <- NULL
+
+# Clean
+final_S6 <- tableS6 %>% 
+  mutate(SocialContext = c("Male-Male Display", rep("", 9), "Courtship Display", rep("", 9)),
+         VisualRealm = c("Luminance  Contrast", rep("", 4), "Chromatic Contrast", rep("", 4), 
+                         "Luminance Contrast", rep("", 4), "Chromatic Contrast", rep("", 4))) %>%
+  select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
+write.csv(tableS6, file = "./output/tables/tableS6.csv", row.names = FALSE)
+
 ```

### 2021-12-02 ms.Rmd
added proper sleect

--- 2021-12-02 ms.Rmd
@@
 
+# Table S6
+tableS6 <- as.data.frame(rbind(robust_localAdapt_DispBgrdAC_s6,
+                               robust_DispBgrdCC_s6,
+                               robust_CourtshipAC_s6,
+                              robust_CourtshipCC_s6))
+tableS6$Parameter <- rownames(tableS6)
+rownames(tableS6) <- NULL
+
+# Clean
+final_S6 <- tableS6 %>% 
+  mutate(SocialContext = c("Male-Male Display", rep("", 9), "Courtship Display", rep("", 9)),
+         VisualRealm = c("Luminance  Contrast", rep("", 4), "Chromatic Contrast", rep("", 4), 
+                         "Luminance Contrast", rep("", 4), "Chromatic Contrast", rep("", 4))) %>%
+  dplyr::select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
+write.csv(tableS6, file = "./output/tables/tableS6.csv", row.names = FALSE)
+
 ```

### 2021-12-02 ms.Rmd
Added export of correct object

--- 2021-12-02 ms.Rmd
@@
 
+# Table S6
+tableS6 <- as.data.frame(rbind(robust_localAdapt_DispBgrdAC_s6,
+                               robust_DispBgrdCC_s6,
+                               robust_CourtshipAC_s6,
+                              robust_CourtshipCC_s6))
+tableS6$Parameter <- rownames(tableS6)
+rownames(tableS6) <- NULL
+
+# Clean
+final_S6 <- tableS6 %>% 
+  mutate(SocialContext = c("Male-Male Display", rep("", 9), "Courtship Display", rep("", 9)),
+         VisualRealm = c("Luminance  Contrast", rep("", 4), "Chromatic Contrast", rep("", 4), 
+                         "Luminance Contrast", rep("", 4), "Chromatic Contrast", rep("", 4))) %>%
+  dplyr::select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
+write.csv(final_S6, file = "./output/tables/tableS6.csv", row.names = FALSE)
+
 ```

### 2021-12-02 supp.Rmd
Added new Table S6 with interaction model

@@
 
+```{r tableS6_new, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of interaction models comparing luminance and chromatic contrast of Hawaiian and Kenyan lizards against a Hawaiian and Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
+Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for models were fitted using robust variance estimators to correct standard errors for non-independence, and we used Satterthwaite's degrees of freedom.
+Significant parameter estimates are those where the 95% CIs do not include zero.
+Relevant contrasts that compare Hawaiian and Kenyan populations both againts the same Kenyan background are indicated by '*'."}
+source("./R/func.R")
+final_S6 <-  read.csv(file = "./output/tables/tableS6.csv")
+
+  # Table
+ Tab_S6 <- flextable(final_S6) %>%  table_style6() %>% fit_to_width(max_width = 7)
+ table1
+
+```
 # Additional Supplemental Results

### 2021-12-02 R/func.R
Added new function for Table S6

@@ -395,3 +395,26 @@ table_style14  <- function(table){
   
   return(table)
 }
+
+
+table_styleS6  <- function(table){
+  table <- table  %>% 
+    align(align = "center", part = "header") %>% 
+    align(align = "center", part = "body") %>% 
+    flextable::compose(part = "header", j = 1, value = as_paragraph(as_b("Social Context"))) %>% 
+    flextable::compose(part = "header", j = 2, value = as_paragraph(as_b("Visual Spectrum"))) %>% 
+    flextable::compose(part = "header", j = 3, value = as_paragraph(as_b("Parameter"))) %>% 
+    flextable::compose(part = "header", j = 4, value = as_paragraph(as_b("Estimate"))) %>% 
+    flextable::compose(part = "header", j = 5, value = as_paragraph(as_b("SE"))) %>%
+    flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("df"))) %>% 
+    flextable::compose(part = "header", j = 7, value = as_paragraph(as_b("95% CI"), as_sub(as_b("lower")))) %>% 
+    flextable::compose(part = "header", j = 8, value = as_paragraph(as_b("95% CI"), as_sub(as_b("upper")))) %>%
+    flextable::compose(part = "body", j = 3, i = c(1,6,11,16), value = as_paragraph("Intercept", as_sub(" Hawaii-Kenya"))) %>%
+    flextable::compose(part = "body", j = 3, i = c(1,6,11,16)+1, value = as_paragraph("Snout-vent length (SVL)")) %>%
+    flextable::compose(part = "body", j = 3, i = c(1,6,11,16)+2, value = as_paragraph("Population", as_sub(" Hawaii"),"-", "Background", as_sub(" Hawaii"))) %>% 
+    flextable::compose(part = "body", j = 3, i = c(1,6,11,16)+3, value = as_paragraph("Population", as_sub(" Kenya"),"-", "Background", as_sub(" Hawaii"))) %>% 
+    flextable::compose(part = "body", j = 3, i = c(1,6,11,16)+4, value = as_paragraph("Population", as_sub(" Kenya"),"-", "Background", as_sub(" Kenya"))) %>% 
+    font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header")
+  return(table)
+}

### 2021-12-06 ms.Rmd
Adding in interaction F test to make it a bit easier as the reviewer will probably think a litle more like this. But, doesn't make sense to have interaction becasue, by definition it won't matter becasue we are just shifting the envirobmenbt while the popultion of individuals remains the same, hence why only one comparison is really needed

@@
 ##########
-  # Fit the main effect model
+  # Fit the interaction model
+  DispBgrdAC_s6 <- lmer(dL ~ sc_svl  + Population*Background_pop + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+  F_AC_Disp_pop_bgk <- Wald_test(DispBgrdAC_s6, constraints = constrain_zero(5), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
   DispBgrdAC_s6 <- lmer(dL ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=DisplayBgrd)
@@
 ##########
-    # Fit the main effect model
+    # Fit the interaction  model
+      DispBgrdCC_s6 <- lmer(dS ~ sc_svl + Population*Background_pop  + (1|ID), data=DisplayBgrd)
+      summary(DispBgrdCC_s6)
+
+    # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+      F_CC_Disp_pop_bgk <- Wald_test(DispBgrdCC_s6, constraints = constrain_zero(5), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+      # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
     DispBgrdCC_s6 <- lmer(dS ~ sc_svl + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=DisplayBgrd)
@@
 
-  # Fit main effects model
+  # Fit interaction model
+    CourtshipAC_s6 <- lmer(dL ~ sc_svl  + Population*Background_pop  + (1|ID), data=CourtshipBgrd)
+    summary(CourtshipAC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+      F_AC_Court_pop_bgk <- Wald_test(CourtshipAC_s6, constraints = constrain_zero(5), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
     CourtshipAC_s6 <- lmer(dL ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=CourtshipBgrd)
@@
   # Fit main effect model
+      CourtshipCC_s6 <- lmer(dS ~ sc_svl  + Population*Background_pop  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+      F_CC_Court_pop_bgk <- Wald_test(CourtshipCC_s6, constraints = constrain_zero(5), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
       CourtshipCC_s6 <- lmer(dS ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=CourtshipBgrd)
       summary(CourtshipCC_s6)
-
   # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 

### 2021-12-06 ms.Rmd
Added inetarction models

@@
       summary(DispBgrdCC_s6)
-
-    # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+    clubSandwich::conf_int(DispBgrdCC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+    # Interaction test.
+Does the difference between Hawaii and Kenya depend on the background.
+It shouldn't.
+Wald F-test.
+Although, somthing might be up with this one - effect size tiny but goes from non-sig to signficant.
       F_CC_Disp_pop_bgk <- Wald_test(DispBgrdCC_s6, constraints = constrain_zero(5), 
@@
       summary(CourtshipCC_s6)
+
   # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 

### 2022-01-22 ms.Rmd
Did text changes to first 3 paragraphs, but this can be reverted back to Sci Advances

@@
 # Abstract
-Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid evolutionary change.
-We demonstrate such character release for color signals in Jackson’s chameleons (*Trioceros j.
-xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) where there are few lizard predators.
+Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid phenotypic change [@Losos2004; @RN4; @RN5; @Stastny2017].
+In particular, conspicuous social and sexual signals are predicted to experience significant character release when natural selection via predation is relaxed [@RN7; @RN4].
+However, we have few good examples of this phenomenon in the wild and none in species with dynamic colour change.
+Here, we show that Jackson’s chameleons (*Trioceros j.
+xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu), where there are few lizard predators, experienced significant character release of color signals.
 Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator color states.
-Hawaiian chameleon display colors were more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
-These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
+Hawaiian chameleon display colors were also more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
+These results demonstrate for the first time that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
 
@@
 
-When species encounter novel environments, such as during human-mediated introductions, evolutionary change can occur more rapidly than Darwin ever thought possible [@RN1; @RN2; @RN3; @RN4].
+When species encounter novel environments, such as during human-mediated introductions, phenotypic change can occur more rapidly than Darwin ever thought possible [@RN1; @RN2; @RN3; @RN4].
 In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4; @RN5].
@@
 In 1972, a shipment of Jackson’s chameleons (*Trioceros jacksonii xanthopholus*; Fig.
-\@ref(fig:figure1) and fig.
-S1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
+\@ref(fig:figure1) and Supplementary Fig.
+1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
 The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11].
-Jackson’s chameleons have a high intrinsic population growth rate (9--12 month generation time and live birth of up to 50 young @RN24), enabling their rapid establishment and potentially rapid evolution (approx.
+Jackson’s chameleons have a high intrinsic population growth rate (9-12 month generation time and live birth of up to 50 young @RN24), enabling their rapid establishment and potentially rapid evolution (approx.
 50 -- 65 generations).
-On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014.
+On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators absent or rare, Supplementary Information), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014.
 This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.
@@
 S5).
-We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25) against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
 

### 2022-01-22 supp.Rmd
All updated.
I've updated the new additinal text to the Sci Advances MS where we made changes to the supplement. There are a few added sentences. We can revert to this if we sent back to Sci Advances

@@
 Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006.
+We did this work during what we expected to be the reproductive season, based on local knowledge.
+The timing is different because of seasonal/geographic differences between Kenya and Hawaii.
+Importantly, the chameleons all exhibited very strong behavioral responses during courtship and male-male competition, consistent with reproductive behavior.
 In Hawaii, chameleons were collected from a single population in the forested Ko’olau Mountains north of Honolulu, between upper Makiki Valley and the summit of Tantalus Mountain, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
@@
 S\@ref(fig:Figcolsig2)A, B).
-Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.
+Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.Jackson’s chameleons are a generalist species which can be locally abundant and relatively mobile.
+This population is from the region (slopes of Mount Kenya) identified as the source population [@Tilbury2018].
 
@@
   # Table
- Tab_S6 <- flextable(final_S6) %>%  table_style6() %>% fit_to_width(max_width = 7)
- table1
+ Tab_S6 <- flextable(final_S6) %>%  table_styleS6() %>% fit_to_width(max_width = 7)
+ Tab_S6
 

### 2022-01-22 ms_Nature.Rmd
Fixed up the Nature MS. New text added, and methods now merged with the main MS. New headers added. Plus, additional sentences added by Martin

--- 2021-08-25 biblio.Rmd
@@
 title: "Invasive chameleons released from predation display more conspicuous colors"
-bibliography: supp_refs.bib
+bibliography: ms_refs.bib
 csl: science.csl
@@
     bibliography: refs.bib
+editor_options: 
+  chunk_output_type: console
 ---
@@
 
+#########################################################***** NEW
+# Background Analysis - Founder           
+#########################################################***** NEW
+DisplayBgrd <- read.table("./data/Display_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+svl_dat <- read.csv("./data/SVL2.csv")
+
+DisplayBgrd <- DisplayBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, 
+                                               if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>%
+                 as.data.frame() #exclude bottom flank.
+
+DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
+
+DisplayBgrd <- DisplayBgrd %>% 
+                mutate(pop_background = interaction(Population, Background_pop)) %>%  
+                left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+                mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
+#unique(DisplayBgrd$pop_background)
+###################################
+# Male-male display
+###################################
+##########
+# Achromatic Contrast - dL
+##########
+  # Fit the interaction model
+  DispBgrdAC_s6 <- lmer(dL ~ sc_svl  + Population*Background_pop + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+  F_AC_Disp_pop_bgk <- Wald_test(DispBgrdAC_s6, constraints = constrain_zero(5), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
+  DispBgrdAC_s6 <- lmer(dL ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC_s6)
+
+  # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
+  robust_localAdapt_DispBgrdAC_s6 <- clubSandwich::conf_int(DispBgrdAC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+##########
+# Chromatic Contrast - dS
+##########
+    # Fit the interaction  model
+      DispBgrdCC_s6 <- lmer(dS ~ sc_svl + Population*Background_pop  + (1|ID), data=DisplayBgrd)
+      summary(DispBgrdCC_s6)
+    clubSandwich::conf_int(DispBgrdCC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+    # Interaction test.
+Does the difference between Hawaii and Kenya depend on the background.
+It shouldn't.
+Wald F-test.
+Although, somthing might be up with this one - effect size tiny but goes from non-sig to signficant.
+      F_CC_Disp_pop_bgk <- Wald_test(DispBgrdCC_s6, constraints = constrain_zero(5), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+      # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
+    DispBgrdCC_s6 <- lmer(dS ~ sc_svl + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=DisplayBgrd)
+          summary(DispBgrdCC_s6)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+      robust_DispBgrdCC_s6 <- clubSandwich::conf_int(DispBgrdCC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+###################
+# Social Context: Courtship
+##################
+CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+CourtshipBgrd <- CourtshipBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
+
+# Filter out only Hawaiian and add in SVL for IDs
+CourtshipBgrd <- CourtshipBgrd %>% 
+               mutate(pop_background = interaction(Population, Background_pop)) %>%  
+                left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
+##########
+## Achromatic Contrast - dL
+##########
+
+  # Fit interaction model
+    CourtshipAC_s6 <- lmer(dL ~ sc_svl  + Population*Background_pop  + (1|ID), data=CourtshipBgrd)
+    summary(CourtshipAC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+      F_AC_Court_pop_bgk <- Wald_test(CourtshipAC_s6, constraints = constrain_zero(5), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
+    CourtshipAC_s6 <- lmer(dL ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=CourtshipBgrd)
+    summary(CourtshipAC_s6)
+
+   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+   robust_CourtshipAC_s6 <- clubSandwich::conf_int(CourtshipAC_s6, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+##########
+# Chromatic - dS
+##########
+
+  # Fit main effect model
+      CourtshipCC_s6 <- lmer(dS ~ sc_svl  + Population*Background_pop  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+      F_CC_Court_pop_bgk <- Wald_test(CourtshipCC_s6, constraints = constrain_zero(5), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
+      CourtshipCC_s6 <- lmer(dS ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC_s6)
+
+  # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+  robust_CourtshipCC_s6 <- clubSandwich::conf_int(CourtshipCC_s6, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
 #########################################################
@@
 
-^2^ Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
-
-^3^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
-
-^4^ School of BioSciences, The University of Melbourne, Parkville, Victoria 3010, Australia
+^2^    School of Animal, Plant and Environmental Sciences, University of the Witwatersrand, Wits 2050, South Africa
+
+^3^ Department of Natural Sciences, Hawaii Pacific University, Honolulu, Hawaii, USA
+
+^4^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2602, Australia
+
+^5^ School of BioSciences, The University of Melbourne, Parkville, Victoria 3010, Australia
 
 # Abstract
-Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid evolutionary change.
-We demonstrate such character release for color signals in Jackson’s chameleons (*Trioceros j.
-xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) where there are few lizard predators.
+Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid phenotypic change [@Losos2004; @RN4; @RN5; @Stastny2017].
+In particular, conspicuous social and sexual signals are predicted to experience significant character release when natural selection via predation is relaxed [@RN7; @RN4].
+However, we have few good examples of this phenomenon in the wild and none in species with dynamic colour change.
+Here, we show that Jackson’s chameleons (*Trioceros j.
+xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu), where there are few lizard predators, experienced significant character release of color signals.
 Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator color states.
-Consistent with local adaptation, the display colors of Hawaiian chameleons were more conspicuous in their local than ancestral habitats.
-These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
-
-When species encounter novel environments, such as during human-mediated introductions, evolutionary change can occur more rapidly than Darwin ever thought possible [@RN1ms; @RN2ms; @RN3ms; @RN4ms].
-In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4ms; @RN5ms].
-Such character release should be most evident for social and sexual signals; released from constraints on conspicuousness imposed by predation risk, signals could elaborate in many directions [e.g., @RN4ms; @RN6ms; @RN7ms; @RN8ms].
-However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7ms; @RN9ms; @RN10ms].
+Hawaiian chameleon display colors were also more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
+These results demonstrate for the first time that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
+
+When species encounter novel environments, such as during human-mediated introductions, phenotypic change can occur more rapidly than Darwin ever thought possible [@RN1; @RN2; @RN3; @RN4].
+In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4; @RN5].
+Such character release should be most evident for social and sexual signals; released from constraints on conspicuousness imposed by predation risk, signals could elaborate in many directions [e.g., @RN4; @RN6; @RN7; @RN8].
+However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7; @RN9; @RN10].
 This is especially true for species with dynamic color change, for which we are not aware of any examples.
@@
 In 1972, a shipment of Jackson’s chameleons (*Trioceros jacksonii xanthopholus*; Fig.
-\@ref(fig:figure1) and fig.
-S1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
-The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11ms].
-Jackson’s chameleons have a high intrinsic population growth rate (9--12 month generation time and live birth of up to 50 young @RN24ms), enabling their rapid establishment and potentially rapid evolution (approx.
+\@ref(fig:figure1) and Supplementary Fig.
+1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
+The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11].
+Jackson’s chameleons have a high intrinsic population growth rate (9-12 month generation time and live birth of up to 50 young @RN24), enabling their rapid establishment and potentially rapid evolution (approx.
 50 -- 65 generations).
-On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014ms.
+On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators absent or rare, Supplementary Information), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014.
 This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.
 
-Chameleons are famous for their rapid color change [@RN12ms; @RN13ms; @RN14ms].
+Chameleons are famous for their rapid color change [@RN12; @RN13; @RN14].
 When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green during courtship or to signal dominance (Fig.
@@
 S1 and S3).
-At other times, they are a duller green and can darken to brown or become mottled green and brown, particularly in response to perceived predation risk (Fig.
+At other times, they are a duller green, brown or become mottled green-brown, particularly in response to predators (Fig.
 \@ref(fig:figure1) and fig.
 S4A).
-Their different color states can appear more or less conspicuous depending on the context, background, and viewer (e.g., conspecific or predator).
-If introduced chameleons in Hawaii experience character release from reduced predation, we expect Hawaiian chameleons to have more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators.
+Their different color states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
+If introduced chameleons in Hawaii experience character release from reduced predation, we expect males to have more conspicuous display colors in response to conspecifics and to be less cryptic in response to bird and snake predators.
 We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.
@@
 S5).
-We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; supplementary text).
-
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
+
+# Results and Discussion
 In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.
 \@ref(fig:figure2), Fig.
-\@ref(fig:figure3)A and table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
-This result was consistent for all body regions measured (supplementary text).
-However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.
-\@ref(fig:figure3)B and table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
+\@ref(fig:figure3)A and Supplementary Table 2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
+This result was consistent for all body regions (Supplementary Information).
+However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of color signals (Fig.
+\@ref(fig:figure3)B and Supplementary Table 2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
 
@@
 \@ref(fig:figure1)).
-Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15ms or changes in the spacing of guanine crystals within iridophores @RN14, both of which strongly affect luminance.
+Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15 or changes in the spacing of guanine crystals within iridophores @RN14, both of which strongly affect luminance.
 Accordingly, character release was observed in the overall luminance of signals rather than their hue.
@@
 Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.
-\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, *P* = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, table S3).
-Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, table S3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, *P* = `r p_value(back_pred_CC_Pred$p_val)`).
-Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.
-\@ref(fig:figure3)D and table S4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`, JND values > 5).
+\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, *P* = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, Supplementary Table 3).
+Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, Supplementary Table 3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, *P* = `r p_value(back_pred_CC_Pred$p_val)`).
+Hawaiian chameleons were significantly more conspicuous to birds than were Kenyan conspecifics (Fig.
+\@ref(fig:figure3)D and Supplementary Table 4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`, JND values > 5).
 By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.
-\@ref(fig:figure3)D and table S4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1; JND values ~ 2).
-The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18ms].
-The poor color discrimination of snakes may also explain why chameleons from Hawaii show little change relative to their Kenyan counterparts in their chromatic response to snakes, despite the absence of snakes on Hawaii.
-
-Character release may reflect rapid evolutionary change, phenotypic plasticity or a combination of these processes [@RN19ms; @RN20ms].
-Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments.
+\@ref(fig:figure3)D and Supplementary Table 4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1; JND values ~ 2).
+The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18].
+The poor color discrimination of snakes may also explain why Hawaiian and Kenyan chameleons differ little their chromatic response to snakes, despite the absence of snakes on Hawaii.
+
+Character release may reflect rapid evolutionary change, phenotypic plasticity or a combination of these processes [@RN19; @RN20].
+Rapid evolutionary change results in local adaptation, whereby signals become more conspicuous to conspecifics in their local environment than in ancestral environments.
 Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.
-\@ref(fig:figure4) and table S5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
-Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26ms].
+\@ref(fig:figure4) and Supplementary Table 5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
+These results are consistent with local adaptation, in which phenotypic plasticity, which can also evolve rapidly [@RN26], may also play a role.
 
 To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
-While there was evidence for differences between body regions (see supplementary text, fig.
-S9, tables S12 and S13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (fig.
-S9).
+While there was evidence for differences between body regions (Supplementary Information, Supplementary Fig.
+9, Supplementary Tables 12 and 13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (Supplementary Fig.
+9).
 Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
-While it is possible that differences between Hawaiian and Kenyan chameleons could reflect a founder effect, the observed differences are all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the local environment.
-
-We have documented a surprisingly strong effect of recent release from predation on a conspicuous, dynamic visual signal.
+An alternate explanation is that differences might be due to a founder effect because chameleons might have been selectively collected for the pet trade.
+While we acknowledge this possibility, it is exceedingly unlikely.
+Chameleons are almost always encountered in a camouflaged color state because they are solitary and only use display colours during relatively brief social encounters which precludes the possibility that hunters selected the most brightly coloured individuals.
+Nevertheless, even if pet traders were selecting for larger animals, our analyses controlled for any relationship body size might have on colour.
+Finally, the traits we quantified, and their observed differences, were all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the environment.
+
+Differences in colouration between populations more broadly can be explained by environmental effects such as diet or climatic variables (e.g., temperature, rainfall/humidity).
+However, the measured colours and colour change represent a dynamic response to specific stimuli (conspecific, model predator) in contrast to the ‘fixed’ colour signals of many birds, fishes and lizards that have been shown to be affected by diet and other environmental factors.
+Additionally, the mechanisms underlying the rapid colour response of chameleons do not involve dietary pigments, and luminance change is partially structural [@RN14].
+Therefore, environmental effects (other than predation) are unlikely to account for major differences in the color responses we observed between populations.
+The degree to which colour change capacity is plastic in chameleons and other organisms with dynamic colour change is unknown.
+We cannot exclude the role of phenotypic plasticity.
+Plasticity can evolve and can itself be an adaptive response.
+The population differences we observed are likely an interaction between genetic and environmental factors which can only be disentangled using common garden studies and for the purposes of this study, is a moot point.
+What is significant about our work is a surprisingly strong effect of recent release from predation on a conspicuous, dynamic visual signal.
+
 Oahu has no snakes or raptors that feed on lizards.
-The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22ms].
-In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014ms -- the conditions were highly favorable for character release of social signals.
+The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22].
+In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014 -- the conditions were highly favorable for character release of social signals.
 Jackson’s chameleons likely experience strong sexual selection given their polygynous mating system, elaborate courtship displays and intense male contest competition (Fig.
 \@ref(fig:figure1)) in which males lock and twist their three horns into rivals during intense bouts that sometimes result in physical injury.
-Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
+Consequently, a unique set of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
 To the best of our knowledge, this is also the first example of character release in a dynamic color signal.
 
-Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12ms].
+Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12].
 Our results suggest that even for species capable of dynamic color change there is an upper threshold to signal intensity that is constrained by natural selection.
 This novel finding raises the question of whether other animals that have convergently evolved dynamic color signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
-More generally, our study highlights the opportunities provided by translocations and invasions, which are often otherwise detrimental, to study natural and sexual selection in the wild.
+More generally, our study highlights the opportunities provided by invasions to study natural and sexual selection in the wild.
+
+# Methods
+## Study area and collection of chameleons
+
+We collected most chameleons (Fig.
+S\@ref(fig:Figcolsig2)) at night by spotlighting, and a small proportion of individuals during the day.
+When chameleons were beyond reach in the forest canopy, we used lightweight 6-m extensible graphite composite poles (South Bend Kwik Stix™ 20' Telescopic Fishing Pole), that chameleons could be encouraged to step on to, before lowering them to the ground.
+We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
+We measured snout-vent length (SVL), a standard measure of body size in lizards, from the tip of the snout to the posterior edge of the vent using a plastic ruler to the nearest 1 mm.
+Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006.
+We did this work during what we expected to be the reproductive season, based on local knowledge.
+The timing is different because of seasonal/geographic differences between Kenya and Hawaii.
+Importantly, the chameleons all exhibited very strong behavioral responses during courtship and male-male competition, consistent with reproductive behavior.
+In Hawaii, chameleons were collected from a single population in the forested Ko’olau Mountains north of Honolulu, between upper Makiki Valley and the summit of Tantalus Mountain, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
+400-600 m.
+The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Fig.
+S\@ref(fig:Figcolsig2)C, D).
+Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
+The area experiences high annual rainfall (mean: ca.
+1500 mm) @RN6 and the mean annual temperature is 22-24°C @RN7.
+In Kenya, field work was conducted near the foothills of Mt.
+Kenya, close to the town of Runyenjes.
+The area is partially cleared for agriculture, leaving small stands of forest.
+The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
+S\@ref(fig:Figcolsig2)A, B).
+Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.Jackson’s chameleons are a generalist species which can be locally abundant and relatively mobile.
+This population is from the region (slopes of Mount Kenya) identified as the source population [@Tilbury2018].
+
+## Behavioral trials
+In order to quantify display coloration, we staged encounters between conspecifics (male-male and male-female) and model predators (bird and snake) and measured their subsequent display coloration (details below).
+All chameleons were used in both predator and social trials although in some cases we were unable to quantify all social signals for a particular individual.
+We conducted all conspecific-social trials first, in case anti-predator trials unduly influenced their social interactions.
+Our analyses focused on comparing color responses of chameleons to conspecifics and predators to ensure that the color state corresponded to specific stimuli.
+We did not, and could not, measure a ‘neutral’ color state because it is impossible to meaningfully gauge a neutral state in a color-changing organism such as a chameleon.
+However, for predator trials, we ensured that chameleons were specifically responding to the predators by conducting seven trials with a just a branch with no predator attached, as a control.
+Chameleons did not respond to the branch (see also @RN18ms).
+Furthermore, during trials with predators, chameleons focused on the predator itself, and exhibited classic anti-predator behavior (e.g., contrasting stress color, body inflation, open-mouth threat).
+
+Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
+We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand (Fig.
+S\@ref(fig:Figcolsig3)).
+This structure was about 1.5 m high and each arm was approximately 60 cm.
+The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Fig.
+S\@ref(fig:Figcolsig3)).
+All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
+The signaling environment was therefore the same for all animals.
+
+We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
+This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Fig.
+S\@ref(fig:Figcolsig)A, B).
+In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
+This would be accompanied by head shakes and gapping, leading to horn-locking (Fig.
+S\@ref(fig:Figcolsig)B) if males were not obviously asymmetric.
+Once dominance was settled, the subordinate would turn brown (Fig.
+S\@ref(fig:Figcolsig)B and Fig.
+S\@ref(fig:Figcolsig3)) and flee.
+In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
+S\@ref(fig:Figcolsig)C).
+
+The day following social trials we presented each chameleon with a snake (Fig.
+S\@ref(fig:Figcolsig4)A) and bird (Fig.
+S\@ref(fig:Figcolsig4)B) model predator model (separately).
+Lizards were drawn at random and the order of presentation (snake or bird first) was random.
+Both trials were conducted on the same day but were separated by at least three hours.
+A trial began when a chameleon was removed from its cloth bag and placed on the triangular stand.
+In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Fig.
+S\@ref(fig:Figcolsig4)A).
+In the case of the bird trials, the bird was attached to an ca.
+2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Fig.
+S\@ref(fig:Figcolsig4)B).
+Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
+Once chameleons changed color, we quantified their spectral reflectance (details below).
+The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
+The bird was a mounted African Cuckoo-hawk (*Aviceda cuculoides*).
+
+## Color measurement
+
+During all behavior trials, we measured the color of each of three body regions: top flank, mid flank, tail base and gular region (Fig.
+S\@ref(fig:Figcolsig5)).
+We measured each body region once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
+The four body regions were measured in random sequence to avoid any bias of order.
+When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
+We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal during measurement.
+The probe holder ensured that measurements were taken at a standard angle (45°) and distance (5 mm) for a 3 x 5 mm area and excluded all ambient light.
+All measurements were relative to a dark, and a certified 99% white reflectance standard (Labsphere, North Sutton, NH, U.S.A.).
+If chameleons started to change color in response to our handling, we returned them to the branch and allowed them to continue a social trial before measuring reflectance of remaining body regions.
+In the case of a predator trial, we would once again present them with the predator model until they returned to their anti-predator display.
+This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behavior following handling.
+
+We obtained spectral reflectance measurements of male dominant color state in male-male contests for 34 individuals from Hawaii and 35 from Kenya, courtship color state for 25 individuals from Hawaii and 32 from Kenya; color state in response to the bird for 56 individuals from Hawaii and 23 from Kenya and color state in response to the snake for 59 individuals from Hawaii and 38 from Kenya.
+
+We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signaling environment.
+We took 120 reflectance measures of the background at each location (Hawaii and Kenya), comprising 40 measurements of the top surface of leaves, 40 of the bottom surface of leaves and 40 stems and branches of suitable width for chameleons.
+To reduce the number of backgrounds, we averaged the reflectance of the 20 darker measurements and 20 lighter measurements for each of the top leaf and bottom leaf measurements and averaged brown stems and green stems.
+This resulted in 6 average background spectra for each location, representing the range of background colors.
+
+```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Location of body regions we measured for spectral reflectance**. See details in text."}
+figs5 <- image_read("./output/figures/FigS5.jpg")
+figs5
+
+```
+
+## Visual modeling
+
+We visualized and quality-checked all spectra, adjusted very minor negative reflectance values caused by electrical noise by lifting the curves by the maximum negative value, then smoothed all spectra with LOESS smoothing of 0.16 using the plotsmooth function in `Pavo` vers.
+1.0 [@RN29].
+We also took the median reflectance for all leaves (top, bottom) and stems (2 measures/stem) for each population (Hawaii, Kenya).
+
+We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` @RN29.
+This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
+Color contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of color discrimination for stimuli viewed simultaneously under ideal viewing conditions @RN17.
+Under natural conditions, discrimination thresholds are likely to be >1 and vary depending on the species and conditions [@RN12;@RN13; @RN14].
+Additionally, perceived conspicuousness may not scale linearly with color distance (JNDs) @RN15.
+Nevertheless, the models provide a reasonable approximation of relative conspicuousness that can be compared between populations and trial types (social and predator).
+
+We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region and color state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
+We first calculated the quantum catch of each photoreceptor under standard daylight illumination (D65; Commission internationale de l’éclairage – CIE) using the ‘vismodel’ function.
+We applied the von Kries chromatic adaptation (vonkries = TRUE) and used the average background reflectance of the relevant background (Hawaii or Kenya) as the adapting background.
+The signal in each photoreceptor class was proportional to the natural logarithm of the quantum catch, in accordance with Fechner's law (“fi” in pavo).
+Next, we calculated the chromatic contrast between pairs of spectra using the function ‘coldist’.
+
+The visual models require information on the spectral sensitivities and photoreceptor noise (Weber fraction), which is a function of the relative density of each photoreceptor type within the retina, for the relevant receiver.
+Chameleons have four spectral classes of single cone enabling tetrachromatic vision @RN16.
+We used spectral sensitivities for the congeneric flap-necked chameleon (*Chameleo dilepis*) @RN16 and a standard Weber fraction of 0.1 @RN17 to account for the receptor noise of the LWS photoreceptor.
+We calculated noise for the other photoreceptor classes using the relative photoreceptor ratio of 1: 1.6: 3.3: 3.2 for the four single cones (UVS, SWS, MWS, and LWS, respectively).
+*Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Fig.
+S\@ref(fig:Figcolsig6SPec)).
+
+```{r Figcolsig6SPec, fig.align = 'center', fig.width = 12, fig.height = 10, fig.cap = "**Spectral sensitivities of chameleons (A), bird (B) and snake (C) used for visual models (details in text)**."}
+sens_cham_c_dilepis <- read_excel("./data/Spectral sensitivities.xls", sheet = "C_dilepis")[,1:6]
+        sens_birdUV <- read_excel("./data/Spectral sensitivities.xls", sheet = "birdUV")
+         sens_snake <- read_excel("./data/Spectral sensitivities.xls", sheet = "snake")[,1:4]
+
+cham <- sens_cham_c_dilepis %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
+bird <- sens_birdUV %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
+snake <- sens_snake %>% pivot_longer(cols = c(2:4)) %>% select(wavelength, name, value) %>%  data.frame()
+
+plot_spec <- function(data, tri = FALSE){
+        data %>% ggplot(aes(x = wavelength, y = value)) + 
+      geom_line(aes(color = name), size = 1.2) + 
+      theme_bw() +
+    scale_colour_manual(values = rev(c("#8D25CC", "#3770C9", "#93C572", "#FFA500", "#DBD8D6"))) +
+      labs(x = "Wavelength (nm)", y = "Absorbance", color = "") + 
+        theme(legend.position  = c(0.90, 0.65),
+            panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(),
+            axis.text = element_text(size = 16),
+            axis.title = element_text(size = 24),
+            legend.text = element_text(size = 16),
+            legend.title = element_text(size = 20))}
+
+p1 <- plot_spec(cham) + theme(legend.position  = "none")
+p2 <- plot_spec(bird) 
+p3 <- plot_spec(snake) + scale_colour_manual(values = rev(c("#8D25CC", "#93C572", "#FFA500"))) + theme(legend.position  = "none")
+
+plot_grid(p1, p2,p3,nrow = 3, ncol = 1, labels = c("A", "B", "C"), label_size = 30,
+  label_fontface = "bold", rel_widths=c(1,1))
+
+```
+
+For birds, we used an ultraviolet sensitive (UVS) visual system @RN18 characteristic of raptors and other birds of prey @RN19.
+A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement @RN20.
+We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21 (Fig.
+S\@ref(fig:Figcolsig6SPec)B).
+
+For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake @RN22.
+We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22 (Fig.
+S\@ref(fig:Figcolsig6SPec)C).
+
+In each case, visual pigment absorbance curves were multiplied by the transmission spectra of ocular media (lens and cornea) and oil droplets associated with the corresponding photoreceptor class and normalized to equal area under the curve to satisfy the assumption of equal stimulation by white light @RN18.
+
+Luminance contrast was calculated using the sensitivity function for the LWS photoreceptor (with transmission of the oil droplet associated with the double cone in chameleons and birds) for each visual system, assuming a Weber fraction of 0.05 because double cones used for luminance perception in birds and chameleons are by far the most abundant photoreceptor type in the retina [@RN16; @RN23].
+
+## Statistical Analysis
+
+Chromatic and luminance contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN24 in combination with the package `clubSandwich` @RN25.
+For each individual, measurements for each body region (top flank, mid-flank, tail base and gular region) were contrasted against each of six average background spectra (i.e., average top and bottom of lighter leaves, average top and bottom of darker leaves, average green and average brown stems).
+While we present sensitivity analyses for all body regions (see below - *Additional Supplemental Results*), we were mainly interested in overall signal differences displayed by chameleons.
+As such, for our main analyses we averaged estimates (marginalized) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see *Additional Supplemental Results*).
+In all cases, our results and conclusions were not impacted (see supplementary code).
+Deriving multiple contrasts for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other.
+To ensure that non-independence of data did not impact our inferences, we included individual ID as a random effect to account for repeated measures on each individual and used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction.
+
+First, we tested whether Hawaiian chameleons were more conspicuous (as perceived by chameleons) in social contexts than Kenyan chameleons.
+We ran separate linear models with chromatic and luminance contrast of male display or courtship coloration against the local background as the response variable.
+We included population of origin and z-transformed snout-vent length (SVL) as fixed effects.
+Next, we tested whether Kenyan chameleons were more cryptic (less conspicuous against the local background) in response to predators.
+For these models, chromatic or luminance contrast against the local background (as perceived by the corresponding predator type) was the response variable and predator type (i.e, snake or bird), population of origin (Hawaii or Kenya) and their interaction were fixed effects.
+Finally, to evaluate evidence for local adaptation, we tested whether the chromatic or luminance contrast of Hawaiian chameleons during social interactions (male-male and male-female) varied depending on whether contrast was against a native Kenyan background or their current introduced habitat background in Hawaii.
+For all models, interaction terms were first tested using Wald tests.
+If found to be non-significant (p > 0.1) we dropped the interaction and fitted a main effects model.
+We then tested the significance of the main effects using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.
 
@@
 
-```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social and predatory stimuli.
+```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social stimuli.
 Male chameleons experience intense sexual selection**.
-During the breeding season they change from a neutral state (green) to a highly conspicuous bright yellow display signal.
+During the breeding season they change from dull green to a highly conspicuous bright yellow display signal.
 They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
-**(A)** A typical dominant male in display coloration.
+**(A)** A dominant male in display coloration.
 **(B)** A subordinate male that lost a contest and turned from bright yellow to brown.
-**(C)** Two males fighting, both are in display coloration and at a stage where they are relatively even matched.
-**(D)** A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male."}
+**(C)** Two males fighting, both are in display coloration and relatively even matched.
+**(D)** A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
+See supplementary text for additional photos, including in response to a snake."}
 library(magick)
@@
 
+# Table S6
+tableS6 <- as.data.frame(rbind(robust_localAdapt_DispBgrdAC_s6,
+                               robust_DispBgrdCC_s6,
+                               robust_CourtshipAC_s6,
+                              robust_CourtshipCC_s6))
+tableS6$Parameter <- rownames(tableS6)
+rownames(tableS6) <- NULL
+
+# Clean
+final_S6 <- tableS6 %>% 
+  mutate(SocialContext = c("Male-Male Display", rep("", 9), "Courtship Display", rep("", 9)),
+         VisualRealm = c("Luminance  Contrast", rep("", 4), "Chromatic Contrast", rep("", 4), 
+                         "Luminance Contrast", rep("", 4), "Chromatic Contrast", rep("", 4))) %>%
+  dplyr::select(SocialContext, VisualRealm, Parameter, beta, SE, df, CI_L, CI_U)
+write.csv(final_S6, file = "./output/tables/tableS6.csv", row.names = FALSE)
+
 ```
 
-```{r setup, include=FALSE, purl = FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
-## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
-  ## and rounded to 2 digits      ## and rounded to 2 digits
-  options(digits = 2)
-```
-
-```{r getlibrary, echo = FALSE, results = "hide"}
-# load libraries
-devtools::install_github("karthik/wesanderson")
-pacman::p_load(flextable, lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, readxl, pavo)
-source("./R/func.R")
-```
-
-```{r display_supp, results = "hide"}
-source("./R/func.R")
-#########################################################
-# Male-male Competition Displays
-#########################################################
-  # Load data and clean a bit
-  Display_1 <- read.table("./data/Display_1.csv", 
-                            header=TRUE, sep=",", dec=".", strip.white=TRUE)
-
-  # Exclude bottom flank.
-  Display_1 <- Display_1 %>% 
-                filter(!BodyRegion == "botflank") %>% as.data.frame()
-
-  # Create as factor
-      Display_1$BodyRegion <- factor(Display_1$BodyRegion)
-
-  # Center SVL to make intercept interpretable
-  Display_1$sc_SVL <- with(Display_1, scale(SVL, scale=FALSE))
-
-##########
-## Achromatic Contrast - dL
-##########
-
-    # Fit the model that estimates changes across body regions
-    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + 
-                               Population:BodyRegion + (1|ID), data=Display_1)
-    summary(ConspicDisplayAC)
-    ConspicDisplayAC_main <- lmer(dL ~ sc_SVL + Population + BodyRegion +  (1|ID), data=Display_1)
-
-    # Wald Table - Test for interaction, plus refit model and test sig of each factor. Uses RVEs
-    wald_table_ConspicDisplayAC <-  wald_table(model_main =  ConspicDisplayAC_main , model_inter = ConspicDisplayAC)
-
-    # Robust variance estimators (RVE). Mostly all consistent
-    Display_subset <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
-    robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, cluster = Display_subset$ID, vcov="CR2")
-
-    # Fit model for linear hypothesis tests
-           Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
-            ConspicDisplayAC2 <- lmer(dL ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Display_1)
-            summary(ConspicDisplayAC2)
-
-     # Fit RVE
-      robust_ConspicDisplayAC2 <- coef_test(ConspicDisplayAC2, cluster = Display_subset$ID, vcov="CR2")
-
-     # Contrasts hypothesis tests, but use RVE 
-    result1 <- Wald_test(ConspicDisplayAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
-    result1$p_val_bonfer <- sapply(p.adjust(result1$p_val, "bonferroni"), function(x) p_value(x))
-
-##########
-## Chromatic Contrast - dL
-##########
-
- # Fit the model that esatimtes changes across body regions
-    ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
-    ConspicDisplayCC_main <- lmer(dS ~ sc_SVL + Population + BodyRegion + (1|ID), data=Display_1)
-
-# Wald Table - Test for interaction, plus refit model and test sig of each factor. Uses RVEs
-    wald_table_ConspicDisplayCC <-  wald_table(model_main =  ConspicDisplayCC_main, model_inter = ConspicDisplayCC)
-
- # Robust variance estimators (RVE). Mostly all consistent
-             Display_subset <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
-    robust_ConspicDisplayCC <- coef_test(ConspicDisplayCC, cluster = Display_subset$ID, vcov="CR2")
-
-  # Fit model for linear hypothesis tests
-    Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
-     ConspicDisplayCC2 <- lmer(dS ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Display_1)
-     summary(ConspicDisplayCC2)
-
- # Contrasts hypothesis tests, but use RVE 
-                 result2 <- Wald_test(ConspicDisplayCC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
-    result2$p_val_bonfer <- sapply(p.adjust(result2$p_val, "bonferroni"), function(x) p_value(x))
-
-#########################################################
-# Courtship Displays
-#########################################################
-
-Courtship <- read.table("./data/Courtship_1.2.csv", 
-                          header=TRUE, sep=",", dec=".", strip.white=TRUE)
-
-# Center SVL to make intercept interpretable
-Courtship$sc_SVL <- with(Courtship, scale(SVL, scale=FALSE))
-
-##########
-## Achromatic Contrast - dL
-##########
-  # Fit model
-    ConspicCourtshipAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + 
-                                 Population:BodyRegion + (1|ID), data=Courtship)
-    summary(ConspicCourtshipAC)
-
-  # Main effects Model
-    ConspicCourtshipAC_main <- lmer(dL ~ sc_SVL + Population + BodyRegion + (1|ID), data=Courtship)
-
-   # Wald Table - Test for interaction, plus refit model and test sig of each factor. Uses RVEs
-    wald_table_ConspicCourtshipAC <-  wald_table(model_main =  ConspicCourtshipAC_main, model_inter = ConspicCourtshipAC)  
-  # Robust variance estimators (RVE). Mostly all consistent
-             Courtship_subset <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
-    robust_ConspicCourtshipAC <- coef_test(ConspicCourtshipAC, cluster = Courtship_subset$ID, vcov="CR2")
-
-  # Refit model for contrasts
-     Courtship$PopBRInt <- interaction(Courtship$Population,Courtship$BodyRegion)
-    ConspicCourtshipAC2 <- lmer(dL ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Courtship)
-
-  # Contrasts hypothesis tests, but use RVE 
-                 result3 <- Wald_test(ConspicCourtshipAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
-    result3$p_val_bonfer <- sapply(p.adjust(result3$p_val, "bonferroni"), function(x) p_value(x))
-
-##########
-## Chromatic Contrast - dL
-#########
-
-    # Fit model, interaction
-        ConspicCourtshipCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Courtship)
-
-    # Fit main effects
-    ConspicCourtshipCC_main <- lmer(dS ~ sc_SVL + Population + BodyRegion + (1|ID), data=Courtship)
-
-    # Wald Table - Test for interaction, plus refit model and test sig of each factor. Uses RVEs
-    wald_table_ConspicCourtshipCC <-  wald_table(model_main =  ConspicCourtshipCC_main, model_inter = ConspicCourtshipCC)
-
-    # Robust variance estimators (RVE). Mostly all consistent
-             Courtship_subset <- Courtship %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
-    robust_ConspicCourtshipCC <- coef_test(ConspicCourtshipCC, cluster = Courtship_subset$ID, vcov="CR2")
-
-    # Refit model to get the contrasts between coefficienst we want
-     Courtship$PopBRInt <- interaction(Courtship$Population,Courtship$BodyRegion)
-    ConspicCourtshipCC2 <- lmer(dS ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Courtship)
-
-    # Contrasts hypothesis tests, but use RVE 
-            result4 <- Wald_test(ConspicCourtshipCC2, 
-                                 constraints = constrain_pairwise(c(2:9)), 
-                                 vcov="CR2", tidy = TRUE)
-      string <- c("PopBRIntKenya.gular - PopBRIntHawaii.gular|PopBRIntKenya.midflank - PopBRIntHawaii.midflank|PopBRIntKenya.tailbase - PopBRIntHawaii.tailbase|PopBRIntKenya.topflank - PopBRIntHawaii.topflank")
-      result4 <- result4[grep(string, result4$hypothesis),]
-    result4$p_val_bonfer <- sapply(p.adjust(result4$p_val, "bonferroni"), function(x) p_value(x))
-
-#########################################################
-# Displays to Predators, birds and snake
-#########################################################
-   Predator <- read.table("./data/Predator_2_3.csv", 
-                         header=TRUE, sep=",", dec=".", strip.white=TRUE)
-
-  # Clean up
-             Predator <- Predator %>% filter(!BodyRegion == "botflank" & !BodyRegion =='gularbot') %>% as.data.frame() #exclude bottom flank.
-  Predator$BodyRegion <- factor(Predator$BodyRegion) #updates levels for the variable BodyRegion
-
-##########
-## Achromatic Contrast - dL
-##########
-
-   # Fit model
-      PredatorAC <- lmer(dL ~ Predator + Population + BodyRegion + Predator:BodyRegion + Predator:Population + 
-                           (1|ID), data=Predator)
-      summary(PredatorAC)
-
-  PredatorAC_main <- lmer(dL ~ Predator + Population + BodyRegion  + (1|ID), data=Predator)
-
-  # Wald table RVE
-
-    Wald_table_PredatorAC <- wald_table_p(PredatorAC_main, PredatorAC)
-
-   # Robust variance estimators (RVE). Mostly all consistent
-             Predator_subset <- Predator %>% filter(!is.na(Population), !is.na(BodyRegion))
-           robust_PredatorAC <- coef_test(PredatorAC, cluster = Predator_subset$ID, vcov="CR2")
-
-  # Fit model for linear hypothesis tests
-    Predator$PopBRInt <- interaction(Predator$Population,Predator$Predator, Predator$BodyRegion)
-     PredatorAC2 <- lmer(dL ~ -1 + PopBRInt + (1|ID), data=Predator)
-     summary(PredatorAC2)
-
- # Contrasts hypothesis tests, but use RVE 
-                 result5 <- Wald_test(PredatorAC2, constraints = constrain_pairwise(c(1:16)), vcov="CR2", tidy = TRUE)
- string <- c("PopBRIntKenya.snake.midflank - PopBRIntHawaii.snake.midflank|PopBRIntKenya.snake.gular - PopBRIntHawaii.snake.gular|PopBRIntKenya.snake.tailbase - PopBRIntHawaii.snake.tailbase|PopBRIntKenya.snake.topflank - PopBRIntHawaii.snake.topflank|PopBRIntKenya.bird.midflank - PopBRIntHawaii.bird.midflank|PopBRIntKenya.bird.gular - PopBRIntHawaii.bird.gular|PopBRIntKenya.bird.tailbase - PopBRIntHawaii.bird.tailbase|PopBRIntKenya.bird.topflank - PopBRIntHawaii.bird.topflank")
-    result5 <- result5[grep(string, result5$hypothesis),]
-    result5$p_val_bonfer <- sapply(p.adjust(result5$p_val, "bonferroni"), function(x) p_value(x))
-
-##########
-## Chromatic Contrast - dS
-########
-
-  # Fit Model
-        PredatorCC <- lmer(dS ~ Predator + Population + BodyRegion + Predator:BodyRegion + Predator:Population + (1|ID), data=Predator)
-        summary(PredatorCC)
-          anova(PredatorAC)
-
-    PredatorCC_main <- lmer(dS ~ Predator + Population + BodyRegion  + (1|ID), data=Predator)
-
-  # Wald table RVE
-
-    Wald_table_PredatorCC <- wald_table_p(PredatorCC_main, PredatorCC)
-
-  # Robust variance estimators (RVE). Mostly all consistent
-             Predator_subset <- Predator %>% filter(!is.na(Population), !is.na(BodyRegion))
-           robust_PredatorCC <- coef_test(PredatorCC, cluster = Predator_subset$ID, vcov="CR2")
-
-  # Fit model for linear hypothesis tests
-    Predator$PopBRInt <- interaction(Predator$Population,Predator$Predator, Predator$BodyRegion)
-          PredatorCC2 <- lmer(dS ~ -1 + PopBRInt + (1|ID), data=Predator)
-          summary(PredatorCC2)
-
- # Contrasts hypothesis tests, but use RVE 
-                 result6 <- Wald_test(PredatorCC2, constraints = constrain_pairwise(c(1:16), with_zero = TRUE), vcov="CR2", tidy = TRUE)
-                   string <- c("PopBRIntKenya.snake.midflank - PopBRIntHawaii.snake.midflank|PopBRIntKenya.snake.gular - PopBRIntHawaii.snake.gular|PopBRIntKenya.snake.tailbase - PopBRIntHawaii.snake.tailbase|PopBRIntKenya.snake.topflank - PopBRIntHawaii.snake.topflank|PopBRIntKenya.bird.midflank - PopBRIntHawaii.bird.midflank|PopBRIntKenya.bird.gular - PopBRIntHawaii.bird.gular|PopBRIntKenya.bird.tailbase - PopBRIntHawaii.bird.tailbase|PopBRIntKenya.bird.topflank - PopBRIntHawaii.bird.topflank")
-    result6 <- result6[grep(string, result6$hypothesis),]
-    result6$p_val_bonfer <- sapply(p.adjust(result6$p_val, "bonferroni"), function(x) p_value(x))
-
-#########################################################
-# Testing local adaptation contrasting backgrounds between hawaii and Kenya
-#########################################################
-##################
-# Social Context: Male-male competition
-##################
-DisplayBgrd <- read.table("./data/Display_5.csv", 
-                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
-svl_dat <- read.csv("./data/SVL2.csv")
-
-DisplayBgrd <- DisplayBgrd %>% 
-                filter(!BodyRegion == "botflank") %>% 
-                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
-                as.data.frame() #exclude bottom flank.
-DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
-
-# Filter out only Hawaiian and add in SVL for IDs
-DisplayBgrd$pop_back_region <- with(DisplayBgrd, interaction(Background_pop, BodyRegion))
-DisplayBgrd_Hawaii <- DisplayBgrd %>% 
-               filter(Population == "Hawaii") %>% 
-               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
-               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
-                filter(!is.na(SVL))
-
-##########
-# Achromatic Contrast - dL
-##########
-
-  # Main effercts
-   DispBgrdAC_main <- lmer(dL ~  sc_svl + Background_pop + BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
-    summary(DispBgrdAC_main)
-
-  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE).
-Given that there is no interaction, we can just present the main effects model with teh figure.
-Plus wald tests
-  robust_localAdapt_DispBgrdAC <- clubSandwich::conf_int(DispBgrdAC_main, cluster = DisplayBgrd_Hawaii$ID, vcov =  "CR2")
-
-  # Add Wald test to test signifiacnce of each factor using roust variance 
-  background_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(3), 
-                                  cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
-
-  BodyRegion_AC_Disp <- Wald_test(DispBgrdAC_main, constraints = constrain_zero(4:6), 
-                                  cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
-
-##########
-# Chromatic Contrast - dS
-##########
-
-     # Fit the inetarction model
-          DispBgrdCC_inter <- lmer(dS ~ sc_svl + Background_pop  + BodyRegion + Background_pop:BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
-          summary(DispBgrdCC_inter)
-
-    # Fit the main effect model
-          DispBgrdCC_main <- lmer(dS ~ sc_svl + Background_pop  + BodyRegion + (1|ID), data=DisplayBgrd_Hawaii)
-          summary(DispBgrdCC_main)
-
-    # Wald test to see if interaction is important. Yes, body regions way in how they change from a Hawaiian and Kenyan population. Fig S4, this makes sense. 
-        intearction_CC_Disp <- Wald_test(DispBgrdCC_inter, constraints = constrain_zero(7:9), 
-                                  cluster = DisplayBgrd_Hawaii$ID, vcov = "CR2")
-
-    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-      robust_DispBgrdCC <- clubSandwich::conf_int(DispBgrdCC_inter, cluster = DisplayBgrd_Hawaii$ID, vcov =  "CR2")
-
- # Get contrasts to understand the specific changes better.
-Refit model in slight different form.
-This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
-The model above is just a contrast-based model.
-    DisplayBgrd_pop_back_regionCC <- lmer(dS ~ -1 + sc_svl + pop_back_region + (1|ID), data = DisplayBgrd_Hawaii) 
-    # Fit Robust model
-    robust_DisplayBgrd_pop_back_regionCC <- clubSandwich::conf_int(DisplayBgrd_pop_back_regionCC, 
-                                                            cluster = DisplayBgrd_Hawaii$ID, 
-                                                            vcov =  "CR2")
-
-    # Contrasts hypothesis tests, but use RVE 
-      result_DisplayBgrd_pop_back_regionCC <- Wald_test(DisplayBgrd_pop_back_regionCC, 
-                                                 constraints = constrain_pairwise(c(1:9)), 
-                                                 vcov="CR2", tidy = TRUE)
-
-      result_DisplayBgrd_pop_back_regionCC$p_val_bonfer <- 
-                                  sapply(p.adjust(result_DisplayBgrd_pop_back_regionCC$p_val, "bonferroni"), 
-                                                                  function(x) p_value(x))
-     # Grab contrasts that we need from the robust esatimtes
-      string <- c("pop_back_regionKenya.gular - pop_back_regionHawaii.gular|pop_back_regionKenya.midflank - pop_back_regionHawaii.midflank|pop_back_regionKenya.tailbase - pop_back_regionHawaii.tailbase|pop_back_regionKenya.topflank - pop_back_regionHawaii.topflank")
-      result_DisplayBgrd_pop_back_regionCC <- result_DisplayBgrd_pop_back_regionCC[grep(string, result_DisplayBgrd_pop_back_regionCC$hypothesis),]
-###################
-# Social Context: Courtship
-##################
-CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
-                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
-
-CourtshipBgrd <- CourtshipBgrd %>% 
-                filter(!BodyRegion == "botflank") %>% 
-                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
-                as.data.frame() #exclude bottom flank.
-CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
-
-# Filter out only Hawaiian and add in SVL for IDs
-CourtshipBgrd$pop_back_region <- with(CourtshipBgrd, interaction(Background_pop, BodyRegion))
-CourtshipBgrd_Hawaii <- CourtshipBgrd %>% 
-               filter(Population == "Hawaii") %>% 
-               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
-               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
-               filter(!is.na(SVL))
-
-##########
-## Achromatic Contrast - dL
-##########
-
-  # Fit main effects model
-    CourtshipAC_main <- lmer(dL ~ sc_svl + Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
-    summary(CourtshipAC_main)
-
-  # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
-   robust_CourtshipAC <- clubSandwich::conf_int(CourtshipAC_main, cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
-
-   # Wald tests
-           background_AC_Court <- Wald_test(CourtshipAC_main, constraints = constrain_zero(3), 
-                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
-
-           BodyRegion_AC_Court <- Wald_test(CourtshipAC_main, constraints = constrain_zero(4:6), 
-                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
-
-##########
-# Chromatic - dS
-##########
-
-  # Fit main effect model
-      CourtshipCC_inter <- lmer(dS ~ sc_svl + Background_pop + BodyRegion + Background_pop:BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
-      summary(CourtshipCC_inter)
-
-  # Fit main effect model
-      CourtshipCC_main <- lmer(dS ~ sc_svl + Background_pop + BodyRegion + (1|ID), data=CourtshipBgrd_Hawaii)
-      summary(CourtshipCC_main)
-
-  # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
-  robust_CourtshipCC <- clubSandwich::conf_int(CourtshipCC_inter, cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
-
-  # Wald test of interaction; Not important, so we'll simplify and assume just additive effects. This makes sense when we look at the raw data.
-  intearction_CC_Court <- Wald_test(CourtshipCC_inter, constraints = constrain_zero(7:9), 
-                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
-
-   # Wald tests
-           background_CC_Court <- Wald_test(CourtshipCC_main, constraints = constrain_zero(3), 
-                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
-
-           BodyRegion_CC_Court <- Wald_test(CourtshipCC_main, constraints = constrain_zero(4:6), 
-                                  cluster = CourtshipBgrd_Hawaii$ID, vcov = "CR2")
-
-    # Get contrasts to understand the specific changes better.
-Refit model in slight different form.
-This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
-The model above is just a contrast-based model.
-
-          CourtshipBgrd_pop_back_BR_CC <- lmer(dS ~ -1 +  sc_svl + pop_back_region + (1|ID), data = CourtshipBgrd_Hawaii)
-
-     # Fit Robust model
-          robust_CourtshipBgrd_pop_back_BR_CC <- clubSandwich::conf_int(CourtshipBgrd_pop_back_BR_CC, 
-                                                              cluster = CourtshipBgrd_Hawaii$ID, vcov =  "CR2")
-
-     # Contrasts hypothesis tests, but use RVE 
-          result_CourtshipBgrd_pop_back_BR_CC <- Wald_test(CourtshipBgrd_pop_back_BR_CC, 
-                                                   constraints = constrain_pairwise(c(1:9)),
-                                                   cluster = CourtshipBgrd_Hawaii$ID,
-                                                   vcov="CR2", tidy = TRUE)
-             result_CourtshipBgrd_pop_back_BR_CC$p_val_bonfer <- sapply(p.adjust(result_CourtshipBgrd_pop_back_BR_CC$p_val, "bonferroni"), 
-                                                                    function(x) p_value(x))
-
-       # Grab contrasts that we need from the robust esatimtes
-      string <- c("pop_back_regionKenya.gular - pop_back_regionHawaii.gular|pop_back_regionKenya.midflank - pop_back_regionHawaii.midflank|pop_back_regionKenya.tailbase - pop_back_regionHawaii.tailbase|pop_back_regionKenya.topflank - pop_back_regionHawaii.topflank")
-
-         result_CourtshipBgrd_pop_back_BR_CC <- result_CourtshipBgrd_pop_back_BR_CC[
-                                            grep(string, result_CourtshipBgrd_pop_back_BR_CC$hypothesis),]
-
-```
-
-Martin J. Whiting, Brenden S. Holland, J. Scott Keogh, Daniel W.A.Noble, Katrina J. Rankin and Devi Stuart-Fox
-
-Correspondance: Martin J. Whiting  
-Email: martin.whiting@mq.edu.au
-
-**The file includes**:
-
-Supplementary text  
-Figures S1 -- S9  
-Tables S1 -- S13  
-Supplementary References
-
-# Expanded Hypotheses and Predictions
-
-We provide an expanded table of the key hypothesis and associated predictions. We also identify the data needed to test these predictions.
-
-```{r tablehyp, tab.cap = "Hypothesis and associated predictions tested in this paper in relation to the population of interest, the receiver visual system used in visual modelling, and the background against which the contrast of the signal was calculated for chromatic and luminance contrast."}
-  source("./R/func.R") 
-  hyp <-  read.csv(file = "./output/tables/Table_hyp_pred.csv")
-
-  # Table
-  hyp2 <- flextable(hyp) %>% table_style12()
-
-  hyp2
-```
-
-# Potential bird predators of chameleons on Oahu
-We are not aware of any evidence of bird predation of chameleons on Oahu.
-We consulted with several ornithologists in Hawaii regarding feeding behavior of the few predatory bird species that consume vertebrate prey, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons @RN1.
-Among the larger, predatory birds, two are owls (*Asio flammeus sandwichensis*, *Tyto alba*), which are extremely rare on Oahu.
-Both species are highly unlikely to feed on chameleons because they feed out in the open rather than in forest canopy where chameleons occur, and typically feed on rodents, birds and insects @RN2.
-Smaller to medium raptors include the Merlin (*Falco columbarius*) and the peregrine falcon (*Falco peregrinus*) which are both considered rare non-breeding vagrants in the Hawaiian Islands, and feed on small-medium birds in more open habitats @RN2.
-There are also eight species of egret and heron, which typically have generalist diets.
-Of these, six are considered rare or accidental @RN1, are associated with aquatic habitats, and are better known for eating fishes @RN3.
-The cattle egret (*Bubulcus ibis*) is a generalist predator that is mostly insectivorous but does eat lizards opportunistically [@RN4;@RN5].
-However, it forages on the ground in more open habitats, often following domestic animals and agricultural or landscaping machinery @RN5 and therefore, is not found in forested habitats and is highly unlikely to encounter chameleons.
-Among other potential predators, the Hawaiian Hawk or ʻio (*Buteo solitarius*) is restricted to the Big Island, while the Hen Harrier (*Circus cyaneus*) is considered an accidental or rare visitor to the Hawaiian islands.
-Finally, among corvids, the Hawaiian Crow (*Corvus hawaiiensis*) is extinct in the wild while the Common Raven (*Corvus corax*) is extremely rare @RN1.
-In summary, we could not identify any bird species that could be regarded as a major predator of chameleons.
-
-# Supplemental Materials and Methods
-## Study area and collection of chameleons
-
-We collected most chameleons (Fig.
-S\@ref(fig:Figcolsig2)) at night by spotlighting, and a small proportion of individuals during the day.
-When chameleons were beyond reach in the forest canopy, we used lightweight 6-m extensible graphite composite poles (South Bend Kwik Stix™ 20' Telescopic Fishing Pole), that chameleons could be encouraged to step on to, before lowering them to the ground.
-We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
-We measured snout-vent length (SVL), a standard measure of body size in lizards, from the tip of the snout to the posterior edge of the vent using a plastic ruler to the nearest 1 mm.
-Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006.
-In Hawaii, chameleons were collected from a single population in the forested Ko’olau Mountains north of Honolulu, between upper Makiki Valley and the summit of Tantalus Mountain, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
-400-600 m.
-The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Fig.
-S\@ref(fig:Figcolsig2)C,D).
-Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
-The area experiences high annual rainfall (mean: ca.
-1500 mm) @RN6 and the mean annual temperature is 22-24°C @RN7.
-In Kenya, field work was conducted near the foothills of Mt.
-Kenya, close to the town of Runyenjes.
-The area is partially cleared for agriculture, leaving small stands of forest.
-The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
-S\@ref(fig:Figcolsig2)A,B).
-Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.
-
-```{r Figcolsig, fig.align = 'center', fig.cap = "**Additional examples of chameleon color signals and behavior in response to different social contexts**.
-**(A)** Two different males in conspicuous display color adopted at the start of a contest and retained by the dominant male.
-**(B)** Two examples of males engaged in horn locking.
-Males in the bottom panel are evenly matched at this stage and both are in full display color.
-In the top image, the male on the left is starting to change to a subordinate color.
-**(C)** Conspicuous yellow-green male courtship color.
-When females reject males, they develop a heavily contrasting pattern."}
-figs1 <- image_read("./output/figures/FigS1.png")
-figs1
-
-```
-
-```{r Figcolsig2, fig.align = 'center', fig.cap = "**Representative photos of habitat in Kenya (A, B) and Hawaii (C, D)**.
-In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines."}
-figs2 <- image_read("./output/figures/FigS2.jpg")
-figs2
-
-```
-
-## Behavioral trials
-In order to quantify display coloration, we staged encounters between conspecifics (male-male and male-female) and model predators (bird and snake) and measured their subsequent display coloration (details below).
-All chameleons were used in both predator and social trials although in some cases we were unable to quantify all social signals for a particular individual.
-We conducted all conspecific-social trials first, in case anti-predator trials unduly influenced their social interactions.
-Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
-We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand (Fig.
-S\@ref(fig:Figcolsig3)).
-This structure was about 1.5 m high and each arm was approximately 60 cm.
-The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Fig.
-S\@ref(fig:Figcolsig3)).
-All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
-The signaling environment was therefore the same for all animals.
-
-```{r Figcolsig3, fig.align = 'center', fig.cap = "**The experimental setup for behavioral trials**.
-We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male.
-The male on the left won this contest and is in full display color while the male on the right has lost this contest and turned brown (subordinate)."}
-figs3 <- image_read("./output/figures/FigS3.jpg")
-figs3
-
-```
-
-We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
-This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Fig.
-S\@ref(fig:Figcolsig)A,B).
-In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
-This would be accompanied by head shakes and gapping, leading to horn-locking (Fig.
-S\@ref(fig:Figcolsig)B) if males were not obviously asymmetric.
-Once dominance was settled, the subordinate would turn brown (Fig.
-S\@ref(fig:Figcolsig)B and Fig.
-S\@ref(fig:Figcolsig3)) and flee.
-In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
-S\@ref(fig:Figcolsig)C).
-
-The day following social trials we presented each chameleon with a snake (Fig.
-S\@ref(fig:Figcolsig4)A) and bird (Fig.
-S\@ref(fig:Figcolsig4)B) model predator model (separately).
-Lizards were drawn at random and the order of presentation (snake or bird first) was random.
-Both trials were conducted on the same day but were separated by at least three hours.
-A trial began when a chameleon was removed from its cloth bag and placed on the triangular stand.
-In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Fig.
-S\@ref(fig:Figcolsig4)A).
-In the case of the bird trials, the bird was attached to an ca.
-2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Fig.
-S\@ref(fig:Figcolsig4)B).
-Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
-Once chameleons changed color, we quantified their spectral reflectance (details below).
-We also conducted seven trials with a just a branch with no predator attached, as a control for the model predator.
-The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
-The bird was a mounted African Cuckoo-hawk (*Aviceda cuculoides*).
-
-```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator**.
-**(A)** A model snake (*Dispholidus typus*) causing a characteristic contrasting color change and **(B)** a model bird (*Aviceda cuculoides*) which was flown over the top of the chameleon."}
-figs4 <- image_read("./output/figures/FigS4.jpg")
-figs4
-
-```
-
-## Color measurement
-
-During all behavior trials, we measured the color of each of three body regions: top flank, mid flank, tail base and gular region (Fig.
-S\@ref(fig:Figcolsig5)).
-We measured each body region once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
-The four body regions were measured in random sequence to avoid any bias of order.
-When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
-We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal during measurement.
-The probe holder ensured that measurements were taken at a standard angle (45°) and distance (5 mm) for a 3 x 5 mm area and excluded all ambient light.
-All measurements were relative to a dark, and a certified 99% white reflectance standard (Labsphere, North Sutton, NH, U.S.A.).
-If chameleons started to change color in response to our handling, we returned them to the branch and allowed them to continue a social trial before measuring reflectance of remaining body regions.
-In the case of a predator trial, we would once again present them with the predator model until they returned to their anti-predator display.
-This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behavior following handling.
-
-We obtained spectral reflectance measurements of male dominant color state in male-male contests for 34 individuals from Hawaii and 35 from Kenya, courtship color state for 25 individuals from Hawaii and 32 from Kenya; color state in response to the bird for 56 individuals from Hawaii and 23 from Kenya and color state in response to the snake for 59 individuals from Hawaii and 38 from Kenya.
-
-We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signaling environment.
-We took 120 reflectance measures of the background at each location (Hawaii and Kenya), comprising 40 measurements of the top surface of leaves, 40 of the bottom surface of leaves and 40 stems and branches of suitable width for chameleons.
-To reduce the number of backgrounds, we averaged the reflectance of the 20 darker measurements and 20 lighter measurements for each of the top leaf and bottom leaf measurements and averaged brown stems and green stems.
-This resulted in 6 average background spectra for each location, representing the range of background colors.
-
-```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Location of body regions we measured for spectral reflectance**. See details in text."}
-figs5 <- image_read("./output/figures/FigS5.jpg")
-figs5
-
-```
-
-## Visual modeling
-
-We visualized and quality-checked all spectra, adjusted very minor negative reflectance values caused by electrical noise by lifting the curves by the maximum negative value, then smoothed all spectra with LOESS smoothing of 0.16 using the plotsmooth function in `Pavo` vers.
-1.0 [@RN29].
-We also took the median reflectance for all leaves (top, bottom) and stems (2 measures/stem) for each population (Hawaii, Kenya).
-
-We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` @RN29.
-This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
-Color contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of color discrimination for stimuli viewed simultaneously under ideal viewing conditions @RN17.
-Under natural conditions, discrimination thresholds are likely to be >1 and vary depending on the species and conditions [@RN12;@RN13; @RN14].
-Additionally, perceived conspicuousness may not scale linearly with color distance (JNDs) @RN15.
-Nevertheless, the models provide a reasonable approximation of relative conspicuousness that can be compared between populations and trial types (social and predator).
-
-We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region and color state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
-We first calculated the quantum catch of each photoreceptor under standard daylight illumination (D65; Commission internationale de l’éclairage – CIE) using the ‘vismodel’ function.
-We applied the von Kries chromatic adaptation (vonkries = TRUE) and used the average background reflectance of the relevant background (Hawaii or Kenya) as the adapting background.
-The signal in each photoreceptor class was proportional to the natural logarithm of the quantum catch, in accordance with Fechner's law (“fi” in pavo).
-Next, we calculated the chromatic contrast between pairs of spectra using the function ‘coldist’.
-
-The visual models require information on the spectral sensitivities and photoreceptor noise (Weber fraction), which is a function of the relative density of each photoreceptor type within the retina, for the relevant receiver.
-Chameleons have four spectral classes of single cone enabling tetrachromatic vision @RN16.
-We used spectral sensitivities for the congeneric flap-necked chameleon (*Chameleo dilepis*) @RN16 and a standard Weber fraction of 0.1 @RN17 to account for the receptor noise of the LWS photoreceptor.
-We calculated noise for the other photoreceptor classes using the relative photoreceptor ratio of 1: 1.6: 3.3: 3.2 for the four single cones (UVS, SWS, MWS, and LWS, respectively).
-*Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Fig.
-S\@ref(fig:Figcolsig6SPec)).
-
-```{r Figcolsig6SPec, fig.align = 'center', fig.width = 12, fig.height = 10, fig.cap = "**Spectral sensitivities of chameleons (A), bird (B) and snake (C) used for visual models (details in text)**."}
-sens_cham_c_dilepis <- read_excel("./data/Spectral sensitivities.xls", sheet = "C_dilepis")[,1:6]
-        sens_birdUV <- read_excel("./data/Spectral sensitivities.xls", sheet = "birdUV")
-         sens_snake <- read_excel("./data/Spectral sensitivities.xls", sheet = "snake")[,1:4]
-
-cham <- sens_cham_c_dilepis %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
-bird <- sens_birdUV %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
-snake <- sens_snake %>% pivot_longer(cols = c(2:4)) %>% select(wavelength, name, value) %>%  data.frame()
-
-plot_spec <- function(data, tri = FALSE){
-        data %>% ggplot(aes(x = wavelength, y = value)) + 
-      geom_line(aes(color = name), size = 1.2) + 
-      theme_bw() +
-    scale_colour_manual(values = rev(c("#8D25CC", "#3770C9", "#93C572", "#FFA500", "#DBD8D6"))) +
-      labs(x = "Wavelength (nm)", y = "Absorbance", color = "") + 
-        theme(legend.position  = c(0.90, 0.65),
-            panel.grid.major = element_blank(), 
-            panel.grid.minor = element_blank(),
-            axis.text = element_text(size = 16),
-            axis.title = element_text(size = 24),
-            legend.text = element_text(size = 16),
-            legend.title = element_text(size = 20))}
-
-p1 <- plot_spec(cham) + theme(legend.position  = "none")
-p2 <- plot_spec(bird) 
-p3 <- plot_spec(snake) + scale_colour_manual(values = rev(c("#8D25CC", "#93C572", "#FFA500"))) + theme(legend.position  = "none")
-
-plot_grid(p1, p2,p3,nrow = 3, ncol = 1, labels = c("A", "B", "C"), label_size = 30,
-  label_fontface = "bold", rel_widths=c(1,1))
-
-```
-
-For birds, we used an ultraviolet sensitive (UVS) visual system @RN18 characteristic of raptors and other birds of prey @RN19.
-A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement @RN20.
-We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21 (Fig.
-S\@ref(fig:Figcolsig6SPec)B).
-
-For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake @RN22.
-We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22 (Fig.
-S\@ref(fig:Figcolsig6SPec)C).
-
-In each case, visual pigment absorbance curves were multiplied by the transmission spectra of ocular media (lens and cornea) and oil droplets associated with the corresponding photoreceptor class and normalized to equal area under the curve to satisfy the assumption of equal stimulation by white light @RN18.
-
-Luminance contrast was calculated using the sensitivity function for the LWS photoreceptor (with transmission of the oil droplet associated with the double cone in chameleons and birds) for each visual system, assuming a Weber fraction of 0.05 because double cones used for luminance perception in birds and chameleons are by far the most abundant photoreceptor type in the retina [@RN16; @RN23].
-
-## Statistical Analysis
-
-Chromatic and luminance contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN24 in combination with the package `clubSandwich` @RN25.
-For each individual, measurements for each body region (top flank, mid-flank, tail base and gular region) were contrasted against each of six average background spectra (i.e., average top and bottom of lighter leaves, average top and bottom of darker leaves, average green and average brown stems).
-While we present sensitivity analyses for all body regions (see below - *Additional Supplemental Results*), we were mainly interested in overall signal differences displayed by chameleons.
-As such, for our main analyses we averaged estimates (marginalized) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see *Additional Supplemental Results*).
-In all cases, our results and conclusions were not impacted (see supplementary code).
-Deriving multiple contrasts for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other.
-To ensure that non-independence of data did not impact our inferences, we included individual ID as a random effect to account for repeated measures on each individual and used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction.
-
-First, we tested whether Hawaiian chameleons were more conspicuous (as perceived by chameleons) in social contexts than Kenyan chameleons.
-We ran separate linear models with chromatic and luminance contrast of male display or courtship coloration against the local background as the response variable.
-We included population of origin and z-transformed snout-vent length (SVL) as fixed effects.
-Next, we tested whether Kenyan chameleons were more cryptic (less conspicuous against the local background) in response to predators.
-For these models, chromatic or luminance contrast against the local background (as perceived by the corresponding predator type) was the response variable and predator type (i.e, snake or bird), population of origin (Hawaii or Kenya) and their interaction were fixed effects.
-Finally, to evaluate evidence for local adaptation, we tested whether the chromatic or luminance contrast of Hawaiian chameleons during social interactions (male-male and male-female) varied depending on whether contrast was against a native Kenyan background or their current introduced habitat background in Hawaii.
-For all models, interaction terms were first tested using Wald tests.
-If found to be non-significant (p > 0.1) we dropped the interaction and fitted a main effects model.
-We then tested the significance of the main effects using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.
-
-# Supplemental Results
-## Body Size Comparisons between Hawaii and Kenya
-```{r svl, results = "hide"}
-SVLdat <- read.table("data/SVL2.csv", 
-                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
-SVLdat$SVL2 <- with(SVLdat, scale(SVL))
-
-# Fit the model
-SVL_pop <- t.test(SVL ~ Population, data=SVLdat)
-summary(SVL_pop)
-
-# Body size
-svl <- data.frame(SVLdat %>% group_by(Population) %>% summarise(mean = mean(SVL), sd = sd(SVL), n = n()))
-
-```
-Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, *n* = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]` mm, *n* = `r svl[2,"n"]`; t = `r SVL_pop$statistic`, *P* <0.001).
-SVL did not significantly predict luminance contrast against the background (Table S\@ref(tab:table1)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
-Snout-vent length was a weak predictor of chromatic contrast (Table S\@ref(tab:table1)): larger chameleons had slightly higher chromatic contrast against the background.
-However, the populations did not differ in chromatic contrast during male displays (courtship and contests).
-
-## Supporting Supplementary Tables for Main Text
-
-```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing mean differences between Kenyan and Hawaiian lizards in luminance and chromatic contrast against the local background in two social contexts (male-male competition and courtship).
-Model estimates, standard errors (SE), Satterthwaite's degrees of freedom (df) and 95% confidence internals (95% CI).
-Standard errors are corrected by using robust variance estimators."}
-source("./R/func.R")
-final_T1 <-  read.csv(file = "./output/tables/Table1_MS.csv")
-
-  # Table
- table1 <- flextable(final_T1) %>%  table_style6() %>% fit_to_width(max_width = 7)
- table1
-
-```
-
-```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic mean differences between Kenyan and Hawaiian lizards in response to snakes and birds.
-Model estimates, standard errors (SE), Satterthwaite's degrees of freedom (df) and 95% confidence internals (95% CI).
-Full models are provided and standard errors are corrected by using robust variance estimators."}
-  final_T2 <- read.csv(file = "./output/tables/Table2_MS.csv")
-  # Table
- table2 <- flextable(final_T2) %>%
-            table_style7(.) %>% fit_to_width(max_width = 7)
-
- table2
-```
-
-```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean luminance and chromatic contrast in Hawaii and Kenya for both predators (snakes and birds).
-Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
-Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).
-We also applied a Bonferroni correction to p-values to control for multiple comparisons."}
- contrast <- read.csv(file = "./output/tables/Table3_MS.csv")
- table3 <- flextable(contrast) %>% table_style8() %>% fit_to_width(max_width = 7)
- table3
-```
-
-```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
-Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for models were fitted using robust variance estimators to correct standard errors for non-independence, and we used Satterthwaite's degrees of freedom.
-Significant parameter estimates are those where the 95% CIs do not include zero."}
- tb4 <- read.csv(file = "./output/tables/Table4_MS.csv")
- table4 <- flextable(tb4) %>% table_style9() %>% fit_to_width(max_width = 7)
- table4
-```
-
-# Additional Supplemental Results
-## Body regions are perceived differently by predators but exhibited similar effect sizes
-
-Chromatic contrast of color signals against the background varied by predator type, population and body region (Fig.
-S\@ref(fig:FigS3)A and Table S\@ref(tab:tableS3)).
-Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Fig.
-S\@ref(fig:FigS3)A).
-There was no significant interaction between predator type ('bird' vs 'snake') and population was not significant for luminance contrast against the background, but there was a marginally non-significant population and body region effect (Fig.
-S\@ref(fig:FigS3)B and Table S\@ref(tab:tableS3)).
-Effects did not vary by predator type (Table S\@ref(tab:tableS3)).
-Nonetheless, effect sizes across body regions were generally in the same direction (Fig.
-S\@ref(fig:FigS3)B and Table S\@ref(tab:tableS4)).
-Overall, there were significant differences between Kenya and Hawaii in luminance contrast of the gular and tailbase regions, but not for mid- and top-flank (Table S\@ref(tab:tableS5)).
-In contrast, no significant differences between Kenya and Hawaii existed across any body region for snakes, whereas significant differences existed between gular, midflank and tailbase for birds (Table S\@ref(tab:tableS6)).
-
-```{r FigS3, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "All four body regions (gular, mid-flank, tailbase and topflank) and their luminance (B) and chromatic contrast (A) to Hawaiian and Kenyan backgrounds (each population against their own background) are shown averaged across stems and leaves found in their habitat.
-Just noticable differences (JNDs) were calculated for both bird and snake predators.
-Mean JNDs are provided along with 95% confidence intervals."}
-
-# Load the data
-Predator <- read.table("./data/Predator_2_3.csv", 
-                       header=TRUE, sep=",", dec=".", strip.white=TRUE)
-
-# Clean up
- Predator <- Predator %>% filter(!BodyRegion == "botflank" & !BodyRegion =='gularbot') %>% as.data.frame() #exclude bottom flank.
-Predator$BodyRegion <- factor(Predator$BodyRegion) #updates levels for the variable BodyRegion
-
-# Calculate summary data
-summary_data_pred <- Predator %>%
-                group_by(Predator, Population, BodyRegion) %>%
-                summarise( Mean_dS = mean(dS, na.rm = T), 
-                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
-                             Upper_dS = Mean_dS + 1.96*SE_dS,
-                             Lower_dS = Mean_dS - 1.96*SE_dS, 
-                           Mean_dL = mean(dL, na.rm = T), 
-                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
-                             Upper_dL = Mean_dL + 1.96*SE_dL,
-                             Lower_dL = Mean_dL - 1.96*SE_dL) %>%
-                mutate(Predator = if_else(Predator == "bird", "Birds", "Snakes"))
-
-# PLot 
-p3 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dL, group = BodyRegion, color = BodyRegion)) + facet_wrap(~Predator) +
-  ylim(8, 30) + 
-  geom_point(size = 3, position = position_dodge(width=0.5)) +
-  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
-  geom_line(position = position_dodge(width=0.5)) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("") + 
-  ylab("Luminance contrast (JND)") +
-  theme_bw() +
-  labs(color = "Body Region") + 
-  theme(legend.position = c(0.90, 0.85), 
-        strip.background = element_blank(),
-    strip.text.x = element_text(size = 24, face = "bold"),
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-p4 <- ggplot(summary_data_pred, aes(x = Population, y = Mean_dS, group = BodyRegion, color = BodyRegion)) + facet_wrap(~Predator) +
-  ylim(1, 9) + 
-  geom_point(size = 3, position = position_dodge(width=0.5)) +
-  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
-  geom_line(position = position_dodge(width=0.5)) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("") + 
-  ylab("Chromatic contrast (JND)") +
-  labs(colour = "Body Region") + 
-  theme_bw() +
-  theme(legend.position = c(0.90, 0.85),
-    strip.background = element_blank(),
-    strip.text.x = element_text(size = 24, face = "bold"),
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-plot_grid(p4, p3, nrow = 2, ncol = 1, labels = c("A", "B"), label_size = 30,
-  label_fontface = "bold", rel_widths=c(1,1))
-
-```
-
-```{r tableS3, tab.cap = "Wald F tests comparing the significance of model parameters for predator models.
-Wald F tests were computed using robust variance estimators with a Satterthwaite correction to the degrees of freedom.
-Interactions were first tested, and models refitted, to test the significance of main effects."}
-source("./R/func.R")
-fulltableP <- rbind(Wald_table_PredatorAC,
-                    Wald_table_PredatorCC)
-
-fulltableP <- fulltableP %>% mutate(JND = c("Luminance Contrast", rep("",4), "Chromatc Contrast", rep("",4)),
-                                            p_val = sapply(p_val, function(x) p_value(x))) %>% 
-                                      dplyr::select(JND, Variable, Fstat, df_num, df_denom, p_val)
-
-S4 <- flextable(fulltableP) %>%
-        table_style3(.) %>% fit_to_width(max_width = 7)
-S4
-```
-
-```{r tableS4, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter from linear mixed effects models.
-Standard errors were corrected using robust variance estimators and Satterthwaite's degrees of freedom correction was used for calculated significance."}
-source("./R/func.R")
-robust_sc_pred_table <- rbind(robust_PredatorAC,
-                         robust_PredatorCC) %>% 
-                    mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)),
-                           JND = c("Luminance Contrast", rep("", 9), 
-                                   "Chromatic Contrast", rep("", 9)))
-
-robust_sc_pred_table$parameter <- rownames(robust_sc_pred_table)
-
-robust_sc_pred_table <- robust_sc_pred_table %>% select(JND, parameter, beta, SE, tstat, df, p_Satt)
-
-tableS6 <- flextable(robust_sc_pred_table) %>%  table_style5() %>% fit_to_width(max_width = 7)
-tableS6
-```
-
-```{r tableS5, tab.cap = "Pairwise comparisons of luminance contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
-Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with pairwise Wald tests for contrast significance.
-The Wald test uses a robust variance correction on the standard errors."}
-source("./R/func.R")
-# Split snakes and birds
-result5_snake <-  result5[grep("snake", result5$hypothesis), ] %>% mutate(Species = c("Snake", rep("",3)))
- result5_bird <- result5[grep("bird", result5$hypothesis), ] %>% mutate(Species = c("Bird", rep("",3)))
-
- # Bind together in order
-  final_r5 <- rbind(result5_snake,result5_bird)
-
-  # Clean
-  final_r5 <- final_r5 %>% 
-              select(-test, -delta, -df_num) %>%
-              mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
-                     p_val = sapply(p_val, function(x) p_value(x))) %>%
-              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
-
-ft5 <- flextable(final_r5) %>%
-        table_style(.) %>% fit_to_width(max_width = 7)
-
- ft5
-
-```
-
-```{r tableS6, tab.cap = "Pairwise comparisons of chromatic contrast between Kenyan and Hawaiian chameleon displays to a snake and bird visual system.
-Four different body regions ('gular', 'midflank', 'tailbase', topflank') are provided along with a pairwise Wald test for contrast significance.
-The Wald test uses a robust variance correction on the standard errors."}
-source("./R/func.R")
-# Split snakes and birds
-
-result6_snake <- result6[grep("snake", result6$hypothesis), ] %>% mutate(Species = c("Snake", rep("",3)))
- result6_bird <- result6[grep("bird", result6$hypothesis), ] %>% mutate(Species = c("Bird", rep("",3)))
-
-# Bind together in order
- final_r6 <- rbind(result6_snake,result6_bird)
-
-# Clean
- final_r6 <- final_r6 %>% 
-              select(-test, -delta, -df_num) %>%
-              mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
-                     p_val = sapply(p_val, function(x) p_value(x))) %>%
-              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
-
- # Table
- ft6 <- flextable(final_r6) %>%
-        table_style(.) %>% fit_to_width(max_width = 7)
-
- ft6
-```
-
-## All body regions exhibit similar contrasts between Kenyan and Hawaiian backgrounds suggesting local adaptation of entire visual signal
-
-We tested whether different body regions of Hawaiian chameleons exhibited different contrasts between Hawaiian habitat background compared to their ancestral, Kenyan background.
-We started off testing models comparing the interaction between background environment and body region for both chromatic and luminance contrast.
-However, for luminance contrast, these models had poor convergence.
-Based on the raw data (Fig.
-S\@ref(fig:FigS4)B and S\@ref(fig:FigS4)D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
-As such, we simplified models and only present the main effects of body region and background environment for luminance contrast.
-However, inspection of the raw data (Fig.
-S\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C) suggested that body regions may differ in how chromatically contrasting they were against the background environment of Hawaii and Kenya.
-In addition, our main results suggested that predators may deferentially drive chromatic signal conspicuousness, and given different body regions maybe more exposed to bird compared to snake, predators, we fitted an interaction and main effects model for chromatic contrast.
-
-The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, *P* = `r p_value(intearction_CC_Court$p_val)`, Fig.
-S\@ref(fig:FigS4)A, S\@ref(fig:FigS4)C and Table S\@ref(tab:tableS7)).
-However, in all cases there was a tendency for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Fig.
-S\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C).
-Tailbase was the only body region that showed a non-significant change (Table S\@ref(tab:tableS8)).
-
-The magnitude of luminance contrast showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court$p_val)`, Fig.
-S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, *P* = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Fig.
-S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)).
-
-```{r FigS4, fig.align = 'center', fig.width = 13, fig.height = 12, fig.cap = "Local adaptation.
-All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat.
-Mean JNDs are provided along with 95% confidence intervals.
-**(A & B)** represent male-male contest displays whereas **(C & D)** represent courtship displays."}
-# Courship data
-
-CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
-                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
-
-CourtshipBgrd <- CourtshipBgrd %>% 
-                filter(!BodyRegion == "botflank") %>% 
-                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
-                as.data.frame() #exclude bottom flank.
-CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
-
-# Display data
-DisplayBgrd <- read.table("./data/Display_5.csv", 
-                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
-
-DisplayBgrd <- DisplayBgrd %>% 
-                filter(!BodyRegion == "botflank") %>% 
-                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
-                as.data.frame() #exclude bottom flank.
-DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
-
-##########
-# Data
-##########
-courtship_summary <- CourtshipBgrd %>% 
-                    group_by(Population, Background_pop, BodyRegion) %>%
-                summarise( Mean_dS = mean(dS, na.rm = T), 
-                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
-                             Upper_dS = Mean_dS + 1.96*SE_dS,
-                             Lower_dS = Mean_dS - 1.96*SE_dS, 
-                           Mean_dL = mean(dL, na.rm = T), 
-                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
-                             Upper_dL = Mean_dL + 1.96*SE_dL,
-                             Lower_dL = Mean_dL - 1.96*SE_dL) %>% filter(Population == "Hawaii") %>% as.data.frame() %>% mutate(Social_Context = "Courtship")
-
-display_summary <- DisplayBgrd %>%
-                group_by(Population, Background_pop, BodyRegion) %>%
-                summarise( Mean_dS = mean(dS, na.rm = T), 
-                             SE_dS = sd(dS)/sqrt(length(unique(ID))),
-                             Upper_dS = Mean_dS + 1.96*SE_dS,
-                             Lower_dS = Mean_dS - 1.96*SE_dS, 
-                           Mean_dL = mean(dL, na.rm = T), 
-                             SE_dL = sd(dL)/sqrt(length(unique(ID))),
-                             Upper_dL = Mean_dL + 1.96*SE_dL,
-                             Lower_dL = Mean_dL - 1.96*SE_dL) %>% filter(Population == "Hawaii") %>% as.data.frame() %>% mutate(Social_Context = "Male-male contest")
-
-##########
-# Plots
-##########
-
-#plots
-pS4.1 <- ggplot(courtship_summary, aes(x = Background_pop, y = Mean_dS, color = BodyRegion, group = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.5)) +
-  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
-  geom_line(position = position_dodge(width=0.5)) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("Background Environment") + 
-  ylab("Chromatic contrast (JND)") +
-  theme_bw() +
-  labs(colour = "Body Region") +
-  theme(legend.position = "none",
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-pS4.2 <- ggplot(courtship_summary, aes(x = Background_pop, y = Mean_dL, color = BodyRegion, group = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.5)) +
-  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
-  geom_line(position = position_dodge(width=0.5)) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("Background Environment") + 
-  ylab("Luminance contrast (JND)") +
-  theme_bw() +
-  labs(colour = "Body Region") +
-  theme(legend.position = "none",
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-pS4.3 <- ggplot(display_summary, aes(x = Background_pop, y = Mean_dL, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.5)) +
-  geom_errorbar(aes(ymin = Lower_dL, ymax = Upper_dL), position = position_dodge(width=0.5), width = 0) + 
-  geom_line(position = position_dodge(width=0.5)) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("Background Environment") + 
-  ylab("Luminance contrast (JND)") +
-  theme_bw() +
-  theme(legend.position = c(0.80, 0.85),
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-pS4.4 <- ggplot(display_summary, aes(x = Background_pop, y = Mean_dS, group = BodyRegion, color = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.5)) +
-  geom_errorbar(aes(ymin = Lower_dS, ymax = Upper_dS), position = position_dodge(width=0.5), width = 0) + 
-  geom_line(position = position_dodge(width=0.5)) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("Background Environment") + 
-  ylab("Chromatic contrast (JND)") +
-  theme_bw() +
-  theme(legend.position = "none",
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-#pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
-plot_grid(pS4.4, pS4.3, pS4.1, pS4.2 , labels = c("A", "B", "C", "D"), label_size = 30,
-  label_fontface = "bold", rel_widths=c(1,1))
-
-```
-
-```{r tableS7, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df), and 95% confidence intervals for each parameter.
-Main effects and interaction models are presented.
-Only main effect models are presented for luminance contrast (see text for details), whereas interaction models are presented for chromatic contrast.
-Pairwise contrasts are presented for chromatic contrast models for ease of interpretation."}
-source("./R/func.R")
-table7 <- rbind(robust_localAdapt_DispBgrdAC,
-                robust_DispBgrdCC,
-                robust_CourtshipAC,
-                robust_CourtshipCC) %>% as.data.frame()
-
-table7$Paramater <- rownames(table7)
-rownames(table7) <- NULL
-
-table7 <- table7 %>% 
-              mutate(Context = c("Male-Male Display", rep("", 14), "Courtship Display", rep("", 14)),
-                     JND = c("Luminance Contrast", rep("", 5), "Chromatic Contrast", rep("", 8), 
-                             "Luminance Contrast", rep("", 5), "Chromatic Contrast", rep("", 8))) %>% 
-                      dplyr::select(Context, JND, Paramater, beta, SE, df, CI_L, CI_U)
-
-Table7 <- flextable(table7) %>% table_style10() %>% fit_to_width(max_width = 7)
-Table7
-```
-
-```{r tableS8, tab.cap = "Pairwise Wald F tests comparing changes in chromatic contrast between different body regions for Hawaiian lizards against both Hawaiian and Kenyan backgrounds during male-male and courtship displays.
-F statistic, degrees of freedom (DF) and p-values are provided (p).
-A Bonferroni correction to p-values is also provided to control for multiple testing."}
-
-source("./R/func.R")
-tab8 <- rbind(result_DisplayBgrd_pop_back_regionCC,
-              result_CourtshipBgrd_pop_back_BR_CC) %>%  as.data.frame()
-
-tab8 <- tab8 %>% mutate(Context = c("Male-Male Display", rep("",3), "Courtship Display", rep("",3)),
-                        p_val = sapply(p_val, function(x) p_value(x))) %>% 
-                        select(Context,c(1:8), -test, -delta, -df_num)
-
-Table8 <- flextable(tab8) %>% table_style11() %>% fit_to_width(max_width = 7)
-Table8
-```
-
-## Hawaiian chameleons show greater individual change between display and bird color states than Kenyan chameleons
-
-```{r, results='hide'}
-#########################################################
-# Comparing displays
-#########################################################
-source("./R/func.R")
- # Load data from Devi
-       Display_bird_diff <- read_excel("data/Display_bird_diff.xlsx")
-
-       Diff <- Display_bird_diff
-
-      names(Diff) # header rows R is reading
-
-    Diff$BodyRegion <- factor(Diff$BodyRegion) #updates levels for the variable BodyRegion
-    str(diff.difftime) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
-    Diff$ID <- factor(Diff$ID) #updates levels for the variable ID
-    str(diff.difftime) #how R is reading the stucture of the object, double checking the levels for body region after updating levels
-
-# Center SVL to make intercept interpretable
-    Diff$sc_SVL <- with(Diff, as.vector(scale(SVL, scale=FALSE)))
-
-# Filter missing SVL data
-  Diff <- Diff %>% filter(!is.na(SVL)) # filter out misssing SVL data
-
-# create as factor
-    Diff$BodyRegion <- factor(Diff$BodyRegion)
-# Add interaction for contrasts
-    Diff$Pop_BodReg <- with(Diff, interaction(Population, BodyRegion))
-
-# Sample sizes
-    N_diff <- data.frame(Diff %>% group_by(Population) %>% summarise(n = length(unique(ID))))
-
-##################
-# Achromatic contrast
-##################
-
-# Fit model testing this idea above. First just test for body region interaction. But, needs RVE
-    Diff_dL_BR_int <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
-    summary(Diff_dL_BR_int)
-
-    # Significance of interaction
-      robust_Diff_dl_int <- Wald_test(Diff_dL_BR_int, constraints = constrain_zero(7:9), 
-                                    cluster = Diff$ID, vcov = "CR2")
-# Main effects BR
-    Diff_dL_BR_me <- lmer(dL_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
-    summary(Diff_dL_BR_me)
-
-    # Sig of Body Region
-        robust_Diff_dL_BR_me <- Wald_test(Diff_dL_BR_me, constraints = constrain_zero(4:6), 
-                                    cluster = Diff$ID, vcov = "CR2")
-
-# Then fit marginalised model so that body regions are averaged
-  Diff_dL <- lmer(dL_diff_abs ~ sc_SVL + Population + (1|ID), data=Diff)
-  summary(Diff_dL)
-
-# Check residuals of model
- # hist(residuals(Diff_dL))
- # plot(Diff_dL) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
-
-# Refit mdoel using robust standard errors
-  robust_Diff_dl_BR <- clubSandwich::conf_int(Diff_dL, cluster = Diff$ID, vcov =  "CR2")
-
-# Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
-
-  robust_Diff_dl_pop <- Wald_test(Diff_dL, constraints = constrain_zero(3), 
-                                  cluster = Diff$ID, vcov = "CR2")
-   robust_Diff_dl_svl <- Wald_test(Diff_dL, constraints = constrain_zero(2), 
-                                  cluster = Diff$ID, vcov = "CR2")
-
- # Get contrasts to understand the specific changes better.
-Refit model in slight different form.
-This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
-The model above is just a contrast-based model.
-
-          Diff_pop_back_BR_AC <- lmer(dL_diff_abs ~ -1 +  sc_SVL + Pop_BodReg + (1|ID), data = Diff)
-
-     # Fit Robust model
-          robust_Diff_pop_back_BR_AC <- clubSandwich::conf_int(Diff_pop_back_BR_AC, 
-                                                              cluster = Diff$ID, vcov =  "CR2")
-
-     # Contrasts hypothesis tests, but use RVE 
-          result_robust_Diff_pop_back_BR_AC <- Wald_test(Diff_pop_back_BR_AC, 
-                                                   constraints = constrain_pairwise(c(2:9)),
-                                                   cluster = Diff$ID,
-                                                   vcov="CR2", tidy = TRUE)
-             result_robust_Diff_pop_back_BR_AC$p_val_bonfer <- sapply(p.adjust(result_robust_Diff_pop_back_BR_AC$p_val, "bonferroni"), 
-                                                                    function(x) p_value(x))
-
-       # Grab contrasts that we need from the robust esatimtes
-      string <- c("Pop_BodRegKenya.gular - Pop_BodRegHawaii.gular|Pop_BodRegKenya.midflank - Pop_BodRegHawaii.midflank|Pop_BodRegKenya.tailbase - Pop_BodRegHawaii.tailbase|Pop_BodRegKenya.topflank - Pop_BodRegHawaii.topflank")
-
-         result_robust_Diff_pop_back_BR_AC <- result_robust_Diff_pop_back_BR_AC[
-                                            grep(string, result_robust_Diff_pop_back_BR_AC$hypothesis),]
-
-##################
-# Chromatic contrast
-##################
-
-  # Fit model testing this idea above. First just test for body region interaction. But, needs RVE
-    Diff_dS_BR_int <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion + BodyRegion:Population + (1|ID), data=Diff)
-    summary(Diff_dS_BR_int)
-
-    # Significance of interaction
-      robust_Diff_dS_int <- Wald_test(Diff_dS_BR_int, constraints = constrain_zero(7:9), 
-                                    cluster = Diff$ID, vcov = "CR2")
-  # Main effects BR
-    Diff_dS_BR_me <- lmer(dS_diff_abs ~ sc_SVL + Population + BodyRegion  + (1|ID), data=Diff)
-    summary(Diff_dS_BR_me)
-
-    # Sig of Body Region
-      robust_Diff_dS_BR_me <- Wald_test(Diff_dS_BR_me, constraints = constrain_zero(4:6), 
-                                    cluster = Diff$ID, vcov = "CR2")
-
-# Fit model testing this idea above
-  Diff_dS <- lmer(dS_diff_abs ~  SVL + Population + (1|ID), data=Diff)
-  summary(Diff_dS)
-
-# Check residuals of model
-  #hist(residuals(Diff_dS))
-  #plot(Diff_dS) # Some patterns in residuals, but likely driven by the absolute. Note that results are essentially the same if using absolute or raw, signed values.
-
-# Refit mdoel using robust standard errors
-  robust_Diff_dS <- clubSandwich::conf_int(Diff_dS, cluster = Diff$ID, vcov =  "CR2")
-
-# Wald tests - Test significance of pop dofferences and SVL differences averaging over body regions
-   robust_Diff_dS_pop <- Wald_test(Diff_dS, constraints = constrain_zero(3), 
-                                  cluster = Diff$ID, vcov = "CR2")
-   robust_Diff_dS_svl <- Wald_test(Diff_dS, constraints = constrain_zero(2), 
-                                  cluster = Diff$ID, vcov = "CR2")
-
-   # Get contrasts to understand the specific changes better.
-Refit model in slight different form.
-This is essentially an interaction model anyway because it's estimating the mean for Hawaiian and Kenyan background for each body region.
-The model above is just a contrast-based model.
-
-          Diff_pop_back_BR_CC <- lmer(dS_diff_abs ~ -1 +  sc_SVL + Pop_BodReg + (1|ID), data = Diff)
-
-     # Fit Robust model
-          robust_Diff_pop_back_BR_CC <- clubSandwich::conf_int(Diff_pop_back_BR_CC, 
-                                                              cluster = Diff$ID, vcov =  "CR2")
-
-     # Contrasts hypothesis tests, but use RVE 
-          result_robust_Diff_pop_back_BR_CC <- Wald_test(Diff_pop_back_BR_CC, 
-                                                   constraints = constrain_pairwise(c(2:9)),
-                                                   cluster = Diff$ID,
-                                                   vcov="CR2", tidy = TRUE)
-             result_robust_Diff_pop_back_BR_CC$p_val_bonfer <- sapply(p.adjust(result_robust_Diff_pop_back_BR_CC$p_val, "bonferroni"), 
-                                                                    function(x) p_value(x))
-
-       # Grab contrasts that we need from the robust esatimtes
-      string <- c("Pop_BodRegKenya.gular - Pop_BodRegHawaii.gular|Pop_BodRegKenya.midflank - Pop_BodRegHawaii.midflank|Pop_BodRegKenya.tailbase - Pop_BodRegHawaii.tailbase|Pop_BodRegKenya.topflank - Pop_BodRegHawaii.topflank")
-
-         result_robust_Diff_pop_back_BR_CC <- result_robust_Diff_pop_back_BR_CC[
-                                            grep(string, result_robust_Diff_pop_back_BR_CC$hypothesis),]
-
-```
-To gauge the extent of individual color change, we calculated the absolute difference in JNDs between for each individual that had both display (male-male contest) and bird response data (*n* = `r N_diff[1,2]` individuals from Hawaii, *n* = `r N_diff[2,2]` Kenya).
-We chose these two color states to maximize sample size and because chameleons and birds have similar tetrachromatic visual systems.
-We tested whether the absolute difference in luminance or chromatic contrast between these two color states differed between populations and body regions.
-We included SVL, population, body region and the interaction between population and body region as fixed effects.
-ID was included as a random effect becasue we had measures for more than one body region per individual.
-
-Body regions varied in their overall effects but mainly showed a decrease in absolute JND difference between display and antipredator color states in Kenya compared to Hawaii (Fig.
-S\@ref(fig:figDisplayPred)).
-While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, *P* = `r p_value(robust_Diff_dS_int[,6])`; luminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, *P* = `r p_value(robust_Diff_dl_int[,6])`) (Fig.
-S\@ref(fig:figDisplayPred) and Table S\@ref(tab:tableS9)).
-However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dS_BR_me[,6])`; luminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dL_BR_me[,6])`) (Fig.
-S\@ref(fig:figDisplayPred); Table S\@ref(tab:tableS9) and Table S\@ref(tab:tableS10)).
-
-```{r, figDisplayPred, fig.align = 'center', fig.width = 13, fig.height = 7, fig.cap = " Extent of color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (*n* = 30) and Kenyan (*n* = 13) chameleons.
-**(A)** Average chromatic and **(B)** luminance contrast are provided for each body region"}
-##########
-# Data
-##########
-abs_display_pred <- Diff %>%
-                group_by(Population, BodyRegion) %>%
-                summarise( Mean_dS_abs = mean(dS_diff_abs, na.rm = T), 
-                             SE_dS_abs = sd(dS_diff_abs)/sqrt(length(unique(ID))),
-                             Upper_dS_abs = Mean_dS_abs + 1.96*SE_dS_abs,
-                             Lower_dS_abs = Mean_dS_abs - 1.96*SE_dS_abs, 
-                           Mean_dL_abs = mean(dL_diff_abs, na.rm = T), 
-                             SE_dL_abs = sd(dL_diff_abs)/sqrt(length(unique(ID))),
-                             Upper_dL_abs = Mean_dL_abs + 1.96*SE_dL_abs,
-                             Lower_dL_abs = Mean_dL_abs - 1.96*SE_dL_abs) %>% as.data.frame()
-##########
-# Plots
-##########
-
-#plots
-p1 <- ggplot(abs_display_pred, aes(x = Population, y = Mean_dS_abs, color = BodyRegion, group = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.5)) +
-  geom_errorbar(aes(ymin = Lower_dS_abs, ymax = Upper_dS_abs), position = position_dodge(width=0.5), width = 0) + 
-  geom_line(position = position_dodge(width=0.5)) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("Population") + 
-  ylab("Chromatic Change (absolute JND)") +
-  theme_bw() +
-  labs(colour = "Body Region") +
-  theme(legend.position = "none",
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-p2 <- ggplot(abs_display_pred, aes(x = Population, y = Mean_dL_abs, color = BodyRegion, group = BodyRegion)) +
-  geom_point(size = 3, position = position_dodge(width=0.5)) +
-  geom_errorbar(aes(ymin = Lower_dL_abs, ymax = Upper_dL_abs), position = position_dodge(width=0.5), width = 0) + 
-  geom_line(position = position_dodge(width=0.5)) + 
-  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
-  xlab("Population") + 
-  ylab("Luminance Change (absolute JND)") +
-  theme_bw() +
-  labs(colour = "Body Region") +
-  theme(legend.position = c(0.80, 0.85),
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(),
-    axis.text=element_text(size=18),
-    axis.title=element_text(size=20,face="bold"),
-    legend.text =element_text(size = 12),
-    legend.title =element_text(size = 12, face = "bold"))
-
-#pdf(file="./output/figures/Figure2.pdf", height = 10, width = 18)
-plot_grid(p1, p2, labels = c("A", "B"), label_size = 30,
-  label_fontface = "bold", rel_widths=c(1,1))
-```
-
-```{r tableS9, tab.cap = "Estimated model parameters, standard errors (SE), degrees of freedom (df) along with 95% confidence interval for each parameter.
-Full interaction models are presented for chromatic and luminance change (absolute JND)."}
-source("./R/func.R")
-tab9 <- read.csv("./output/tables/Table5_MS.csv")
-
-Table9 <- flextable(tab9) %>% table_style13() %>% fit_to_width(max_width = 7)
-Table9
-```
-
-```{r tableS10, tab.cap = "Pairwise Wald F tests for chromatic and luminance change between different body regions for color plasticity between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states) for Hawaiian (*n* = 30) and Kenyan (*n* = 13) chameleons.
-F statistic, degrees of freedom (DF) and p-values are provided (p).
-A Bonferroni correction to p-values is also provided to control for multiple testing."}
-source("./R/func.R")
-tabS10 <- rbind(result_robust_Diff_pop_back_BR_AC, result_robust_Diff_pop_back_BR_CC)
-
-tabS10 <- tabS10 %>% mutate(VisRealm = c("Luminance Change", rep("",3), "Chromatic Change", rep("",3))) %>% 
-                        select(VisRealm, hypothesis, Fstat, df_denom, p_val, p_val_bonfer, -test, -delta, -df_num)
-
-Tables10 <- flextable(tabS10) %>% table_style14() %>% fit_to_width(max_width = 7)
-Tables10
-```
-
-# References
+**Acknowledgments**: This research was approved by the Kenya Wildlife Service (R.
+Bagine, KWS 5001) and Ministry of Science & Technology (Research Permit No.
+MOS&T 13/36C 208).
+**Funding**: This work was supported by the University of the Witwatersrand.
+**Authors contributions**: MJW conceived the study.
+MJW and DSF designed the study with help from JSK and BH.
+MJW, BSH, JSK and DSF collected the data.
+MJW, DSF, KJR and DWAN performed data curation, visualisation and exploratory analyses.
+DWAN performed final analyses.
+MJW wrote the first draft, with help from DWAN and DSF.
+All authors contributed substantially to subsequent versions.
+**Competing interests**: we (the authors) declare no competing interests.
+**Data and materials availability**: the data and code are available at OSF.
+
+**List of Supplementary materials**
+Materials and Methods
+Figures S1 – S9
+Tables S1 – S13
+References (24-48)
+
+# References and Notes

### 2022-01-22 ms_Nature.Rmd
Updated all "Supplemental" Fig and Tables in methods

@@
 
-We collected most chameleons (Fig.
-S\@ref(fig:Figcolsig2)) at night by spotlighting, and a small proportion of individuals during the day.
+We collected most chameleons (Supplementary Fig.
+1) at night by spotlighting, and a small proportion of individuals during the day.
 When chameleons were beyond reach in the forest canopy, we used lightweight 6-m extensible graphite composite poles (South Bend Kwik Stix™ 20' Telescopic Fishing Pole), that chameleons could be encouraged to step on to, before lowering them to the ground.
@@
 400-600 m.
-The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Fig.
-S\@ref(fig:Figcolsig2)C, D).
+The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Supplementary Fig.
+2C, D).
 Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
@@
 The area is partially cleared for agriculture, leaving small stands of forest.
-The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
-S\@ref(fig:Figcolsig2)A, B).
+The vegetation was a mix of native trees and shrubs interspersed with exotic species (Supplementary Fig.
+2A, B).
 Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.Jackson’s chameleons are a generalist species which can be locally abundant and relatively mobile.
@@
 Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
-We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand (Fig.
-S\@ref(fig:Figcolsig3)).
+We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand (Supplementary Fig.
+3).
 This structure was about 1.5 m high and each arm was approximately 60 cm.
-The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Fig.
-S\@ref(fig:Figcolsig3)).
+The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Supplementary Fig.
+3).
 All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
@@
 We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
-This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Fig.
-S\@ref(fig:Figcolsig)A, B).
+This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Supplementary Fig.
+1A, B).
 In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
-This would be accompanied by head shakes and gapping, leading to horn-locking (Fig.
-S\@ref(fig:Figcolsig)B) if males were not obviously asymmetric.
-Once dominance was settled, the subordinate would turn brown (Fig.
-S\@ref(fig:Figcolsig)B and Fig.
-S\@ref(fig:Figcolsig3)) and flee.
-In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
-S\@ref(fig:Figcolsig)C).
-
-The day following social trials we presented each chameleon with a snake (Fig.
-S\@ref(fig:Figcolsig4)A) and bird (Fig.
-S\@ref(fig:Figcolsig4)B) model predator model (separately).
+This would be accompanied by head shakes and gapping, leading to horn-locking Supplementary Fig.
+1B) if males were not obviously asymmetric.
+Once dominance was settled, the subordinate would turn brown (Supplementary Fig.
+1B and Supplementary Fig.
+3) and flee.
+In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Supplementary Fig.
+1C).
+
+The day following social trials we presented each chameleon with a snake (Supplementary Fig.
+4A) and bird (Supplementary Fig.
+4B) model predator model (separately).
 Lizards were drawn at random and the order of presentation (snake or bird first) was random.
@@
 A trial began when a chameleon was removed from its cloth bag and placed on the triangular stand.
-In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Fig.
-S\@ref(fig:Figcolsig4)A).
+In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Supplementary Fig.
+4A).
 In the case of the bird trials, the bird was attached to an ca.
-2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Fig.
-S\@ref(fig:Figcolsig4)B).
+2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Supplementary Fig.
+4B).
 Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
@@
 
-During all behavior trials, we measured the color of each of three body regions: top flank, mid flank, tail base and gular region (Fig.
-S\@ref(fig:Figcolsig5)).
+During all behavior trials, we measured the color of each of three body regions: top flank, mid-flank, tail base and gular region (Supplementary Fig.
+5).
 We measured each body region once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
@@
 
-```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Location of body regions we measured for spectral reflectance**. See details in text."}
-figs5 <- image_read("./output/figures/FigS5.jpg")
-figs5
-
-```
-
 ## Visual modeling
@@
 We calculated noise for the other photoreceptor classes using the relative photoreceptor ratio of 1: 1.6: 3.3: 3.2 for the four single cones (UVS, SWS, MWS, and LWS, respectively).
-*Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Fig.
-S\@ref(fig:Figcolsig6SPec)).
-
-```{r Figcolsig6SPec, fig.align = 'center', fig.width = 12, fig.height = 10, fig.cap = "**Spectral sensitivities of chameleons (A), bird (B) and snake (C) used for visual models (details in text)**."}
-sens_cham_c_dilepis <- read_excel("./data/Spectral sensitivities.xls", sheet = "C_dilepis")[,1:6]
-        sens_birdUV <- read_excel("./data/Spectral sensitivities.xls", sheet = "birdUV")
-         sens_snake <- read_excel("./data/Spectral sensitivities.xls", sheet = "snake")[,1:4]
-
-cham <- sens_cham_c_dilepis %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
-bird <- sens_birdUV %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
-snake <- sens_snake %>% pivot_longer(cols = c(2:4)) %>% select(wavelength, name, value) %>%  data.frame()
-
-plot_spec <- function(data, tri = FALSE){
-        data %>% ggplot(aes(x = wavelength, y = value)) + 
-      geom_line(aes(color = name), size = 1.2) + 
-      theme_bw() +
-    scale_colour_manual(values = rev(c("#8D25CC", "#3770C9", "#93C572", "#FFA500", "#DBD8D6"))) +
-      labs(x = "Wavelength (nm)", y = "Absorbance", color = "") + 
-        theme(legend.position  = c(0.90, 0.65),
-            panel.grid.major = element_blank(), 
-            panel.grid.minor = element_blank(),
-            axis.text = element_text(size = 16),
-            axis.title = element_text(size = 24),
-            legend.text = element_text(size = 16),
-            legend.title = element_text(size = 20))}
-
-p1 <- plot_spec(cham) + theme(legend.position  = "none")
-p2 <- plot_spec(bird) 
-p3 <- plot_spec(snake) + scale_colour_manual(values = rev(c("#8D25CC", "#93C572", "#FFA500"))) + theme(legend.position  = "none")
-
-plot_grid(p1, p2,p3,nrow = 3, ncol = 1, labels = c("A", "B", "C"), label_size = 30,
-  label_fontface = "bold", rel_widths=c(1,1))
-
-```
+*Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Supplementary Fig.
+6).
 
@@
 A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement @RN20.
-We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21 (Fig.
-S\@ref(fig:Figcolsig6SPec)B).
+We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21 (Supplementary Fig.
+6B).
 
 For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake @RN22.
-We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22 (Fig.
-S\@ref(fig:Figcolsig6SPec)C).
+We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22 (Supplementary Fig.
+6C).
 

### 2022-01-22 ms_Nature.Rmd
Figure Legends fixed. Need to do figures

@@
 They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
-**(A)** A dominant male in display coloration.
-**(B)** A subordinate male that lost a contest and turned from bright yellow to brown.
-**(C)** Two males fighting, both are in display coloration and relatively even matched.
-**(D)** A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
-See supplementary text for additional photos, including in response to a snake."}
+a, A dominant male in display coloration.
+b, A subordinate male that lost a contest and turned from bright yellow to brown.
+c, Two males fighting, both are in display coloration and relatively even matched.
+d, A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
+See Supplementary information for additional photos, including in response to a snake"}
 library(magick)
@@
 ```{r figure2, fig.width = 14, fig.height= 14, fig.cap="**Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya**.
-The context for measurement was **(A)** male contest displays; **(B)** courtship; **(C)** bird predator; and **(D)** snake predator.
-More details in online supplementary text and Fig.
-S5."}
+The context for measurement was a, male contest displays; b, courtship; c, bird predator; and d, snake predator.
+More details in online Supplementary Information and Supplementary Fig.
+5"}
 library(magick)
@@
 ```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "**Chromatic and luminance contrast in Just Noticeable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves)**.
-Luminance **(A)** and chromatic **(B)** contrast of male chameleons during male-male contests and female courtship.
+Luminance a, and chromatic b, contrast of male chameleons during male-male contests and female courtship.
 JNDs are calculated based on the chameleon visual system.
-Luminance **(C)** and chromatic **(D)** contrasts of male chameleons during snake  and birdpredator encounters.
+Luminance c, and chromatic d, contrast of male chameleons during snake and bird predator encounters.
 JNDs are calculated based on the snake and bird visual systems.
@@
 
-```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "**Evidence for local adaptation of social color signals in both (A) chromatic- and (B) luminance contrast for Hawaiian chameleons**.
+```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "**Evidence for local adaptation of social color signals in both a, chromatic- and b, luminance contrast for Hawaiian chameleons**.
 For local adaptation, signals are predicted to be more conspicuous against their own background.

### 2022-01-22 ms_Nature.Rmd
Updated yaml header

@@
 title: "Invasive chameleons released from predation display more conspicuous colors"
-bibliography: ms_refs.bib
-csl: science.csl
+bibliography: supp_refs.bib
+csl: nature.csl
 output: 

### 2022-01-22 ms_Nature.Rmd
Updated all refs so they grab the correct ones now. Note that I changed to supplement bib file because that bib contains all refs. The fact MS and supp are now merged means that all MS refs have a "ms" after them to denote they are part of the MS. I've added ms to all RN numbers for MS text

@@
 # Abstract
-Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid phenotypic change [@Losos2004; @RN4; @RN5; @Stastny2017].
-In particular, conspicuous social and sexual signals are predicted to experience significant character release when natural selection via predation is relaxed [@RN7; @RN4].
+Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid phenotypic change [@Losos2004; @RN4ms; @RN5ms; @Stastny2017].
+In particular, conspicuous social and sexual signals are predicted to experience significant character release when natural selection via predation is relaxed [@RN7ms; @RN4ms].
 However, we have few good examples of this phenomenon in the wild and none in species with dynamic colour change.
@@
 
-When species encounter novel environments, such as during human-mediated introductions, phenotypic change can occur more rapidly than Darwin ever thought possible [@RN1; @RN2; @RN3; @RN4].
-In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4; @RN5].
-Such character release should be most evident for social and sexual signals; released from constraints on conspicuousness imposed by predation risk, signals could elaborate in many directions [e.g., @RN4; @RN6; @RN7; @RN8].
-However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7; @RN9; @RN10].
+When species encounter novel environments, such as during human-mediated introductions, phenotypic change can occur more rapidly than Darwin ever thought possible [@RN1ms; @RN2ms; @RN3ms; @RN4ms].
+In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4ms; @RN5ms].
+Such character release should be most evident for social and sexual signals; released from constraints on conspicuousness imposed by predation risk, signals could elaborate in many directions [e.g., @RN4ms; @RN6ms; @RN7ms; @RN8ms].
+However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7ms; @RN9ms; @RN10ms].
 This is especially true for species with dynamic color change, for which we are not aware of any examples.
@@
 1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
-The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11].
-Jackson’s chameleons have a high intrinsic population growth rate (9-12 month generation time and live birth of up to 50 young @RN24), enabling their rapid establishment and potentially rapid evolution (approx.
+The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11ms].
+Jackson’s chameleons have a high intrinsic population growth rate (9-12 month generation time and live birth of up to 50 young @RN24ms), enabling their rapid establishment and potentially rapid evolution (approx.
 50 -- 65 generations).
-On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators absent or rare, Supplementary Information), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014.
+On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators absent or rare, Supplementary Information), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014ms.
 This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.
 
-Chameleons are famous for their rapid color change [@RN12; @RN13; @RN14].
+Chameleons are famous for their rapid color change [@RN12ms; @RN13ms; @RN14ms].
 When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green during courtship or to signal dominance (Fig.
@@
 S5).
-We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
 
@@
 \@ref(fig:figure1)).
-Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15 or changes in the spacing of guanine crystals within iridophores @RN14, both of which strongly affect luminance.
+Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15ms or changes in the spacing of guanine crystals within iridophores @RN14ms, both of which strongly affect luminance.
 Accordingly, character release was observed in the overall luminance of signals rather than their hue.
@@
 \@ref(fig:figure3)D and Supplementary Table 4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1; JND values ~ 2).
-The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18].
+The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18ms].
 The poor color discrimination of snakes may also explain why Hawaiian and Kenyan chameleons differ little their chromatic response to snakes, despite the absence of snakes on Hawaii.
 
-Character release may reflect rapid evolutionary change, phenotypic plasticity or a combination of these processes [@RN19; @RN20].
+Character release may reflect rapid evolutionary change, phenotypic plasticity or a combination of these processes [@RN19ms; @RN20ms].
 Rapid evolutionary change results in local adaptation, whereby signals become more conspicuous to conspecifics in their local environment than in ancestral environments.
@@
 \@ref(fig:figure4) and Supplementary Table 5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
-These results are consistent with local adaptation, in which phenotypic plasticity, which can also evolve rapidly [@RN26], may also play a role.
+These results are consistent with local adaptation, in which phenotypic plasticity, which can also evolve rapidly [@RN26ms], may also play a role.
 
@@
 However, the measured colours and colour change represent a dynamic response to specific stimuli (conspecific, model predator) in contrast to the ‘fixed’ colour signals of many birds, fishes and lizards that have been shown to be affected by diet and other environmental factors.
-Additionally, the mechanisms underlying the rapid colour response of chameleons do not involve dietary pigments, and luminance change is partially structural [@RN14].
+Additionally, the mechanisms underlying the rapid colour response of chameleons do not involve dietary pigments, and luminance change is partially structural [@RN14ms].
 Therefore, environmental effects (other than predation) are unlikely to account for major differences in the color responses we observed between populations.
@@
 Oahu has no snakes or raptors that feed on lizards.
-The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22].
-In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014 -- the conditions were highly favorable for character release of social signals.
+The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22ms].
+In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014ms -- the conditions were highly favorable for character release of social signals.
 Jackson’s chameleons likely experience strong sexual selection given their polygynous mating system, elaborate courtship displays and intense male contest competition (Fig.
@@
 
-Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12].
+Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12ms].
 Our results suggest that even for species capable of dynamic color change there is an upper threshold to signal intensity that is constrained by natural selection.

### 2022-01-22 ms_Nature.Rmd
Fixed awknowlefgements etc.

@@
 
-Chromatic and luminance contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN24 in combination with the package `clubSandwich` @RN25.
+Chromatic and luminance contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package [@RN24] in combination with the package `clubSandwich` [@RN25].
 For each individual, measurements for each body region (top flank, mid-flank, tail base and gular region) were contrasted against each of six average background spectra (i.e., average top and bottom of lighter leaves, average top and bottom of darker leaves, average green and average brown stems).
-While we present sensitivity analyses for all body regions (see below - *Additional Supplemental Results*), we were mainly interested in overall signal differences displayed by chameleons.
-As such, for our main analyses we averaged estimates (marginalized) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see *Additional Supplemental Results*).
-In all cases, our results and conclusions were not impacted (see supplementary code).
+While we present sensitivity analyses for all body regions (see Supplemental Results), we were mainly interested in overall signal differences displayed by chameleons.
+As such, for our main analyses we averaged estimates (marginalized) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplemental Results).
+In all cases, our results and conclusions were not impacted (see Supplemental Results).
 Deriving multiple contrasts for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other.
@@
 They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
-a, A dominant male in display coloration.
-b, A subordinate male that lost a contest and turned from bright yellow to brown.
-c, Two males fighting, both are in display coloration and relatively even matched.
-d, A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
+**a**, A dominant male in display coloration.
+**b**, A subordinate male that lost a contest and turned from bright yellow to brown.
+**c**, Two males fighting, both are in display coloration and relatively even matched.
+**d**, A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
 See Supplementary information for additional photos, including in response to a snake"}
@@
 
-img65 <- image_read("./output/figures/fig1.pdf")
+img65 <- image_read("./output/figures/fig1_ms_Nature.png")
 img65
@@
 ```{r figure2, fig.width = 14, fig.height= 14, fig.cap="**Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya**.
-The context for measurement was a, male contest displays; b, courtship; c, bird predator; and d, snake predator.
+The context for measurement was **a**, male contest displays; **b**, courtship; **c**, bird predator; and **d**, snake predator.
 More details in online Supplementary Information and Supplementary Fig.
@@
 library(magick)
-img <- image_read("./output/figures/Figure3.png")
+img <- image_read("./output/figures/Figure3_Nature.png")
 img
@@
 ```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "**Chromatic and luminance contrast in Just Noticeable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves)**.
-Luminance a, and chromatic b, contrast of male chameleons during male-male contests and female courtship.
+Luminance **a**, and chromatic **b**, contrast of male chameleons during male-male contests and female courtship.
 JNDs are calculated based on the chameleon visual system.
-Luminance c, and chromatic d, contrast of male chameleons during snake and bird predator encounters.
+Luminance **c**, and chromatic **d**, contrast of male chameleons during snake and bird predator encounters.
 JNDs are calculated based on the snake and bird visual systems.
@@
 
-  img2 <- image_read("./output/figures/Figure1.png")
+  img2 <- image_read("./output/figures/Figure1_ms_Nature.png")
   img2
@@
 } else {
-  img3 <- image_read("./output/figures/Figure2.png")
+  img3 <- image_read("./output/figures/Figure2_ms_Nature.png")
   img3
@@
 
-**Acknowledgments**: This research was approved by the Kenya Wildlife Service (R.
+**Acknowledgments** 
+This research was approved by the Kenya Wildlife Service (R.
 Bagine, KWS 5001) and Ministry of Science & Technology (Research Permit No.
 MOS&T 13/36C 208).
-**Funding**: This work was supported by the University of the Witwatersrand.
-**Authors contributions**: MJW conceived the study.
+Financial support was provided by the University of the Witwatersrand to MJW.
+We thank Idle Farah and Patrick Malonza of the National Museums of Kenya for facilitating the research, Joseph Gathua for field support and Jan Stipala for use of his field vehicle.
+We thank the Ditsong National Museum of Natural History (formerly Transvaal Museum) for the stuffed African Cuckoo Hawk.
+Ben Phillips, Rick Shine and Tobias Uller provided feedback on earlier drafts of this manuscript.
+We thank Fonti Kar for help curating data and Drew Allen for statistical advice.
+
+*Contributions**
+MJW conceived the study.
 MJW and DSF designed the study with help from JSK and BH.
@@
 All authors contributed substantially to subsequent versions.
-**Competing interests**: we (the authors) declare no competing interests.
-**Data and materials availability**: the data and code are available at OSF.
-
-**List of Supplementary materials**
-Materials and Methods
-Figures S1 – S9
-Tables S1 – S13
-References (24-48)
+
+*Corresponding author**
+Correspondence to Martin Whiting.
+
+*Ethics declarations**
+
+*Competing interests**
+We (the authors) declare no competing interests.
+
+*Extended data figures and tables**
+The data and code are available at OSF.
+
+*Supplementary information**
+This file includes supplementary notes, 9 Supplementary Figures and 13 Supplementary Tables.
 

### 2022-01-22 ms_Nature.Rmd
Added some more detail on stats for new Table S6

@@
 For these models, chromatic or luminance contrast against the local background (as perceived by the corresponding predator type) was the response variable and predator type (i.e, snake or bird), population of origin (Hawaii or Kenya) and their interaction were fixed effects.
-Finally, to evaluate evidence for local adaptation, we tested whether the chromatic or luminance contrast of Hawaiian chameleons during social interactions (male-male and male-female) varied depending on whether contrast was against a native Kenyan background or their current introduced habitat background in Hawaii.
 For all models, interaction terms were first tested using Wald tests.
@@
 We then tested the significance of the main effects using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.
+
+Finally, to evaluate evidence for local adaptation, we tested whether the chromatic or luminance contrast of Hawaiian chameleons during social interactions (male-male and male-female) varied depending on whether they were contrasted against a native Kenyan background or their current introduced habitat background in Hawaii.
+In other words, we pamaterised these models to include the interaction between population and background, such that each population (Kenya and Hawaii) was compared against each background (Kenyan and Hawaiian).
+Doing so decouples the confounding effect of lizard display from their background.
+We present full interaction models given that we were, *a priori*, interested in specific comparisons.
 

### 2022-01-22 ms_Nature.Rmd
I don't think we need any more detail on new analyses really. I think it's all clear, but I made reference to table S6 now.

@@
 Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.
-\@ref(fig:figure4) and Supplementary Table 5; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
+\@ref(fig:figure4) and Supplementary Tables 5 and 6; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
 These results are consistent with local adaptation, in which phenotypic plasticity, which can also evolve rapidly [@RN26ms], may also play a role.

### 2022-01-22 ms_Nature.Rmd
Just fixes to title page

@@
 ---
-title: "Invasive chameleons released from predation display more conspicuous colors"
+title: "Title: Invasive chameleons released from predation display more conspicuous colors"
 bibliography: supp_refs.bib
@@
 
-Martin J. Whiting^1,2^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel W.A.Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
-
+**Authors** Martin J. Whiting^1,2^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel W.A.Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
+
+**Affiliations**
 ^1^ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
@@
 
+# Main
 When species encounter novel environments, such as during human-mediated introductions, phenotypic change can occur more rapidly than Darwin ever thought possible [@RN1ms; @RN2ms; @RN3ms; @RN4ms].
@@
 When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green during courtship or to signal dominance (Fig.
-1, figs.
-S1 and S3).
+1, Supplementary Fig.
+1 and 3).
 At other times, they are a duller green, brown or become mottled green-brown, particularly in response to predators (Fig.
-\@ref(fig:figure1) and fig.
-S4A).
+\@ref(fig:figure1) and Supplementary Fig.
+4A).
 Their different color states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
@@
 
-To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (fig.
-S4), or a control (stick).
+To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Supplementary Fig.
+4), or a control (stick).
 We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
-\@ref(fig:figure2) and fig.
-S5).
+\@ref(fig:figure2) and Supplementary Fig.
+5).
 We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
@@
 The area experiences high annual rainfall (mean: ca.
-1500 mm) @RN6 and the mean annual temperature is 22-24°C @RN7.
+1500 mm) [@RN6] and the mean annual temperature is 22-24°C [@RN7].
 In Kenya, field work was conducted near the foothills of Mt.
@@
 2A, B).
-Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.Jackson’s chameleons are a generalist species which can be locally abundant and relatively mobile.
+Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm [@RN8].Jackson’s chameleons are a generalist species which can be locally abundant and relatively mobile.
 This population is from the region (slopes of Mount Kenya) identified as the source population [@Tilbury2018].
@@
 
-We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` @RN29.
+We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` [@RN29].
 This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
-Color contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of color discrimination for stimuli viewed simultaneously under ideal viewing conditions @RN17.
+Color contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of color discrimination for stimuli viewed simultaneously under ideal viewing conditions [@RN17].
 Under natural conditions, discrimination thresholds are likely to be >1 and vary depending on the species and conditions [@RN12;@RN13; @RN14].
-Additionally, perceived conspicuousness may not scale linearly with color distance (JNDs) @RN15.
+Additionally, perceived conspicuousness may not scale linearly with color distance (JNDs) [@RN15].
 Nevertheless, the models provide a reasonable approximation of relative conspicuousness that can be compared between populations and trial types (social and predator).
@@
 The visual models require information on the spectral sensitivities and photoreceptor noise (Weber fraction), which is a function of the relative density of each photoreceptor type within the retina, for the relevant receiver.
-Chameleons have four spectral classes of single cone enabling tetrachromatic vision @RN16.
-We used spectral sensitivities for the congeneric flap-necked chameleon (*Chameleo dilepis*) @RN16 and a standard Weber fraction of 0.1 @RN17 to account for the receptor noise of the LWS photoreceptor.
+Chameleons have four spectral classes of single cone enabling tetrachromatic vision [@RN16].
+We used spectral sensitivities for the congeneric flap-necked chameleon (*Chameleo dilepis*) [@RN16] and a standard Weber fraction of 0.1 @RN17 to account for the receptor noise of the LWS photoreceptor.
 We calculated noise for the other photoreceptor classes using the relative photoreceptor ratio of 1: 1.6: 3.3: 3.2 for the four single cones (UVS, SWS, MWS, and LWS, respectively).
@@
 
-For birds, we used an ultraviolet sensitive (UVS) visual system @RN18 characteristic of raptors and other birds of prey @RN19.
-A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement @RN20.
-We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21 (Supplementary Fig.
+For birds, we used an ultraviolet sensitive (UVS) visual system [@RN18] characteristic of raptors and other birds of prey [@RN19].
+A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement [@RN20].
+We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds [@RN21] (Supplementary Fig.
 6B).
 
-For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake @RN22.
-We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22 (Supplementary Fig.
+For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake [@RN22].
+We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively [@RN22] (Supplementary Fig.
 6C).
 
-In each case, visual pigment absorbance curves were multiplied by the transmission spectra of ocular media (lens and cornea) and oil droplets associated with the corresponding photoreceptor class and normalized to equal area under the curve to satisfy the assumption of equal stimulation by white light @RN18.
+In each case, visual pigment absorbance curves were multiplied by the transmission spectra of ocular media (lens and cornea) and oil droplets associated with the corresponding photoreceptor class and normalized to equal area under the curve to satisfy the assumption of equal stimulation by white light [@RN18].
 
@@
 
-*Contributions**
+**Contributions**
 MJW conceived the study.
@@
 
-*Corresponding author**
+**Corresponding author**
 Correspondence to Martin Whiting.
 
-*Ethics declarations**
-
-*Competing interests**
+**Ethics declarations**
+
+**Competing interests**
 We (the authors) declare no competing interests.
 
-*Extended data figures and tables**
+**Extended data figures and tables**
 The data and code are available at OSF.
 
-*Supplementary information**
+**Supplementary information**
 This file includes supplementary notes, 9 Supplementary Figures and 13 Supplementary Tables.

### 2022-01-22 ms_Nature.Rmd
Changed all occurances of color to colour

@@
 Here, we show that Jackson’s chameleons (*Trioceros j.
-xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu), where there are few lizard predators, experienced significant character release of color signals.
-Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator color states.
-Hawaiian chameleon display colors were also more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
+xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu), where there are few lizard predators, experienced significant character release of colour signals.
+Hawaiian chameleons displayed more conspicuous colour signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator colour states.
+Hawaiian chameleon display colours were also more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
 These results demonstrate for the first time that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
@@
 However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7ms; @RN9ms; @RN10ms].
-This is especially true for species with dynamic color change, for which we are not aware of any examples.
-Here, we examine character release of a dynamic color signal by taking advantage of an unintended evolutionary ‘experiment’: accidental translocation of chameleons to a novel, low predation environment.
+This is especially true for species with dynamic colour change, for which we are not aware of any examples.
+Here, we examine character release of a dynamic colour signal by taking advantage of an unintended evolutionary ‘experiment’: accidental translocation of chameleons to a novel, low predation environment.
 
@@
 
-Chameleons are famous for their rapid color change [@RN12ms; @RN13ms; @RN14ms].
+Chameleons are famous for their rapid colour change [@RN12ms; @RN13ms; @RN14ms].
 When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green during courtship or to signal dominance (Fig.
@@
 4A).
-Their different color states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
-If introduced chameleons in Hawaii experience character release from reduced predation, we expect males to have more conspicuous display colors in response to conspecifics and to be less cryptic in response to bird and snake predators.
-We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.
+Their different colour states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
+If introduced chameleons in Hawaii experience character release from reduced predation, we expect males to have more conspicuous display colours in response to conspecifics and to be less cryptic in response to bird and snake predators.
+We should also see evidence of local adaptation: colour signals should be more conspicuous to conspecifics against the local than ancestral background.
 
@@
 4), or a control (stick).
-We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
+We measured the spectral reflectance of male colour states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
 \@ref(fig:figure2) and Supplementary Fig.
 5).
-We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
+We then modelled how conspicuous these colours would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
 
 # Results and Discussion
-In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.
+In support of predictions of character release, male display colouration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.
 \@ref(fig:figure2), Fig.
@@
 This result was consistent for all body regions (Supplementary Information).
-However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of color signals (Fig.
+However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of colour signals (Fig.
 \@ref(fig:figure3)B and Supplementary Table 2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
 
-The importance of luminance contrast is consistent with the nature and mechanism of rapid color change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.
+The importance of luminance contrast is consistent with the nature and mechanism of rapid colour change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.
 \@ref(fig:figure1)).
-Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15ms or changes in the spacing of guanine crystals within iridophores @RN14ms, both of which strongly affect luminance.
+Specifically, colour change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15ms or changes in the spacing of guanine crystals within iridophores @RN14ms, both of which strongly affect luminance.
 Accordingly, character release was observed in the overall luminance of signals rather than their hue.
@@
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18ms].
-The poor color discrimination of snakes may also explain why Hawaiian and Kenyan chameleons differ little their chromatic response to snakes, despite the absence of snakes on Hawaii.
+The poor colour discrimination of snakes may also explain why Hawaiian and Kenyan chameleons differ little their chromatic response to snakes, despite the absence of snakes on Hawaii.
 
@@
 
-To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
+To gauge the extent of colour plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) colour states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
 While there was evidence for differences between body regions (Supplementary Information, Supplementary Fig.
@@
 9).
-Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
+Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator colour states.
 An alternate explanation is that differences might be due to a founder effect because chameleons might have been selectively collected for the pet trade.
 While we acknowledge this possibility, it is exceedingly unlikely.
-Chameleons are almost always encountered in a camouflaged color state because they are solitary and only use display colours during relatively brief social encounters which precludes the possibility that hunters selected the most brightly coloured individuals.
+Chameleons are almost always encountered in a camouflaged colour state because they are solitary and only use display colours during relatively brief social encounters which precludes the possibility that hunters selected the most brightly coloured individuals.
 Nevertheless, even if pet traders were selecting for larger animals, our analyses controlled for any relationship body size might have on colour.
@@
 Additionally, the mechanisms underlying the rapid colour response of chameleons do not involve dietary pigments, and luminance change is partially structural [@RN14ms].
-Therefore, environmental effects (other than predation) are unlikely to account for major differences in the color responses we observed between populations.
+Therefore, environmental effects (other than predation) are unlikely to account for major differences in the colour responses we observed between populations.
 The degree to which colour change capacity is plastic in chameleons and other organisms with dynamic colour change is unknown.
@@
 Oahu has no snakes or raptors that feed on lizards.
-The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22ms].
+The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited colour vision [@RN22ms].
 In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014ms -- the conditions were highly favorable for character release of social signals.
@@
 Consequently, a unique set of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
-To the best of our knowledge, this is also the first example of character release in a dynamic color signal.
-
-Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12ms].
-Our results suggest that even for species capable of dynamic color change there is an upper threshold to signal intensity that is constrained by natural selection.
-This novel finding raises the question of whether other animals that have convergently evolved dynamic color signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
+To the best of our knowledge, this is also the first example of character release in a dynamic colour signal.
+
+Rapid colour change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12ms].
+Our results suggest that even for species capable of dynamic colour change there is an upper threshold to signal intensity that is constrained by natural selection.
+This novel finding raises the question of whether other animals that have convergently evolved dynamic colour signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
 More generally, our study highlights the opportunities provided by invasions to study natural and sexual selection in the wild.
@@
 ## Behavioral trials
-In order to quantify display coloration, we staged encounters between conspecifics (male-male and male-female) and model predators (bird and snake) and measured their subsequent display coloration (details below).
+In order to quantify display colouration, we staged encounters between conspecifics (male-male and male-female) and model predators (bird and snake) and measured their subsequent display colouration (details below).
 All chameleons were used in both predator and social trials although in some cases we were unable to quantify all social signals for a particular individual.
 We conducted all conspecific-social trials first, in case anti-predator trials unduly influenced their social interactions.
-Our analyses focused on comparing color responses of chameleons to conspecifics and predators to ensure that the color state corresponded to specific stimuli.
-We did not, and could not, measure a ‘neutral’ color state because it is impossible to meaningfully gauge a neutral state in a color-changing organism such as a chameleon.
+Our analyses focused on comparing colour responses of chameleons to conspecifics and predators to ensure that the colour state corresponded to specific stimuli.
+We did not, and could not, measure a ‘neutral’ colour state because it is impossible to meaningfully gauge a neutral state in a colour-changing organism such as a chameleon.
 However, for predator trials, we ensured that chameleons were specifically responding to the predators by conducting seven trials with a just a branch with no predator attached, as a control.
 Chameleons did not respond to the branch (see also @RN18ms).
-Furthermore, during trials with predators, chameleons focused on the predator itself, and exhibited classic anti-predator behavior (e.g., contrasting stress color, body inflation, open-mouth threat).
+Furthermore, during trials with predators, chameleons focused on the predator itself, and exhibited classic anti-predator behavior (e.g., contrasting stress colour, body inflation, open-mouth threat).
 
@@
 We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
-This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Supplementary Fig.
+This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing colour and behavior (Supplementary Fig.
 1A, B).
-In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
+In male-male trials both individuals would adopt a lemon-yellow display colouration and advance towards each other.
 This would be accompanied by head shakes and gapping, leading to horn-locking Supplementary Fig.
@@
 Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
-Once chameleons changed color, we quantified their spectral reflectance (details below).
+Once chameleons changed colour, we quantified their spectral reflectance (details below).
 The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
@@
 
-## Color measurement
-
-During all behavior trials, we measured the color of each of three body regions: top flank, mid-flank, tail base and gular region (Supplementary Fig.
+## colour measurement
+
+During all behavior trials, we measured the colour of each of three body regions: top flank, mid-flank, tail base and gular region (Supplementary Fig.
 5).
-We measured each body region once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
+We measured each body region once only, because chameleons change colour rapidly and we had to reduce the handling stress to the animal.
 The four body regions were measured in random sequence to avoid any bias of order.
-When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
+When a chameleon had completely changed colour (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
 We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal during measurement.
@@
 All measurements were relative to a dark, and a certified 99% white reflectance standard (Labsphere, North Sutton, NH, U.S.A.).
-If chameleons started to change color in response to our handling, we returned them to the branch and allowed them to continue a social trial before measuring reflectance of remaining body regions.
+If chameleons started to change colour in response to our handling, we returned them to the branch and allowed them to continue a social trial before measuring reflectance of remaining body regions.
 In the case of a predator trial, we would once again present them with the predator model until they returned to their anti-predator display.
@@
 
-We obtained spectral reflectance measurements of male dominant color state in male-male contests for 34 individuals from Hawaii and 35 from Kenya, courtship color state for 25 individuals from Hawaii and 32 from Kenya; color state in response to the bird for 56 individuals from Hawaii and 23 from Kenya and color state in response to the snake for 59 individuals from Hawaii and 38 from Kenya.
+We obtained spectral reflectance measurements of male dominant colour state in male-male contests for 34 individuals from Hawaii and 35 from Kenya, courtship colour state for 25 individuals from Hawaii and 32 from Kenya; colour state in response to the bird for 56 individuals from Hawaii and 23 from Kenya and colour state in response to the snake for 59 individuals from Hawaii and 38 from Kenya.
 
@@
 To reduce the number of backgrounds, we averaged the reflectance of the 20 darker measurements and 20 lighter measurements for each of the top leaf and bottom leaf measurements and averaged brown stems and green stems.
-This resulted in 6 average background spectra for each location, representing the range of background colors.
+This resulted in 6 average background spectra for each location, representing the range of background colours.
 
@@
 We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` [@RN29].
-This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
-Color contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of color discrimination for stimuli viewed simultaneously under ideal viewing conditions [@RN17].
+This model assumes that colour discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
+colour contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of colour discrimination for stimuli viewed simultaneously under ideal viewing conditions [@RN17].
 Under natural conditions, discrimination thresholds are likely to be >1 and vary depending on the species and conditions [@RN12;@RN13; @RN14].
-Additionally, perceived conspicuousness may not scale linearly with color distance (JNDs) [@RN15].
+Additionally, perceived conspicuousness may not scale linearly with colour distance (JNDs) [@RN15].
 Nevertheless, the models provide a reasonable approximation of relative conspicuousness that can be compared between populations and trial types (social and predator).
 
-We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region and color state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
+We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region and colour state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
 We first calculated the quantum catch of each photoreceptor under standard daylight illumination (D65; Commission internationale de l’éclairage – CIE) using the ‘vismodel’ function.
@@
 First, we tested whether Hawaiian chameleons were more conspicuous (as perceived by chameleons) in social contexts than Kenyan chameleons.
-We ran separate linear models with chromatic and luminance contrast of male display or courtship coloration against the local background as the response variable.
+We ran separate linear models with chromatic and luminance contrast of male display or courtship colouration against the local background as the response variable.
 We included population of origin and z-transformed snout-vent length (SVL) as fixed effects.
@@
 
-```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social stimuli.
+```{r figure1, cache=FALSE, fig.cap = "**Chameleon colour signal change in response to different social stimuli.
 Male chameleons experience intense sexual selection**.
@@
 They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
-**a**, A dominant male in display coloration.
+**a**, A dominant male in display colouration.
 **b**, A subordinate male that lost a contest and turned from bright yellow to brown.
-**c**, Two males fighting, both are in display coloration and relatively even matched.
-**d**, A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
+**c**, Two males fighting, both are in display colouration and relatively even matched.
+**d**, A courting male in full display colour while the female has turned to a contrasting colour, rejecting the male.
 See Supplementary information for additional photos, including in response to a snake"}
@@
 #plots
-p1 <- ggplot(summary_data, aes(x = Population, y = Mean_dS, group = Colour, color = Colour)) +
+p1 <- ggplot(summary_data, aes(x = Population, y = Mean_dS, group = Colour, colour = Colour)) +
     ylim(2.5, 6.5) + 

### 2022-01-22 ms_Nature.Rmd
color in title

@@
 ---
-title: "Title: Invasive chameleons released from predation display more conspicuous colors"
+title: "Title: Invasive chameleons released from predation display more conspicuous colours"
 bibliography: supp_refs.bib

### 2022-01-23 supp_Nature.Rmd
Removed the methods from supplement. Seeing if I can auto format the figures with some new yaml header stuff.

--- 2021-12-02 supp.Rmd
@@
 ---
-title: "Invasive chameleons released from predation display more conspicuous colors"
+title: "Invasive chameleons released from predation display more conspicuous colours"
 date: "`r Sys.Date()`"
@@
     reference_docx: template.docx
+    language:
+      label:
+        fig: !expr function(x) sprintf("**Figure %s.** ", x)
   bookdown::html_document2:
@@
 Supplementary text  
-Figures S1 -- S9  
-Tables S1 -- S13  
+Figures 1 -- 9  
+Tables 1 -- 13  
 Supplementary References
@@
 # Supplemental Materials and Methods
-## Study area and collection of chameleons
-
-We collected most chameleons (Fig.
-S\@ref(fig:Figcolsig2)) at night by spotlighting, and a small proportion of individuals during the day.
-When chameleons were beyond reach in the forest canopy, we used lightweight 6-m extensible graphite composite poles (South Bend Kwik Stix™ 20' Telescopic Fishing Pole), that chameleons could be encouraged to step on to, before lowering them to the ground.
-We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
-We measured snout-vent length (SVL), a standard measure of body size in lizards, from the tip of the snout to the posterior edge of the vent using a plastic ruler to the nearest 1 mm.
-Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006.
-In Hawaii, chameleons were collected from a single population in the forested Ko’olau Mountains north of Honolulu, between upper Makiki Valley and the summit of Tantalus Mountain, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
-400-600 m.
-The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Fig.
-S\@ref(fig:Figcolsig2)C, D).
-Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
-The area experiences high annual rainfall (mean: ca.
-1500 mm) @RN6 and the mean annual temperature is 22-24°C @RN7.
-In Kenya, field work was conducted near the foothills of Mt.
-Kenya, close to the town of Runyenjes.
-The area is partially cleared for agriculture, leaving small stands of forest.
-The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
-S\@ref(fig:Figcolsig2)A, B).
-Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm @RN8.
 
@@
 
-## Behavioral trials
-In order to quantify display coloration, we staged encounters between conspecifics (male-male and male-female) and model predators (bird and snake) and measured their subsequent display coloration (details below).
-All chameleons were used in both predator and social trials although in some cases we were unable to quantify all social signals for a particular individual.
-We conducted all conspecific-social trials first, in case anti-predator trials unduly influenced their social interactions.
-Our analyses focused on comparing color responses of chameleons to conspecifics and predators to ensure that the color state corresponded to specific stimuli.
-We did not, and could not, measure a ‘neutral’ color state because it is impossible to meaningfully gauge a neutral state in a color-changing organism such as a chameleon.
-However, for predator trials, we ensured that chameleons were specifically responding to the predators by conducting seven trials with a just a branch with no predator attached, as a control.
-Chameleons did not respond to the branch (see also @RN18ms).
-Furthermore, during trials with predators, chameleons focused on the predator itself, and exhibited classic anti-predator behavior (e.g., contrasting stress color, body inflation, open-mouth threat).
-
-Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
-We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand (Fig.
-S\@ref(fig:Figcolsig3)).
-This structure was about 1.5 m high and each arm was approximately 60 cm.
-The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Fig.
-S\@ref(fig:Figcolsig3)).
-All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
-The signaling environment was therefore the same for all animals.
-
 ```{r Figcolsig3, fig.align = 'center', fig.cap = "**The experimental setup for behavioral trials**.
@@
 
-We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
-This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Fig.
-S\@ref(fig:Figcolsig)A, B).
-In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
-This would be accompanied by head shakes and gapping, leading to horn-locking (Fig.
-S\@ref(fig:Figcolsig)B) if males were not obviously asymmetric.
-Once dominance was settled, the subordinate would turn brown (Fig.
-S\@ref(fig:Figcolsig)B and Fig.
-S\@ref(fig:Figcolsig3)) and flee.
-In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
-S\@ref(fig:Figcolsig)C).
-
-The day following social trials we presented each chameleon with a snake (Fig.
-S\@ref(fig:Figcolsig4)A) and bird (Fig.
-S\@ref(fig:Figcolsig4)B) model predator model (separately).
-Lizards were drawn at random and the order of presentation (snake or bird first) was random.
-Both trials were conducted on the same day but were separated by at least three hours.
-A trial began when a chameleon was removed from its cloth bag and placed on the triangular stand.
-In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Fig.
-S\@ref(fig:Figcolsig4)A).
-In the case of the bird trials, the bird was attached to an ca.
-2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Fig.
-S\@ref(fig:Figcolsig4)B).
-Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
-Once chameleons changed color, we quantified their spectral reflectance (details below).
-The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
-The bird was a mounted African Cuckoo-hawk (*Aviceda cuculoides*).
-
 ```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator**.
@@
 
-## Color measurement
-
-During all behavior trials, we measured the color of each of three body regions: top flank, mid flank, tail base and gular region (Fig.
-S\@ref(fig:Figcolsig5)).
-We measured each body region once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal.
-The four body regions were measured in random sequence to avoid any bias of order.
-When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
-We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal during measurement.
-The probe holder ensured that measurements were taken at a standard angle (45°) and distance (5 mm) for a 3 x 5 mm area and excluded all ambient light.
-All measurements were relative to a dark, and a certified 99% white reflectance standard (Labsphere, North Sutton, NH, U.S.A.).
-If chameleons started to change color in response to our handling, we returned them to the branch and allowed them to continue a social trial before measuring reflectance of remaining body regions.
-In the case of a predator trial, we would once again present them with the predator model until they returned to their anti-predator display.
-This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behavior following handling.
-
-We obtained spectral reflectance measurements of male dominant color state in male-male contests for 34 individuals from Hawaii and 35 from Kenya, courtship color state for 25 individuals from Hawaii and 32 from Kenya; color state in response to the bird for 56 individuals from Hawaii and 23 from Kenya and color state in response to the snake for 59 individuals from Hawaii and 38 from Kenya.
-
-We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signaling environment.
-We took 120 reflectance measures of the background at each location (Hawaii and Kenya), comprising 40 measurements of the top surface of leaves, 40 of the bottom surface of leaves and 40 stems and branches of suitable width for chameleons.
-To reduce the number of backgrounds, we averaged the reflectance of the 20 darker measurements and 20 lighter measurements for each of the top leaf and bottom leaf measurements and averaged brown stems and green stems.
-This resulted in 6 average background spectra for each location, representing the range of background colors.
-
 ```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Location of body regions we measured for spectral reflectance**. See details in text."}
@@
 ```
-
-## Visual modeling
-
-We visualized and quality-checked all spectra, adjusted very minor negative reflectance values caused by electrical noise by lifting the curves by the maximum negative value, then smoothed all spectra with LOESS smoothing of 0.16 using the plotsmooth function in `Pavo` vers.
-1.0 [@RN29].
-We also took the median reflectance for all leaves (top, bottom) and stems (2 measures/stem) for each population (Hawaii, Kenya).
-
-We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` @RN29.
-This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
-Color contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of color discrimination for stimuli viewed simultaneously under ideal viewing conditions @RN17.
-Under natural conditions, discrimination thresholds are likely to be >1 and vary depending on the species and conditions [@RN12;@RN13; @RN14].
-Additionally, perceived conspicuousness may not scale linearly with color distance (JNDs) @RN15.
-Nevertheless, the models provide a reasonable approximation of relative conspicuousness that can be compared between populations and trial types (social and predator).
-
-We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region and color state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
-We first calculated the quantum catch of each photoreceptor under standard daylight illumination (D65; Commission internationale de l’éclairage – CIE) using the ‘vismodel’ function.
-We applied the von Kries chromatic adaptation (vonkries = TRUE) and used the average background reflectance of the relevant background (Hawaii or Kenya) as the adapting background.
-The signal in each photoreceptor class was proportional to the natural logarithm of the quantum catch, in accordance with Fechner's law (“fi” in pavo).
-Next, we calculated the chromatic contrast between pairs of spectra using the function ‘coldist’.
-
-The visual models require information on the spectral sensitivities and photoreceptor noise (Weber fraction), which is a function of the relative density of each photoreceptor type within the retina, for the relevant receiver.
-Chameleons have four spectral classes of single cone enabling tetrachromatic vision @RN16.
-We used spectral sensitivities for the congeneric flap-necked chameleon (*Chameleo dilepis*) @RN16 and a standard Weber fraction of 0.1 @RN17 to account for the receptor noise of the LWS photoreceptor.
-We calculated noise for the other photoreceptor classes using the relative photoreceptor ratio of 1: 1.6: 3.3: 3.2 for the four single cones (UVS, SWS, MWS, and LWS, respectively).
-*Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Fig.
-S\@ref(fig:Figcolsig6SPec)).
 
@@
 
-For birds, we used an ultraviolet sensitive (UVS) visual system @RN18 characteristic of raptors and other birds of prey @RN19.
-A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement @RN20.
-We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds @RN21 (Fig.
-S\@ref(fig:Figcolsig6SPec)B).
-
-For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake @RN22.
-We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively @RN22 (Fig.
-S\@ref(fig:Figcolsig6SPec)C).
-
-In each case, visual pigment absorbance curves were multiplied by the transmission spectra of ocular media (lens and cornea) and oil droplets associated with the corresponding photoreceptor class and normalized to equal area under the curve to satisfy the assumption of equal stimulation by white light @RN18.
-
-Luminance contrast was calculated using the sensitivity function for the LWS photoreceptor (with transmission of the oil droplet associated with the double cone in chameleons and birds) for each visual system, assuming a Weber fraction of 0.05 because double cones used for luminance perception in birds and chameleons are by far the most abundant photoreceptor type in the retina [@RN16; @RN23].
-
-## Statistical Analysis
-
-Chromatic and luminance contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN24 in combination with the package `clubSandwich` @RN25.
-For each individual, measurements for each body region (top flank, mid-flank, tail base and gular region) were contrasted against each of six average background spectra (i.e., average top and bottom of lighter leaves, average top and bottom of darker leaves, average green and average brown stems).
-While we present sensitivity analyses for all body regions (see below - *Additional Supplemental Results*), we were mainly interested in overall signal differences displayed by chameleons.
-As such, for our main analyses we averaged estimates (marginalized) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see *Additional Supplemental Results*).
-In all cases, our results and conclusions were not impacted (see supplementary code).
-Deriving multiple contrasts for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other.
-To ensure that non-independence of data did not impact our inferences, we included individual ID as a random effect to account for repeated measures on each individual and used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction.
-
-First, we tested whether Hawaiian chameleons were more conspicuous (as perceived by chameleons) in social contexts than Kenyan chameleons.
-We ran separate linear models with chromatic and luminance contrast of male display or courtship coloration against the local background as the response variable.
-We included population of origin and z-transformed snout-vent length (SVL) as fixed effects.
-Next, we tested whether Kenyan chameleons were more cryptic (less conspicuous against the local background) in response to predators.
-For these models, chromatic or luminance contrast against the local background (as perceived by the corresponding predator type) was the response variable and predator type (i.e, snake or bird), population of origin (Hawaii or Kenya) and their interaction were fixed effects.
-Finally, to evaluate evidence for local adaptation, we tested whether the chromatic or luminance contrast of Hawaiian chameleons during social interactions (male-male and male-female) varied depending on whether contrast was against a native Kenyan background or their current introduced habitat background in Hawaii.
-For all models, interaction terms were first tested using Wald tests.
-If found to be non-significant (p > 0.1) we dropped the interaction and fitted a main effects model.
-We then tested the significance of the main effects using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.
-
 # Supplemental Results

### 2022-01-23 supp_Nature.Rmd
Must have been updates to cluBSandwich objects. Needed some fixing for Table S6

@@
 devtools::install_github("karthik/wesanderson")
-pacman::p_load(flextable, lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, readxl, pavo)
+pacman::p_load(tidyverse, flextable, lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich, readxl, pavo)
 source("./R/func.R")
@@
                          robust_PredatorCC) %>% 
-                    mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)),
-                           JND = c("Luminance Contrast", rep("", 9), 
-                                   "Chromatic Contrast", rep("", 9)))
+                    mutate(JND = c("Luminance Contrast", rep("", 9), 
+                                   "Chromatic Contrast", rep("", 9))) %>%  data.frame()
 
@@
 
-robust_sc_pred_table <- robust_sc_pred_table %>% select(JND, parameter, beta, SE, tstat, df, p_Satt)
+robust_sc_pred_table <- robust_sc_pred_table %>% dplyr::select(JND, parameter, beta, SE, tstat, df_Satt, p_Satt)
+row.names(robust_sc_pred_table) <- 1:nrow(robust_sc_pred_table)
+robust_sc_pred_table <- robust_sc_pred_table %>% mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)))
 

### 2022-01-23 supp_Nature.Rmd
Removed uneeded yaml header

@@
     reference_docx: template.docx
-    language:
-      label:
-        fig: !expr function(x) sprintf("**Figure %s.** ", x)
   bookdown::html_document2:

### 2022-01-23 supp_Nature.Rmd
Removed ms refs as they are not longer merged. Not sure why formatting for Table S6 is not working.

@@
 
-[@RN15ms;@RN7ms;@RN11ms;@RN6ms;@RN22ms;@RN2ms;@RN25ms;@RN8ms;@RN3ms;@RN1ms;@RN13ms;@RN18ms;@RN9ms;@RN12ms;@RN26ms;@RN24ms;@RN19ms;@RN5ms;@Measey2014ms;@RN20ms;@RN14ms;@RN4ms;@RN10ms]
-
 # Expanded Hypotheses and Predictions
@@
  Tab_S6
-
-```
+```
+
 # Additional Supplemental Results

### 2022-01-23 supp_Nature.Rmd
Add note on how to change fig and tab

@@
 Supplementary References  
-
+<!-- Note to change the formatting of table and figure legends add a _bookdwon.yml file and manipulate the fig: and tab: settings.
+Details here: https://community.rstudio.com/t/how-to-change-the-figure-table-caption-style-in-bookdown/110397-->
 # Expanded Hypotheses and Predictions
@@
 We are not aware of any evidence of bird predation of chameleons on Oahu.
-We consulted with several ornithologists in Hawaii regarding feeding behavior of the few predatory bird species that consume vertebrate prey, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons @RN1.
+We consulted with several ornithologists in Hawaii regarding feeding behavior of the few predatory bird species that consume vertebrate prey, and we examined the checklist of the birds of Oahu to determine if any species were potential predators of chameleons [@RN1].
 Among the larger, predatory birds, two are owls (*Asio flammeus sandwichensis*, *Tyto alba*), which are extremely rare on Oahu.
-Both species are highly unlikely to feed on chameleons because they feed out in the open rather than in forest canopy where chameleons occur, and typically feed on rodents, birds and insects @RN2.
-Smaller to medium raptors include the Merlin (*Falco columbarius*) and the peregrine falcon (*Falco peregrinus*) which are both considered rare non-breeding vagrants in the Hawaiian Islands, and feed on small-medium birds in more open habitats @RN2.
+Both species are highly unlikely to feed on chameleons because they feed out in the open rather than in forest canopy where chameleons occur, and typically feed on rodents, birds and insects [@RN2].
+Smaller to medium raptors include the Merlin (*Falco columbarius*) and the peregrine falcon (*Falco peregrinus*) which are both considered rare non-breeding vagrants in the Hawaiian Islands, and feed on small-medium birds in more open habitats [@RN2].
 There are also eight species of egret and heron, which typically have generalist diets.
-Of these, six are considered rare or accidental @RN1, are associated with aquatic habitats, and are better known for eating fishes @RN3.
+Of these, six are considered rare or accidental [@RN1], are associated with aquatic habitats, and are better known for eating fishes [@RN3].
 The cattle egret (*Bubulcus ibis*) is a generalist predator that is mostly insectivorous but does eat lizards opportunistically [@RN4;@RN5].
-However, it forages on the ground in more open habitats, often following domestic animals and agricultural or landscaping machinery @RN5 and therefore, is not found in forested habitats and is highly unlikely to encounter chameleons.
+However, it forages on the ground in more open habitats, often following domestic animals and agricultural or landscaping machinery [@RN5] and therefore, is not found in forested habitats and is highly unlikely to encounter chameleons.
 Among other potential predators, the Hawaiian Hawk or ʻio (*Buteo solitarius*) is restricted to the Big Island, while the Hen Harrier (*Circus cyaneus*) is considered an accidental or rare visitor to the Hawaiian islands.
-Finally, among corvids, the Hawaiian Crow (*Corvus hawaiiensis*) is extinct in the wild while the Common Raven (*Corvus corax*) is extremely rare @RN1.
+Finally, among corvids, the Hawaiian Crow (*Corvus hawaiiensis*) is extinct in the wild while the Common Raven (*Corvus corax*) is extremely rare [@RN1].
 In summary, we could not identify any bird species that could be regarded as a major predator of chameleons.

### 2022-01-23 supp_Nature.Rmd
Changes all in text ref to Supplementary Fig. and Supplementary Table

@@
 Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, *n* = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]` mm, *n* = `r svl[2,"n"]`; t = `r SVL_pop$statistic`, *P* <0.001).
-SVL did not significantly predict luminance contrast against the background (Table S\@ref(tab:table1)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
-Snout-vent length was a weak predictor of chromatic contrast (Table S\@ref(tab:table1)): larger chameleons had slightly higher chromatic contrast against the background.
+SVL did not significantly predict luminance contrast against the background (Supplementary Table\@ref(tab:table1)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
+Snout-vent length was a weak predictor of chromatic contrast (Supplementary Table\@ref(tab:table1)): larger chameleons had slightly higher chromatic contrast against the background.
 However, the populations did not differ in chromatic contrast during male displays (courtship and contests).
@@
 
-Chromatic contrast of color signals against the background varied by predator type, population and body region (Fig.
-S\@ref(fig:FigS3)A and Table S\@ref(tab:tableS3)).
-Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Fig.
-S\@ref(fig:FigS3)A).
-There was no significant interaction between predator type ('bird' vs 'snake') and population was not significant for luminance contrast against the background, but there was a marginally non-significant population and body region effect (Fig.
-S\@ref(fig:FigS3)B and Table S\@ref(tab:tableS3)).
-Effects did not vary by predator type (Table S\@ref(tab:tableS3)).
-Nonetheless, effect sizes across body regions were generally in the same direction (Fig.
-S\@ref(fig:FigS3)B and Table S\@ref(tab:tableS4)).
-Overall, there were significant differences between Kenya and Hawaii in luminance contrast of the gular and tailbase regions, but not for mid- and top-flank (Table S\@ref(tab:tableS5)).
-In contrast, no significant differences between Kenya and Hawaii existed across any body region for snakes, whereas significant differences existed between gular, midflank and tailbase for birds (Table S\@ref(tab:tableS6)).
+Chromatic contrast of color signals against the background varied by predator type, population and body region (Supplementary Fig.
+\@ref(fig:FigS3)A and Supplementary Table\@ref(tab:tableS3)).
+Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Supplementary Fig.
+\@ref(fig:FigS3)A).
+There was no significant interaction between predator type ('bird' vs 'snake') and population was not significant for luminance contrast against the background, but there was a marginally non-significant population and body region effect (Supplementary Fig.
+\@ref(fig:FigS3)B and Supplementary Table\@ref(tab:tableS3)).
+Effects did not vary by predator type (Supplementary Table\@ref(tab:tableS3)).
+Nonetheless, effect sizes across body regions were generally in the same direction (Supplementary Fig.
+\@ref(fig:FigS3)B and Supplementary Table\@ref(tab:tableS4)).
+Overall, there were significant differences between Kenya and Hawaii in luminance contrast of the gular and tailbase regions, but not for mid- and top-flank (Supplementary Table\@ref(tab:tableS5)).
+In contrast, no significant differences between Kenya and Hawaii existed across any body region for snakes, whereas significant differences existed between gular, midflank and tailbase for birds (Supplementary Table\@ref(tab:tableS6)).
 
@@
 However, for luminance contrast, these models had poor convergence.
-Based on the raw data (Fig.
-S\@ref(fig:FigS4)B and S\@ref(fig:FigS4)D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
+Based on the raw data (Supplementary Fig.
+\@ref(fig:FigS4)B and S\@ref(fig:FigS4)D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
 As such, we simplified models and only present the main effects of body region and background environment for luminance contrast.
-However, inspection of the raw data (Fig.
-S\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C) suggested that body regions may differ in how chromatically contrasting they were against the background environment of Hawaii and Kenya.
+However, inspection of the raw data (Supplementary Fig.
+\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C) suggested that body regions may differ in how chromatically contrasting they were against the background environment of Hawaii and Kenya.
 In addition, our main results suggested that predators may deferentially drive chromatic signal conspicuousness, and given different body regions maybe more exposed to bird compared to snake, predators, we fitted an interaction and main effects model for chromatic contrast.
 
-The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, *P* = `r p_value(intearction_CC_Court$p_val)`, Fig.
-S\@ref(fig:FigS4)A, S\@ref(fig:FigS4)C and Table S\@ref(tab:tableS7)).
-However, in all cases there was a tendency for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Fig.
-S\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C).
-Tailbase was the only body region that showed a non-significant change (Table S\@ref(tab:tableS8)).
-
-The magnitude of luminance contrast showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court$p_val)`, Fig.
-S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, *P* = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Fig.
-S\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Table S\@ref(tab:tableS7)).
+The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, *P* = `r p_value(intearction_CC_Court$p_val)`, Supplementary Fig.
+\@ref(fig:FigS4)A, S\@ref(fig:FigS4)C and Supplementary Table\@ref(tab:tableS7)).
+However, in all cases there was a tendency for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Supplementary Fig.
+\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C).
+Tailbase was the only body region that showed a non-significant change (Supplementary Table\@ref(tab:tableS8)).
+
+The magnitude of luminance contrast showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court$p_val)`, Supplementary Fig.
+\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Supplementary Table\@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, *P* = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Supplementary Fig.
+\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Supplementary Table\@ref(tab:tableS7)).
 
@@
 
-Body regions varied in their overall effects but mainly showed a decrease in absolute JND difference between display and antipredator color states in Kenya compared to Hawaii (Fig.
-S\@ref(fig:figDisplayPred)).
-While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, *P* = `r p_value(robust_Diff_dS_int[,6])`; luminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, *P* = `r p_value(robust_Diff_dl_int[,6])`) (Fig.
-S\@ref(fig:figDisplayPred) and Table S\@ref(tab:tableS9)).
-However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dS_BR_me[,6])`; luminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dL_BR_me[,6])`) (Fig.
-S\@ref(fig:figDisplayPred); Table S\@ref(tab:tableS9) and Table S\@ref(tab:tableS10)).
+Body regions varied in their overall effects but mainly showed a decrease in absolute JND difference between display and antipredator color states in Kenya compared to Hawaii (Supplementary Fig.
+\@ref(fig:figDisplayPred)).
+While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, *P* = `r p_value(robust_Diff_dS_int[,6])`; luminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, *P* = `r p_value(robust_Diff_dl_int[,6])`) (Supplementary Fig.
+\@ref(fig:figDisplayPred) and Supplementary Table\@ref(tab:tableS9)).
+However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dS_BR_me[,6])`; luminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dL_BR_me[,6])`) (Supplementary Fig.
+\@ref(fig:figDisplayPred); Supplementary Table\@ref(tab:tableS9) and Supplementary Table\@ref(tab:tableS10)).
 
@@
 ```
-# Extended acknowledgments
-We thank Idle Farah and Patrick Malonza of the National Museums of Kenya for facilitating the research, Joseph Gathua for field support and Jan Stipala for use of his field vehicle.
-We thank the Ditsong National Museum of Natural History (formerly Transvaal Museum) for the stuffed African Cuckoo Hawk.
-Ben Phillips, Rick Shine and Tobias Uller provided feedback on earlier drafts of this manuscript.
-We thank Fonti Kar for help curating data and Drew Allen for statistical advice.
 

### 2022-01-23 supp_Nature.Rmd
Update to style file

@@
 bibliography: supp_refs.bib
-csl: science.csl
+csl: nature.csl
 output: 

### 2022-01-23 R/func.R
chnaged font for tables

@@ -102,8 +102,8 @@ table_style  <- function(table){
             value = as_paragraph("Kenya", as_sub(" tailbase"), " - ", "Hawaii", as_sub(" tailbase"))) %>% 
     flextable::compose(part = "body", i = 8, j = 2, 
             value = as_paragraph("Kenya", as_sub(" topflank"), " - ", "Hawaii", as_sub(" topflank"))) %>% 
-    font(fontname="Times", part = "body") %>% 
-    font(fontname="Times", part = "header") 
+    font(fontname="Calibri", part = "body") %>% 
+    font(fontname="Calibri", part = "header") 
   return(table)
 }
 
@@ -118,8 +118,8 @@ table_style2 <- function(table){
     flextable::compose(part = "header", j = 5, value = as_paragraph(as_b("df"), as_sub(as_b(" num")))) %>%
     flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("df"), as_sub(as_b(" denom")))) %>% 
     flextable::compose(part = "header", j = 7, value = as_paragraph(as_b("p"))) %>% 
-    font(fontname="Times", part = "body") %>% 
-    font(fontname="Times", part = "header")
+    font(fontname="Calibri", part = "body") %>% 
+    font(fontname="Calibri", part = "header")
   return(table)
 }
 
@@ -135,8 +135,8 @@ table_style3 <- function(table){
     flextable::compose(part = "header", j = 6, value = as_paragraph(as_b("p"))) %>% 
     flextable::compose(part = "body", j = 2, i = c(4,9), value = as_paragraph("Predator", "* BodyRegion")) %>% 
     flextable::compose(part = "body", j = 2, i = c(4,9)+1, value = as_paragraph("Predator", "* Population")) %>% 
-    font(fontname="Times", part = "body") %>% 
-    font(fontname="Times", part = "header")
+    font(fontname="Calibri", part = "body") %>% 
+    font(fontname="Calibri", part = "header")
   return(table)
 }
 
@@ -162,8 +162,8 @@ table_style4 <- function(table){
     flextable::compose(part = "body", j = 3, i = c(1, 10, 19, 28)+7, value = as_paragraph("Population", as_sub(" Kenya"), "* BodyRegion", as_sub(" tailbase"))) %>%
[…]

### 2022-01-23 supp_Nature.Rmd
export from dplyr select to avoid conflict

@@
                filter(Population == "Hawaii") %>% 
-               left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               left_join(dplyr::select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
                mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
@@
 Details here: https://community.rstudio.com/t/how-to-change-the-figure-table-caption-style-in-bookdown/110397-->
+
 # Expanded Hypotheses and Predictions
@@
 
-cham <- sens_cham_c_dilepis %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
-bird <- sens_birdUV %>% pivot_longer(cols = c(2:6)) %>% select(wavelength, name, value) %>%  data.frame()
-snake <- sens_snake %>% pivot_longer(cols = c(2:4)) %>% select(wavelength, name, value) %>%  data.frame()
+cham <- sens_cham_c_dilepis %>% pivot_longer(cols = c(2:6)) %>% dplyr::select(wavelength, name, value) %>%  data.frame()
+bird <- sens_birdUV %>% pivot_longer(cols = c(2:6)) %>% dplyr::select(wavelength, name, value) %>%  data.frame()
+snake <- sens_snake %>% pivot_longer(cols = c(2:4)) %>% dplyr::select(wavelength, name, value) %>%  data.frame()
 
@@
 Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, *n* = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]` mm, *n* = `r svl[2,"n"]`; t = `r SVL_pop$statistic`, *P* <0.001).
-SVL did not significantly predict luminance contrast against the background (Supplementary Table\@ref(tab:table1)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
-Snout-vent length was a weak predictor of chromatic contrast (Supplementary Table\@ref(tab:table1)): larger chameleons had slightly higher chromatic contrast against the background.
+SVL did not significantly predict luminance contrast against the background (Supplementary Table \@ref(tab:ms_tableS2)); thus the higher luminance contrast of Hawaiian chameleons during male-male contests and courtship cannot be explained by body size differences between the population.
+Snout-vent length was a weak predictor of chromatic contrast (Supplementary Table \@ref(tab:ms_tableS2)): larger chameleons had slightly higher chromatic contrast against the background.
 However, the populations did not differ in chromatic contrast during male displays (courtship and contests).
@@
 
-```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing mean differences between Kenyan and Hawaiian lizards in luminance and chromatic contrast against the local background in two social contexts (male-male competition and courtship).
+```{r ms_tableS2, tab.cap = "Results of models comparing mean differences between Kenyan and Hawaiian lizards in luminance and chromatic contrast against the local background in two social contexts (male-male competition and courtship).
 Model estimates, standard errors (SE), Satterthwaite's degrees of freedom (df) and 95% confidence internals (95% CI).
@@
 
-```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic mean differences between Kenyan and Hawaiian lizards in response to snakes and birds.
+```{r ms_tableS3, tab.cap = "Results of models comparing luminance and chromatic mean differences between Kenyan and Hawaiian lizards in response to snakes and birds.
 Model estimates, standard errors (SE), Satterthwaite's degrees of freedom (df) and 95% confidence internals (95% CI).
@@
 
-```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean luminance and chromatic contrast in Hawaii and Kenya for both predators (snakes and birds).
+```{r ms_tableS4, tab.cap = "Planned contrasts between mean luminance and chromatic contrast in Hawaii and Kenya for both predators (snakes and birds).
 Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper).
@@
 
-```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
+```{r ms_tableS5, tab.cap = "Results of models comparing luminance and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
 Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for models were fitted using robust variance estimators to correct standard errors for non-independence, and we used Satterthwaite's degrees of freedom.
@@
 
-```{r tableS6_new, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Results of interaction models comparing luminance and chromatic contrast of Hawaiian and Kenyan lizards against a Hawaiian and Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
+```{r ms_tableS6, tab.cap = "Results of interaction models comparing luminance and chromatic contrast of Hawaiian and Kenyan lizards against a Hawaiian and Kenyan background (background = Kenya) in both social contexts (male-male competition and courtship).
 Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for models were fitted using robust variance estimators to correct standard errors for non-independence, and we used Satterthwaite's degrees of freedom.
@@
 Chromatic contrast of color signals against the background varied by predator type, population and body region (Supplementary Fig.
-\@ref(fig:FigS3)A and Supplementary Table\@ref(tab:tableS3)).
+\@ref(fig:FigS3)A and Supplementary Table \@ref(tab:tableS3)).
 Whereas all body regions in Kenyan chameleons showed a reduced contrast against the background for birds there were small differences among populations in how snakes perceived color signals (Supplementary Fig.
@@
 There was no significant interaction between predator type ('bird' vs 'snake') and population was not significant for luminance contrast against the background, but there was a marginally non-significant population and body region effect (Supplementary Fig.
-\@ref(fig:FigS3)B and Supplementary Table\@ref(tab:tableS3)).
-Effects did not vary by predator type (Supplementary Table\@ref(tab:tableS3)).
+\@ref(fig:FigS3)B and Supplementary Table \@ref(tab:tableS3)).
+Effects did not vary by predator type (Supplementary Table \@ref(tab:tableS3)).
 Nonetheless, effect sizes across body regions were generally in the same direction (Supplementary Fig.
-\@ref(fig:FigS3)B and Supplementary Table\@ref(tab:tableS4)).
-Overall, there were significant differences between Kenya and Hawaii in luminance contrast of the gular and tailbase regions, but not for mid- and top-flank (Supplementary Table\@ref(tab:tableS5)).
-In contrast, no significant differences between Kenya and Hawaii existed across any body region for snakes, whereas significant differences existed between gular, midflank and tailbase for birds (Supplementary Table\@ref(tab:tableS6)).
+\@ref(fig:FigS3)B and Supplementary Table \@ref(tab:tableS4)).
+Overall, there were significant differences between Kenya and Hawaii in luminance contrast of the gular and tailbase regions, but not for mid- and top-flank (Supplementary Table \@ref(tab:tableS5)).
+In contrast, no significant differences between Kenya and Hawaii existed across any body region for snakes, whereas significant differences existed between gular, midflank and tailbase for birds (Supplementary Table \@ref(tab:tableS6)).
 
@@
   final_r5 <- final_r5 %>% 
-              select(-test, -delta, -df_num) %>%
+              dplyr::select(-test, -delta, -df_num) %>%
               mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
                      p_val = sapply(p_val, function(x) p_value(x))) %>%
-              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
+              dplyr::select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
 
@@
  final_r6 <- final_r6 %>% 
-              select(-test, -delta, -df_num) %>%
+              dplyr::select(-test, -delta, -df_num) %>%
               mutate(hypothesis = gsub("PopBRInt", "", hypothesis),
                      p_val = sapply(p_val, function(x) p_value(x))) %>%
-              select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
+              dplyr::select(Species, hypothesis, Fstat, df_denom, p_val, p_val_bonfer)
 
@@
 The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, *P* = `r p_value(intearction_CC_Court$p_val)`, Supplementary Fig.
-\@ref(fig:FigS4)A, S\@ref(fig:FigS4)C and Supplementary Table\@ref(tab:tableS7)).
+\@ref(fig:FigS4)A, S\@ref(fig:FigS4)C and Supplementary Table \@ref(tab:tableS7)).
 However, in all cases there was a tendency for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Supplementary Fig.
 \@ref(fig:FigS4)A and S\@ref(fig:FigS4)C).
-Tailbase was the only body region that showed a non-significant change (Supplementary Table\@ref(tab:tableS8)).
+Tailbase was the only body region that showed a non-significant change (Supplementary Table \@ref(tab:tableS8)).
 
 The magnitude of luminance contrast showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court$p_val)`, Supplementary Fig.
-\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Supplementary Table\@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, *P* = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Supplementary Fig.
-\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Supplementary Table\@ref(tab:tableS7)).
+\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Supplementary Table \@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, *P* = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Supplementary Fig.
+\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Supplementary Table \@ref(tab:tableS7)).
 
@@
                         p_val = sapply(p_val, function(x) p_value(x))) %>% 
-                        select(Context,c(1:8), -test, -delta, -df_num)
+                        dplyr::select(Context,c(1:8), -test, -delta, -df_num)
 
@@
 While there was some evidence for an interaction between population and body region for luminance change (albeit weak) this was not the case for chromatic change (population by body region interaction: chromatic: $F_{`r robust_Diff_dS_int$df_num`,`r robust_Diff_dS_int$df_denom`}$ = `r robust_Diff_dS_int$Fstat`, *P* = `r p_value(robust_Diff_dS_int[,6])`; luminance: $F_{`r robust_Diff_dl_int$df_num`,`r robust_Diff_dl_int$df_denom`}$ = `r robust_Diff_dl_int$Fstat`, *P* = `r p_value(robust_Diff_dl_int[,6])`) (Supplementary Fig.
-\@ref(fig:figDisplayPred) and Supplementary Table\@ref(tab:tableS9)).
+\@ref(fig:figDisplayPred) and Supplementary Table \@ref(tab:tableS9)).
 However, the degree to which body regions varied within poplations was more apparent for luminance than for chromatic change (main effect of body region: chromatic: $F_{`r robust_Diff_dS_BR_me$df_num`,`r robust_Diff_dS_BR_me$df_denom`}$ = `r robust_Diff_dS_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dS_BR_me[,6])`; luminance: $F_{`r robust_Diff_dL_BR_me$df_num`,`r robust_Diff_dL_BR_me$df_denom`}$ = `r robust_Diff_dL_BR_me$Fstat`, *P* = `r p_value(robust_Diff_dL_BR_me[,6])`) (Supplementary Fig.
-\@ref(fig:figDisplayPred); Supplementary Table\@ref(tab:tableS9) and Supplementary Table\@ref(tab:tableS10)).
+\@ref(fig:figDisplayPred); Supplementary Table \@ref(tab:tableS9) and Supplementary Table \@ref(tab:tableS10)).
 
@@
 tabS10 <- tabS10 %>% mutate(VisRealm = c("Luminance Change", rep("",3), "Chromatic Change", rep("",3))) %>% 
-                        select(VisRealm, hypothesis, Fstat, df_denom, p_val, p_val_bonfer, -test, -delta, -df_num)
+                        dplyr::select(VisRealm, hypothesis, Fstat, df_denom, p_val, p_val_bonfer, -test, -delta, -df_num)
 

### 2022-01-23 supp_Nature.Rmd
Fixed a few occurences of "S" for supplement. Tables need to be edited manually. I cannot fix the bookdown issue, and it's really odd!

@@
     toc: no
-    toc_depth: 6
     number_sections: false
@@
     toc_float: yes
-    bibliography: refs.bib
 ---
@@
 ```
+
 Of the individuals used in social interactions (male-male display or male courtship), chameleons from Kenya were significantly larger than those from Hawaii (mean ± standard deviation Hawaii: `r svl[1,"mean"]`$\pm$ `r svl[1,"sd"]` mm, *n* = `r svl[1,"n"]`; Kenya `r svl[2,"mean"]`$\pm$ `r svl[2,"sd"]` mm, *n* = `r svl[2,"n"]`; t = `r SVL_pop$statistic`, *P* <0.001).
@@
  table1 <- flextable(final_T1) %>%  table_style6() %>% fit_to_width(max_width = 7)
+
  table1
-
 ```
@@
  table3 <- flextable(contrast) %>% table_style8() %>% fit_to_width(max_width = 7)
+
  table3
@@
  table4 <- flextable(tb4) %>% table_style9() %>% fit_to_width(max_width = 7)
+
  table4
@@
  Tab_S6 <- flextable(final_S6) %>%  table_styleS6() %>% fit_to_width(max_width = 7)
+
  Tab_S6
@@
 Based on the raw data (Supplementary Fig.
-\@ref(fig:FigS4)B and S\@ref(fig:FigS4)D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
+\@ref(fig:FigS4)B and \@ref(fig:FigS4)D) there was also weak evidence that body regions differed in how they changed when viewed by conspecifics against the Hawaiian and Kenyan backgrounds.
 As such, we simplified models and only present the main effects of body region and background environment for luminance contrast.
 However, inspection of the raw data (Supplementary Fig.
-\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C) suggested that body regions may differ in how chromatically contrasting they were against the background environment of Hawaii and Kenya.
+\@ref(fig:FigS4)A and \@ref(fig:FigS4)C) suggested that body regions may differ in how chromatically contrasting they were against the background environment of Hawaii and Kenya.
 In addition, our main results suggested that predators may deferentially drive chromatic signal conspicuousness, and given different body regions maybe more exposed to bird compared to snake, predators, we fitted an interaction and main effects model for chromatic contrast.
@@
 The magnitude of chromatic contrast differences between Hawaiian chameleons against their own environment and a Kenyan enviroment did depend on the body region (Male-male display: $F_{`r intearction_CC_Disp$df_num`, `r intearction_CC_Disp$df_denom`}$ = `r intearction_CC_Disp$Fstat`, p = `r p_value(intearction_CC_Disp$p_val)`; Courtship Display: $F_{`r intearction_CC_Court$df_num`, `r intearction_CC_Court$df_denom`}$ = `r intearction_CC_Court$Fstat`, *P* = `r p_value(intearction_CC_Court$p_val)`, Supplementary Fig.
-\@ref(fig:FigS4)A, S\@ref(fig:FigS4)C and Supplementary Table \@ref(tab:tableS7)).
+\@ref(fig:FigS4)A, \@ref(fig:FigS4)C and Supplementary Table \@ref(tab:tableS7)).
 However, in all cases there was a tendency for all body regions of Hawaiian populations to decrease in average contrast when viewed by a conspecific against a Hawaiian versus Kenyan background (Supplementary Fig.
-\@ref(fig:FigS4)A and S\@ref(fig:FigS4)C).
+\@ref(fig:FigS4)A and \@ref(fig:FigS4)C).
 Tailbase was the only body region that showed a non-significant change (Supplementary Table \@ref(tab:tableS8)).
@@
 The magnitude of luminance contrast showed a significant overall decrease when viewed by conspecifics against a Kenyan background (Overall Background Effect: Male-male display, $F_{`r background_AC_Disp$df_num`, `r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp$p_val)`, Courtship Display, $F_{`r background_AC_Court$df_num`, `r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court$p_val)`, Supplementary Fig.
-\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Supplementary Table \@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, *P* = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Supplementary Fig.
-\@ref(fig:FigS4)B, S\@ref(fig:FigS4)D and Supplementary Table \@ref(tab:tableS7)).
+\@ref(fig:FigS4)B, \@ref(fig:FigS4)D and Supplementary Table \@ref(tab:tableS7)), and there were significant differences among body regions (Overall Body Region Effect: Male-male display, $F_{`r BodyRegion_AC_Disp$df_num`, `r BodyRegion_AC_Disp$df_denom`}$ = `r BodyRegion_AC_Disp$Fstat`, *P* = `r p_value(BodyRegion_AC_Disp$p_val)`, Courtship Display, $F_{`r BodyRegion_AC_Court$df_num`, `r BodyRegion_AC_Court$df_denom`}$ = `r BodyRegion_AC_Court$Fstat`, *P* = `r p_value(BodyRegion_AC_Court$p_val)`, Supplementary Fig.
+\@ref(fig:FigS4)B, \@ref(fig:FigS4)D and Supplementary Table \@ref(tab:tableS7)).
 

### 2022-01-23 ms_Nature.Rmd
colour again
clubSandiwch objects changed. Needed to update objecs to grab the correct results. lowercase the values

@@
               comparison = c("Snake (Hawaii-Kenya)", "Birds (Hawaii-Kenya)"), 
-                Estimate = c(robust_PredatorAC_pred_pop2[4,1], robust_PredatorAC_pred_pop1[3, 1]),
-                lwr = c(robust_PredatorAC_pred_pop2[4,4], robust_PredatorAC_pred_pop1[3, 4]),
-                upr = c(robust_PredatorAC_pred_pop2[4,5], robust_PredatorAC_pred_pop1[3, 5]),
+                Estimate = c(robust_PredatorAC_pred_pop2[4,2], robust_PredatorAC_pred_pop1[3, 2]),
+                lwr = c(robust_PredatorAC_pred_pop2[4,5], robust_PredatorAC_pred_pop1[3, 5]),
+                upr = c(robust_PredatorAC_pred_pop2[4,6], robust_PredatorAC_pred_pop1[3, 6]),
               Fstat = c(result_pred[5,3], result_pred[2,3]),
@@
               comparison = c("Snake (Hawaii-Kenya)", "Birds (Hawaii-Kenya)"), 
-                Estimate = c(robust_PredatorCC_pred_pop2[4,1], robust_PredatorCC_pred_pop1[3, 1]),
-                lwr = c(robust_PredatorCC_pred_pop2[4,4], robust_PredatorCC_pred_pop1[3, 4]),
-                upr = c(robust_PredatorCC_pred_pop2[4,5], robust_PredatorCC_pred_pop1[3, 5]),
+                Estimate = c(robust_PredatorCC_pred_pop2[4,2], robust_PredatorCC_pred_pop1[3, 2]),
+                lwr = c(robust_PredatorCC_pred_pop2[4,5], robust_PredatorCC_pred_pop1[3, 5]),
+                upr = c(robust_PredatorCC_pred_pop2[4,6], robust_PredatorCC_pred_pop1[3, 6]),
               Fstat = c(result_pred[5,3], result_pred[2,3]),
@@
 \@ref(fig:figure1) and Supplementary Fig.
-4A).
+4a).
 Their different colour states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
@@
 \@ref(fig:figure3)A and Supplementary Table 2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
-This result was consistent for all body regions (Supplementary Information).
+and was independent of the background environment (i.e.,  contrasted with Hawaiian or Kenyan environment) (Supplementary Information and Supplementary Table 6).
 However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of colour signals (Fig.
-\@ref(fig:figure3)B and Supplementary Table 2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
+\@ref(fig:figure3)b and Supplementary Table 2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
 
@@
 Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.
-\@ref(fig:figure3)C; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, *P* = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, Supplementary Table 3).
+\@ref(fig:figure3)c; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, *P* = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, Supplementary Table 3).
 Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, Supplementary Table 3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, *P* = `r p_value(back_pred_CC_Pred$p_val)`).
-Hawaiian chameleons were significantly more conspicuous to birds than were Kenyan conspecifics (Fig.
-\@ref(fig:figure3)D and Supplementary Table 4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`, JND values > 5).
+Hawaiian chameleons had higher chromatic contrast to birds than were Kenyan conspecifics (Fig.
+\@ref(fig:figure3)d and Supplementary Table 4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`).
 By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.
-\@ref(fig:figure3)D and Supplementary Table 4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1; JND values ~ 2).
+\@ref(fig:figure3)d and Supplementary Table 4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1).
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18ms].
@@
 
-## colour measurement
+## Colour measurement
 

### 2022-01-23 ms_Nature.Rmd
Added "are"

@@
 50 -- 65 generations).
-On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators absent or rare, Supplementary Information), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014ms.
+On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators are absent or rare, Supplementary Information), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014ms.
 This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.

### 2022-01-23 ms_Nature.Rmd
Fixed in text code chunks to deal with changes in clubSandwich object

@@
 Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.
-\@ref(fig:figure4) and Supplementary Tables 5 and 6; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,1]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,4]` to `r robust_localAdapt_DispBgrdAC[3,5]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,1]`, 95% CI = `r robust_DispBgrdCC[3,4]` to `r robust_DispBgrdCC[3,5]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,1]`, 95% CI = `r robust_CourtshipAC[3,4]` to `r robust_CourtshipAC[3,5]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,1]`, 95% CI = `r robust_CourtshipCC[3,4]` to `r robust_CourtshipCC[3,5]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
+\@ref(fig:figure4) and Supplementary Tables 5 and 6; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,2]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,5]` to `r robust_localAdapt_DispBgrdAC[3,6]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,2]`, 95% CI = `r robust_DispBgrdCC[3,5]` to `r robust_DispBgrdCC[3,6]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,2]`, 95% CI = `r robust_CourtshipAC[3,5]` to `r robust_CourtshipAC[3,6]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,2]`, 95% CI = `r robust_CourtshipCC[3,5]` to `r robust_CourtshipCC[3,6]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
 These results are consistent with local adaptation, in which phenotypic plasticity, which can also evolve rapidly [@RN26ms], may also play a role.

### 2022-02-02 ms_Nature.Rmd
Revised main MS

@@
 
-**Authors** Martin J. Whiting^1,2^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel W.A.Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
+**Authors** Martin J. Whiting^1,2^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel  W.A. Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
 
@@
 Here, we show that Jackson’s chameleons (*Trioceros j.
-xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu), where there are few lizard predators, experienced significant character release of colour signals.
-Hawaiian chameleons displayed more conspicuous colour signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator colour states.
+xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu), where there are no co-evolved, native lizard predators, experienced significant character release of colour signals.
+Hawaiian chameleons displayed more conspicuous social colour signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator colour states.
 Hawaiian chameleon display colours were also more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
-These results demonstrate for the first time that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
+These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
 
 # Main
-When species encounter novel environments, such as during human-mediated introductions, phenotypic change can occur more rapidly than Darwin ever thought possible [@RN1ms; @RN2ms; @RN3ms; @RN4ms].
-In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4ms; @RN5ms].
+When species encounter novel environments, such as during human-mediated introductions, phenotypic change through a combination of plasticity and evolution, can occur remarkably rapidly [@RN1ms; @RN2ms; @RN3ms; @RN4ms].
+In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the phenotypic landscape [@RN4ms; @RN5ms].
 Such character release should be most evident for social and sexual signals; released from constraints on conspicuousness imposed by predation risk, signals could elaborate in many directions [e.g., @RN4ms; @RN6ms; @RN7ms; @RN8ms].
@@
 50 -- 65 generations).
-On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators are absent or rare, Supplementary Information), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators @Measey2014ms.
+On Oahu, there are few potential predators of chameleons (no snakes or lizard-eating raptors, other potential bird predators are absent or rare, Supplementary Information), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators [@Measey2014ms].
 This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.
@@
 
-To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Supplementary Fig.
+To test these predictions, we presented wild-caught adult male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Supplementary Fig.
 4), or a control (stick).
@@
 \@ref(fig:figure3)A and Supplementary Table 2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
-and was independent of the background environment (i.e.,  contrasted with Hawaiian or Kenyan environment) (Supplementary Information and Supplementary Table 6).
+This result was consistent for all body regions and was driven by the increased luminance of Hawaiian chameleons, rather than differences in the background environment (i.e., Hawaiian chameleons were more conspicuous against either Hawaiian or Kenyan environment) (Supplementary Information and Supplementary Table 6).
 However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of colour signals (Fig.
@@
 \@ref(fig:figure1)).
-Specifically, colour change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15ms or changes in the spacing of guanine crystals within iridophores @RN14ms, both of which strongly affect luminance.
+Specifically, colour change is caused by dispersion or concentration of melanin within melanophore pigment cells [@RN15ms] or changes in the spacing of guanine crystals within iridophores [@RN14ms], both of which strongly affect luminance.
 Accordingly, character release was observed in the overall luminance of signals rather than their hue.
@@
 While there was evidence for differences between body regions (Supplementary Information, Supplementary Fig.
-9, Supplementary Tables 12 and 13), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (Supplementary Fig.
+9, Supplementary Tables 13 and 14), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (Supplementary Fig.
 9).
@@
 Chameleons are almost always encountered in a camouflaged colour state because they are solitary and only use display colours during relatively brief social encounters which precludes the possibility that hunters selected the most brightly coloured individuals.
-Nevertheless, even if pet traders were selecting for larger animals, our analyses controlled for any relationship body size might have on colour.
+Even if pet traders were selecting for larger animals, our analyses controlled for any relationship body size might have on colour.
 Finally, the traits we quantified, and their observed differences, were all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the environment.
@@
 The degree to which colour change capacity is plastic in chameleons and other organisms with dynamic colour change is unknown.
-We cannot exclude the role of phenotypic plasticity.
-Plasticity can evolve and can itself be an adaptive response.
-The population differences we observed are likely an interaction between genetic and environmental factors which can only be disentangled using common garden studies and for the purposes of this study, is a moot point.
-What is significant about our work is a surprisingly strong effect of recent release from predation on a conspicuous, dynamic visual signal.
+While we cannot exclude the role of phenotypic plasticity, plasticity can itself evolve and be an adaptive response.
+The population differences we observed are likely an interaction between genetic and environmental factors which are difficult to disentangle, even with common garden experiments.
+Regardless, we have demonstrated a surprisingly strong effect of recent release from predation on a conspicuous, dynamic visual signal.
 
@@
 The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited colour vision [@RN22ms].
-In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators @Measey2014ms -- the conditions were highly favorable for character release of social signals.
+In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators [@Measey2014ms] -- the conditions were highly favorable for character release of social signals.
 Jackson’s chameleons likely experience strong sexual selection given their polygynous mating system, elaborate courtship displays and intense male contest competition (Fig.
@@
 We measured snout-vent length (SVL), a standard measure of body size in lizards, from the tip of the snout to the posterior edge of the vent using a plastic ruler to the nearest 1 mm.
-Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006.
+Field work was conducted on the island of Oahu, in Hawaii, during January-February 2006 and in Kenya during 5-13 April 2006.
 We did this work during what we expected to be the reproductive season, based on local knowledge.
@@
 2A, B).
-Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm [@RN8].Jackson’s chameleons are a generalist species which can be locally abundant and relatively mobile.
-This population is from the region (slopes of Mount Kenya) identified as the source population [@Tilbury2018].
+Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm [@RN8].
+Jackson’s chameleons are a generalist species which can be locally abundant and relatively mobile.
+This population is from the region (slopes of Mount Kenya) identified as the source population [@Tilbury2018] for the introduction to the Hawaiian Islands.
 
@@
 Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
-We erected a frame consisting of branches from the chameleons’ environment that were tied together in a triangle and which sat atop three vertical branches that formed a stand (Supplementary Fig.
+We erected a frame consisting of branches from the chameleons’ environment that were tied together in a horizontal trianglular perch and which sat atop three vertical branches that formed a stand (Supplementary Fig.
 3).
@@
 3).
-All trials were conducted outdoors, but in the shade under the cover of a roof, with vegetation nearby, to simulate their natural environment.
+All trials were conducted outdoors, but in the shade under the cover of a roof, surrounded by vegetation, to simulate their natural environment.
 The signaling environment was therefore the same for all animals.
@@
 
-During all behavior trials, we measured the colour of each of three body regions: top flank, mid-flank, tail base and gular region (Supplementary Fig.
+During all behavior trials, we measured the colour of each of four body regions: top flank, mid-flank, tail base and gular region (Supplementary Fig.
 5).
-We measured each body region once only, because chameleons change colour rapidly and we had to reduce the handling stress to the animal.
+We measured each body region once only, because chameleons change colour rapidly, and we had to reduce the handling stress to the animal.
 The four body regions were measured in random sequence to avoid any bias of order.
@@
 
-We visualized and quality-checked all spectra, adjusted very minor negative reflectance values caused by electrical noise by lifting the curves by the maximum negative value, then smoothed all spectra with LOESS smoothing of 0.16 using the plotsmooth function in `Pavo` vers.
+We visualized and quality-checked all spectra, adjusted very minor negative reflectance values caused by electrical noise by lifting the curves by the maximum negative value, then smoothed all spectra with LOESS smoothing of 0.16 using the plotsmooth function in *Pavo* vers.
 1.0 [@RN29].
@@
 
-We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package `pavo` [@RN29].
+We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package *pavo* [@RN29].
 This model assumes that colour discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
@@
 Chameleons have four spectral classes of single cone enabling tetrachromatic vision [@RN16].
-We used spectral sensitivities for the congeneric flap-necked chameleon (*Chameleo dilepis*) [@RN16] and a standard Weber fraction of 0.1 @RN17 to account for the receptor noise of the LWS photoreceptor.
+We used spectral sensitivities for the congeneric flap-necked chameleon (*Chameleo dilepis*) [@RN16] and a standard Weber fraction of 0.1 [@RN17] to account for the receptor noise of the LWS photoreceptor.
 We calculated noise for the other photoreceptor classes using the relative photoreceptor ratio of 1: 1.6: 3.3: 3.2 for the four single cones (UVS, SWS, MWS, and LWS, respectively).
@@
 
-Chromatic and luminance contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package [@RN24] in combination with the package `clubSandwich` [@RN25].
+Chromatic and luminance contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the *lme4* package [@RN24] in combination with the package *clubSandwich* [@RN25].
 For each individual, measurements for each body region (top flank, mid-flank, tail base and gular region) were contrasted against each of six average background spectra (i.e., average top and bottom of lighter leaves, average top and bottom of darker leaves, average green and average brown stems).
@@
 **Extended data figures and tables**
-The data and code are available at OSF.
+The data and code are available at OSF (https://osf.io/vxyp5/?view_only=7ef95ab5a95048cd8d9944fd6bee4d4c)
 
 **Supplementary information**
-This file includes supplementary notes, 9 Supplementary Figures and 13 Supplementary Tables.
+This file includes supplementary notes, 9 Supplementary Figures and 14 Supplementary Tables.
 

### 2022-02-02 ms_Nature.Rmd
ref fix

@@
 The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11ms].
-Jackson’s chameleons have a high intrinsic population growth rate (9-12 month generation time and live birth of up to 50 young @RN24ms), enabling their rapid establishment and potentially rapid evolution (approx.
+Jackson’s chameleons have a high intrinsic population growth rate (9-12 month generation time and live birth of up to 50 young [@RN24ms]), enabling their rapid establishment and potentially rapid evolution (approx.
 50 -- 65 generations).

### 2022-02-14 ms_Sci_Adv.Rmd
Renaming Nature files to Sci Advances. Making formatting changes for Sci Adv

--- 2022-02-02 ms_Nature.Rmd
@@
 
-**Authors** Martin J. Whiting^1,2^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel  W.A. Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
-
-**Affiliations**
+**Short title**: Color signal character release in chameleons
+
+**Authors**: Martin J. Whiting^1,2^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel  W.A. Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
+
+**Affiliations**:
 ^1^ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
@@
 # Abstract
-Colonization of novel environments can free invasive populations from the constraints imposed by native-range predators and promote rapid phenotypic change [@Losos2004; @RN4ms; @RN5ms; @Stastny2017].
-In particular, conspicuous social and sexual signals are predicted to experience significant character release when natural selection via predation is relaxed [@RN7ms; @RN4ms].
+Conspicuous social and sexual signals are predicted to experience significant character release when natural selection via predation is relaxed.
 However, we have few good examples of this phenomenon in the wild and none in species with dynamic colour change.
-Here, we show that Jackson’s chameleons (*Trioceros j.
-xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu), where there are no co-evolved, native lizard predators, experienced significant character release of colour signals.
+Here, we show that Jackson’s chameleons inadvertently introduced from Kenya to Hawaii (Oahu), where there are no co-evolved, native lizard predators, experienced significant character release of colour signals.
 Hawaiian chameleons displayed more conspicuous social colour signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator colour states.
@@
 
-# Main
+**One sentence summary**: Relaxing predation results in character release and local adaptation in a dynamic social signal in a chameleon.
+
+# INTRODUCTION
 When species encounter novel environments, such as during human-mediated introductions, phenotypic change through a combination of plasticity and evolution, can occur remarkably rapidly [@RN1ms; @RN2ms; @RN3ms; @RN4ms].
@@
 
-# Results and Discussion
+# RESULTS AND DISCUSSION
 In support of predictions of character release, male display colouration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.
@@
 
-# Methods
+# MATERIALS AND METHODS
 ## Study area and collection of chameleons
@@
 
-**Acknowledgments** 
+**Acknowledgments**: 
 This research was approved by the Kenya Wildlife Service (R.
@@
 Financial support was provided by the University of the Witwatersrand to MJW.
-We thank Idle Farah and Patrick Malonza of the National Museums of Kenya for facilitating the research, Joseph Gathua for field support and Jan Stipala for use of his field vehicle.
+We thank Idle Farah and Patrick Malonza of the National Museums of Kenya for facilitating the research, Joseph Gathua for amazing field support and Jan Stipala for use of his field vehicle.
 We thank the Ditsong National Museum of Natural History (formerly Transvaal Museum) for the stuffed African Cuckoo Hawk.
@@
 
-**Contributions**
+**Funding**: Financial support was provided by the University of the Witwatersrand to MJW.
+
+**Author contributions**
 MJW conceived the study.
@@
 
-**Ethics declarations**
-
-**Competing interests**
+**Competing interests**:
 We (the authors) declare no competing interests.
 
-**Extended data figures and tables**
+**Extended data figures and tables**:
 The data and code are available at OSF (https://osf.io/vxyp5/?view_only=7ef95ab5a95048cd8d9944fd6bee4d4c)
 
-**Supplementary information**
-This file includes supplementary notes, 9 Supplementary Figures and 14 Supplementary Tables.
+**Supplementary Materials**:
+This PDF file includes:
+Supplementary notes
+Figs. S1 to S9
+Tables S1 to S14
 

### 2022-02-14 supp_Sci_Adv.Rmd
Renaming Nature files to Sci Advances. Making formatting changes for Sci Adv

--- 2022-01-23 supp_Nature.Rmd
@@
 
-**The file includes**:
-
-Supplementary text  
-Figures 1 -- 9  
-Tables 1 -- 13  
-Supplementary References  
 <!-- Note to change the formatting of table and figure legends add a _bookdwon.yml file and manipulate the fig: and tab: settings.
@@
 Significant parameter estimates are those where the 95% CIs do not include zero.
-Relevant contrasts that compare Hawaiian and Kenyan populations both againts the same Kenyan background are indicated by '*'."}
+Relevant contrasts that compare Hawaiian and Kenyan populations both against the same Kenyan background are indicated by '*'."}
 source("./R/func.R")

### 2022-02-14 ms_Sci_Adv.Rmd
Adding new Sci Advances formatting

@@
 bibliography: supp_refs.bib
-csl: nature.csl
+csl: science-advances.csl
 output: 
@@
     reference_docx: template.docx
-  bookdown::html_document2:
-    code_folding: hide
-    number_sections: no
-    toc: yes
-    toc_depth: 6
-    toc_float: yes
-    bibliography: refs.bib
 editor_options: 
@@
 **Affiliations**:
-^1^ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia
+^1^ School of Natural Sciences, Macquarie University, Sydney, NSW 2109, Australia; martin.whiting@mq.edu.au
 
@@
 
+[@RN1; @RN2; @RN3; @RN4; @RN5]
 **Acknowledgments**: 

### 2022-02-14 supp_Sci_Adv.Rmd
Adding new Sci Advances formatting

@@
     reference_docx: template.docx
-  bookdown::html_document2:
-    code_folding: hide
-    number_sections: no
-    toc: yes
-    toc_depth: 6
-    toc_float: yes
 ---
@@
 Details here: https://community.rstudio.com/t/how-to-change-the-figure-table-caption-style-in-bookdown/110397-->
+
+**The file includes**:
+\n Supplementary text
+\n Figures S1 – S9
+\n Tables S1 – S14
+\n References 1-5
 

### 2022-02-15 ms_Sci_Adv.Rmd
colour -> color

@@
 ---
-title: "Title: Invasive chameleons released from predation display more conspicuous colours"
+title: "Title: Invasive chameleons released from predation display more conspicuous colors"
 bibliography: supp_refs.bib
@@
 Conspicuous social and sexual signals are predicted to experience significant character release when natural selection via predation is relaxed.
-However, we have few good examples of this phenomenon in the wild and none in species with dynamic colour change.
-Here, we show that Jackson’s chameleons inadvertently introduced from Kenya to Hawaii (Oahu), where there are no co-evolved, native lizard predators, experienced significant character release of colour signals.
-Hawaiian chameleons displayed more conspicuous social colour signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator colour states.
-Hawaiian chameleon display colours were also more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
+However, we have few good examples of this phenomenon in the wild and none in species with dynamic color change.
+Here, we show that Jackson’s chameleons inadvertently introduced from Kenya to Hawaii (Oahu), where there are no co-evolved, native lizard predators, experienced significant character release of color signals.
+Hawaiian chameleons displayed more conspicuous social color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators and showed greater change between display and anti-predator color states.
+Hawaiian chameleon display colors were also more conspicuous in their local than ancestral habitats, consistent with local adaptation of social signals.
 These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.
@@
 However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7ms; @RN9ms; @RN10ms].
-This is especially true for species with dynamic colour change, for which we are not aware of any examples.
-Here, we examine character release of a dynamic colour signal by taking advantage of an unintended evolutionary ‘experiment’: accidental translocation of chameleons to a novel, low predation environment.
+This is especially true for species with dynamic color change, for which we are not aware of any examples.
+Here, we examine character release of a dynamic color signal by taking advantage of an unintended evolutionary ‘experiment’: accidental translocation of chameleons to a novel, low predation environment.
 
@@
 
-Chameleons are famous for their rapid colour change [@RN12ms; @RN13ms; @RN14ms].
+Chameleons are famous for their rapid color change [@RN12ms; @RN13ms; @RN14ms].
 When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green during courtship or to signal dominance (Fig.
@@
 4a).
-Their different colour states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
-If introduced chameleons in Hawaii experience character release from reduced predation, we expect males to have more conspicuous display colours in response to conspecifics and to be less cryptic in response to bird and snake predators.
-We should also see evidence of local adaptation: colour signals should be more conspicuous to conspecifics against the local than ancestral background.
+Their different color states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
+If introduced chameleons in Hawaii experience character release from reduced predation, we expect males to have more conspicuous display colors in response to conspecifics and to be less cryptic in response to bird and snake predators.
+We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.
 
@@
 4), or a control (stick).
-We measured the spectral reflectance of male colour states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
+We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
 \@ref(fig:figure2) and Supplementary Fig.
 5).
-We then modelled how conspicuous these colours would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
+We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
 
 # RESULTS AND DISCUSSION
-In support of predictions of character release, male display colouration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.
+In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.
 \@ref(fig:figure2), Fig.
@@
 This result was consistent for all body regions and was driven by the increased luminance of Hawaiian chameleons, rather than differences in the background environment (i.e., Hawaiian chameleons were more conspicuous against either Hawaiian or Kenyan environment) (Supplementary Information and Supplementary Table 6).
-However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of colour signals (Fig.
+However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of color signals (Fig.
 \@ref(fig:figure3)b and Supplementary Table 2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
 
-The importance of luminance contrast is consistent with the nature and mechanism of rapid colour change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.
+The importance of luminance contrast is consistent with the nature and mechanism of rapid color change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.
 \@ref(fig:figure1)).
-Specifically, colour change is caused by dispersion or concentration of melanin within melanophore pigment cells [@RN15ms] or changes in the spacing of guanine crystals within iridophores [@RN14ms], both of which strongly affect luminance.
+Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells [@RN15ms] or changes in the spacing of guanine crystals within iridophores [@RN14ms], both of which strongly affect luminance.
 Accordingly, character release was observed in the overall luminance of signals rather than their hue.
@@
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18ms].
-The poor colour discrimination of snakes may also explain why Hawaiian and Kenyan chameleons differ little their chromatic response to snakes, despite the absence of snakes on Hawaii.
+The poor color discrimination of snakes may also explain why Hawaiian and Kenyan chameleons differ little their chromatic response to snakes, despite the absence of snakes on Hawaii.
 
@@
 
-To gauge the extent of colour plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) colour states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
+To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
 While there was evidence for differences between body regions (Supplementary Information, Supplementary Fig.
@@
 9).
-Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator colour states.
+Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
 An alternate explanation is that differences might be due to a founder effect because chameleons might have been selectively collected for the pet trade.
 While we acknowledge this possibility, it is exceedingly unlikely.
-Chameleons are almost always encountered in a camouflaged colour state because they are solitary and only use display colours during relatively brief social encounters which precludes the possibility that hunters selected the most brightly coloured individuals.
-Even if pet traders were selecting for larger animals, our analyses controlled for any relationship body size might have on colour.
+Chameleons are almost always encountered in a camouflaged color state because they are solitary and only use display colors during relatively brief social encounters which precludes the possibility that hunters selected the most brightly colored individuals.
+Even if pet traders were selecting for larger animals, our analyses controlled for any relationship body size might have on color.
 Finally, the traits we quantified, and their observed differences, were all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the environment.
 
-Differences in colouration between populations more broadly can be explained by environmental effects such as diet or climatic variables (e.g., temperature, rainfall/humidity).
-However, the measured colours and colour change represent a dynamic response to specific stimuli (conspecific, model predator) in contrast to the ‘fixed’ colour signals of many birds, fishes and lizards that have been shown to be affected by diet and other environmental factors.
-Additionally, the mechanisms underlying the rapid colour response of chameleons do not involve dietary pigments, and luminance change is partially structural [@RN14ms].
-Therefore, environmental effects (other than predation) are unlikely to account for major differences in the colour responses we observed between populations.
-The degree to which colour change capacity is plastic in chameleons and other organisms with dynamic colour change is unknown.
+Differences in coloration between populations more broadly can be explained by environmental effects such as diet or climatic variables (e.g., temperature, rainfall/humidity).
+However, the measured colors and color change represent a dynamic response to specific stimuli (conspecific, model predator) in contrast to the ‘fixed’ color signals of many birds, fishes and lizards that have been shown to be affected by diet and other environmental factors.
+Additionally, the mechanisms underlying the rapid color response of chameleons do not involve dietary pigments, and luminance change is partially structural [@RN14ms].
+Therefore, environmental effects (other than predation) are unlikely to account for major differences in the color responses we observed between populations.
+The degree to which color change capacity is plastic in chameleons and other organisms with dynamic color change is unknown.
 While we cannot exclude the role of phenotypic plasticity, plasticity can itself evolve and be an adaptive response.
@@
 Oahu has no snakes or raptors that feed on lizards.
-The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited colour vision [@RN22ms].
+The main threat for Oahu chameleons is domestic and feral cats, and cats have very limited color vision [@RN22ms].
 In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators [@Measey2014ms] -- the conditions were highly favorable for character release of social signals.
@@
 Consequently, a unique set of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.
-To the best of our knowledge, this is also the first example of character release in a dynamic colour signal.
-
-Rapid colour change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12ms].
-Our results suggest that even for species capable of dynamic colour change there is an upper threshold to signal intensity that is constrained by natural selection.
-This novel finding raises the question of whether other animals that have convergently evolved dynamic colour signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
+To the best of our knowledge, this is also the first example of character release in a dynamic color signal.
+
+Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12ms].
+Our results suggest that even for species capable of dynamic color change there is an upper threshold to signal intensity that is constrained by natural selection.
+This novel finding raises the question of whether other animals that have convergently evolved dynamic color signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained.
 More generally, our study highlights the opportunities provided by invasions to study natural and sexual selection in the wild.
@@
 ## Behavioral trials
-In order to quantify display colouration, we staged encounters between conspecifics (male-male and male-female) and model predators (bird and snake) and measured their subsequent display colouration (details below).
+In order to quantify display coloration, we staged encounters between conspecifics (male-male and male-female) and model predators (bird and snake) and measured their subsequent display coloration (details below).
 All chameleons were used in both predator and social trials although in some cases we were unable to quantify all social signals for a particular individual.
 We conducted all conspecific-social trials first, in case anti-predator trials unduly influenced their social interactions.
-Our analyses focused on comparing colour responses of chameleons to conspecifics and predators to ensure that the colour state corresponded to specific stimuli.
-We did not, and could not, measure a ‘neutral’ colour state because it is impossible to meaningfully gauge a neutral state in a colour-changing organism such as a chameleon.
+Our analyses focused on comparing color responses of chameleons to conspecifics and predators to ensure that the color state corresponded to specific stimuli.
+We did not, and could not, measure a ‘neutral’ color state because it is impossible to meaningfully gauge a neutral state in a color-changing organism such as a chameleon.
 However, for predator trials, we ensured that chameleons were specifically responding to the predators by conducting seven trials with a just a branch with no predator attached, as a control.
 Chameleons did not respond to the branch (see also @RN18ms).
-Furthermore, during trials with predators, chameleons focused on the predator itself, and exhibited classic anti-predator behavior (e.g., contrasting stress colour, body inflation, open-mouth threat).
+Furthermore, during trials with predators, chameleons focused on the predator itself, and exhibited classic anti-predator behavior (e.g., contrasting stress color, body inflation, open-mouth threat).
 
@@
 We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
-This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing colour and behavior (Supplementary Fig.
+This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Supplementary Fig.
 1A, B).
-In male-male trials both individuals would adopt a lemon-yellow display colouration and advance towards each other.
+In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
 This would be accompanied by head shakes and gapping, leading to horn-locking Supplementary Fig.
@@
 Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
-Once chameleons changed colour, we quantified their spectral reflectance (details below).
+Once chameleons changed color, we quantified their spectral reflectance (details below).
 The snake model was molded from a dead boomslang (*Dispholidus typus*) and painted by a professional artist to resemble a typical green form of this species (previously used in [@RN9;@RN10]).
@@
 
-## Colour measurement
-
-During all behavior trials, we measured the colour of each of four body regions: top flank, mid-flank, tail base and gular region (Supplementary Fig.
+## Color measurement
+
+During all behavior trials, we measured the color of each of four body regions: top flank, mid-flank, tail base and gular region (Supplementary Fig.
 5).
-We measured each body region once only, because chameleons change colour rapidly, and we had to reduce the handling stress to the animal.
+We measured each body region once only, because chameleons change color rapidly, and we had to reduce the handling stress to the animal.
 The four body regions were measured in random sequence to avoid any bias of order.
-When a chameleon had completely changed colour (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
+When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 spectrophotometer and a PX2 light source.
 We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal during measurement.
@@
 All measurements were relative to a dark, and a certified 99% white reflectance standard (Labsphere, North Sutton, NH, U.S.A.).
-If chameleons started to change colour in response to our handling, we returned them to the branch and allowed them to continue a social trial before measuring reflectance of remaining body regions.
+If chameleons started to change color in response to our handling, we returned them to the branch and allowed them to continue a social trial before measuring reflectance of remaining body regions.
 In the case of a predator trial, we would once again present them with the predator model until they returned to their anti-predator display.
@@
 
-We obtained spectral reflectance measurements of male dominant colour state in male-male contests for 34 individuals from Hawaii and 35 from Kenya, courtship colour state for 25 individuals from Hawaii and 32 from Kenya; colour state in response to the bird for 56 individuals from Hawaii and 23 from Kenya and colour state in response to the snake for 59 individuals from Hawaii and 38 from Kenya.
+We obtained spectral reflectance measurements of male dominant color state in male-male contests for 34 individuals from Hawaii and 35 from Kenya, courtship color state for 25 individuals from Hawaii and 32 from Kenya; color state in response to the bird for 56 individuals from Hawaii and 23 from Kenya and color state in response to the snake for 59 individuals from Hawaii and 38 from Kenya.
 
@@
 To reduce the number of backgrounds, we averaged the reflectance of the 20 darker measurements and 20 lighter measurements for each of the top leaf and bottom leaf measurements and averaged brown stems and green stems.
-This resulted in 6 average background spectra for each location, representing the range of background colours.
+This resulted in 6 average background spectra for each location, representing the range of background colors.
 
@@
 We modelled how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the receptor-noise limited (RNL) model [@RN27; @RN28] implemented in the R package *pavo* [@RN29].
-This model assumes that colour discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
-colour contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of colour discrimination for stimuli viewed simultaneously under ideal viewing conditions [@RN17].
+This model assumes that color discrimination is limited only by photoreceptor noise and does not account for potential effects of neural processing.
+Color contrasts are expressed in units of just-noticeable-differences (JND), where 1 JND is the theoretical threshold of color discrimination for stimuli viewed simultaneously under ideal viewing conditions [@RN17].
 Under natural conditions, discrimination thresholds are likely to be >1 and vary depending on the species and conditions [@RN12;@RN13; @RN14].
-Additionally, perceived conspicuousness may not scale linearly with colour distance (JNDs) [@RN15].
+Additionally, perceived conspicuousness may not scale linearly with color distance (JNDs) [@RN15].
 Nevertheless, the models provide a reasonable approximation of relative conspicuousness that can be compared between populations and trial types (social and predator).
 
-We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region and colour state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
+We calculated JNDs for chromatic and luminance contrast against each of the relevant backgrounds for each body region and color state, based on the visual system of the relevant receiver (chameleon, bird, snake; see below).
 We first calculated the quantum catch of each photoreceptor under standard daylight illumination (D65; Commission internationale de l’éclairage – CIE) using the ‘vismodel’ function.
@@
 First, we tested whether Hawaiian chameleons were more conspicuous (as perceived by chameleons) in social contexts than Kenyan chameleons.
-We ran separate linear models with chromatic and luminance contrast of male display or courtship colouration against the local background as the response variable.
+We ran separate linear models with chromatic and luminance contrast of male display or courtship coloration against the local background as the response variable.
 We included population of origin and z-transformed snout-vent length (SVL) as fixed effects.
@@
 
-```{r figure1, cache=FALSE, fig.cap = "**Chameleon colour signal change in response to different social stimuli.
+```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social stimuli.
 Male chameleons experience intense sexual selection**.
@@
 They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns.
-**a**, A dominant male in display colouration.
+**a**, A dominant male in display coloration.
 **b**, A subordinate male that lost a contest and turned from bright yellow to brown.
-**c**, Two males fighting, both are in display colouration and relatively even matched.
-**d**, A courting male in full display colour while the female has turned to a contrasting colour, rejecting the male.
+**c**, Two males fighting, both are in display coloration and relatively even matched.
+**d**, A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
 See Supplementary information for additional photos, including in response to a snake"}

### 2022-02-16 ms_Sci_Adv.Rmd
Added in the new analysis, which was in the ms.Rmd script...should have been in here too. Also updated everything with Devi's changes

@@
 
+#########################################################***** NEW
+# Background Analysis - Founder           
+#########################################################***** NEW
+DisplayBgrd <- read.table("./data/Display_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+svl_dat <- read.csv("./data/SVL2.csv")
+
+DisplayBgrd <- DisplayBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, 
+                                               if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>%
+                 as.data.frame() #exclude bottom flank.
+
+DisplayBgrd$BodyRegion <- factor(DisplayBgrd$BodyRegion) #updates levels for the variable BodyRegion
+
+DisplayBgrd <- DisplayBgrd %>% 
+                mutate(pop_background = interaction(Population, Background_pop)) %>%  
+                left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+                mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
+###################################
+# Male-male display
+###################################
+##########
+# Achromatic Contrast - dL
+##########
+  # Fit the interaction model
+  DispBgrdAC_s6 <- lmer(dL ~ sc_svl  + Population*Background_pop + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+  F_AC_Disp_pop_bgk <- Wald_test(DispBgrdAC_s6, constraints = constrain_zero(5), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
+  DispBgrdAC_s6 <- lmer(dL ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=DisplayBgrd)
+  summary(DispBgrdAC_s6)
+
+  # Check robustness of results to non-independence using Robust Variance Estimator (RVE)
+  robust_localAdapt_DispBgrdAC_s6 <- clubSandwich::conf_int(DispBgrdAC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+##########
+# Chromatic Contrast - dS
+##########
+    # Fit the interaction  model
+      DispBgrdCC_s6 <- lmer(dS ~ sc_svl + Population*Background_pop  + (1|ID), data=DisplayBgrd)
+      summary(DispBgrdCC_s6)
+    clubSandwich::conf_int(DispBgrdCC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+    # Interaction test.
+Does the difference between Hawaii and Kenya depend on the background.
+It shouldn't.
+Wald F-test.
+Although, somthing might be up with this one - effect size tiny but goes from non-sig to signficant.
+      F_CC_Disp_pop_bgk <- Wald_test(DispBgrdCC_s6, constraints = constrain_zero(5), 
+                                  cluster = DisplayBgrd$ID, vcov = "CR2")
+
+      # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
+    DispBgrdCC_s6 <- lmer(dS ~ sc_svl + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=DisplayBgrd)
+          summary(DispBgrdCC_s6)
+
+    # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+      robust_DispBgrdCC_s6 <- clubSandwich::conf_int(DispBgrdCC_s6, cluster = DisplayBgrd$ID, vcov =  "CR2")
+
+###################
+# Social Context: Courtship
+##################
+CourtshipBgrd <- read.table("./data/Courtship_5.csv", 
+                        header=TRUE, sep=",", dec=".", strip.white=TRUE)
+
+CourtshipBgrd <- CourtshipBgrd %>% 
+                filter(!BodyRegion == "botflank") %>% 
+                mutate(Background_pop = ifelse(Background2 == "own", Population, if_else(Background2 == "other" & Population == "Hawaii", "Kenya", "Hawaii"))) %>% 
+                as.data.frame() #exclude bottom flank.
+CourtshipBgrd$BodyRegion <- factor(CourtshipBgrd$BodyRegion)
+
+# Filter out only Hawaiian and add in SVL for IDs
+CourtshipBgrd <- CourtshipBgrd %>% 
+               mutate(pop_background = interaction(Population, Background_pop)) %>%  
+                left_join(select(svl_dat, ID, SVL), by = "ID", keep = FALSE) %>% 
+               mutate(sc_svl = scale(SVL, scale = FALSE)) %>% 
+                filter(!is.na(SVL))
+
+##########
+## Achromatic Contrast - dL
+##########
+
+  # Fit interaction model
+    CourtshipAC_s6 <- lmer(dL ~ sc_svl  + Population*Background_pop  + (1|ID), data=CourtshipBgrd)
+    summary(CourtshipAC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+      F_AC_Court_pop_bgk <- Wald_test(CourtshipAC_s6, constraints = constrain_zero(5), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
+    CourtshipAC_s6 <- lmer(dL ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=CourtshipBgrd)
+    summary(CourtshipAC_s6)
+
+   # Check robustness of results to non-indepenedence using Robust Variance Estimator (RVE)
+   robust_CourtshipAC_s6 <- clubSandwich::conf_int(CourtshipAC_s6, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
+##########
+# Chromatic - dS
+##########
+
+  # Fit main effect model
+      CourtshipCC_s6 <- lmer(dS ~ sc_svl  + Population*Background_pop  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC_s6)
+
+  # Interaction test. Does the difference between Hawaii and Kenya depend on the background. It shouldn't. Wald F-test
+      F_CC_Court_pop_bgk <- Wald_test(CourtshipCC_s6, constraints = constrain_zero(5), 
+                                  cluster = CourtshipBgrd$ID, vcov = "CR2")
+
+  # Reparameterise the interaction model to generate the contrasts. It's not surprising that there is no interaction because we are literatlly just shifting the environment. 
+      CourtshipCC_s6 <- lmer(dS ~ sc_svl  + relevel(pop_background, ref = "Hawaii.Kenya")  + (1|ID), data=CourtshipBgrd)
+      summary(CourtshipCC_s6)
+
+  # Robust variance esatimation to correct SE's for fixed effects given spectral curves are used multiple times for generating JNDs for each individuals data. 
+  robust_CourtshipCC_s6 <- clubSandwich::conf_int(CourtshipCC_s6, cluster = CourtshipBgrd$ID, vcov =  "CR2")
+
 ```
@@
 This is especially true for species with dynamic color change, for which we are not aware of any examples.
-Here, we examine character release of a dynamic color signal by taking advantage of an unintended evolutionary ‘experiment’: accidental translocation of chameleons to a novel, low predation environment.
+Here, we examine character release of a dynamic color signal by taking advantage of an unintended 'evolutionary experiment’: accidental translocation of chameleons to a novel, low predation environment.
 
 In 1972, a shipment of Jackson’s chameleons (*Trioceros jacksonii xanthopholus*; Fig.
-\@ref(fig:figure1) and Supplementary Fig.
-1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
+\@ref(fig:figure1) and Fig.
+S1; thought to be 36 or fewer individuals) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade.
 The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11ms].
@@
 When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green during courtship or to signal dominance (Fig.
-1, Supplementary Fig.
-1 and 3).
+1, Fig.
+S1 and S3).
 At other times, they are a duller green, brown or become mottled green-brown, particularly in response to predators (Fig.
-\@ref(fig:figure1) and Supplementary Fig.
-4a).
+\@ref(fig:figure1) and Fig.
+S4a).
 Their different color states vary in conspicuousness depending on the context, background, and viewer (e.g., conspecific or predator).
@@
 
-To test these predictions, we presented wild-caught adult male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Supplementary Fig.
-4), or a control (stick).
+To test these predictions, we presented wild-caught adult male chameleons in both Hawaii and the source population in Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Fig.
+S4), or a control (stick).
 We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, *n* = `r N_display[1,2]` Hawaii, *n* = `r N_display[2,2]` Kenya), courtship (*n* = `r N_Courtship[1,2]` Hawaii, *n* = `r N_Courtship[2,2]` Kenya) and in response to bird (*n* = `r N_Pred[1,3]` Hawaii, *n* = `r N_Pred[2,3]` Kenya) and snake (*n* = `r N_Pred[3,3]` Hawaii, *n* = `r N_Pred[4,3]` Kenya) predators (Fig.
-\@ref(fig:figure2) and Supplementary Fig.
-5).
+\@ref(fig:figure2) and Fig.
+S5).
 We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
+
+F_AC_Disp_pop_bgk # Lum object male-male display
+
+F_CC_Disp_pop_bgk # Chrom object male-male display
+
+F_AC_Court_pop_bgk $ Lum object social
+
+F_CC_Court_pop_bgk  # Chromatic object social
 
@@
 \@ref(fig:figure2), Fig.
-\@ref(fig:figure3)A and Supplementary Table 2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
-This result was consistent for all body regions and was driven by the increased luminance of Hawaiian chameleons, rather than differences in the background environment (i.e., Hawaiian chameleons were more conspicuous against either Hawaiian or Kenyan environment) (Supplementary Information and Supplementary Table 6).
+\@ref(fig:figure3)A and Table S2) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, *P* = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipAC$p_val)`).
+This result was consistent for all body regions and was driven by the increased luminance of Hawaiian chameleons, rather than differences in the background environment.
+Hawaiian chameleons were more conspicuous against either Hawaiian or Kenyan environment (i.e., no population by background interaction effect for luminance: male-male contests: luminance contrast ($dl$); $F_{`r F_AC_Disp_pop_bgk$df_num`,`r F_AC_Disp_pop_bgk$df_denom`}$ = `r F_AC_Disp_pop_bgk$Fstat`, *P* = `r p_value(F_AC_Disp_pop_bgk$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r F_AC_Court_pop_bgk$df_num`,`r F_AC_Court_pop_bgk$df_denom`}$ = `r F_AC_Court_pop_bgk$Fstat`, *P* = `r p_value(F_AC_Court_pop_bgk$p_val)`; See also Supplementary Information and Table S6).
 However, Hawaiian chameleons did not differ from Kenyan chameleons in chromatic contrast of color signals (Fig.
-\@ref(fig:figure3)b and Supplementary Table 2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
+\@ref(fig:figure3)b and Table S2, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, *P* = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, *P* = `r p_value(Population_ConspicCourtshipCC$p_val)`).
 
@@
 Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.
-\@ref(fig:figure3)c; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, *P* = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, Supplementary Table 3).
-Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, Supplementary Table 3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, *P* = `r p_value(back_pred_CC_Pred$p_val)`).
+\@ref(fig:figure3)c; main effect of population; $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, *P* = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type, Table S3).
+Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction, Table s3: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, *P* = `r p_value(back_pred_CC_Pred$p_val)`).
 Hawaiian chameleons had higher chromatic contrast to birds than were Kenyan conspecifics (Fig.
-\@ref(fig:figure3)d and Supplementary Table 4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`).
+\@ref(fig:figure3)d and Table s4; Birds (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, *P* = `r CI_hyp_test[2,8]`).
 By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.
-\@ref(fig:figure3)d and Supplementary Table 4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1).
+\@ref(fig:figure3)d and Table S4; Snakes (Hawaii-Kenya): $\beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, *P* = 1).
 The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18ms].
@@
 Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, although effect sizes for luminance were much larger than for chromatic contrast (Fig.
-\@ref(fig:figure4) and Supplementary Tables 5 and 6; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,2]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,5]` to `r robust_localAdapt_DispBgrdAC[3,6]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,2]`, 95% CI = `r robust_DispBgrdCC[3,5]` to `r robust_DispBgrdCC[3,6]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,2]`, 95% CI = `r robust_CourtshipAC[3,5]` to `r robust_CourtshipAC[3,6]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,2]`, 95% CI = `r robust_CourtshipCC[3,5]` to `r robust_CourtshipCC[3,6]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
+\@ref(fig:figure4) and Tables S5 and S6; male-male contests: luminance ($dl$), $\beta$ = `r robust_localAdapt_DispBgrdAC[3,2]`, 95% CI = `r robust_localAdapt_DispBgrdAC[3,5]` to `r robust_localAdapt_DispBgrdAC[3,6]`, $F_{`r background_AC_Disp$df_num`,`r background_AC_Disp$df_denom`}$ = `r background_AC_Disp$Fstat`, *P* = `r p_value(background_AC_Disp[,6])`, chromatic ($dS$), $\beta$ = `r robust_DispBgrdCC[3,2]`, 95% CI = `r robust_DispBgrdCC[3,5]` to `r robust_DispBgrdCC[3,6]`, $F_{`r background_CC_Disp$df_num`,`r background_CC_Disp$df_denom`}$ = `r background_CC_Disp$Fstat`, *P* = `r p_value(background_CC_Disp[,6])`; courtship: luminance ($dl$), $\beta$ = `r robust_CourtshipAC[3,2]`, 95% CI = `r robust_CourtshipAC[3,5]` to `r robust_CourtshipAC[3,6]`, $F_{`r background_AC_Court$df_num`,`r background_AC_Court$df_denom`}$ = `r background_AC_Court$Fstat`, *P* = `r p_value(background_AC_Court[,6])`, chromatic ($dS$) $\beta$ = `r robust_CourtshipCC[3,2]`, 95% CI = `r robust_CourtshipCC[3,5]` to `r robust_CourtshipCC[3,6]`, $F_{`r background_CC_Court$df_num`,`r background_CC_Court$df_denom`}$ = `r background_CC_Court$Fstat`, *P* = `r p_value(background_CC_Court[,6]) `).
 These results are consistent with local adaptation, in which phenotypic plasticity, which can also evolve rapidly [@RN26ms], may also play a role.
@@
 To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states (absolute difference in JNDs between states; *n* = `r N_diff[1,2]` Hawaii, *n* = `r N_diff[2,2]` Kenya).
-While there was evidence for differences between body regions (Supplementary Information, Supplementary Fig.
-9, Supplementary Tables 13 and 14), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (Supplementary Fig.
-9).
+While there was evidence for differences between body regions (Supplementary Information, Fig.
+S9, Tables S13 and S14), overall, Hawaiian chameleons showed greater chromatic but not luminance change compared to Kenyan chameleons (chromatic: $F_{`r robust_Diff_dS_pop$df_num`,`r robust_Diff_dS_pop$df_denom`}$ = `r robust_Diff_dS_pop$Fstat`, *P* = `r p_value(robust_Diff_dS_pop[,6])`; luminance: $F_{`r robust_Diff_dl_pop$df_num`,`r robust_Diff_dl_pop$df_denom`}$ = `r robust_Diff_dl_pop$Fstat`, *P* = `r p_value(robust_Diff_dl_pop[,6])`); although luminance change showed similar trends across body regions (Fig.
+S9).
 Together, these results indicate that chameleons introduced to Hawaii only 50 years ago have evolved signals that are locally adapted, more conspicuous to conspecifics, and show greater changes between display and anti-predator color states.
-An alternate explanation is that differences might be due to a founder effect because chameleons might have been selectively collected for the pet trade.
-While we acknowledge this possibility, it is exceedingly unlikely.
-Chameleons are almost always encountered in a camouflaged color state because they are solitary and only use display colors during relatively brief social encounters which precludes the possibility that hunters selected the most brightly colored individuals.
+An alternative explanation is that differences might be due to a founder effect.
+Founder effects are expected to be random with respect to the environment unless sampling is deliberately biased; for example, selective collection of chameleons for the pet trade.
+The traits we quantified, and their observed differences, were all in the predicted direction of adaptive change, arguing against a random founder effect.
+The possibility that hunters selected the most brightly colored individuals is similarly unlikely because chameleons are almost always encountered in a camouflaged color state; they are solitary and only use display colors during relatively brief social encounters.
 Even if pet traders were selecting for larger animals, our analyses controlled for any relationship body size might have on color.
-Finally, the traits we quantified, and their observed differences, were all in the predicted direction of adaptive change, whereas founder effects are expected to be random with respect to the environment.
-
-Differences in coloration between populations more broadly can be explained by environmental effects such as diet or climatic variables (e.g., temperature, rainfall/humidity).
+Thus, a founder effect, due to random or deliberately biased sampling, is unlikely to explain phenotypic differences consistent with character release in chameleons.
+
+Differences in coloration between populations could be due to environmental effects such as diet or climatic variables (e.g., temperature, rainfall/humidity).
 However, the measured colors and color change represent a dynamic response to specific stimuli (conspecific, model predator) in contrast to the ‘fixed’ color signals of many birds, fishes and lizards that have been shown to be affected by diet and other environmental factors.
-Additionally, the mechanisms underlying the rapid color response of chameleons do not involve dietary pigments, and luminance change is partially structural [@RN14ms].
+Additionally, the mechanisms underlying the rapid color response of chameleons are not known to dietary pigments, and luminance change is partially structural [@RN14ms].
 Therefore, environmental effects (other than predation) are unlikely to account for major differences in the color responses we observed between populations.
 The degree to which color change capacity is plastic in chameleons and other organisms with dynamic color change is unknown.
-While we cannot exclude the role of phenotypic plasticity, plasticity can itself evolve and be an adaptive response.
-The population differences we observed are likely an interaction between genetic and environmental factors which are difficult to disentangle, even with common garden experiments.
+While we cannot exclude the role of phenotypic plasticity, plasticity can itself evolve and be an adaptive response [@RN26ms].
+The population differences we observed are likely an interaction between genetic and environmental factors which are difficult to disentangle, even with common garden experiments [@Heerwaarden].
 Regardless, we have demonstrated a surprisingly strong effect of recent release from predation on a conspicuous, dynamic visual signal.
@@
 
-We collected most chameleons (Supplementary Fig.
-1) at night by spotlighting, and a small proportion of individuals during the day.
-When chameleons were beyond reach in the forest canopy, we used lightweight 6-m extensible graphite composite poles (South Bend Kwik Stix™ 20' Telescopic Fishing Pole), that chameleons could be encouraged to step on to, before lowering them to the ground.
-We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
-We measured snout-vent length (SVL), a standard measure of body size in lizards, from the tip of the snout to the posterior edge of the vent using a plastic ruler to the nearest 1 mm.
-Field work was conducted on the island of Oahu, in Hawaii, during January-February 2006 and in Kenya during 5-13 April 2006.
-We did this work during what we expected to be the reproductive season, based on local knowledge.
-The timing is different because of seasonal/geographic differences between Kenya and Hawaii.
-Importantly, the chameleons all exhibited very strong behavioral responses during courtship and male-male competition, consistent with reproductive behavior.
+We conducted experiments on chameleons in Hawaii and the likely source population in Kenya.
 In Hawaii, chameleons were collected from a single population in the forested Ko’olau Mountains north of Honolulu, between upper Makiki Valley and the summit of Tantalus Mountain, Oahu, accessible along Tantalus and Roundtop Drives at an elevation of ca.
 400-600 m.
-The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth  (Supplementary Fig.
-2C, D).
+The vegetation in this area consisted of closed-canopy, mid-elevation tropical rain forest, with extensive vines and undergrowth (Fig.
+S2C, D).
 Most of the plants in this area were non-native, and chameleons also occurred in hedge-grows and stands of bamboo.
@@
 Kenya, close to the town of Runyenjes.
+This population is from the region (slopes of Mount Kenya) identified as the source population [@Tilbury2018] for the introduction to the Hawaiian Islands.
 The area is partially cleared for agriculture, leaving small stands of forest.
-The vegetation was a mix of native trees and shrubs interspersed with exotic species (Supplementary Fig.
-2A, B).
+The vegetation was a mix of native trees and shrubs interspersed with exotic species (Fig.
+S2A, B).
 Mean annual temperature is 20.2°C (mean max: 24.8°C) and the mean annual rainfall is 121 mm [@RN8].
-Jackson’s chameleons are a generalist species which can be locally abundant and relatively mobile.
-This population is from the region (slopes of Mount Kenya) identified as the source population [@Tilbury2018] for the introduction to the Hawaiian Islands.
+
+We collected most chameleons (Fig.
+S1) at night by spotlighting, and a small proportion of individuals during the day.
+When chameleons were beyond reach in the forest canopy, we used lightweight 6-m extensible graphite composite poles (South Bend Kwik Stix™ 20’ Telescopic Fishing Pole), that chameleons could be encouraged to step on to, before lowering them to the ground.
+We then placed them in cotton bags with vegetation to cling to, before returning them to the lab.
+We measured snout-vent length (SVL), a standard measure of body size in lizards, from the tip of the snout to the posterior edge of the vent using a plastic ruler to the nearest 1 mm.
+Field work was conducted in Hawaii during January-February 2006 and in Kenya during 5-13 April 2006, when chameleons are expected to show reproductive behavior.
+The timing was different because of seasonal/geographic differences between Kenya and Hawaii.
+Importantly, the chameleons all exhibited very strong behavioral responses during courtship and male-male competition, consistent with reproductive behavior.
 
@@
 Behavioral trials were conducted during the chameleons’ natural activity period (09:00-16:00 h) beginning the day following capture and continuing over a 2-3 day period.
-We erected a frame consisting of branches from the chameleons’ environment that were tied together in a horizontal trianglular perch and which sat atop three vertical branches that formed a stand (Supplementary Fig.
-3).
+We erected a frame consisting of branches from the chameleons’ environment that were tied together in a horizontal trianglular perch and which sat atop three vertical branches that formed a stand (Fig.
+S3).
 This structure was about 1.5 m high and each arm was approximately 60 cm.
-The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Supplementary Fig.
-3).
+The triangular setup allowed chameleons to display at a comfortable distance, and if necessary, to escape from aggressive interactions, particularly during male-male trials (Fig.
+S3).
 All trials were conducted outdoors, but in the shade under the cover of a roof, surrounded by vegetation, to simulate their natural environment.
@@
 We staged trials between males to elicit dominant and subordinate displays and between males and females to elicit courtship displays.
-This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Supplementary Fig.
-1A, B).
+This consisted of placing two chameleons at a comfortable distance at which point they typically responded to the conspecific’s presence by rapidly changing color and behavior (Fig.
+S1A, B).
 In male-male trials both individuals would adopt a lemon-yellow display coloration and advance towards each other.
-This would be accompanied by head shakes and gapping, leading to horn-locking Supplementary Fig.
-1B) if males were not obviously asymmetric.
-Once dominance was settled, the subordinate would turn brown (Supplementary Fig.
-1B and Supplementary Fig.
-3) and flee.
-In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Supplementary Fig.
-1C).
-
-The day following social trials we presented each chameleon with a snake (Supplementary Fig.
-4A) and bird (Supplementary Fig.
-4B) model predator model (separately).
+This would be accompanied by head shakes and gapping, leading to horn-locking Fig.
+S1B) if males were not obviously asymmetric.
+Once dominance was settled, the subordinate would turn brown (Fig.
+S1B and Fig.
+S3) and flee.
+In male-female (courtship) trials, males would turn the same lemon-yellow and approach the female while head shaking and rocking (Fig.
+S1C).
+
+The day following social trials we presented each chameleon with a snake (Fig.
+S4A) and bird (Fig.
+S4B) model predator model (separately).
 Lizards were drawn at random and the order of presentation (snake or bird first) was random.
@@
 A trial began when a chameleon was removed from its cloth bag and placed on the triangular stand.
-In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Supplementary Fig.
-4A).
+In the case of the snake model, one of us slowly moved the model snake towards the chameleon from both above and below it, without ever contacting the chameleon, in a standardized fashion (Fig.
+S4A).
 In the case of the bird trials, the bird was attached to an ca.
-2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Supplementary Fig.
-4B).
+2 m wooden pole and was ‘flown’ past the chameleon from above and below it (Fig.
+S4B).
 Trials lasted no more than 2-min and we scored whether the chameleon flipped to the opposite side of the branch, which is a classic anti-predator behavior by chameleons.
@@
 
-During all behavior trials, we measured the color of each of four body regions: top flank, mid-flank, tail base and gular region (Supplementary Fig.
-5).
+During all behavior trials, we measured the color of each of four body regions: top flank, mid-flank, tail base and gular region (Fig.
+S5).
 We measured each body region once only, because chameleons change color rapidly, and we had to reduce the handling stress to the animal.
@@
 We calculated noise for the other photoreceptor classes using the relative photoreceptor ratio of 1: 1.6: 3.3: 3.2 for the four single cones (UVS, SWS, MWS, and LWS, respectively).
-*Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Supplementary Fig.
-6).
+*Chameleo dilepis* and other chameleons appear to have two populations of MWS cones ($\lambda_{max}$ 481 nm, 497 nm) and three populations of LWS cones ($\lambda_{max}$ 568 nm, 584 nm, 605 nm) so we used a combined sensitivity function for the MWS and LWS cones (Fig.
+S6).
 
@@
 A Weber fraction of 0.06 was used for the LWS photoreceptor based on the most recent empirical measurement [@RN20].
-We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds [@RN21] (Supplementary Fig.
-6B).
+We used a relative photoreceptor ratio of 1:2:3.4:3 for the UVS, SWS, MWS, and LWS photoreceptors respectively, which represent average values for UVS birds [@RN21] (Fig.
+S6B).
 
 For snakes, we assumed a trichromatic visual system characteristic of active diurnal snake species, represented by the spectral sensitivities for the garter snake [@RN22].
-We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively [@RN22] (Supplementary Fig.
-6C).
+We used a Weber fraction of 0.1 for the LWS photoreceptor and a relative photoreceptor ratio of 1:1.6:7.3 for the UVS, SWS, MWS, and LWS photoreceptors, respectively [@RN22] (Fig.
+S6C).
 
@@
 We included population of origin and z-transformed snout-vent length (SVL) as fixed effects.
-Next, we tested whether Kenyan chameleons were more cryptic (less conspicuous against the local background) in response to predators.
+To determine whether differences in conspicuousness were driven by differences in the colour of chameleons or the background, we also fit models that compared the luminance and chromatic contrast of Hawaiian and Kenyan lizards against a Hawaiian and Kenyan backgrounds and examined the interaction between population and background.
+
+Second, we tested whether Kenyan chameleons were more cryptic (less conspicuous against the local background) in response to predators.
 For these models, chromatic or luminance contrast against the local background (as perceived by the corresponding predator type) was the response variable and predator type (i.e, snake or bird), population of origin (Hawaii or Kenya) and their interaction were fixed effects.
@@
 For local adaptation, signals are predicted to be more conspicuous against their own background.
-In this case, Hawaiian chameleons were both significantly more chromatic and brighter (luminance contrast) against their own background compared to the Kenyan background suggesting that selection has acted on increasing local signal conspicuousness."}
+In this case, Hawaiian chameleons were both significantly more chromatic and brighter (luminance contrast) against their own background compared to the Kenyan background indicating increased local signal conspicuousness."}
 rerun=FALSE

### 2022-02-22 ms_Sci_Adv.Rmd
MS updated removing random text
Fixed up formatting issues

@@
 
-**Authors**: Martin J. Whiting^1,2^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel  W.A. Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
+**Authors**: Martin J. Whiting^1,2,*^, Brenden S. Holland^3^, J. Scott Keogh^4^, Daniel  W.A. Noble^4^, Katrina J. Rankin^5^ and Devi Stuart-Fox^5^
 
@@
 We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs; @RN25ms) against the background vegetation of each environment (Hawaii or Kenya; Supplementary Information).
-
-F_AC_Disp_pop_bgk # Lum object male-male display
-
-F_CC_Disp_pop_bgk # Chrom object male-male display
-
-F_AC_Court_pop_bgk $ Lum object social
-
-F_CC_Court_pop_bgk  # Chromatic object social
 
@@
 **d**, A courting male in full display color while the female has turned to a contrasting color, rejecting the male.
-See Supplementary information for additional photos, including in response to a snake"}
+See Supplementary information for additional photos, including in response to a snake.
+Photo credit: Martin Whiting, Macquarie University."}
 library(magick)
@@
 More details in online Supplementary Information and Supplementary Fig.
-5"}
+5.
+Photo credit: Martin Whiting, Macquarie University."}
 library(magick)
@@
 JNDs are calculated based on the snake and bird visual systems.
-Contrasts between means, ‘Beta’, that are bold indicate significant effects."}
+Contrasts between means, ‘Beta’, that are bold indicate significant effects.
+Photo credit: Martin Whiting, Macquarie University."}
 rerun = FALSE
@@
 For local adaptation, signals are predicted to be more conspicuous against their own background.
-In this case, Hawaiian chameleons were both significantly more chromatic and brighter (luminance contrast) against their own background compared to the Kenyan background indicating increased local signal conspicuousness."}
+In this case, Hawaiian chameleons were both significantly more chromatic and brighter (luminance contrast) against their own background compared to the Kenyan background indicating increased local signal conspicuousness.
+Photo credit: Martin Whiting, Macquarie University."}
 rerun=FALSE

### 2022-02-22 supp.Rmd
Fixed up formatting issues

@@
 **(C)** Conspicuous yellow-green male courtship color.
-When females reject males, they develop a heavily contrasting pattern."}
+When females reject males, they develop a heavily contrasting pattern.
+Photo credit: Martin Whiting, Macquarie University."}
 figs1 <- image_read("./output/figures/FigS1.png")
@@
 ```{r Figcolsig2, fig.align = 'center', fig.cap = "**Representative photos of habitat in Kenya (A, B) and Hawaii (C, D)**.
-In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines."}
+In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines.
+Photo credit: Martin Whiting, Macquarie University."}
 figs2 <- image_read("./output/figures/FigS2.jpg")
@@
 We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male.
-The male on the left won this contest and is in full display color while the male on the right has lost this contest and turned brown (subordinate)."}
+The male on the left won this contest and is in full display color while the male on the right has lost this contest and turned brown (subordinate).
+Photo credit: Martin Whiting, Macquarie University."}
 figs3 <- image_read("./output/figures/FigS3.jpg")
@@
 ```{r Figcolsig4, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**We presented chameleons with model snake and bird predators to elicit an anti-predator response, particularly, color change, to test the hypothesis that character release in Hawaii would also result in a signal more conspicuous to a predator**.
-**(A)** A model snake (*Dispholidus typus*) causing a characteristic contrasting color change and **(B)** a model bird (*Aviceda cuculoides*) which was flown over the top of the chameleon."}
+**(A)** A model snake (*Dispholidus typus*) causing a characteristic contrasting color change and **(B)** a model bird (*Aviceda cuculoides*) which was flown over the top of the chameleon.
+Photo credit: Martin Whiting, Macquarie University."}
 figs4 <- image_read("./output/figures/FigS4.jpg")
@@
 
-```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Location of body regions we measured for spectral reflectance**. See details in text."}
+```{r Figcolsig5, fig.align = 'center', fig.width = 12, fig.height = 12, fig.cap = "**Location of body regions we measured for spectral reflectance**.
+See details in text.
+Photo credit: Martin Whiting, Macquarie University."}
 figs5 <- image_read("./output/figures/FigS5.jpg")
